### 2017-01-11 AGU_Journal/ms_Geo_Let.tex
River-Network Climate Portfolio: Initial Push
Here I add the code for my PhD thesis chapter 1 and two rejected
manuscripts.

\documentclass[linenumbers,draft]{AGUJournal}
\usepackage{gensymb}
\usepackage{textcomp}
\usepackage{upgreek}
\usepackage{natbib}
\usepackage[parfill]{parskip}
\journalname{Geophysical Research Letters}
\begin{document}
\title{River networks dampen long-term hydrological signals of climate change}
\authors{K. A. Chezik\affil{1},
	S. C. Anderson\affil{2},
J. W. Moore\affil{1}}

\affiliation{1}{Earth to Ocean Research Group, Department of Biological Sciences, Simon Fraser University, 8888 University Dr., Burnaby, British Columbia V5A 1S6, Canada}
\affiliation{2}{School of Aquatic and Fishery Sciences, University of Washington, Box 455020, Seattle, WA 98195, USA}

\correspondingauthor{K. A. Chezik}{kchezik@sfu.ca}

\begin{keypoints}
\item River networks act as diversified portfolios of climate; by integrating varied climate signals the network substantially dampens downstream climate-driven hydrologic trends.
\item River network climate dampening varied seasonally in the Fraser River basin, declining sharply with the spring freshet, suggesting that basin-wide events homogenize the climate portfolio, thereby degrading the network dampening effect.
\item Rivers that integrate a diverse climate portfolio offer climate change relief to riverine biota that is many times greater than homogenized systems. Leveraging network dampening properties by promoting land use practices that maintain or enhance climate portfolio diversity may be a powerful tool in mitigating climate change impacts. 
\end{keypoints}

\begin{abstract}
River networks may dampen local hydrologic climate change by aggregating a diverse climate portfolio with variable upstream climate signals. Here we examine this hypothesis using flow trend estimates (1970--2007) at 55 hydrometric gauge stations within the Fraser River watershed in British Columbia, Canada. The Fraser River dampened variability in downstream long-term median discharge by 95\% from between -6 and 18\% per-decade among small watersheds to between 0.8 and 1.6\% per-decade near the confluence. A null model approach indicated that dampening of annual median flow trends within the Fraser River was typically 5.4 times greater than would be expected of a watershed exhibiting a consistent basin-wide climate response. Although climate dampening strength varied seasonally, our findings indicate that large free-flowing rivers offer a powerful and largely unappreciated process of climate change mitigation. River networks that integrate a diverse climate portfolio can dampen local extremes and offer climate relief to riverine biota.
\end{abstract}

\section{Introduction} 
As billions are spent adapting to climate change by such means as building dikes or improving stormwater drainage \citep{Narain:2011}, certain habitats and processes provide a natural defense system that mitigates climate change impacts \citep{Jones:2012}. For example, coastal habitats such as oyster reefs, mangrove forests, and eelgrass beds shield 67\% of the United States coastline and 1.4 million people from sea level rise and storms \citep{Arkema:2013}. Protection and enhancement of these natural systems has become a key component of many climate mitigation strategies \citep{Guerry:2015}, as they return multiple cost-effective services in contrast to engineered alternatives \citep{Jones:2012}. For instance, revitalization of the Mississippi delta not only improves storm protection, but also increases production and sustainability of fisheries and wildlife resources, protects the water supply, and reduces the impact of wastewater effluent \citep{Conservation:1998}. The value of climate buffering habitats stands only to increase as average global temperatures climb. Recent international agreements in Paris aim to limit global temperature rise to 1.5\degree{C} \citep{Hulme:2016}, a goal that concedes a minimum doubling of observed warming \citep{Hartmann:2013}; therefore, an urgent challenge and opportunity is discovering systems that naturally buffer against oncoming impacts of climate change.

River systems may offer an unappreciated defense against some aspects of climate change. Climate change is altering global precipitation patterns \citep{Donat:2016} and subsequently river flow regimes \citep{Hartmann:2013,Palmer:2009}. Earlier snowmelt \citep{Rauscher:2008}, reduced snow pack \citep{McCabe:2014}, shifts from snow to rain, and changes in the annual distribution of precipitation are all impacting stream flow \citep{Hartmann:2013}, resulting in flashier flows with increased frequency of flooding \citep{Hirabayashi:2013}, longer periods of drought and critically low flows for wildlife and agriculture \citep{Melillo:2014}. However, we hypothesize that large free-flowing river networks may naturally dampen these signals of climate change by integrating across different manifestations of climate. Processes that aggregate across asynchronous components tend to dampen the aggregate, a process known as the portfolio effect \citep{Doak:1998}. Climate is expressed differently and asynchronously over the landscape \citep{George:2015} and on account of their branching architecture and directional flow, river networks directionally integrate these varied climate manifestations \citep{Peterson:2013}, potentially dampening the local climatic response of sub-catchments. Large free-flowing rivers may have diverse climate portfolios, with sub-catchments acting as assets in their portfolio, thereby dampening local climate impacts. Portfolio effects in rivers are already known to dampen short-term variation \citep{Moore:2015,Yeakel:2014} but attenuation of long-term climate driven shifts have yet to be examined.

Here we examine whether rivers attenuate local long-term climate change through analyses of hydrometric data from one of the largest free-flowing rivers in North America. Located in British Columbia, Canada, the Fraser River drains an area approximately the size of the United Kingdom (\texttildelow 217,000 km\textsuperscript{2}) and in a rare combination is fairly well monitored while also having no dams on its main stem \citep{Vorosmarty:2010} (Figure~\ref{figone}). This watershed drains interior high plateau, coastal mountains, and the Canadian Rockies, and thus integrates a diverse mosaic of landscapes, weather, and climate. As with other mid-latitude rivers \citep{Bindoff:2013}, the Fraser River has expressed increasing trends in discharge volume and variability over the last several decades \citep{Dery:2012,Morrison:2002}. The river's discharge is monitored by hydrometric gauge stations throughout the watershed. We analyzed 38 years of data collected concurrently at 55 of these stations and applied a novel analytical approach to examine whether this large river exhibits signals of long-term climate dampening.

\begin{figure}[h]
	\centerline{\includegraphics[height=4in]{Fig1_Map2.pdf}}
	\caption{Flow gauge sites within the Fraser River basin (dark grey) in British Columbia Canada (light grey). Flow gauge sites (yellow dots) are scaled by the size of the contributing area. The map is projected in Albers BC UTMs to provide an equal area depiction of the region but labels are expressed in WGS84 latitude and longitude.}
	\label{figone}
\end{figure}

\section{Data and Methods}
\subsection{Discharge Data and Flow Metrics}
To assess signals of climate change in hydrology within the Fraser River basin we downloaded daily flow data from Environment Canada's HYDAT database. Within HYDAT we selected hydrometric gauge stations using three criteria: (i) the station was within the Fraser River basin, (ii) the station's flow data were not observed to be dam influenced, and (iii) the station collected data in each month between 1970 and 2007 where no month was missing more than 5 days of data and at least 35 complete years were present. We selected this 38-year range because this period maximized the number of gauge stations in the Fraser River basin (\textit{n}=55) operating concurrently over a long enough time period in which climate trends may be observable. We interpolated missing data using a linear model in log-space, but of the 763,345 days of data only 23 days were interpolated. Because dams control flow and stabilize flow trends, we removed sites where daily flow exhibited unusually high autocorrelation and low variability relative to other sites of similar size, and were observed downstream of, and atypically close to, a dam (\textit{n}=3). The downstream distance of gauge stations from dams was determined using GIS spatial layers available in the \textit{BC Data Catalog}, maintained by the provincial government of British Columbia.

As climate change alters precipitation patterns from snow to rain in temperate regions, rivers are generally predicted to have lower low-flows and higher high-flows, as well as earlier snowmelt \citep{Nijssen:2001}. These manifestations of rising temperatures are impacting species \citep{Xenopoulos:2006} and people \citep{Hirabayashi:2013}. We summarized our daily discharge data into annual and monthly flow metrics that capture these transitions. We used maximum- and minimum-flow estimates to capture trends in extreme discharge and median-flow estimates to capture general annual and seasonal trends. Because the timing of flow is particularly important to the phenology of many organisms, we also calculated the day-of-year in which half the annual flow was reached.

\subsection{River Discharge Trend Analysis}
For a river network to act as a portfolio of climate and dampen the effects of climate change, expressions of climate and their trends over time need to vary throughout the network (Figure~\ref{figtwo}). We estimated annual and monthly flow trends for each response variable at each gauge station using a generalized least squares (GLS) model with an AR1 autocorrelation function: 
\begin{linenomath*}
\begin{equation}
  \mathrm{flow}_{s,t} = a_s + b_s \mathrm{year}_{s,t} + \epsilon_{s,t}, \quad 
  \epsilon_{s,t} \sim \mathcal{N}(\upphi \epsilon_{s,t-1}, \sigma_\mathrm{flow}^2) \label{eq1},
\end{equation}
\end{linenomath*}
where $\mathrm{flow}_{s,t}$ represents a flow metric [log(maximum), log(minimum), log(median), logit(day-of-year to half-annual-flow)] at each site ($s$) and time point ($t$),  and $\mathrm{year}_{s,t}$ represents the time in years (e.g., 1970, 1971, \ldots, 2007) with 1988 subtracted to approximately center the predictor. Parameters $b_{s}$ (slope) and $a_{s}$ (intercept) represent the estimated mean effect of time on $\mathrm{flow}_{s,t}$ and the estimate of $\mathrm{flow}_{s,t}$ at time zero (i.e., 1988) respectively. Error ($\epsilon_{s,t}$) of the current time step for a given site was allowed to be correlated with that of the previous time step by $\upphi$ and was assumed to be normally distributed with a variance of $\sigma^{2}$. To ensure our model estimates and simulations remained within the calendar year we logit transformed our flow-timing response variable (i.e., day-of-year to half-annual-flow) after scaling the data between 0 and 1 (i.e., dividing by 365).

\begin{figure}[h]
	\centerline{\includegraphics[height=3in]{Fig2_Flow-Time.pdf}}
	\caption{Annual median flow at 55 sites within the Fraser River basin. Flow rates are scaled at each site by the root mean square in order to make flow rates comparable among sites. Lines are loess smoothers with a smoothing span of 0.25 predicted each month between 1970 and 2007. Line and point size are scaled to each flow gauge site's watershed area. The data show inter-annual variability within sites with varying magnitudes and directions among sites, illustrating variable long-term change in hydrology across the basin.}
	\label{figtwo}
\end{figure}

\subsection{River Network Trend Attenuation Analysis}
We hypothesized that larger catchments within the Fraser River watershed with more diverse climate portfolios would have dampened long-term hydrological change when compared to their smaller sub-catchments. Therefore, to test whether networks dampen climate signals we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model:
\begin{linenomath*}
\begin{equation}
	\hat{b}_{s} = c + d\sqrt{\mathrm{area}_{s}} + \eta_{s}, \quad
  \eta_{s} \sim \mathcal{N}(0, f(\mathrm{area}_{s})) \label{eq2},
\end{equation}
\end{linenomath*}
where $\hat{b}_{s}$ represents a flow trend at a given site ($s$) and $\mathrm{area}_{s}$ represents site $s$'s watershed area. Fitted $d$ and $c$ parameters represent the mean effect of $\sqrt{\mathrm{watershed area}}$ on flow trends and the mean basin-wide flow trend at a theoretical watershed area of zero, respectively. To quantify attenuation, we modeled the change in trend variability with the change in catchment area using an exponential variance function \citep[p.~211]{Pinheiro:2006}: 
\begin{linenomath*}
\begin{equation}
	f(\mathrm{area}_{s})=\sigma_b^2 \exp(2\updelta\sqrt{\mathrm{area}_{s}}) \label{eq3},
\end{equation}
\end{linenomath*} 
where the variance of the estimated error ($\eta$) changes exponentially with increasing $\sqrt{\mathrm{area}}$. This variance function allowed us to predict the range of flow trend values that would be expected as watershed area increased. A decreasing range of flow trends with increasing watershed area would indicate a climate-dampening network effect.

\subsection{River Discharge Trend Simulation}
In the observed data, small watersheds exhibited greater variability around their trend estimates than large watersheds, likely because of greater short-term variation in small catchments \citep{Moore:2015}. This relationship between watershed size and trend certainty may pull small watershed trend estimates away from zero, thereby creating the appearance of decreasing trend variability among sites as watershed size increases. Using a null-hypothesis framework we simulated time-series for each site with no underlying trend. These simulations were generated by a random walk process with the observed standard deviation and autocorrelation parameters ($\hat{\upsigma}$, $\hat{\upphi}$) at our sites. Using the same GLS model and AR1 correlation structure as applied to the observed data (eq.~\ref{eq1}), we estimated trends for 1000 simulated time series at each site for each response variable (e.g., Figure S3). This simulation process, a form of parametric bootstrapping, created distributions of null-expectations with which we compared our observed results.

We then applied equation~\ref{eq2} to each of the 1000 basin-wide simulations, resulting in 1 observed variance exponent parameter ($\hat{\updelta}$) and 1000 simulated $\hat{\updelta}$ for each flow metric. By comparing our observed attenuation with our basin-wide null-model simulations we addressed the potential that more variable flows in smaller catchments contribute to the observed pattern of flow trends as a function of watershed area (Figure~\ref{figthree}). This null-model approach asks how likely our observation is due to a sampling effect versus a network portfolio effect.

\begin{figure}[h]
	\centerline{\includegraphics[height=4.5in]{Fig3_Annual-Funnel.pdf}}
	\caption{Annual flow trend attenuation within the Fraser River basin. (\textbf{Left}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}). Blue represents observed attenuation; orange and red represent simulated attenuation that is weaker and stronger than observed. (\textbf{Right}) Density plots show null model simulated variance exponents ($\hat{\updelta}$) and the proportion on either side of the observed variance exponent (blue). Flow metrics include long-term flow-timing shifts (change per decade in day-of-year (DOY) to half annual flow), where decreasing trends suggest more annual flow is occurring earlier in the year (a), and the percent change per decade in minimum (b), maximum (c) and median (d) annual flow. Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
	\label{figthree}
\end{figure}

\subsection{Dampening Summary Statistics}
We calculated attenuation certainty as the percentage of simulated $\hat{\updelta}$ that were less than the observed $\hat{\updelta}$. We calculated attenuation strength as the ratio between the standard deviation at the smallest and largest watersheds as defined by the exponential variance function (i.e.,\\$\sqrt{f(\mathrm{area}_1)} / \sqrt{f(\mathrm{area}_2)}$ ). To compare the degree of attenuation observed with the null-simulated attenuation, we compared this ratio with the same ratio calculated from the null-simulated data (i.e., $\left( \sqrt{f(\mathrm{area}_1)} / \sqrt{f(\mathrm{area}_2)} \right) / \left( \sqrt{f(\mathrm{area}_{\mathrm{null} 1})} / \sqrt{f(\mathrm{area}_{\mathrm{null} 2})} \right)$ ). Finally, to estimate how the day-of-year to half-annual-flow was changing, we estimated the steepest point of each site's logistic curve and multiplied this slope value by 365 to revert from our 0--1 scale to our original annual range of 0--365. To get a decadal rate we multiplied these results by 3.8 (number of decades in 38 years). We estimated the steepest part of the logistic curve using the \textit{divide by four} rule \citep{Gelman:2008}, where dividing the coefficient by four equals the first derivative (slope) of the logistic curve at its steepest point.

\section{Results}
\subsection{Hydrologic Trend Diversity}
Long-term changes in flow metrics varied substantially across the watershed (Figure~\ref{figthree}). For instance, annual maximum flow decreased in 38 sites but increased in 17 over the last four decades (-28 to 15\% change$\cdot$decade\textsuperscript{-1}, std. dev. = 6\%). Annual median flow showed similar variability but a greater tendency towards rising flow trends with 49 increasing and only six decreasing (-5 to 25\% change$\cdot$decade\textsuperscript{-1} std. dev. = 5\%). Thus, within the Fraser River basin, there are a variety of long-term trends in flow, suggesting a diversity of climate responses.

\subsection{Annual Hydrologic Trend Dampening}
Long-term changes in hydrology were less variable in larger catchments, as predicted, demonstrating between 85 and 92\% dampening of flow trends moving from headwater catchments to the network's confluence. In other words, the largest catchments were approximately 10 times more stable in their climate response than the smallest catchments. For instance, trends ranged between a 6\% reduction and 19\% increase in median flow per decade among small watersheds, while large watersheds ranged between 0.8 and 1.6\% change per decade. Similar attenuation was seen in flow timing with a 94\% reduction in the day-of-year to half-annual-flow trends such that small watersheds were reaching half their annual flow between 8 days later and 25 days earlier while large watersheds have only shifted 2 to 4 days earlier between 1970 and 2007.

A null model approach illustrated that support for a river network portfolio effect was greater than 90\% for three of four annual flow trend metrics and as high as 98\% for median-annual-flow trends (Figure~\ref{figthree}). Trend attenuation was 3.1 (0.7--9.1 this and hereafter are 90\% intervals), 1.6 (0.3--4.6), 3.4 (1.0--8.5) and 5.4(1.5--14.6) times greater than the null model for annual-flow timing, and minimum, maximum and median annual-flow respectively. These statistics provide strong support for the hypothesis that river network portfolio effects contribute to climate dampening in large rivers.

\subsection{Seasonal Hydrologic Trend Dampening}
Regional flow-trends and attenuation strength varied by season. Analysis of monthly trends revealed that winter flows have been getting higher and summer flows have been getting lower over time, but that these trends have been more variable in smaller catchments (Figure~\ref{figfour}, S1, S2). For example, winter maximum flows in small catchments ranged from nearly no change to dramatic 44\% increases per decade with decreases only as large as 15\%. Despite these extreme locations, on average the basin has only experienced moderate winter maximum flow increases of 5--10\%. Summer flows exhibited the opposite, with decreases in high flows over time. However, there was weak evidence of attenuation in the spring (Figure~\ref{figfour}, S1, S2). For example, in the spring the likelihood of network-driven attenuation of maximum flow trends was as low as 24\% (Figure~\ref{figfour}). The deterioration of the network's attenuation strength was coincident with the spring freshet, when snowmelt drives high flows across the basin. Synchronization events such as the freshet may subvert river network dampening mechanisms by homogenizing the region's response to climate shifts.

\begin{figure}[htp]
	\centerline{\includegraphics[height=5.2in]{Fig4_Max-Monthly-Density-Funnel.pdf}}
	\caption{Monthly maximum flow trend attenuation within the Fraser River basin. (\textbf{Left}) Fraser River's basin-wide maximum-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulations. Observed values falling further from the center of the density distribution suggest greater evidence for changes in maximum flow and a greater shift in magnitude. (\textbf{Center}) Observed monthly Fraser River maximum-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates. Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed. (\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}) for four seasonally representative months.These reflect months in the prior columns and describe the variation in percent change per decade of maximum flow among sites. Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
	\label{figfour}
\end{figure}

\section{Discussion}
Climate change is already causing economic and conservation challenges for river systems worldwide \citep{Palmer:2009}. For example, sockeye salmon in the Fraser River have seen as much as 80\% pre-spawn mortality in years with later run-off and elevated temperatures, leading managers to reduce harvest rates in these years. In the absence of climate dampening the impacts of these temperature extremes could have been dramatically worse. As climate change progresses, river hydrology will continue to shift, further stressing riverine ecosystems and subsequently demanding responsive management. However, river networks provide an underappreciated defense against these climate-change impacts as they are uniquely organized to leverage locally-filtered expressions of climate into stability. A simple product of form and gravity, stability emerges as river networks integrate landscape heterogeneity, creating a downstream portfolio of climate that smooths local extremes and tempers long-term flow trends. Therefore, larger rivers should have flow regimes that are less sensitive to local climate trends.

Despite these dampening processes, river networks are still vulnerable to climate change. In the Fraser River, the furthest downstream reaches are still exhibiting small but detectable signals of hydrological shifts. Late fall, winter and early spring flows have generally increased while late spring, summer and early fall flows have decreased (Figure~\ref{figfour}, S1, S2). These findings generally support previous studies demonstrating earlier spring runoff  \citep{Dery:2012}, decreasing summer and fall flows  \citep{Stahl:2006} and increased winter flows  \citep{Healey:2011}. Together, this body of work indicates basin-wide effects of climate change, where shifts in winter precipitation from snow to rain is resulting in increased winter runoff and decreased snow pack contributions to summer flows. Our work suggests that because the basin is responding heterogeneously to the shifting climate and integrating climate trends, the downstream impacts of climate change on flow have been greatly reduced. In contrast to previously studied types of climate mitigation where habitats modify climate drivers (e.g., carbon storage) or physically absorb climate change impacts (e.g., mangroves), network climate dampening smooths out extreme climatic trends thereby reducing but not eliminating the impact of local extremes.

Our study provides insight into the spatial scaling of climate change. We focused on a vast watershed with a remarkably diverse climate portfolio, yet we would still expect that smaller rivers could dampen some degree of climate variability. For instance, based on our data, we predict that rivers draining a catchment of 60,000 km\textsuperscript{2} would have 66\% less variability in the rate of change in hydrology than a smaller river draining ~5000 km\textsuperscript{2}. While this dampening is less than the 10-fold decrease observed in the entire 220,000 km\textsuperscript{2} Fraser River, smaller watersheds can offer a defense against the impacts of climate change.

There have been extensive efforts to predict the hydrologic sensitivity of rivers due to climate change through downscaled climate projections \citep{Nijssen:2001}. Here we demonstrate that large rivers may have an inherent ability to absorb local climate change. But just as the destruction of coastal habitats degrades their natural capacity to mitigate sea-level rise or storm events \citep{Arkema:2013}, the management of river basins likely impacts the ability for river networks to attenuate climate trends. For example, dams synchronize the flow regimes of rivers  \citep{Poff:2007} and logging increases the magnitude and frequency of extreme events such as high and low flows through increased runoff rates \citep{Zhang:2014}. These anthropogenic activities may magnify the impacts of climatic shifts such as the transition from snow- to rain-dominated precipitation in some basins. The climate cost of these watershed activities should be considered in environmental decision making. Climate change is a global challenge; here we suggest large rivers dampen local change by leveraging landscape diversity into a portfolio of climate, an important tool in the climate-mitigation toolbox.

\acknowledgments
All the data and R scripts used in this study are freely available on github at https://github.com/kchezik/River-Network-Flow-Trends. This study could not have been done without the decades of flow data collected by those at Environment Canada. K.A. Chezik and J.W. Moore were supported by the Liber Ero Chair of Coastal Science and Management and Simon Fraser University. S.C. Anderson was supported by the David H. Smith Conservation Research Fellowship.

\begin{thebibliography}{30}
\providecommand{\natexlab}[1]{#1}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi

\bibitem[{\textit{Arkema et~al.}(2013)\textit{Arkema, Guannel, Verutes, Wood,
  Guerry, Ruckelshaus, Kareiva, Lacayo, and Silver}}]{Arkema:2013}
Arkema, K.~K., G.~Guannel, G.~Verutes, S.~A. Wood, A.~Guerry, M.~Ruckelshaus,
  P.~Kareiva, M.~Lacayo, and J.~M. Silver (2013), Coastal habitats shield
  people and property from sea-level rise and storms, \textit{Nature Climate
  Change}, \textit{3}(10), 913--918.

\bibitem[{\textit{Bindoff et~al.}(2013)\textit{Bindoff, Stott, AchutaRao,
  Gillett, Gutzler, Hansingo, Hegerl, Hu, Jain, Mokhov, Overland, Perlwitz,
  Sebbari, and Zhang}}]{Bindoff:2013}
Bindoff, N.~L., P.~A. Stott, M.~R. AchutaRao, N.~P. Gillett, D.~Gutzler,
  K.~Hansingo, G.~C. Hegerl, Y.~Hu, S.~Jain, I.~I. Mokhov, J.~Overland,
  J.~Perlwitz, R.~Sebbari, and X.~Zhang (2013), Detection and attribution of
  climate change: from global to regional, \textit{Tech. rep.},
  Intergovernmental Panel on Climate Change.

\bibitem[{\textit{D{\'e}ry et~al.}(2012)\textit{D{\'e}ry,
  Hern{\'a}ndez-Henr{\'\i}quez, Owens, Parkes, and Petticrew}}]{Dery:2012}
D{\'e}ry, S.~J., M.~A. Hern{\'a}ndez-Henr{\'\i}quez, P.~N. Owens, M.~W. Parkes,
  and E.~L. Petticrew (2012), A century of hydrological variability and trends
  in the {F}raser {R}iver basin, \textit{Environmental Research Letters},
  \textit{7}(2), 024,019.

\bibitem[{\textit{Doak et~al.}(1998)\textit{Doak, Bigger, Harding, Marvier,
  O'malley, and Thomson}}]{Doak:1998}
Doak, D.~F., D.~Bigger, E.~Harding, M.~Marvier, R.~O'malley, and D.~Thomson
  (1998), The statistical inevitability of stability-diversity relationships in
  community ecology, \textit{The American Naturalist}, \textit{151}(3),
  264--276 %@ 0003--0147.

\bibitem[{\textit{Donat et~al.}(2016)\textit{Donat, Lowry, Alexander, O'Gorman,
  and Maher}}]{Donat:2016}
Donat, M.~G., A.~L. Lowry, L.~V. Alexander, P.~A. O'Gorman, and N.~Maher
  (2016), More extreme precipitation in the world's dry and wet regions,
  \textit{Nature Climate Change}, \textit{6}(5), 508--513,
  \doi{10.1038/nclimate2941}.

\bibitem[{\textit{Gelman and Hill}(2008)}]{Gelman:2008}
Gelman, A., and J.~Hill (2008), Data analysis using regression and multilevel
  hierarchical models, \doi{10.1080/02664760801949043}.

\bibitem[{\textit{George et~al.}(2015)\textit{George, Thompson~III, and
  Faaborg}}]{George:2015}
George, A.~D., F.~R. Thompson~III, and J.~Faaborg (2015), Using {L}i{DAR} and
  remote microclimate loggers to downscale near-surface air temperatures for
  site-level studies, \textit{Remote Sensing Letters}, \textit{6}(12),
  924--932.

\bibitem[{\textit{Guerry et~al.}(2015)\textit{Guerry, Polasky, Lubchenco,
  Chaplin-Kramer, Daily, Griffin, Ruckelshaus, Bateman, Duraiappah, Elmqvist,
  Feldman, Folke, Hoekstra, Kareiva, Keeler, Li, McKenzie, Ouyang, Reyers,
  Ricketts, Rockstr{\"o}m, Tallis, and Vira}}]{Guerry:2015}
Guerry, A.~D., S.~Polasky, J.~Lubchenco, R.~Chaplin-Kramer, G.~C. Daily,
  R.~Griffin, M.~Ruckelshaus, I.~J. Bateman, A.~Duraiappah, T.~Elmqvist, M.~W.
  Feldman, C.~Folke, J.~Hoekstra, P.~M. Kareiva, B.~L. Keeler, S.~Li,
  E.~McKenzie, Z.~Ouyang, B.~Reyers, T.~H. Ricketts, J.~Rockstr{\"o}m,
  H.~Tallis, and B.~Vira (2015), Natural capital and ecosystem services
  informing decisions: From promise to practice, \textit{Proceedings of the
  National Academy of Sciences}, \textit{112}(24), 7348--7355,
  \doi{10.1073/pnas.1503751112}.

\bibitem[{\textit{Hartmann et~al.}(2013)\textit{Hartmann, Klein, Rusticucci,
  Alexander, Bronnimann, Charabi, Dentener, Dlugokencky, Easterling, Kaplan,
  Soden, Thorne, Wild, and Zhai}}]{Hartmann:2013}
Hartmann, D.~L., A.~M.~G. Klein, M.~Rusticucci, L.~V. Alexander, S.~Bronnimann,
  Y.~Charabi, F.~J. Dentener, D.~R. Dlugokencky, A.~Easterling, B.~J. Kaplan,
  B.~J. Soden, P.~W. Thorne, M.~Wild, and P.~M. Zhai (2013), Climate change
  2013: The physical science basis., \textit{Tech. rep.}, Intergovernmental
  Panel on Climate Change.

\bibitem[{\textit{Healey and Bradford}(2011)}]{Healey:2011}
Healey, M., and M.~Bradford (2011), The cumulative impacts of climate change on
  {F}raser {R}iver sockeye salmon (\textit{{O}ncorhynchus nerka}) and
  implications for management, \textit{Can. J. Fish. Aquat. Sci.},
  \textit{68}(4), 718--737.

\bibitem[{\textit{Hirabayashi et~al.}(2013)\textit{Hirabayashi, Mahendran,
  Koirala, Konoshima, Yamazaki, Watanabe, Kim, and Kanae}}]{Hirabayashi:2013}
Hirabayashi, Y., R.~Mahendran, S.~Koirala, L.~Konoshima, D.~Yamazaki,
  S.~Watanabe, H.~Kim, and S.~Kanae (2013), Global flood risk under climate
  change, \textit{Nature Climate Change}, \textit{3}, 816--821.

\bibitem[{\textit{Hulme}(2016)}]{Hulme:2016}
Hulme, M. (2016), 1.5$\degree${C} and climate research after the {P}aris
  {A}greement, \textit{Nature Climate Change}, \textit{6}(3), 222--224.

\bibitem[{\textit{Jones et~al.}(2012)\textit{Jones, Hole, and
  Zavaleta}}]{Jones:2012}
Jones, H.~P., D.~G. Hole, and E.~S. Zavaleta (2012), Harnessing nature to help
  people adapt to climate change, \textit{Nature Climate Change},
  \textit{2}(7), 504--509.

\bibitem[{\textit{McCabe and Wolock}(2014)}]{McCabe:2014}
McCabe, G.~J., and D.~M. Wolock (2014), Spatial and temporal patterns in
  conterminous {U}nited {S}tates streamflow characteristics,
  \textit{Geophysical Research Letters}, \textit{41}(19), 6889--6897.

\bibitem[{\textit{Melillo et~al.}(2014)\textit{Melillo, Richmond, and
  Yohe}}]{Melillo:2014}
Melillo, J.~M., T.~Richmond, and G.~W. Yohe (2014), Climate change impacts in
  the {U}nited {S}tates: the third national climate assessment, \textit{US
  Global change research program}, \textit{841}.

\bibitem[{\textit{Moore et~al.}(2015)\textit{Moore, Beakes, Nesbitt, Yeakel,
  Patterson, Thompson, Phillis, Braun, Favaro, Scott et~al.}}]{Moore:2015}
Moore, J.~W., M.~P. Beakes, H.~K. Nesbitt, J.~D. Yeakel, D.~A. Patterson, L.~A.
  Thompson, C.~Phillis, D.~C. Braun, C.~Favaro, D.~C. Scott, et~al. (2014),
  Emergent stability in a large free-flowing watershed, \textit{Ecology}, 
  \textit{96}(2), 340--347.

\bibitem[{\textit{Morrison et~al.}(2002)\textit{Morrison, Quick, and
  Foreman}}]{Morrison:2002}
Morrison, J., M.~C. Quick, and M.~G. Foreman (2002), Climate change in the
  {F}raser {R}iver watershed: flow and temperature projections, \textit{Journal
  of Hydrology}, \textit{263}(1), 230--244.

\bibitem[{\textit{Narain et~al.}(2011)\textit{Narain, Margulis, and
  Essam}}]{Narain:2011}
Narain, U., S.~Margulis, and T.~Essam (2011), Estimating costs of adaptation to
  climate change, \textit{Climate Policy}, \textit{11}(3), 1001--1019.

\bibitem[{\textit{Nijssen et~al.}(2001)\textit{Nijssen, O'Donnell, Hamlet, and
  Lettenmaier}}]{Nijssen:2001}
Nijssen, B., G.~M. O'Donnell, A.~F. Hamlet, and D.~P. Lettenmaier (2001),
  Hydrologic sensitivity of global rivers to climate change, \textit{Climatic
  Change}, \textit{50}(1-2), 143--175.

\bibitem[{\textit{Palmer et~al.}(2009)\textit{Palmer, Lettenmaier, Poff,
  Postel, Richter, and Warner}}]{Palmer:2009}
Palmer, M.~A., D.~P. Lettenmaier, N.~L. Poff, S.~L. Postel, B.~Richter, and
  R.~Warner (2009), Climate change and river ecosystems: protection and
  adaptation options, \textit{Environmental Management}, \textit{44}(6),
  1053--1068.

\bibitem[{\textit{Peterson et~al.}(2013)\textit{Peterson, Ver~Hoef, Isaak,
  Falke, Fortin, Jordan, McNyset, Monestiez, Ruesch, Sengupta
  et~al.}}]{Peterson:2013}
Peterson, E.~E., J.~M. Ver~Hoef, D.~J. Isaak, J.~A. Falke, M.-J. Fortin, C.~E.
  Jordan, K.~McNyset, P.~Monestiez, A.~S. Ruesch, A.~Sengupta, et~al. (2013),
  Modelling dendritic ecological networks in space: an integrated network
  perspective, \textit{Ecology Letters}, \textit{16}(5), 707--719.

\bibitem[{\textit{Pinheiro and Bates}(2006)}]{Pinheiro:2006}
Pinheiro, J., and D.~Bates (2006), Mixed-effects models in {S} and {S-PLUS},
  \textit{Springer Science \& Business Media}.

\bibitem[{\textit{Poff et~al.}(2007)\textit{Poff, Olden, Merritt, and
  Pepin}}]{Poff:2007}
Poff, L.~N., J.~D. Olden, D.~M. Merritt, and D.~M. Pepin (2007), Homogenization
  of regional river dynamics by dams and global biodiversity implications,
  \textit{Proceedings of the National Academy of Sciences}, \textit{104}(14),
  5732--5737, \doi{10.1073/pnas.0609812104}.

\bibitem[{\textit{Rauscher et~al.}(2008)\textit{Rauscher, Pal, Diffenbaugh, and
  Benedetti}}]{Rauscher:2008}
Rauscher, S.~A., J.~S. Pal, N.~S. Diffenbaugh, and M.~M. Benedetti (2008),
  Future changes in snowmelt-driven runoff timing over the western {US},
  \textit{Geophysical Research Letters}, \textit{35}(16).

\bibitem[{\textit{\relax{Louisiana Coastal Wetlands Conservation and
  Restoration Task Force and the Wetlands Conservation and Restoration
  Authority}}(1998)}]{Conservation:1998}
\relax{Louisiana Coastal Wetlands Conservation and Restoration Task Force and
  the Wetlands Conservation and Restoration Authority} (1998), Coast 2050:
  {T}oward a sustainable coastal louisiana, \textit{Tech. rep.}, Louisiana
  Department of Natural Resources.

\bibitem[{\textit{Stahl and Moore}(2006)}]{Stahl:2006}
Stahl, K., and R.~Moore (2006), Influence of watershed glacier coverage on
  summer streamflow in {B}ritish {C}olumbia, {C}anada, \textit{Water Resources
  Research}, \textit{42}(6).

\bibitem[{\textit{V{\"o}r{\"o}smarty et~al.}(2010)\textit{V{\"o}r{\"o}smarty,
  McIntyre, Gessner, Dudgeon, Prusevich, Green, Glidden, Bunn, Sullivan,
  Liermann et~al.}}]{Vorosmarty:2010}
V{\"o}r{\"o}smarty, C.~J., P.~McIntyre, M.~O. Gessner, D.~Dudgeon,
  A.~Prusevich, P.~Green, S.~Glidden, S.~E. Bunn, C.~A. Sullivan, C.~R.
  Liermann, et~al. (2010), Global threats to human water security and river
  biodiversity, \textit{Nature}, \textit{467}(7315), 555--561.

\bibitem[{\textit{Xenopoulos and Lodge}(2006)}]{Xenopoulos:2006}
Xenopoulos, M.~A., and D.~M. Lodge (2006), Going with the flow: using
  species-discharge relationships to forecast losses in fish biodiversity,
  \textit{Ecology}, \textit{87}(8), 1907--1914.

\bibitem[{\textit{Yeakel et~al.}(2014)\textit{Yeakel, Moore, Guimaraes, and
  Aguiar}}]{Yeakel:2014}
Yeakel, J. D., J. W.~Moore, P. R.~Guimaraes, and M. A. M. de~Aguiar (2014), Synchronisation and
  stability in river metapopulation networks, \textit{Ecology Letters},
  \textit{17}(3), 273--283.

\bibitem[{\textit{Zhang and Wei}(2014)}]{Zhang:2014}
Zhang, M., and X.~Wei (2014), Alteration of flow regimes caused by large-scale
  forest disturbance: a case study from a large watershed in the interior of
  {B}ritish {C}olumbia, {C}anada, \textit{Ecohydrology}, \textit{7}(2),
  544--556, \doi{10.1002/eco.1374}.

\end{thebibliography}

%\bibliography{/Users/kylechezik/Documents/lit}

\end{document}

### 2017-01-11 Code_01 Data_Organization.Rmd
River-Network Climate Portfolio: Initial Push
Here I add the code for my PhD thesis chapter 1 and two rejected
manuscripts.

---
title: "Flow Data Organization"
author: "Kyle Chezik"
date: "June 24, 2015"
output: pdf_document
---

###Load Original Raw Data.
```{r,echo=FALSE,message=FALSE}
load("01_Data-Flow-Orig.RData")
load("01_Data-Wtshd-Orig.RData")
names(WS.2) = c("Station.ID","Lat","Long","Area")
head(Flow.2)
head(WS.2)

PDO = read.csv("http://www.ncdc.noaa.gov/teleconnections/pdo/data.csv",skip = 1, header = T)
PDO = PDO[which(PDO$Date>196912),]
ENSO = read.csv("http://www.ncdc.noaa.gov/teleconnections/enso/indicators/soi/data.csv",skip = 1, header = T)
ENSO = ENSO[which(ENSO$Date>196912),]

#Load in libraries.
library(dplyr);library(plyr);library(tidyr);library(MARSS)
```

###Subset and Constrain Data
```{r, echo=FALSE, message=F}
# Remove Dam Influenced Sites.
Flow.2 = Flow.2[which(Flow.2$Station.ID %in% c("08JC001","08JC002","08ME002")==F),]

#Limit the HYDAT flow data to years between 1970 and 2007.
Data = Flow.2[which(as.numeric(format(Flow.2$Date, "%Y"))>=1970 & as.numeric(format(Flow.2$Date, "%Y"))<=2007),]
# Data <- Data %>% arrange(Station.ID, Date) # ensure correct order
# Data <- Data %>% group_by(Station.ID) %>%
	# mutate(gap = c(NA, Date[2:n()] - Date[1:(n() - 1)])) %>%
	# mutate(any_NAs = ifelse(max(gap > 1, na.rm = TRUE), TRUE, FALSE)) %>%
	# mutate(long_NA_string = ifelse(max(gap > 7, na.rm = TRUE), TRUE, FALSE)) %>%
	# as.data.frame()

#Split up the date for processesing.
Data$Year = as.numeric(format(Data$Date,"%Y")) #Add a column for the year.
Data$nMonth = as.numeric(format(Data$Date,"%m")) #Add a column for the month.
Data$Month = as.factor(format(Data$Date,"%b")) #Add a column for the month.
Data$Day = as.numeric(format(Data$Date,"%d")) #Add a column for the day.
Data$DOY = as.numeric(format(Data$Date,"%j")) #Add a column for the day.
head(Data)

#Create a series of dates between January 1st of 1970 and December 31st of 2007. This will be used as a ruler for determining missing days in each month of each year.
Date.Cont = seq(as.Date("1970-01-01"), as.Date("2007-12-31"), by = "day")

#Determine the number of days missing in each month of each year by flow gauge station (DIMM = Days In Month Missing).
years.continuous <- as.numeric(format(Date.Cont,"%Y"))
months.continuous <- as.numeric(format(Date.Cont,"%m"))
DIMM <- Data %>% group_by(Station.ID, Year, Month) %>%
	dplyr::summarise(Days.Missing =
			length(Date.Cont[which(years.continuous == unique(Year) & 
							  months.continuous == unique(nMonth))]) - n()) %>%
	as.data.frame()

#Remove months with more then 5 days missing.
DIM.Limited.Months = DIMM[which(DIMM$Days.Missing<=5),]

#Determine the number of months in each year for each flow gauge site.
DIM.Limited.Months = ddply(DIM.Limited.Months,c("Station.ID","Year"),function(x){
	Months = dim(x)[1]
	data.frame(Months)
})

#Remove years with fewer than 12 months.
DIM.Limited.Years1 = DIM.Limited.Months[which(DIM.Limited.Months$Months==12),]

#Isolate data with Years that have 12 months in which no month is missing more than 5 days.
Data = match_df(Data,DIM.Limited.Years1,on = c("Station.ID","Year"))

#Determine the number of years included in each site.
DIM.Limited.Years2 = ddply(DIM.Limited.Years1,c("Station.ID"),function(x){
	Years = dim(x)[1]
	data.frame(Years)
})

#Remove stations with fewer than 35 years of data.
Stations.Final = DIM.Limited.Years2[which(DIM.Limited.Years2$Years>=35),]

#Isolate data such that all included sites have at least 35 of the possible 38 years between 1970 and 2007, and each year has 12 full months where each month is missing no more than five days.
Data = match_df(Data,Stations.Final,on = "Station.ID")
save(Data,file = "02a_Data_Clean.RData")
```
###Create Monthly Dataset
```{r,echo=F,message=F}
#Summarise data monthly mean, minnimum and maximums.
load("02a_Data_Clean.RData")
M.Data <- Data %>% group_by(Station.ID, Year, nMonth, Month) %>%
	dplyr::summarise(median.M.flow = median(Flow.Data),
 						max.M.flow = max(Flow.Data),
 						min.M.flow = min(Flow.Data))

M.Data = left_join(M.Data,WS.2,by = c("Station.ID"))

save(M.Data,file = "03_Data_Monthly.RData")
```
###Create Annual Dataset
```{r,echo=F,message=F}
#Fill in missing data for the 'day of year' analysis.
load("02a_Data_Clean.RData") #Load processesed data.

all_expected_dat <- expand.grid(unique(Data$Station.ID), Date.Cont,
	stringsAsFactors = FALSE)
names(all_expected_dat) <- c("Station.ID", "Date")
Data <- left_join(all_expected_dat, Data)
Data <- Data %>% arrange(Station.ID, Date)
Data$Year = as.numeric(format(Data$Date,"%Y")) #Add a column for the year.
Data$nMonth = as.numeric(format(Data$Date,"%m")) #Add a column for the month.
Data$Month = as.factor(format(Data$Date,"%b")) #Add a column for the month.
Data$Day = as.numeric(format(Data$Date,"%d")) #Add a column for the day.
Data$DOY = as.numeric(format(Data$Date,"%j")) #Add a column for the day.

#Determine the number of missing days in each Station, Year, Month combination. Retain those that have more than 0.
Missing <- Data %>% group_by(Station.ID, Year, Month) %>%
	dplyr::summarise(Days.Missing =
    length(Date.Cont[which(years.continuous == unique(Year) & 
      months.continuous == unique(nMonth))]) - n()) %>%
	dplyr::filter(Days.Missing > 0)

library(doParallel)
registerDoParallel(cores = 7)
#Create a list for each station and year combination where a month is missing data. These lists include data for the year prior and after the year in which data is missing. This creates a vector/dataset in the wide format which is necessary for the auto-regressive state-space model.
Sites = dlply(Missing,c("Station.ID","Year"),function(x){
	Period = seq(unique(x$Year)-1,unique(x$Year+1))
	Site = subset(Data,Station.ID==x$Station.ID & Year%in%Period)
	Potential.Dates = seq(as.Date(paste(Period[1],"-01-01",sep="")), as.Date(paste(Period[3],"-12-31",sep="")), by = "day")
	Date.Rule = data.frame(Station.ID = rep(unique(x$Station.ID),length(Potential.Dates)),Date = Potential.Dates)
	Site.Missing = full_join(Site, Date.Rule, by = c("Date","Station.ID"))
	Site = spread(Site.Missing[,c(1:3)],Date,Flow.Data)[1,]
	as.numeric(as.vector(Site[,c(2:ncol(Site))]))
})

pdf("interpolation-checks-approx.pdf", width = 9, height = 6)
par(mfrow = c(3, 3), mar = c(3, 3, 1, 1), cex = 0.5)
plyr::l_ply(Sites, function(i) {
	ind <- seq_along(i)
	d <- data.frame(flow = i, ind = ind)
	d$approx <- approx(d$ind[!is.na(d$flow)], y = log(d$flow[!is.na(d$flow)]),
		xout = d$ind)$y %>% exp()
  plot(d$ind, d$flow, type = "o", pch = 20, cex = 0.6)
  # lines(d$ind, d$approx, col = "blue")
  pred_nas <- d[which(is.na(d$flow)), c("ind", "approx")]
  points(pred_nas$ind, pred_nas$approx, col = "red", cex = 1, pch = 20)
})
dev.off()

# Interpolate the missing values via a linear interpolation (in log space):
Preds = ldply(names(Sites),function(x){
	Year = as.numeric(strsplit(x,split = "\\.")[[1]][2])
	Site = strsplit(x,split = "\\.")[[1]][1]
	Potential.Dates = seq(as.Date(paste(Year-1,"-01-01",sep="")), as.Date(paste(Year+1,"-12-31",sep="")), by = "day")
	x.Dates = Data[which(Data$Station.ID==Site & Data$Year==Year),"Date"]
	rows = which(Potential.Dates%in%x.Dates==F & as.numeric(format(Potential.Dates,"%Y"))==Year)

	ind <- seq_along(Sites[[x]])
	d <- data.frame(flow = Sites[[x]], ind = ind)
	d$approx <- approx(d$ind[!is.na(d$flow)], y = log(d$flow[!is.na(d$flow)]),
		xout = d$ind)$y %>% exp()
	d <- d[rows, ]

	data.frame(Station.ID = rep(Site,length(rows)),
		Date = Potential.Dates[rows],
		Flow.Data = d$approx)
})
Preds$Year = as.numeric(format(Preds$Date,"%Y")) #Add a column for the year.
Preds$nMonth = as.numeric(format(Preds$Date,"%m")) #Add a column for the month.
Preds$Month = as.factor(format(Preds$Date,"%b")) #Add a column for the month.
Preds$Day = as.numeric(format(Preds$Date,"%d")) #Add a column for the day.
Preds$DOY = as.numeric(format(Preds$Date,"%j")) #Add a column for the day.

save(Preds,file = "02b_Missing_Data_Predictions.RData")

#Combine predictions with the original data.
load("02b_Missing_Data_Predictions.RData")
Data.Preds = full_join(Data,Preds)
Data.Preds$Station.ID = as.factor(Data.Preds$Station.ID)
Data.Preds$Month = as.factor(Data.Preds$Month)

#Summarise annual data by the day of year to half annual flow.
# Get the row numbers of when cumulative flow volume hits half the total yearly volume:
accum_flow <- Data.Preds %>%
	mutate(row_number = 1:length(Year)) %>%
	group_by(Station.ID, Year) %>%
	mutate(half_total_flow = sum(Flow.Data) / 2) %>%
	summarise(row_number = max(row_number[cumsum(Flow.Data) < half_total_flow])) %>%
	as.data.frame()
# Select those rows:
accum_flow <- Data.Preds[accum_flow$row_number, ] %>%
	inner_join(accum_flow) %>%
	select(-row_number) %>%
	rename(DOY2 = DOY)
Y.Data = left_join(accum_flow, WS.2, by = c("Station.ID"))
Y.Data$Station.ID = as.factor(Y.Data$Station.ID)

flow.stats = ddply(Data.Preds, c("Station.ID","Year"),summarize,Median.F = median(Flow.Data),Min.F = min(Flow.Data),Max.F = max(Flow.Data))
Y.Data = left_join(flow.stats,Y.Data, by = c("Station.ID","Year"))

save(Y.Data,file = "03_Data_Annual.RData")
```

### 2017-01-11 Code_01b Outlier ID.Rmd
River-Network Climate Portfolio: Initial Push
Here I add the code for my PhD thesis chapter 1 and two rejected
manuscripts.

What's going on with the funnels?

```{r load-stuff, echo=FALSE}
load("03_Data_Annual.RData")
load("03_Data_Monthly.RData")
Y.Data = plyr::mutate(Y.Data, Year.Center = Year-1969)
Y.Data = plyr::ddply(Y.Data, "Station.ID", plyr::mutate, Med.F.sc = scale(Median.F, center = F),Max.F.sc = scale(Max.F, center = F),Min.F.sc = scale(Min.F, center = F))
M.Data = plyr::mutate(M.Data, Year.Center = Year-1969)

# this function fits slopes to real data
fit_slopes <- function(flow.dat, response) {
  area_dat <- unique(flow.dat[,c("Station.ID", "Area")])
  equation <- as.formula(paste(response,"~Year.Center"))
  models <- plyr::ddply(flow.dat,c("Station.ID"), function(x){
    library(nlme)
    mod <- gls(equation, correlation = corAR1(), data = x)
    slope <- coef(mod)[[2]]
    intercept <- coef(mod)[[1]]
    se <- summary(mod)$tTable[2,2]
    sigma <- mod$sigma
    phi <- coef(mod$model[[1]], unconstrained = F)[[1]]
    data.frame(intercept,slope,se,sigma,phi)
})
  models <- plyr::join(models,area_dat, by = "Station.ID")
  models
}

# this function takes a data frame from fit_slopes() and generates a simulated
# time series (WITHOUT SLOPE) plus a gls() fit to that simulated data
sim_slopes <- function(slope.dat, yrs, return_ts = FALSE) {
  out <- plyr::adply(slope.dat, 1, function(x) {
    y <- as.numeric(arima.sim(n = length(yrs),
        model = list(order = c(1, 0, 0), ar = x$phi),
        mean = 0, sd = x$sigma)) + x$intercept
    mod_sim <- gls(y~yrs, correlation = corAR1())
    slope_sim <- coef(mod_sim)[[2]]
    se_sim <- summary(mod_sim)$tTable[2,2]
    if (!return_ts) {
      data.frame(slope_sim, se_sim)
    } else {
      data.frame(y, yrs)
    }
  })
  if (return_ts) out <- out[,c("Station.ID", "yrs", "y")]
  out
}

# this function fits a gls model with a variance structure for the residual 
# error
# slopes and area are both vectors
fit_var <- function(slopes, area) {
  scale_factor <- sd(slopes)
  # scale_factor <- 1
  scaled_slope <- slopes / scale_factor
  m <- gls(scaled_slope~1, weights = varExp(form= ~sqrt(area)/1e4))
  varexp <- m$model[[1]][[1]]
  sigma <- m$sigma * scale_factor
  intercept <- coef(m)[[1]] * scale_factor
  data.frame(varexp, sigma, intercept)
}

# wrapper function
null_sim <- function(flow.dat, response, iter, capture_slopes_i = -999) {
  yrs <- unique(flow.dat$Year.Center)
  real_slopes <- fit_slopes(flow.dat, response)
  example_ts <- sim_slopes(slope.dat = real_slopes, yrs = yrs,
    return_ts = TRUE)
  sim_varexp <- plyr::ldply(seq_len(iter), function(i) {
    simulated_slopes <- sim_slopes(slope.dat = real_slopes, yrs = yrs)
    if (i %in% capture_slopes_i) {
    	saveRDS(simulated_slopes, file = paste0("sim-slopes-", response, "-", i, ".rds"))
    }
    out <- fit_var(simulated_slopes$slope_sim, area = simulated_slopes$Area)
    out$.n <- i
    out
  })
  real_varexp <- fit_var(real_slopes$slope, area = real_slopes$Area)
  list(real_varexp = real_varexp, real_slopes = real_slopes,
    sim_varexp = sim_varexp, example_ts = example_ts)
}
```

```{r run-stuff, cache=TRUE}
set.seed(1)
out_min <- null_sim(Y.Data, "Min.F.sc", iter = 9, capture_slopes_i = 1:9)
out_med <- null_sim(Y.Data, "Med.F.sc", iter = 9, capture_slopes_i = 1:9)
```

```{r, fig.width=11, fig.height=11, cache=TRUE}
library(ggplot2)
# real:
ggplot(Y.Data, aes(Year.Center, Min.F.sc)) + geom_line() + 
	facet_wrap(~Station.ID, scales = "free_y")
ggplot(Y.Data, aes(Year.Center, Med.F.sc)) + geom_line() + 
	facet_wrap(~Station.ID, scales = "free_y")
```

Plot some replicates:

```{r}

# intercept_real + 1.96 * sqrt(sigma_real^2 * exp(2*sqrt(models$Area)*x$varexp))
check_funnels <- function(response, sim_var_df) {
	par(mfrow = c(3, 3), cex = 0.7, mar = c(2, 2, 1, 1))
	for(i in 1:9) {
		x <- readRDS(paste0("sim-slopes-", response, "-", i, ".rds"))
		a <- seq(sqrt(min(x$Area)), sqrt(max(x$Area)), length.out = 50L) / 1e4
		xx <- sim_var_df[i,]
		u <- xx$intercept + 1.96 * sqrt(xx$sigma^2 * exp(2*a*xx$varexp))
		l <- xx$intercept - 1.96 * sqrt(xx$sigma^2 * exp(2*a*xx$varexp))
		plot(sqrt(x$Area), x$slope_sim, ylim = range(c(l, u, x$slope_sim)), type = "n")
		x$.n <- 1:nrow(x)
		text(sqrt(x$Area), x$slope_sim, labels = x$.n)
		lines(a * 1e4, u)
		lines(a * 1e4, l)
		mtext(i)
	}
}
check_funnels("Med.F.sc", out_med$sim_varexp)
check_funnels("Min.F.sc", out_min$sim_varexp)
```

This doesn't look quite as extreme as in your plots, Kyle, but 1 out of 9 is expanding with area.

Here's an example culprit:

```{r}
x <- readRDS(paste0("sim-slopes-", "Med.F.sc", "-", 6, ".rds"))
x[c(2, 3, 10, 34),]
```

Look at those huge sigmas and phis for rows 2 and 3 compared to 10 and 34 (typical sites).

And here's what it can create:

```{r}
x$.n <- 1:nrow(x)
plot(sqrt(x$Area), x$slope_sim, type = "n")
text(sqrt(x$Area), x$slope_sim, labels = x$.n)
```

A log Area axis helps show what the model sees:

```{r}
# log axis:
plot(log(x$Area), x$slope_sim, type = "n")
text(log(x$Area), x$slope_sim, labels = x$.n)
```

These are some underlying real time series causing this:

```{r}
library(ggplot2)
q <- subset(Y.Data, Station.ID %in% c("08JC002", "08JC001"))
ggplot(q, aes(Year.Center, Med.F.sc)) + geom_line() + 
	facet_wrap(~Station.ID, scales = "free_y")
```

So it's the area-sigma-phi relationship that's causing this.

For min flow, big sites have low variability and little autocorrelation:

```{r}
ggplot(out_min$real_slopes, aes(log(Area), sigma, size = phi)) + geom_point() +
	ggtitle("Minimum flow")
```

For median flow, there are a few moderately big sites that have high variability and fairly high autocorrelation:

```{r}
ggplot(out_med$real_slopes, aes(log(Area), sigma, size = phi)) + geom_point() +
		ggtitle("Median flow")
```

It's those same two sites:

```{r}
subset(out_med$real_slopes, log(Area) > 10 & sigma > 0.35)
```

The other sorta culprit is the 3rd one here:

```{r}
subset(out_med$real_slopes, log(Area) > 9 & sigma > 0.35)
```

### 2017-01-11 drafts/Nature Climate Change/methods.tex
River-Network Climate Portfolio: Initial Push
Here I add the code for my PhD thesis chapter 1 and two rejected
manuscripts.

\documentclass[12pt]{article}
\usepackage{geometry} 
\geometry{letterpaper}
\usepackage{graphicx}
%\usepackage{pdfpages}
%\usepackage{lscape}
%\usepackage{pdflscape}
%\usepackage{setspace}
%\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{upgreek}

\title{}
\author{}
\date{}

\begin{document}
%\begin{spacing}{1.1}
%\doublespacing
%\onehalfspacing  
\maketitle
%\tableofcontents
%\thispagestyle{empty}
We calculated the probability that we were observing attenuation as the percentage of simulated $\hat{\updelta}$ that were greater than the observed $\hat{\updelta}$ (a negative $\hat{\updelta}$ indicates decreasing variance with increasing area, Fig.~3). We calculated attenuation strength as the ratio between the standard deviation at the smallest and largest  watersheds as defined by the exponential variance function  (i.e.,\ $\sqrt{f(\mathrm{area}_1)} / \sqrt{f(\mathrm{area}_2)}$). To compare the degree of attenuation observed with the null-simulated attenuation, we compared this ratio with the same ratio calculated from each of the null simulations (i.e.,\ $\left( \sqrt{f(\mathrm{area}_1)} / \sqrt{f(\mathrm{area}_2)} \right) / \left( \sqrt{f(\mathrm{area}_{\mathrm{null} 1})} / \sqrt{f(\mathrm{area}_{\mathrm{null} 2})} \right)$). Finally, to estimate how the day-of-year to half-annual-flow was changing, we calculated the steepest point of each site's logistic curve and multiplied this slope value by 365 to revert from our 0--1 scale to our original annual range of 0--365. To get a decadal rate we multiplied these results by 3.8 (number of decades in 38 years). 
\textit{
Could keep this sentence in or just leave it out. This is just a shortcut to calculate the first derivative. Maybe we don't need to describe the shortcut: 
}
We estimated the steepest part of the logistic curve using the \textit{divide by four} ruleADD REF HERE (dividing the coefficient by four equals the first derivative of the logistic curve at its steepest point).

%\end{spacing}
\end{document}

### 2017-01-11 Code_02 Null-Sim_ARIMA-Mod.R

@@ -1,0 +1,66 @@
+# test gls model and arima simulation
+# prove that they are interchangeable
+# answer: yes.
+#
+# gls() with arima residuals is interchangeable with fitted arima model
+# with same order and drift
+
+set.seed(1)
+library(forecast) # for Arima()
+library(dplyr)
+
+out <- plyr::rdply(200, function() {
+
+  y <- as.numeric(arima.sim(n = 60L, model = list(model = c(1, 0, 0),
+        ar = 0.3), sd = 0.2))
+
+  m1_arima <- Arima(y, order = c(1L, 0L, 0L), include.drift = TRUE)
+  phi_arima <- coef(m1_arima)[[1]]
+  intercept_arima <- coef(m1_arima)[[2]]
+  drift_arima <- coef(m1_arima)[[3]]
+  sigma_arima <- sqrt(m1_arima$sigma2)
+
+  m1 <- gls(y~seq_along(y), correlation = corAR1())
+  phi <- coef(m1$model[[1]], unconstrained = FALSE)[[1]]
+  sigma <- m1$sigma
+  intercept <- coef(m1)[[1]]
+  slope <- coef(m1)[[2]]
+
+  ysim <- arima.sim(n = length(y), model = list(order = c(1, 0, 0),
+      ar = phi), sd = sigma) %>% as.numeric()
+
+  m2 <- gls(ysim~seq_along(ysim), correlation = corAR1())
+  phi_sim <- coef(m2$model[[1]], unconstrained = FALSE)[[1]]
+  sigma_sim <- m2$sigma
+  intercept_sim <- coef(m2)[[1]]
+
+  data.frame(phi, sigma, phi_sim, sigma_sim, phi_arima, intercept_arima,
+    sigma_arima, drift_arima, slope, intercept)
+})
+
+out <- out %>% mutate(phi_re = (phi_sim - phi) / phi,
+  sigma_re = (sigma_sim - sigma) / sigma)
+
+par(mfrow = c(3, 2))
+hist(out$phi)
+abline(v = 0.3, col = "red")
+hist(out$sigma)
+abline(v = 0.2, col = "red")
+hist(out$phi_sim)
+abline(v = 0.3, col = "red")
+hist(out$sigma_sim)
+abline(v = 0.2, col = "red")
+hist(out$phi_re)
+abline(v = 0, col = "red")
+hist(out$sigma_re)
+abline(v = 0, col = "red")
+
+par(mfrow = c(2, 2))
[…]

### 2017-01-11 Code_02_Post_Analysis_Null_Test.R

@@ -1,0 +1,167 @@
+load("03_Data_Annual.RData")
+load("03_Data_Monthly.RData")
+logit = function(p){log(p/(1-p))}
+Y.Data = plyr::ddply(Y.Data,"Station.ID",plyr::mutate, med.log.sd = scale(Median.F, center = F), max.log.sd = scale(Max.F, center = F), min.log.sd = scale(Min.F, center = F), DOY2.logit = logit(DOY2/365), Year.Center = Year-1988)
+M.Data = plyr::ddply(M.Data, "Station.ID", plyr::mutate, med.log.sd = scale(median.M.flow, center = F),max.log.sd = scale(max.M.flow, center = F), min.log.sd = scale(min.M.flow, center = F), Year.Center = Year-1988)
+#Scaling the data by site has no impact on the results, rather it only stands to reduce the likelyhood of computational issues during model bootstraping.
+
+# this function fits slopes to real data
+fit_slopes <- function(flow.dat, response) {
+	area_dat <- unique(flow.dat[,c("Station.ID", "Area")])
+	equation <- as.formula(paste(response,"~Year.Center"))
+	models <- plyr::ddply(flow.dat,"Station.ID", function(x){
+		library(nlme)
+		mod <- gls(equation, correlation = corAR1(), data = x)
+		slope <- coef(mod)[[2]]
+		intercept <- coef(mod)[[1]]
+		se <- summary(mod)$tTable[2,2]
+		sigma <- mod$sigma
+		phi <- coef(mod$model[[1]], unconstrained = F)[[1]]
+		data.frame(intercept,slope,se,sigma,phi)
+	})
+	models <- plyr::join(models,area_dat, by = "Station.ID")
+	models
+}
+
+# this function takes a data frame from fit_slopes() and generates a simulated
+# time series (WITHOUT SLOPE) plus a gls() fit to that simulated data
+sim_slopes <- function(slope.dat, yrs, return_ts = FALSE){
+	out <- plyr::adply(slope.dat, 1, function(x) {
+		y <- as.numeric(arima.sim(n = length(yrs),
+															model = list(order = c(1, 0, 0), ar = x$phi),
+															mean = 0, sd = x$sigma)) + x$intercept
[…]

### 2017-01-11 Code_02b lme_mod.R

@@ -1,0 +1,60 @@
+library(dplyr);library(nlme);library(ggplot2)
+load("02a_Data_Clean.RData")
+names(Data) <- tolower(names(Data))
+dat = Data; rm(Data)
+
+#Split up the date for processesing.
+dat = dat %>% group_by(station.id) %>% 
+	mutate(.,flow.scale = scale(flow.data,center = F))
+dat$year.center = dat$year-1988
+
+doParallel::registerDoParallel(cores = parallel::detectCores()-1)
+
+lmeCoef = plyr::ddply(dat,"station.id",function(x){
+	#browser()
+	print(unique(x$station.id))
+	mod = lme(log(flow.scale) ~ year.center, random = ~1+year|doy, data = x, control = lmeControl(maxIter = 5000L, msMaxIter = 5000L))
+	int = mod$coefficients$fixed[[1]]
+	slope = mod$coefficients$fixed[[2]]
+	se = summary(mod)[[20]][4]
+	data.frame(int,slope,se)
+},.parallel = F)
+
+load("01_Data-Wtshd-Orig.RData")
+names(WS.2) = c("station.id","lat","long","area")
+lmeCoef = left_join(lmeCoef,WS.2,by = "station.id")
+
+limits = aes(ymax = slope + se, ymin = slope - se)
+ggplot(lmeCoef, aes(sqrt(area),slope)) +
+	geom_point() +
+	geom_errorbar(limits) + 
+	theme_classic()
+
+m <- gls(slope~sqrt(area), data = lmeCoef, weights = varExp(form= ~sqrt(area)/1e3), 
+				 control = glsControl(maxIter = 1000L, msMaxIter = 1000L))
+summary(m)
+
+
+
+
+#Analaysis by single site.
+samp.dat = filter(dat,station.id == "08MH103")
+mod = lme(log(flow.scale) ~ year.center, random = ~1+year|doy, data = samp.dat, control = lmeControl(maxIter = 5000L, msMaxIter = 5000L))
+summary(mod)
+mod.fit = data.frame(mod$fitted)
+names(mod.fit) = c("fixed", "rand")
+samp.dat = bind_cols(samp.dat, mod.fit)
+
+ggplot(samp.dat, aes(year,log(flow.scale), by = factor(doy), colour = doy)) +
+	geom_point() +
+	geom_line(aes(year, rand)) +
[…]

### 2017-01-11 Code_04 Annual_Funnel_Plot.R

@@ -1,0 +1,229 @@
+load("out_doy2.RData")
+load("out_min.RData")
+load("out_max.RData")
+load("out_med.RData")
+
+funnel.plots = function(x, col1="#737373", col2="#252525", col3="#737373",col4="#252525", pcol="#737373", lcol="#969696", shade1="99", shade2="10", axis.L, yaxis.break = F, iter = 200, xaxis = F, d.xaxis = F, pane, xcoord, ycoord){
+	library(magrittr);library(wesanderson)
+	#Real Data: Coefficients and Area
+	Real.Dat = dplyr::tbl_df(x$real_slopes)%>%plyr::mutate(sqrt.Area = sqrt(Area))%>%plyr::arrange(Area)%>%plyr::mutate(., se.upper = slope+se, se.lower = slope-se); Area = sort(x$real_slopes$Area)
+	#Initial Plot Region Parameters
+	par(mar = c(0,5,0.2,0), family = "serif", bg = "white",fg = "white")
+
+	#Control Panel
+	Line.Width = 1
+	Axis.Label.Size = 1.1
+	Lab.Loc = 3.1
+	vline = 4
+	
+	#Identify Plot Limits and Create Plotting Region with Axis.
+	if(yaxis.break == TRUE){
+		ylimits = pretty(c(min(Real.Dat$slope),max(Real.Dat$slope)))
+		label = round((ylimits/4*38*365)/3.8,0)
+	} else{
+		ylimits = pretty(c(min(Real.Dat$se.lower), max(Real.Dat$se.upper)))
+		label = round(exp(ylimits*10)*100-100,0)
+	}
+		plot(slope~sqrt.Area, xlim = c(0,500), ylim = range(ylimits), xlab = "", ylab = "", axes = F, data = Real.Dat)
+		par(fg = "black")
+		axis(2, lwd = Line.Width, cex.axis = Axis.Label.Size, las = 1, mgp = c(2,0.70,0), at = ylimits, labels = label)
+		mtext(axis.L, side=2, line=Lab.Loc, cex = Axis.Label.Size, las = 0)  #y-axis label.
+	
+	if(xaxis == TRUE){
+		axis(1, lwd = Line.Width, cex.axis = Axis.Label.Size, labels = c("0","10000","40000","90000","160000","250000"), at = c(0,100,200,300,400,500), mgp = c(2,0.70, 0), outer = T, hadj = 1)
[…]

### 2017-01-11 Code_05 Monthly_Plot.R

@@ -1,0 +1,153 @@
+load("max_month.RData")
+load("med_month.RData")
+load("min_month.RData")
+
+#Global Control Panel	
+Line.Width = 1
+Axis.Label.Size = 1.1
+par(family = "serif", bg = "white", fg = "black")
+Lab.Loc = 3.1
+vline = 4
+shade1="99"; shade2="10"
+axisLabCol = "black"
+
+funnel.plots = function(dat, col1="#F2AD00", col2="#67A9CF", col3="#F98400", col4="#FF0000", col5="#525252", pcol="#737373", lcol="#969696",funnel.y = "", axis.L, yaxis.break = F, iter = 200, xaxis = F, d.xaxis = F, pane, xcoord, ycoord, y.axis = F){
+	
+	#Real Data: Coefficients and Area
+	Real.Dat = dplyr::tbl_df(dat$real_slopes)%>%plyr::mutate(sqrt.Area = sqrt(Area))%>%plyr::arrange(Area)%>%plyr::mutate(., se.upper = slope+se, se.lower = slope-se); Area = sort(dat$real_slopes$Area)
+	
+	#Initial Plot Region Parameters
+	par(mar = c(0,5,0.2,0), bg = "white",fg = "white")
+	
+	#Create Plotting Region
+	ylimits = pretty(range(scale(c(Real.Dat$se.upper,Real.Dat$se.lower),center = dat$real_varexp$intercept,scale = F)))
+	label = round(exp(ylimits*10)*100-100,0)
+	plot(slope~sqrt.Area, xlim = c(0,500), ylim = range(ylimits), xlab = "", ylab = "", axes = F, data = Real.Dat)
+	
+	#Add y-axis and labels
+	#par(fg = "white")
+	par(fg = "black")
+	axis(2, lwd = Line.Width, cex.axis = Axis.Label.Size, las = 1, at = ylimits, labels = label, col.axis = axisLabCol)
+	legend("topright", axis.L, bty="n", xjust = 0, cex = Axis.Label.Size)
+	#mtext(axis.L, side=2, line=Lab.Loc, cex = Axis.Label.Size, las = 0)  #y-axis label.
+	if( y.axis == T) {
+		mtext(funnel.y, side=2, at = 0.05, line=Lab.Loc, cex = Axis.Label.Size, las = 0)  #y-axis label.
+	}
+	
+	#Add x-axis w/wo labels.
[…]

### 2017-01-11 Code_06 Flow_Time_Plot.R

@@ -1,0 +1,77 @@
+#library(dplyr)
+load("03_Data_Annual.RData")
+
+Y.Data = Y.Data %>%
+	group_by(Station.ID) %>%
+	mutate(Med.sc = as.vector(scale(Median.F, center = F)),
+				 Min.sc = as.vector(scale(Min.F, center = F)),
+				 Max.sc = as.vector(scale(Max.F, center = F)))
+
+Flow.Time = function(time, response, weight, group, alpha = 0.25, seq.inc = 0.083, Te.cex = 0.7, xlab= "time", ylab ="response"){
+	#vector scaled 0-1 for function 
+	col = ((weight-min(weight))/(diff(range(weight))))^0.3
+	
+	#color ramp function
+	library(RColorBrewer)
+	Blues = brewer.pal(9, "Blues")#; RdYlBu = RdYlBu[-c(5,6,7)]
+	FUN = colorRamp(Blues[c(2:7)], bias=1)
+	
+	#apply function
+	cols = FUN(col)
+	cols = rgb(cols, maxColorValue=256)
+	cols = paste(cols, "99", sep = "")
+	
+	#Determination of point/line size and legend point data.
+	Area.width = ((((weight-min(weight))/(diff(range(weight))))^0.3)+0.01)*8 #Line width.
+	Area.size = (((weight-min(weight))/(diff(range(weight))))^0.3)*3 #Point size.
+	Area.values = c(500,1500,10000,100000,200000) #Area values of interest for the legends.
+	Legend.Area.size = (((Area.values-min(weight))/(diff(range(weight))))^0.3)*3 
+	#Equivalent size given legend point size values.
+	Legend.Area.size.lines = (((Area.values-min(weight))/(diff(range(weight))))^0.3)*8
+	#Equivalent size given legend line size values.
+	Legend.x = seq(1999,2006,length.out = 5) #x values for line/point legend.
+	Legend.y = rep(max(pretty(response)), 5) #y values for line/point legend.
+	Legend.x.col = seq(1999,2006,length.out = 20000) #x values for color legend.
+	Legend.y.col = rep(max(pretty(response)), 20000) #y values for color legend.
+	
+	#color legend point data scaled from 0-1.
+	legend.col = (Legend.x.col-min(Legend.x.col))/(diff(range(Legend.x.col)))
[…]

### 2017-01-11 Code_07 Supplementary_Plots.R

@@ -1,0 +1,90 @@
+#supplementary plots
+library(dplyr)
+load("out_med.RData")
+load("03_Data_Annual.RData")
+logit = function(p){log(p/(1-p))}
+Y.Data = plyr::ddply(Y.Data,"Station.ID",plyr::mutate, med.log.sd = scale(Median.F, center = F), max.log.sd = scale(Max.F, center = F), min.log.sd = scale(Min.F, center = F), DOY2.logit = logit(DOY2/365), Year.Center = Year-1988)
+
+
+dat_sim = out_med$example_ts; names(dat_sim)[2] = "Year.Center"
+dat = dplyr::left_join(dat_sim, Y.Data, by = c("Station.ID", "Year.Center"))
+Adjuster = Y.Data %>%
+	group_by(Station.ID) %>%
+	summarise(adjust = sqrt(mean(Median.F^2)))
+
+dat = dplyr::left_join(Adjuster, dat, by = "Station.ID")
+dat = dat %>% mutate(med_sim_unscale = exp(y)*adjust)
+
+ggplot(dat, aes(Year, med_sim_unscale)) +
+	geom_line(colour = "#F2AD00") +
+	geom_line(aes(Year.Center + 1988, Median.F), colour = "#67A9CF", alpha = 0.5) +
+	facet_wrap(~Station.ID, scales = "free_y") +
+	theme_classic(base_size = 9) +
+	annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
+	annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf) +
+	theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm")) +
+	labs(x = "Year", y = expression("Simulated & Observed Annual Median-Flow (m"^3%.%"sec"^-1*")"%.%"year"^-1))
+
+ggsave("FigS3_site_simulation.pdf", width = 11, height = 8.5)
+
+#clean up NA's in simulated data where no observed data existed.
+dat$Area = apply(dat, 1, function(x){
+	#browser()
+	if(is.na(x["Area"])){
+		as.numeric(na.omit(unique(dat[which(dat$Station.ID==x["Station.ID"]),"Area"])))
+	} else as.numeric(x["Area"])
+})
+
+# this function fits slopes to real data
+fit_slopes <- function(flow.dat, response) {
[…]

### 2017-01-11 Code_08 ClimatePortfolio.R

@@ -1,0 +1,141 @@
+library(readr); library(tidyverse);library(nlme);library(multidplyr)
+setwd("~/sfuvault/Simon_Fraser_University/PhD_Research/Projects/Data/Original_Data/GIS_Data/BC Climate Data/Fraser_Annual_1970_2007")
+
+#Read in all the data and compile into a single dataframe.
+df <- dir() %>% 
+	lapply(read_csv) %>% 
+	bind_rows
+
+#Select only variables of interest.
+df = df %>% select(Year, ID1, Latitude, Longitude, Elevation, MAT, MAP,PAS, EMT, EXT)
+names(df) = tolower(names(df))
+
+#Calcuate a climate trend at each location for each climate variable. Retain only the slope coefficients.
+cluster = create_cluster(8) #Initialize cores to be used.
+df = df %>% partition(id1, cluster = cluster) #Partition the data across nodes.
+df %>% cluster_library("nlme") #Load necessary libraries across cores.
+df %>% cluster_library("dplyr") #Load necessary libraries across cores.
+
+#Perform Analysis
+result = df %>% group_by(id1) %>% do({
+	mod.mat = gls(mat~year, data = ., correlation = corAR1())
+	mod.map = gls(map~year, data = ., correlation = corAR1())
+	mod.pas = gls(pas~year, data = ., correlation = corAR1())
+	mod.emt = gls(pas~year, data = ., correlation = corAR1())
+	mod.ext = gls(pas~year, data = ., correlation = corAR1())
+	
+	data_frame(matS = coef(mod.mat)[[2]], mapS = coef(mod.map)[[2]], pasS = coef(mod.pas)[[2]],
+						 emtS = coef(mod.emt)[[2]], extS = coef(mod.ext)[[2]])
+})
+result = collect(result) #Gather results from cores.
+
+#Add back in lat/long values.
+
+result = df %>% select(id1, latitude, longitude) %>% left_join(., result, by = "id1")
+
+#For QGIS viewing.
+#result = read.csv(file = "~/Desktop/climate_slopes.csv")
+#result$matS =  result$matS / 10^7
+#result$mapS =  result$mapS / 10^7
+#result$pasS =  result$pasS / 10^7
+#result$emtS =  result$emtS / 10^7
[…]

### 2017-01-11 Shiny_App1/app.R
Flow Regime Shiny App
This Shiny App allows you to play with a loess smoother of annual flow
data for 55 flow gauge sites within the Fraser River basin. The size of
points and lines are proportional to the contributing area for that
site.

@@ -1,0 +1,109 @@
+library(shiny) #Call shiny library
+
+#######################################
+#### Initialize the User Interface ####
+#######################################
+
+ui <- fluidPage(
+	includeCSS("style.css"),
+	fluidRow(
+		column(2),
+		column(4,tags$h4("Fraser River Annual Flow", align = "center"), style = "height: 70px; border-right: 1px solid #EEEEEE; top: 25px; position: relative"),
+		column(4, sliderInput(inputId = "num",
+													label = "Smoother Period (Years)",
+													value = 10, min = 3, max = 15, ticks = F)
+					 , align = "center", offset = 1, style = "font-family: 'Raleway'; postion: relative; top: 30px"),
+		column(2)
+	),
+	tags$hr(),
+	fluidRow(
+		column(12, plotOutput("plot"), align = "center")
+	)
+)
+
+#Load Data
+load("app_data.RData")
+
+#Set initial parameters
+time = Y.Data$Year; response = Y.Data$Med.sc; weight = Y.Data$Area; group = Y.Data$Station.ID; xlab = "Year"; ylab = expression("Scaled Median Flow (m"^3%.%"sec"^-1*")"%.%"year"^-1); seq.inc = 0.1; Te.cex = 1.3
+
+##################################
+#### Colour and Scale of Data ####
+##################################
+
+#vector scaled 0-1 for function 
+col = ((weight-min(weight))/(diff(range(weight))))^0.3
+
+#color ramp function
+library(RColorBrewer)
+Blues = brewer.pal(9, "Blues")#; RdYlBu = RdYlBu[-c(5,6,7)]
+FUN = colorRamp(Blues[c(2:7)], bias=1)
+
+#apply function
+cols = FUN(col)
+cols = rgb(cols, maxColorValue=256)
+cols = paste(cols, "99", sep = "")
+
+#####################################
+#### Legend Placement and Colour ####
+#####################################
+
+#Determination of point/line size and legend point data.
+Area.width = ((((weight-min(weight))/(diff(range(weight))))^0.3)+0.01)*8 #Line width.
+Area.size = (((weight-min(weight))/(diff(range(weight))))^0.3)*3 #Point size.
+Area.values = c(500,1500,10000,100000,200000) #Area values of interest for the legends.
[…]

### 2017-02-28 Code_01 Data_Organization.Rmd
Major Overhaul
Updated the methods for data summary. Removed redundant or unnecessary
code files. Recoded the Annual funnel plot and added a climate
portfolio analysis file.

@@
 ###Load Original Raw Data.
-```{r,echo=FALSE,message=FALSE}
-load("01_Data-Flow-Orig.RData")
-load("01_Data-Wtshd-Orig.RData")
-names(WS.2) = c("Station.ID","Lat","Long","Area")
-head(Flow.2)
-head(WS.2)
-
-PDO = read.csv("http://www.ncdc.noaa.gov/teleconnections/pdo/data.csv",skip = 1, header = T)
-PDO = PDO[which(PDO$Date>196912),]
-ENSO = read.csv("http://www.ncdc.noaa.gov/teleconnections/enso/indicators/soi/data.csv",skip = 1, header = T)
-ENSO = ENSO[which(ENSO$Date>196912),]
-
+```{r, message=FALSE, cache=TRUE, include=FALSE}
 #Load in libraries.
-library(dplyr);library(plyr);library(tidyr);library(MARSS)
+library(plyr);library(tidyverse);library(MARSS);library(zoo);library(lubridate)
 ```
@@
 ###Subset and Constrain Data
-```{r, echo=FALSE, message=F}
+```{r, echo=FALSE, message=F, dependson=1}
+load("01_Data-Flow-Orig.RData")
 # Remove Dam Influenced Sites.
-Flow.2 = Flow.2[which(Flow.2$Station.ID %in% c("08JC001","08JC002","08ME002")==F),]
+Flow.2 = filter(Flow.2, !(Station.ID %in% c("08JC001","08JC002","08ME002")))
 
 #Limit the HYDAT flow data to years between 1970 and 2007.
-Data = Flow.2[which(as.numeric(format(Flow.2$Date, "%Y"))>=1970 & as.numeric(format(Flow.2$Date, "%Y"))<=2007),]
-# Data <- Data %>% arrange(Station.ID, Date) # ensure correct order
-# Data <- Data %>% group_by(Station.ID) %>%
-	# mutate(gap = c(NA, Date[2:n()] - Date[1:(n() - 1)])) %>%
-	# mutate(any_NAs = ifelse(max(gap > 1, na.rm = TRUE), TRUE, FALSE)) %>%
-	# mutate(long_NA_string = ifelse(max(gap > 7, na.rm = TRUE), TRUE, FALSE)) %>%
-	# as.data.frame()
+Data = Flow.2 %>% filter(year(Date)>=1970, year(Date)<=2007)
 
 #Split up the date for processesing.
-Data$Year = as.numeric(format(Data$Date,"%Y")) #Add a column for the year.
-Data$nMonth = as.numeric(format(Data$Date,"%m")) #Add a column for the month.
-Data$Month = as.factor(format(Data$Date,"%b")) #Add a column for the month.
-Data$Day = as.numeric(format(Data$Date,"%d")) #Add a column for the day.
-Data$DOY = as.numeric(format(Data$Date,"%j")) #Add a column for the day.
-head(Data)
+Data = Data %>% mutate(Year = year(Date), nMonth = month(Date), Month = month(Date, label = T), Day = day(Date), DOY = yday(Date))
 
-#Create a series of dates between January 1st of 1970 and December 31st of 2007. This will be used as a ruler for determining missing days in each month of each year.
-Date.Cont = seq(as.Date("1970-01-01"), as.Date("2007-12-31"), by = "day")
+#Determine the number of days missing in each month of each year by flow gauge station (DIMM = Days In Month Missing). Remove months with more then 5 days missing.
+DIMM = Data %>% group_by(Station.ID, Year, Month) %>%
+	summarize(Days.Missing = unique(days_in_month(Date)-n())) %>% 
+	filter(Days.Missing<=5) %>% ungroup()
 
-#Determine the number of days missing in each month of each year by flow gauge station (DIMM = Days In Month Missing).
-years.continuous <- as.numeric(format(Date.Cont,"%Y"))
-months.continuous <- as.numeric(format(Date.Cont,"%m"))
-DIMM <- Data %>% group_by(Station.ID, Year, Month) %>%
-	dplyr::summarise(Days.Missing =
-			length(Date.Cont[which(years.continuous == unique(Year) & 
-							  months.continuous == unique(nMonth))]) - n()) %>%
-	as.data.frame()
-
-#Remove months with more then 5 days missing.
-DIM.Limited.Months = DIMM[which(DIMM$Days.Missing<=5),]
-
-#Determine the number of months in each year for each flow gauge site.
-DIM.Limited.Months = ddply(DIM.Limited.Months,c("Station.ID","Year"),function(x){
-	Months = dim(x)[1]
-	data.frame(Months)
-})
-
-#Remove years with fewer than 12 months.
-DIM.Limited.Years1 = DIM.Limited.Months[which(DIM.Limited.Months$Months==12),]
+#Determine the number of months in each year for each flow gauge site. Remove years with fewer than 12 months.
+DIM.Limited.Months = DIMM %>% group_by(Station.ID, Year) %>% 
+	summarize(Months = n()) %>% filter(Months == 12)
 
 #Isolate data with Years that have 12 months in which no month is missing more than 5 days.
-Data = match_df(Data,DIM.Limited.Years1,on = c("Station.ID","Year"))
+Data = plyr::match_df(Data,DIM.Limited.Months,on = c("Station.ID","Year"))
 
-#Determine the number of years included in each site.
-DIM.Limited.Years2 = ddply(DIM.Limited.Years1,c("Station.ID"),function(x){
-	Years = dim(x)[1]
-	data.frame(Years)
-})
-
-#Remove stations with fewer than 35 years of data.
-Stations.Final = DIM.Limited.Years2[which(DIM.Limited.Years2$Years>=35),]
+#Determine the number of years included in each site. Remove stations with fewer than 35 years of data.
+Stations.Final = DIM.Limited.Months %>% summarise(Years = n()) %>% 
+	filter(Years>=35)
 
 #Isolate data such that all included sites have at least 35 of the possible 38 years between 1970 and 2007, and each year has 12 full months where each month is missing no more than five days.
-Data = match_df(Data,Stations.Final,on = "Station.ID")
+Data = plyr::match_df(Data,Stations.Final,on = "Station.ID")
 save(Data,file = "02a_Data_Clean.RData")
 ```
-###Create Monthly Dataset
-```{r,echo=F,message=F}
-#Summarise data monthly mean, minnimum and maximums.
-load("02a_Data_Clean.RData")
-M.Data <- Data %>% group_by(Station.ID, Year, nMonth, Month) %>%
-	dplyr::summarise(median.M.flow = median(Flow.Data),
- 						max.M.flow = max(Flow.Data),
- 						min.M.flow = min(Flow.Data))
 
-M.Data = left_join(M.Data,WS.2,by = c("Station.ID"))
-
-save(M.Data,file = "03_Data_Monthly.RData")
-```
-###Create Annual Dataset
-```{r,echo=F,message=F}
+###Interpolate Mising Data
+```{r,echo=F,message=F, dependson=1}
 #Fill in missing data for the 'day of year' analysis.
@@
 
-all_expected_dat <- expand.grid(unique(Data$Station.ID), Date.Cont,
-	stringsAsFactors = FALSE)
-names(all_expected_dat) <- c("Station.ID", "Date")
-Data <- left_join(all_expected_dat, Data)
-Data <- Data %>% arrange(Station.ID, Date)
-Data$Year = as.numeric(format(Data$Date,"%Y")) #Add a column for the year.
-Data$nMonth = as.numeric(format(Data$Date,"%m")) #Add a column for the month.
-Data$Month = as.factor(format(Data$Date,"%b")) #Add a column for the month.
-Data$Day = as.numeric(format(Data$Date,"%d")) #Add a column for the day.
-Data$DOY = as.numeric(format(Data$Date,"%j")) #Add a column for the day.
+#Split up the date for processesing.
+Data = Data %>% mutate(Year = year(Date), nMonth = month(Date), Month = month(Date, label = T), Day = day(Date), DOY = yday(Date))
 
 #Determine the number of missing days in each Station, Year, Month combination. Retain those that have more than 0.
-Missing <- Data %>% group_by(Station.ID, Year, Month) %>%
-	dplyr::summarise(Days.Missing =
-    length(Date.Cont[which(years.continuous == unique(Year) & 
-      months.continuous == unique(nMonth))]) - n()) %>%
-	dplyr::filter(Days.Missing > 0)
+Missing = Data %>% group_by(Station.ID, Year, nMonth) %>% 
+	summarize(Days.Missing = unique(days_in_month(Date)-n())) %>% 
+	filter(Days.Missing > 0)
 
@@
 This creates a vector/dataset in the wide format which is necessary for the auto-regressive state-space model.
-Sites = dlply(Missing,c("Station.ID","Year"),function(x){
+Sites = plyr::dlply(Missing,c("Station.ID","Year"),function(x){
+	#browser()
 	Period = seq(unique(x$Year)-1,unique(x$Year+1))
-	Site = subset(Data,Station.ID==x$Station.ID & Year%in%Period)
+	Site = subset(Data, Station.ID==x$Station.ID & Year%in%Period)
 	Potential.Dates = seq(as.Date(paste(Period[1],"-01-01",sep="")), as.Date(paste(Period[3],"-12-31",sep="")), by = "day")
@@
 # Interpolate the missing values via a linear interpolation (in log space):
-Preds = ldply(names(Sites),function(x){
+Preds = plyr::ldply(names(Sites),function(x){
 	Year = as.numeric(strsplit(x,split = "\\.")[[1]][2])
@@
 })
-Preds$Year = as.numeric(format(Preds$Date,"%Y")) #Add a column for the year.
-Preds$nMonth = as.numeric(format(Preds$Date,"%m")) #Add a column for the month.
-Preds$Month = as.factor(format(Preds$Date,"%b")) #Add a column for the month.
-Preds$Day = as.numeric(format(Preds$Date,"%d")) #Add a column for the day.
-Preds$DOY = as.numeric(format(Preds$Date,"%j")) #Add a column for the day.
 
-save(Preds,file = "02b_Missing_Data_Predictions.RData")
+#Split up the date for processesing.
+Preds = Preds %>% mutate(Year = year(Date), nMonth = month(Date), Month = month(Date, label = T), Day = day(Date), DOY = yday(Date))
 
-#Combine predictions with the original data.
+Data.Preds = bind_rows(Data, Preds)
+save(Data.Preds, file = "02b_Missing_Data_Predictions.RData")
+```
+
+###Create Annual Dataset
+```{r,echo=F,message=F,dependson=1}
+#Load Data
+load("01_Data-Wtshd-Orig.RData"); names(WS.2) = c("Station.ID","Lat","Long","Area")
 load("02b_Missing_Data_Predictions.RData")
-Data.Preds = full_join(Data,Preds)
-Data.Preds$Station.ID = as.factor(Data.Preds$Station.ID)
-Data.Preds$Month = as.factor(Data.Preds$Month)
+
+#Calculate Rolling Average to smooth out daily extremes.
+flow.stats = Data.Preds %>% group_by(Station.ID) %>% 
+	arrange(Date) %>%
+	do({
+		z = zoo(.$Flow.Data, .$Date)
+		x = rollapply(data = z, width = 5, by = 1, FUN = mean)
+		data.frame(Date = index(x), mean5day = coredata(x))
+	})
 
@@
 # Get the row numbers of when cumulative flow volume hits half the total yearly volume:
-accum_flow <- Data.Preds %>%
+accum_flow <- Data.Preds %>% left_join(., flow.stats, by  = c("Station.ID", "Date")) %>% 
 	mutate(row_number = 1:length(Year)) %>%
@@
 	inner_join(accum_flow) %>%
-	select(-row_number) %>%
-	rename(DOY2 = DOY)
-Y.Data = left_join(accum_flow, WS.2, by = c("Station.ID"))
-Y.Data$Station.ID = as.factor(Y.Data$Station.ID)
+	select(-row_number)
 
-flow.stats = ddply(Data.Preds, c("Station.ID","Year"),summarize,Median.F = median(Flow.Data),Min.F = min(Flow.Data),Max.F = max(Flow.Data))
-Y.Data = left_join(flow.stats,Y.Data, by = c("Station.ID","Year"))
+# Summarize flow data.
+flow.stats = left_join(Data.Preds, flow.stats, by = c("Station.ID","Date")) %>% 
+	group_by(Station.ID, Year) %>% 
+	summarise(Median.F = median(Flow.Data),
+									 Min.F = min(mean5day, na.rm = T),
+									 Max.F = max(mean5day, na.rm = T))
 
+Y.Data = left_join(accum_flow, flow.stats, by = c("Station.ID","Year"))
+Y.Data = left_join(Y.Data, WS.2, by = c("Station.ID"))
+names(Y.Data)[8] = "DOY2"
 save(Y.Data,file = "03_Data_Annual.RData")
 ```
+
+###Create Monthly Dataset
+```{r,echo=F,message=F,dependson=1}
+#Summarise data monthly mean, minnimum and maximums.
+load("01_Data-Wtshd-Orig.RData"); names(WS.2) = c("Station.ID","Lat","Long","Area")
+load("02b_Missing_Data_Predictions.RData")
+df = Data.Preds %>% group_by(Station.ID) %>% arrange(Date) %>%
+	do({
+		z = zoo(.$Flow.Data, .$Date)
+		x = rollapply(data = z, width = 5, by = 1, FUN = mean)
+		data.frame(Date = index(x), mean5day = coredata(x))
+	})
+
+M.Data = left_join(Data.Preds, df, by = c("Station.ID","Date")) %>% 
+	group_by(Station.ID, Year, nMonth, Month) %>% 
+	summarise(median.M.flow = median(Flow.Data),
+ 						max.M.flow = max(mean5day, na.rm = T),
+ 						min.M.flow = min(mean5day, na.rm = T))
+#Join location and area data to flow data.
+M.Data = left_join(M.Data, WS.2, by = "Station.ID")
+save(M.Data,file = "03_Data_Monthly.RData")
+```

### 2017-02-28 Code_02b lme_mod.R

@@ -1,60 +1,0 @@
-library(dplyr);library(nlme);library(ggplot2)
-load("02a_Data_Clean.RData")
-names(Data) <- tolower(names(Data))
-dat = Data; rm(Data)
-
-#Split up the date for processesing.
-dat = dat %>% group_by(station.id) %>% 
-	mutate(.,flow.scale = scale(flow.data,center = F))
-dat$year.center = dat$year-1988
-
-doParallel::registerDoParallel(cores = parallel::detectCores()-1)
-
-lmeCoef = plyr::ddply(dat,"station.id",function(x){
-	#browser()
-	print(unique(x$station.id))
-	mod = lme(log(flow.scale) ~ year.center, random = ~1+year|doy, data = x, control = lmeControl(maxIter = 5000L, msMaxIter = 5000L))
-	int = mod$coefficients$fixed[[1]]
-	slope = mod$coefficients$fixed[[2]]
-	se = summary(mod)[[20]][4]
-	data.frame(int,slope,se)
-},.parallel = F)
-
-load("01_Data-Wtshd-Orig.RData")
-names(WS.2) = c("station.id","lat","long","area")
-lmeCoef = left_join(lmeCoef,WS.2,by = "station.id")
-
-limits = aes(ymax = slope + se, ymin = slope - se)
-ggplot(lmeCoef, aes(sqrt(area),slope)) +
-	geom_point() +
-	geom_errorbar(limits) + 
-	theme_classic()
-
-m <- gls(slope~sqrt(area), data = lmeCoef, weights = varExp(form= ~sqrt(area)/1e3), 
-				 control = glsControl(maxIter = 1000L, msMaxIter = 1000L))
-summary(m)
-
-
-
-
-#Analaysis by single site.
-samp.dat = filter(dat,station.id == "08MH103")
-mod = lme(log(flow.scale) ~ year.center, random = ~1+year|doy, data = samp.dat, control = lmeControl(maxIter = 5000L, msMaxIter = 5000L))
-summary(mod)
-mod.fit = data.frame(mod$fitted)
-names(mod.fit) = c("fixed", "rand")
-samp.dat = bind_cols(samp.dat, mod.fit)
-
-ggplot(samp.dat, aes(year,log(flow.scale), by = factor(doy), colour = doy)) +
-	geom_point() +
-	geom_line(aes(year, rand)) +
[…]

### 2017-02-28 Code_03_Post_Analysis_Null_Test.R

@@ -1,13 +1,42 @@
+library(tidyverse); library(stringr)
+#Load/Organize Annual Data
 load("03_Data_Annual.RData")
-load("03_Data_Monthly.RData")
 logit = function(p){log(p/(1-p))}
-Y.Data = plyr::ddply(Y.Data,"Station.ID",plyr::mutate, med.log.sd = scale(Median.F, center = F), max.log.sd = scale(Max.F, center = F), min.log.sd = scale(Min.F, center = F), DOY2.logit = logit(DOY2/365), Year.Center = Year-1988)
-M.Data = plyr::ddply(M.Data, "Station.ID", plyr::mutate, med.log.sd = scale(median.M.flow, center = F),max.log.sd = scale(max.M.flow, center = F), min.log.sd = scale(min.M.flow, center = F), Year.Center = Year-1988)
+Y.Data = Y.Data %>% group_by(Station.ID) %>% mutate(med.log.sd = scale(Median.F, center = F), max.log.sd = scale(Max.F, center = F), min.log.sd = scale(Min.F, center = F), DOY2.logit = logit(DOY2/365), Year.Center = Year-1988)
+#Load/Orgnize Montly Data
+load("03_Data_Monthly.RData")
+M.Data = M.Data %>% group_by(Station.ID) %>% mutate(med.log.sd = scale(median.M.flow, center = F), max.log.sd = scale(max.M.flow, center = F), min.log.sd = scale(min.M.flow, center = F), Year.Center = Year-1988)
 #Scaling the data by site has no impact on the results, rather it only stands to reduce the likelyhood of computational issues during model bootstraping.
 
+#Incorporate the Climate Data
+setwd("~/sfuvault/Simon_Fraser_University/PhD_Research/Projects/Data/Original_Data/GIS_Data/BC_Climate_Data/ClimateTrends/zonal_stats")
+vars = unlist(lapply(strsplit(dir(), "-"), function(x) x[1]))
+AnnualVars = vars[which(nchar(vars) == 3)]
+MonthlyVars = vars[which(nchar(vars) > 3 & nchar(vars)<7)]
+#Annual
+for(i in AnnualVars){
+	cDat = read.table(paste(i,"-stats.txt",sep = ""), header = T, as.is = T) %>% select(file, std)
+	names(cDat) = c("Station.ID", paste(i,".sd", sep = ""))
[…]

### 2017-02-28 Code_04 Annual_Funnel_Plot.R

@@ -1,229 +1,0 @@
-load("out_doy2.RData")
-load("out_min.RData")
-load("out_max.RData")
-load("out_med.RData")
-
-funnel.plots = function(x, col1="#737373", col2="#252525", col3="#737373",col4="#252525", pcol="#737373", lcol="#969696", shade1="99", shade2="10", axis.L, yaxis.break = F, iter = 200, xaxis = F, d.xaxis = F, pane, xcoord, ycoord){
-	library(magrittr);library(wesanderson)
-	#Real Data: Coefficients and Area
-	Real.Dat = dplyr::tbl_df(x$real_slopes)%>%plyr::mutate(sqrt.Area = sqrt(Area))%>%plyr::arrange(Area)%>%plyr::mutate(., se.upper = slope+se, se.lower = slope-se); Area = sort(x$real_slopes$Area)
-	#Initial Plot Region Parameters
-	par(mar = c(0,5,0.2,0), family = "serif", bg = "white",fg = "white")
-
-	#Control Panel
-	Line.Width = 1
-	Axis.Label.Size = 1.1
-	Lab.Loc = 3.1
-	vline = 4
-	
-	#Identify Plot Limits and Create Plotting Region with Axis.
-	if(yaxis.break == TRUE){
-		ylimits = pretty(c(min(Real.Dat$slope),max(Real.Dat$slope)))
-		label = round((ylimits/4*38*365)/3.8,0)
-	} else{
-		ylimits = pretty(c(min(Real.Dat$se.lower), max(Real.Dat$se.upper)))
-		label = round(exp(ylimits*10)*100-100,0)
-	}
-		plot(slope~sqrt.Area, xlim = c(0,500), ylim = range(ylimits), xlab = "", ylab = "", axes = F, data = Real.Dat)
-		par(fg = "black")
-		axis(2, lwd = Line.Width, cex.axis = Axis.Label.Size, las = 1, mgp = c(2,0.70,0), at = ylimits, labels = label)
-		mtext(axis.L, side=2, line=Lab.Loc, cex = Axis.Label.Size, las = 0)  #y-axis label.
-	
-	if(xaxis == TRUE){
-		axis(1, lwd = Line.Width, cex.axis = Axis.Label.Size, labels = c("0","10000","40000","90000","160000","250000"), at = c(0,100,200,300,400,500), mgp = c(2,0.70, 0), outer = T, hadj = 1)
[…]

### 2017-02-28 Code_06 Annual_Funnel_Plot.R

@@ -1,0 +1,188 @@
+
+
+#Function for scaling values between 0 and 1.
+zero_one = function(x) ((x-min(x)))/(diff(range(x)))
+
+#Function for simplifying data and adding climate index value.
+arrange.data = function(df, response){
+	tp = df[[grep(response, names(df))]]$real_slopes
+	tp = tp %>% select(8:ncol(tp)) %>% apply(.,2,function(x) zero_one(x)) %>% 
+		apply(.,1,function(x) sum(x)) %>% mutate(tp, std.clim = .)
+	names(tp)[grep(response,names(tp))] = "resp"
+	tp %>% mutate(sqrt.resp = sqrt(resp)) %>% mutate(se.upper = slope+se, se.lower = slope-se) %>% arrange(resp)
+}
+
+#Add colour gradient to climate index.
+color.gradient = function(df, pal = "GnBu", shade = "99"){
+	#vector scaled 0-1 for function
+	col = zero_one(df$std.clim)
+	#color ramp function
+	RdYlBu = brewer.pal(9, pal)[3:9]
+	FUN = colorRamp(RdYlBu, bias=1)
+	#apply function
+	cols = FUN(col)
+	cols = rgb(cols, maxColorValue=256)
+	paste(cols, shade, sep = "")
+}
+
+#Axis label determination
+axis.Fun = function(df, yaxis.DOY=F){
+	if(yaxis.DOY == T) {
+		ylimits <<- pretty(c(min(df$slope), max(df$slope)))
+		label <<- round((ylimits/4*38*365)/3.8,0)
+	} else {
+		ylimits <<- pretty(c(min(df$se.lower), max(df$se.upper)))
+		label <<- round(exp(ylimits*10)*100-100,0)
+	}
+}
+
+#Draw x and y axis.
+axis.Draw = function(LW, ALS, ALX, ALY, LL, YLim, YLab, xAxis = T){
+	if(xAxis == T){
+		axis(1, lwd = LW, cex.axis = ALS, outer = T, hadj = 1,
+				 labels = c("0","10000","40000","90000","160000","250000"), at = c(0,100,200,300,400,500))
+		mtext(ALX, side = 1, line = LL, cex = ALS) # x-axis label.
+	} else axis(1, lwd = LW, tick = TRUE, labels = FALSE, tck = 0)
+	axis(2, lwd = LW, cex.axis = ALS, at = YLim, labels = YLab)
+	mtext(ALY, side=2, line=LL, cex = ALS, las = 0)  #y-axis label.
+}
+
[…]

### 2017-02-28 Code_06 AttenStats.R

@@ -1,0 +1,120 @@
+#Attenuation Stats.
+
+#This function finds the predicted attenuation values at any two catchment areas and returns the percent decrease in predicted range values when moving from small catchments to a large catchment.
+pred.F = function(x, Area){
+	prediction.1= x$real_varexp$slope*(sqrt(Area) - mean(sqrt(Area))) + 1.96 * sqrt(x$real_varexp$sigma^2 * exp(2*(sqrt(Area)/1e3)*x$real_varexp$varexp))
+	prediction.2= x$real_varexp$slope*(sqrt(Area) - mean(sqrt(Area))) - 1.96 * sqrt(x$real_varexp$sigma^2 * exp(2*(sqrt(Area)/1e3)*x$real_varexp$varexp))
+	percent.reduced = (1-sqrt(diff(c(prediction.1[2],prediction.2[2]))^2/diff(c(prediction.1[1],prediction.2[1]))^2))*100
+	data.frame(percent.reduced)}
+
+pred.F(out_max,range(out_max$real_slopes$Area))
+pred.F(out_min,range(out_max$real_slopes$Area))
+pred.F(out_med,range(out_max$real_slopes$Area))
+pred.F(out_doy2,range(out_max$real_slopes$Area))
+pred.F(out_med,c(5000,60000))
+
+#This function produces lower and upper range estimates for trend values among smaller catchments (Small_Area) and then at a larger catchment (Large_Area). Values are returned as the percent change per decade.
+pred.F2 = function(x){
+	upper = x$real_varexp$intercept + x$real_varexp$slope*(sqrt(x$real_slopes$Area) - mean(sqrt(x$real_slopes$Area))) + 1.96 * sqrt(x$real_varexp$sigma^2 * exp(2*(sqrt(x$real_slopes$Area)/1e3)*x$real_varexp$varexp))
+	upper.est = round((exp(sort(range(upper),decreasing = T)*10)-1)*100,2)
+	
+	lower = x$real_varexp$intercept + x$real_varexp$slope*(sqrt(x$real_slopes$Area) - mean(sqrt(x$real_slopes$Area))) - 1.96 * sqrt(x$real_varexp$sigma^2 * exp(2*(sqrt(x$real_slopes$Area)/1e3)*x$real_varexp$varexp))
+	lower.est = round((exp(range(lower)*10)-1)*100,2)
[…]

### 2017-02-28 Code_07 Monthly_Plot.R

@@ -11,10 +11,20 @@ vline = 4
 shade1="99"; shade2="10"
 axisLabCol = "black"
 
-funnel.plots = function(dat, col1="#F2AD00", col2="#67A9CF", col3="#F98400", col4="#FF0000", col5="#525252", pcol="#737373", lcol="#969696",funnel.y = "", axis.L, yaxis.break = F, iter = 200, xaxis = F, d.xaxis = F, pane, xcoord, ycoord, y.axis = F){
+funnel.plots = function(dat, response, col1="#F2AD00", col2="#67A9CF", col3="#F98400", col4="#FF0000", col5="#525252", pcol="#737373", lcol="#969696",funnel.y = "", axis.L, yaxis.break = F, iter = 200, xaxis = F, d.xaxis = F, pane, xcoord, ycoord, y.axis = F){
+	library(tidyverse);library(wesanderson)
+	#browser()
 	
-	#Real Data: Coefficients and Area
-	Real.Dat = dplyr::tbl_df(dat$real_slopes)%>%plyr::mutate(sqrt.Area = sqrt(Area))%>%plyr::arrange(Area)%>%plyr::mutate(., se.upper = slope+se, se.lower = slope-se); Area = sort(dat$real_slopes$Area)
+	#Identify predictor variable data for the given response variable.
+	slot = grep(pattern = response, x = names(dat)); dat = dat[[slot]]
+	#Change predictor variable column name to generic value.
+	column = grep(pattern = response, x = names(dat$real_slopes))
+	names(dat$real_slopes)[column] = "resp"
+	
+	#Real Data: Coefficients and Response Variable isolation and transformation.
+	Real.Dat = dat$real_slopes %>% mutate(sqrt.resp = sqrt(resp)) %>% arrange(resp) %>% 
+		mutate(se.upper = slope+se, se.lower = slope-se)
+	resp = sort(Real.Dat$resp)
 	
 	#Initial Plot Region Parameters
 	par(mar = c(0,5,0.2,0), bg = "white",fg = "white")
@@ -22,7 +32,7 @@ funnel.plots = function(dat, col1="#F2AD00", col2="#67A9CF", col3="#F98400", col
 	#Create Plotting Region
[…]

### 2017-02-28 Code_08 ClimatePortfolio.R

@@ -1,141 +1,0 @@
-library(readr); library(tidyverse);library(nlme);library(multidplyr)
-setwd("~/sfuvault/Simon_Fraser_University/PhD_Research/Projects/Data/Original_Data/GIS_Data/BC Climate Data/Fraser_Annual_1970_2007")
-
-#Read in all the data and compile into a single dataframe.
-df <- dir() %>% 
-	lapply(read_csv) %>% 
-	bind_rows
-
-#Select only variables of interest.
-df = df %>% select(Year, ID1, Latitude, Longitude, Elevation, MAT, MAP,PAS, EMT, EXT)
-names(df) = tolower(names(df))
-
-#Calcuate a climate trend at each location for each climate variable. Retain only the slope coefficients.
-cluster = create_cluster(8) #Initialize cores to be used.
-df = df %>% partition(id1, cluster = cluster) #Partition the data across nodes.
-df %>% cluster_library("nlme") #Load necessary libraries across cores.
-df %>% cluster_library("dplyr") #Load necessary libraries across cores.
-
-#Perform Analysis
-result = df %>% group_by(id1) %>% do({
-	mod.mat = gls(mat~year, data = ., correlation = corAR1())
-	mod.map = gls(map~year, data = ., correlation = corAR1())
-	mod.pas = gls(pas~year, data = ., correlation = corAR1())
-	mod.emt = gls(pas~year, data = ., correlation = corAR1())
-	mod.ext = gls(pas~year, data = ., correlation = corAR1())
-	
-	data_frame(matS = coef(mod.mat)[[2]], mapS = coef(mod.map)[[2]], pasS = coef(mod.pas)[[2]],
-						 emtS = coef(mod.emt)[[2]], extS = coef(mod.ext)[[2]])
-})
-result = collect(result) #Gather results from cores.
-
-#Add back in lat/long values.
-
-result = df %>% select(id1, latitude, longitude) %>% left_join(., result, by = "id1")
-
-#For QGIS viewing.
-#result = read.csv(file = "~/Desktop/climate_slopes.csv")
-#result$matS =  result$matS / 10^7
-#result$mapS =  result$mapS / 10^7
-#result$pasS =  result$pasS / 10^7
-#result$emtS =  result$emtS / 10^7
[…]

### 2017-03-01 tmb-example.R
Add example TMB variance function regression

@@ -1,0 +1,37 @@
+library(TMB)
+compile("varianceFunction.cpp")
+dyn.load(dynlib("varianceFunction"))
+
+set.seed(1234)
+b0 <- 0.5
+b1 <- 0.3
+b2 <- 0.1
+sigma0 <- 0.1
+sigma1 <- -0.2
+sigma2 <- 0.3
+N <- 2000
+s1 <- sample(seq(1, 10), N, replace = TRUE)
+s2 <- sample(seq(1, 10), N, replace = TRUE)
+
+# This is the way that nlme::varExp() has the function set up:
+# var * exp(2 * t * v)
+# This one is set up a little bit differently but has the same general shape:
+u <- rnorm(N, b0 + b1 * s1 + b2 * s2, sqrt(exp(sigma0 + sigma1 * s1 + sigma2 * s2)))
+# matching it exactly:
+# u <- rnorm(N, b0 + b1 * s, sqrt(sigma0^2 * exp(2 * sigma1 * s)))
+# Not using this one right now because I'd have to tweak the model matrix below 
+# and I'm out of time. 
+plot(s1, u)
+plot(s2, u)
+
+mm <- model.matrix(~ s1 + s2)
+
+obj <- MakeADFun(data = list(x_ij = mm, y_i = u),
+  parameters = list(b_j = c(0, 0, 0), sigma_j = c(0, 0, 0)),
+  DLL = "varianceFunction")
+
+opt <- nlminb(start=obj$env$last.par.best, objective=obj$fn, gradient=obj$gr)
+opt$par
+# If you need standard errors and therefore confidence intervals... exp(sigma1 + 1.96 * SE)... :
+rep <- sdreport(obj)
+rep

### 2017-03-01 varianceFunction.cpp

@@ -1,0 +1,28 @@
+#include <TMB.hpp>
+
+template<class Type>
+Type objective_function<Type>::operator() () {
+  // data:
+  DATA_MATRIX(x_ij);
+  DATA_VECTOR(y_i);
+  
+  // parameters:
+  PARAMETER_VECTOR(b_j)
+  PARAMETER_VECTOR(sigma_j);
+  
+  int n_data = y_i.size(); // get number of data points to loop over
+  
+  // Linear predictor
+  vector<Type> linear_predictor_i(n_data);
+  vector<Type> linear_predictor_sigma_i(n_data);
+  linear_predictor_i = x_ij*b_j;
+  linear_predictor_sigma_i = sqrt(exp(x_ij*sigma_j));
+  
+  Type nll = 0.0; // initialize negative log likelihood
+  
+  for(int i = 0; i < n_data; i++){
+    nll -= dnorm(y_i(i), linear_predictor_i(i) , linear_predictor_sigma_i(i), true);
+  }
+  
+  return nll;
+}

### 2017-03-01 Code_03_Post_Analysis_Null_Test.R
Homogenizing Code

@@ -154,6 +154,48 @@ out_med_month <- plyr::dlply(M.Data, "nMonth", function(x)
 save(out_med_month, file = "med_month.RData")
 
 
+
+
+
+set.seed(123)
+vars = c("Area","emt.sd","ext.sd","map.sd","mat.sd","pas.sd")
+real_slopes = fit_slopes(Y.Data, "DOY2.logit", "Area", vars)
+zero_one = function(x) ((x-min(x)))/(diff(range(x)))
+real_slopes = real_slopes %>% select(8:ncol(real_slopes)) %>% apply(.,2,function(x) zero_one(x)) %>% 
+	apply(.,1,function(x) sum(x)) %>% mutate(real_slopes, std.clim = .)
+real_slopes = real_slopes %>% mutate(area_std = sqrt(Area)/sd(sqrt(Area)), clim_std = std.clim/sd(std.clim))
+
+ m <- gls(slope~sqrt(Area), weights = varExp(form = ~sqrt(area_std)+clim_std),
+				 control = glsControl(maxIter = 1000L, msMaxIter = 1000L), data = real_slopes)
+
+plot(slope~sqrt(Area), real_slopes)
+#Add Simulation Lines to Plotting Region.
+line.sims = function(df, col, LW, shade = "10", transp = T, intercept, slope, resp.sqrt, sigma, varexp){
+	apply(df,1,function(y){
+		if(transp == T) colur = paste(col,shade,sep="")
+		else colur = col
+		upper = intercept + slope*(resp.sqrt - mean(resp.sqrt)) + 1.96 * sqrt(sigma^2 * exp(2*(resp.sqrt/1e3)*y[1]))
+		lines(resp.sqrt, upper, col = colur, lwd = LW)
+		lower = intercept + slope*(resp.sqrt - mean(resp.sqrt)) - 1.96 * sqrt(sigma^2 * exp(2*(resp.sqrt/1e3)*y[1]))
+		lines(resp.sqrt, lower, col = colur, lwd = LW)
+	})}
+
+varexp <- m$model[[1]][[1]][1]
+sigma <- m$sigma
+intercept <- coef(m)[[1]]
+slope = m$coefficients[[2]]
+slopePval = summary(m)[[18]][8]
+df1 = data.frame(varexp, sigma, intercept, slope, slopePval)
[…]

### 2017-03-01 Code_06 Annual_Funnel_Plot.R

@@ -1,9 +1,11 @@
-
+#########################################################################################################
+################## Plot Annual Flow Trends Against Area and Display Climate Index Data ##################
+#########################################################################################################
 
 #Function for scaling values between 0 and 1.
 zero_one = function(x) ((x-min(x)))/(diff(range(x)))
 
-#Function for simplifying data and adding climate index value.
+#Function for simplifying data and adding climate index values.
 arrange.data = function(df, response){
 	tp = df[[grep(response, names(df))]]$real_slopes
 	tp = tp %>% select(8:ncol(tp)) %>% apply(.,2,function(x) zero_one(x)) %>% 

### 2017-03-01 Code_07 Monthly_Plot.R

@@ -1,177 +1,269 @@
-load("max_month.RData")
-load("med_month.RData")
-load("min_month.RData")
+###########################################################################################################
+######################################## Sub-Functions ####################################################
+###########################################################################################################
 
-#Global Control Panel	
-Line.Width = 1
-Axis.Label.Size = 1.1
-par(family = "serif", bg = "white", fg = "black")
-Lab.Loc = 3.1
-vline = 4
-shade1="99"; shade2="10"
-axisLabCol = "black"
 
-funnel.plots = function(dat, response, col1="#F2AD00", col2="#67A9CF", col3="#F98400", col4="#FF0000", col5="#525252", pcol="#737373", lcol="#969696",funnel.y = "", axis.L, yaxis.break = F, iter = 200, xaxis = F, d.xaxis = F, pane, xcoord, ycoord, y.axis = F){
-	library(tidyverse);library(wesanderson)
-	#browser()
-	
-	#Identify predictor variable data for the given response variable.
-	slot = grep(pattern = response, x = names(dat)); dat = dat[[slot]]
-	#Change predictor variable column name to generic value.
-	column = grep(pattern = response, x = names(dat$real_slopes))
-	names(dat$real_slopes)[column] = "resp"
-	
-	#Real Data: Coefficients and Response Variable isolation and transformation.
-	Real.Dat = dat$real_slopes %>% mutate(sqrt.resp = sqrt(resp)) %>% arrange(resp) %>% 
-		mutate(se.upper = slope+se, se.lower = slope-se)
-	resp = sort(Real.Dat$resp)
-	
-	#Initial Plot Region Parameters
-	par(mar = c(0,5,0.2,0), bg = "white",fg = "white")
-	
-	#Create Plotting Region
-	ylimits = pretty(range(scale(c(Real.Dat$se.upper,Real.Dat$se.lower),center = dat$real_varexp$intercept,scale = F)))
-	label = round(exp(ylimits*10)*100-100,0)
[…]

### 2017-03-01 Code_01 Data_Organization.Rmd

@@
 M.Data = left_join(Data.Preds, df, by = c("Station.ID","Date")) %>% 
+	mutate(flow.data_mean = if_else(is.na(.$mean5day), .$Flow.Data, .$mean5day)) %>%
+	select(-mean5day) %>% 
 	group_by(Station.ID, Year, nMonth, Month) %>% 
-	summarise(median.M.flow = median(Flow.Data),
- 						max.M.flow = max(mean5day, na.rm = T),
- 						min.M.flow = min(mean5day, na.rm = T))
+	summarise(median.M.flow = median(flow.data_mean),
+ 						max.M.flow = max(flow.data_mean, na.rm = T),
+ 						min.M.flow = min(flow.data_mean, na.rm = T))
 #Join location and area data to flow data.

### 2017-03-01 Code_05 Map_Plot.Rmd

---
title: "Figures"
author: "Kyle Chezik"
date: "July 10, 2015"
output: html_document
---

Figure 1: A map of the study area, sites and 
```{r Load Data, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, dependson=c(1,2,5)}
#Call necessary packages.
library(rgdal);library(dplyr)

#Load Data
load("01a_Data-Wtshd-Simple.RData")
load("03_Data_Annual.RData")
WS.MetaD = semi_join(WS.MetaD,Y.Data,by = "Station.ID")
BC.lim <- readOGR(dsn="Albers", layer="BC_Limit_Albers") #Call the BC boundary polygon.
Fraser.WS = readOGR(dsn="Albers", layer="Fraser_WS_Albers") # Call the Fraser basin polygon.

#Read in the polyline files for the different tributaries.
Tribs = readOGR(dsn="Albers", layer="Fraser_Tribs_Albers")
Tribs.5 = subset(Tribs, STRMRDR >4)
Tribs.6 = subset(Tribs, STRMRDR >5)
Tribs.7 = subset(Tribs, STRMRDR >6)
Tribs.8 = subset(Tribs, STRMRDR >7)
Tribs.9 = subset(Tribs, STRMRDR >8)
```

```{r Map Plot, echo=FALSE, message=FALSE, warning=FALSE}
#Create values to describe the size of points that represents relative contributing area.
WS.MetaD = mutate(WS.MetaD, Area.size = ((((Area-min(Area))/(diff(range(Area))))+0.02)^0.4)*3.5)
Legend.size = ((with(WS.MetaD, (c(500,1500,10000,100000,200000)-min(Area))/(diff(range(Area))))+0.02)^0.4)*3.5 #Point sizes for area legend.
Area.legend.values = c("500", "1500", "10000", "100000", "200000") #Values to be displayed in the area legend.
Area.legend.y = c(1140000, 1090000, 1050000, 1018000, 990000) #x-coordinate locations for area-legend points.
Area.legend.x = rep(1450000, 5) #y-coordinate location for area-legend points.

pdf(file="Fig1_Map2.pdf", bg="white",height=5.5, width =6 ,pointsize=12) #Set pdf arguments.

# Control Panel
Te.cex = 0.9 #Text size.
Te.cex.sub = 0.9 #Legend Sub-text size.

par(oma=c(3.25,4.5,0.25,0.25)) #Size of the outer margin surrounding the two plots.

par(family = 'serif', fg = NA, las = 1, mar = c(0.25,0.2,0,0), mex = 1, mfrow = c(1,1))
plot(BC.lim, ylim = c(480000, 1225000),xlim=c(850000, 1595000), las=1, lwd=0.25, col = "gray") #Plot the BC limits.
par(fg = "black", cex.axis = Te.cex, mgp = c(4,0.65,0)) # (fg) Add outlines to boarders, (cex.axis) Change size of axis labels, (mgp) Change the axis margin line before the y axis is printed.

utmcoor = SpatialPoints(cbind(X=c(750000,1000000,1250000,1500000,1750000), Y=c(1200000,1000000,800000,600000,400000)), proj4string = CRS("+proj=aea +lat_1=50 +lat_2=58.5 +lat_0=45 +lon_0=-126 +x_0=1000000 +y_0=0 +datum=NAD83 +units=m +no_defs +ellps=GRS80 +towgs84=0,0,0"))
longlatcoor = spTransform(utmcoor, CRS("+proj=longlat"))
long = as.character(round(longlatcoor$X,0))
labelsX=parse(text=paste(long, "^o ", "*W", sep=""))
lat = as.character(round(longlatcoor$Y,0))
labelsY=parse(text=paste(lat, "^o ", "*N", sep=""))

#For BC Albers UTM axis' labels.
#axis(2, at = c(1200000,1000000,800000,600000,400000), labels = c("1200000","1000000","800000","600000","400000") ,lwd = 0.5, line = 0.2, outer = TRUE) # y-axis.
#axis(1,at=c(750000,1000000,1250000,1500000,1750000), labels = c("750000","1000000","1250000","1500000","1750000") ,lwd = 0.5, line = 0, outer = TRUE) #x-axis

axis(2, at = c(1200000,1000000,800000,600000,400000), labels = labelsY ,lwd = 0.5, line = 0.2, outer = TRUE) # y-axis.
axis(1,at=c(750000,1000000,1250000,1500000,1750000), labels = labelsX ,lwd = 0.5, line = 0, outer = TRUE) #x-axis

#For BC Albers axis' labels.
#mtext("Albers BC UTM Easting", side = 1, line = 2, cex = Te.cex+0.2, outer = TRUE) # x-axis label.
#mtext("Albers BC UTM Northing", side = 2, line = 4.2, cex = Te.cex+0.2, las = 0, outer = TRUE)  #y-axis label.

mtext("Longitute", side = 1, line = 2, cex = Te.cex+0.2, outer = TRUE) # x-axis label.
mtext("Latitude", side = 2, line = 3.2, cex = Te.cex+0.2, las = 0, outer = TRUE)  #y-axis label.

par(fg = NA) # Remove outline from boarders/points/text etc.
plot(Fraser.WS, col="grey25", add=TRUE) #Plot the Fraser watershed area.
plot(Tribs.5, axes=TRUE, add=TRUE, lwd=0.5, col = "#525252") #Plot tribs...
plot(Tribs.6, axes=TRUE, add=TRUE, lwd=0.5, col = "#737373")
plot(Tribs.7, axes=TRUE, add=TRUE, lwd=0.5, col = "#737373")
plot(Tribs.8, axes=TRUE, add=TRUE, lwd=0.5, col = "#969696")
plot(Tribs.9, axes=TRUE, add=TRUE, lwd=0.5, col = "#969696")

# plot(Tribs.5, axes=TRUE, add=TRUE, lwd=0.01, col = "#525252") #Plot tribs...
# plot(Tribs.6, axes=TRUE, add=TRUE, lwd=0.01, col = "#737373")
# plot(Tribs.7, axes=TRUE, add=TRUE, lwd=0.01, col = "#737373")
# plot(Tribs.8, axes=TRUE, add=TRUE, lwd=0.01, col = "#969696")
# plot(Tribs.9, axes=TRUE, add=TRUE, lwd=0.01, col = "#969696")

pt.col = paste("#F2AD00","99",sep = "")
points(WS.MetaD$UTM_E, WS.MetaD$UTM_N, bg = pt.col, pch = 21, cex = WS.MetaD$Area.size) #Add points...
Dble.pt = subset(WS.MetaD,Area<50000)
points(Dble.pt$UTM_E, Dble.pt$UTM_N, bg = pt.col, pch = 21, cex = Dble.pt$Area.size) #Add points again

par(fg = "black") #Add outlines to boarders.
points(Area.legend.x[c(1:5)], Area.legend.y[c(1:5)], pch = 21, cex = Legend.size[c(5:1)], bg = "black")
text(Area.legend.x[c(1:5)]+70000, Area.legend.y[c(1:5)], labels = Area.legend.values[c(5:1)], cex = Te.cex-0.1)
text(1520000, 1180000, labels = expression("      Flow Gauge \n Wtshd Area (km"^2*")"), cex = Te.cex-0.1)

dev.off()
```

### 2017-03-01 Code_03_Post_Analysis_Null_Test.R
Example TMB applied to Fraser data.
Simply used the existing framework and plugged in the Fraser flow trend
data along with covariates to estimate the variance exponents for the
Area and Climate Index variables.

@@ -154,48 +154,6 @@ out_med_month <- plyr::dlply(M.Data, "nMonth", function(x)
 save(out_med_month, file = "med_month.RData")
 
 
-
-
-
-set.seed(123)
-vars = c("Area","emt.sd","ext.sd","map.sd","mat.sd","pas.sd")
-real_slopes = fit_slopes(Y.Data, "DOY2.logit", "Area", vars)
-zero_one = function(x) ((x-min(x)))/(diff(range(x)))
-real_slopes = real_slopes %>% select(8:ncol(real_slopes)) %>% apply(.,2,function(x) zero_one(x)) %>% 
-	apply(.,1,function(x) sum(x)) %>% mutate(real_slopes, std.clim = .)
-real_slopes = real_slopes %>% mutate(area_std = sqrt(Area)/sd(sqrt(Area)), clim_std = std.clim/sd(std.clim))
-
- m <- gls(slope~sqrt(Area), weights = varExp(form = ~sqrt(area_std)+clim_std),
-				 control = glsControl(maxIter = 1000L, msMaxIter = 1000L), data = real_slopes)
-
-plot(slope~sqrt(Area), real_slopes)
-#Add Simulation Lines to Plotting Region.
-line.sims = function(df, col, LW, shade = "10", transp = T, intercept, slope, resp.sqrt, sigma, varexp){
-	apply(df,1,function(y){
-		if(transp == T) colur = paste(col,shade,sep="")
-		else colur = col
-		upper = intercept + slope*(resp.sqrt - mean(resp.sqrt)) + 1.96 * sqrt(sigma^2 * exp(2*(resp.sqrt/1e3)*y[1]))
-		lines(resp.sqrt, upper, col = colur, lwd = LW)
-		lower = intercept + slope*(resp.sqrt - mean(resp.sqrt)) - 1.96 * sqrt(sigma^2 * exp(2*(resp.sqrt/1e3)*y[1]))
-		lines(resp.sqrt, lower, col = colur, lwd = LW)
-	})}
-
-varexp <- m$model[[1]][[1]][1]
-sigma <- m$sigma
-intercept <- coef(m)[[1]]
-slope = m$coefficients[[2]]
-slopePval = summary(m)[[18]][8]
-df1 = data.frame(varexp, sigma, intercept, slope, slopePval)
[…]

### 2017-03-01 tmb-example.R

@@ -35,3 +35,47 @@ opt$par
 # If you need standard errors and therefore confidence intervals... exp(sigma1 + 1.96 * SE)... :
 rep <- sdreport(obj)
 rep
+
+
+
+# Replicate with Fraser Data.
+library(tidyverse)
+load("out_doy2.RData")
+
+#Function for scaling values between 0 and 1.
+zero_one = function(x) ((x-min(x)))/(diff(range(x)))
+#Function for standardizing values.
+std = function(x) (x/sd(x))
+
+u = out_doy2$Area$real_slopes$slope
+s1 = out_doy2$Area$real_slopes %>% mutate(STDsqrtArea = std(sqrt(Area))) %>% .$STDsqrtArea
+s2 = out_doy2$Area$real_slopes %>% select(8:ncol(.)) %>% apply(.,2,function(x) zero_one(x)) %>% 
+	apply(.,1,function(x) sum(x)) %>% std(.)
+mm = model.matrix(~ s1 + s2)
+
+library(TMB)
+compile("varianceFunction.cpp")
+dyn.load(dynlib("varianceFunction"))
+
+obj <- MakeADFun(data = list(x_ij = mm, y_i = u),
+								 parameters = list(b_j = c(0, 0, 0), sigma_j = c(0, 0, 0)),
+								 DLL = "varianceFunction")
+
+opt <- nlminb(start=obj$env$last.par.best, objective=obj$fn, gradient=obj$gr)
+opt$par
+
+#			b_j           b_j           b_j       sigma_j       sigma_j       sigma_j 
+#-2.197504e-03  2.525132e-04  1.253969e-04 -1.308270e+01 -1.327247e+00  5.401138e-01
+
+rep <- sdreport(obj)
+rep
+
+# sdreport(.) result
+# 						Estimate   Std. Error
+# b_j     -2.197504e-03 0.0004539609
+# b_j      2.525132e-04 0.0001232716
+# b_j      1.253969e-04 0.0002550198
+# sigma_j -1.308270e+01 0.4312540804
+# sigma_j -1.327247e+00 0.1735828737
+# sigma_j  5.401138e-01 0.2108288820
+# Maximum gradient component: 0.02134764

### 2017-03-01 tmb-example.R
Added nlme varComb models with two varExp components.

@@ -79,3 +79,7 @@ rep
 # sigma_j -1.327247e+00 0.1735828737
 # sigma_j  5.401138e-01 0.2108288820
 # Maximum gradient component: 0.02134764
+
+library(nlme)
+mod = gls(slope~s1 + s2, data = out_doy2$Area$real_slopes, weights = varComb(varExp(form = ~s1), varExp(form = ~s2)))
+mod = gls(slope~s1 + s2, data = out_doy2$Area$real_slopes, weights = varExp(form = ~s2))

### 2017-03-01 tmb-example.R
Pushing DOY2 Data

@@ -64,14 +64,14 @@ obj <- MakeADFun(data = list(x_ij = mm, y_i = u),
 opt <- nlminb(start=obj$env$last.par.best, objective=obj$fn, gradient=obj$gr)
 opt$par
 
-#			b_j           b_j           b_j       sigma_j       sigma_j       sigma_j 
+#b_j           b_j           b_j       sigma_j       sigma_j       sigma_j 
 #-2.197504e-03  2.525132e-04  1.253969e-04 -1.308270e+01 -1.327247e+00  5.401138e-01
 
 rep <- sdreport(obj)
 rep
 
 # sdreport(.) result
-# 						Estimate   Std. Error
+# Estimate   Std. Error
 # b_j     -2.197504e-03 0.0004539609
 # b_j      2.525132e-04 0.0001232716
 # b_j      1.253969e-04 0.0002550198

### 2017-03-01 tmb-example.R
Plot all the climate data individually.

@@ -41,16 +41,29 @@ rep
 # Replicate with Fraser Data.
 library(tidyverse)
 load("out_doy2.RData")
+load("out_med.RData")
+load("out_max.RData")
+load("out_min.RData")
+
+x = out_med
+
+y = x$Area$real_slopes%>% select(8:ncol(.)) %>% apply(.,2,function(x) zero_one(x)) %>% 
+	apply(.,1,function(x) sum(x)) %>% mutate(x$Area$real_slopes, clim.sd = .) %>%  gather(., climate, index, emt.sd,ext.sd,map.sd,mat.sd,pas.sd,clim.sd)
+
+ggplot(y, aes(index, slope)) +
+	geom_point() +
+	facet_wrap(~climate, scales = "free_x")
 
 #Function for scaling values between 0 and 1.
 zero_one = function(x) ((x-min(x)))/(diff(range(x)))
 #Function for standardizing values.
 std = function(x) (x/sd(x))
 
-u = out_doy2$Area$real_slopes$slope
-s1 = out_doy2$Area$real_slopes %>% mutate(STDsqrtArea = std(sqrt(Area))) %>% .$STDsqrtArea
-s2 = out_doy2$Area$real_slopes %>% select(8:ncol(.)) %>% apply(.,2,function(x) zero_one(x)) %>% 
+u = x$Area$real_slopes$slope
+s1 = x$Area$real_slopes %>% mutate(STDsqrtArea = std(sqrt(Area))) %>% .$STDsqrtArea
+s2 = x$Area$real_slopes %>% select(8:ncol(.)) %>% apply(.,2,function(x) zero_one(x)) %>% 
 	apply(.,1,function(x) sum(x)) %>% std(.)
+s2 = x$Area$real_slopes$mat.sd %>% std(.)
 mm = model.matrix(~ s1 + s2)
 
 library(TMB)
@@ -81,5 +94,5 @@ rep
 # Maximum gradient component: 0.02134764
 
 library(nlme)
-mod = gls(slope~s1 + s2, data = out_doy2$Area$real_slopes, weights = varComb(varExp(form = ~s1), varExp(form = ~s2)))
-mod = gls(slope~s1 + s2, data = out_doy2$Area$real_slopes, weights = varExp(form = ~s2))
+mod = gls(slope~s1 + s2, data = x$Area$real_slopes, weights = varComb(varExp(form = ~s1), varExp(form = ~s2)))
[…]

### 2017-03-07 Code_05 Map_Plot.Rmd
Updated Post Analysis Null Test
Added the potential to calculate exponent values for both climate and
area.

@@
 #Call necessary packages.
-library(rgdal);library(dplyr)
+library(rgdal);library(tidyverse);library(sp);library(maptools)
 
@@
 ```
+
+Here I'm simply creating a sites shapefile with the adjusted size values for the point sizes. These will then be used to inform point sizes in QGIS.
+```{r Map Plot, echo=FALSE, message=FALSE, warning=FALSE}
+x = c(seq(1,3.5,by = 0.5),max(WS.MetaD$Area.size))
+round(((((x/3.5)^(1/0.4))-0.02) * diff(range(WS.MetaD$Area)))+min(WS.MetaD$Area),0)
+
+WS.MetaD = mutate(WS.MetaD, Area.size = ((((Area-min(Area))/(diff(range(Area))))+0.02)^0.4)*3.5)
+BCAlbers = CRS("+init=epsg:3005")
+sp = SpatialPointsDataFrame(data = WS.MetaD, coords = WS.MetaD[,c("UTM_E","UTM_N")], proj4string = BCAlbers)
+writeOGR(sp, dsn = "mapPlot", layer = "sites", driver="ESRI Shapefile", overwrite_layer = T, morphToESRI = T)
+
+```

### 2017-03-07 Code_03_Post_Analysis_Null_Test.R

@@ -1,4 +1,4 @@
-library(tidyverse); library(stringr)
+library(tidyverse); library(stringr);library(attenPlot)
 #Load/Organize Annual Data
 load("03_Data_Annual.RData")
 logit = function(p){log(p/(1-p))}
@@ -20,6 +20,7 @@ for(i in AnnualVars){
 	Y.Data = dplyr::left_join(Y.Data, cDat, by = "Station.ID")
 }
 Y.Data = Y.Data %>% ungroup(); rm(cDat, AnnualVars, i, logit, vars)
+
 #Monthly
 vars = unique(str_extract(string = MonthlyVars, pattern = "([^0-9]){3,}"))
 M.Data = lapply(vars,function(x){
@@ -32,6 +33,10 @@ M.Data = lapply(vars,function(x){
 		M.Data <<- left_join(M.Data, clim, by = c("Station.ID","nMonth"))
 })[[length(vars)]] %>% ungroup()
 rm(vars, MonthlyVars)
+
+remove = c("08MH006","08MH076","08MH090","08MH056","08LG016","08LG048","08MH103","08MH001","08MH016","08MH029")
+M.Data = filter(M.Data, Station.ID%in%remove == F)
+Y.Data = filter(Y.Data, Station.ID%in%remove == F)
 
 setwd("~/sfuvault/Simon_Fraser_University/PhD_Research/Projects/River-Network-Flow-Trends")
 # this function fits slopes to real data
@@ -74,12 +79,25 @@ sim_slopes <- function(slope.dat, yrs, return_ts = FALSE){
 
 # this function fits a gls model with a variance structure for the residual error
 # slopes and area are both vectors
-fit_var <- function(slopes, ind) {
+fit_var <- function(slopes, ind, clim, var2 = F) {
 	scale_factor <- var(slopes)
 	scaled_slope <- slopes / scale_factor
 	x <- sqrt(ind)
 	x <- x - mean(x)
+	v = sqrt(ind)/sd(sqrt(ind))
+	w = clim/sd(clim)
 	tryCatch({
+		if(var2 == T){
+			m <- gls(scaled_slope~x, weights = varComb(varExp(form= ~v), varExp(form= ~w)), 
+							 control = glsControl(maxIter = 1000L, msMaxIter = 1000L))
+			varexpA <- m$model[[1]][[1]][[1]]
+			varexpC <- m$model[[1]][[2]][[1]]
+			sigma <- m$sigma * scale_factor
[…]

### 2017-03-07 Code_08 Area_ClimatePortfolioPlot.R

@@ -1,0 +1,99 @@
+load("out_doy2.RData")
+library(tidyverse); library(attenPlot)
+df = out_doy2$Area$real_slopes
+df = df %>% select(8:ncol(.)) %>% apply(.,2,function(x) zero_one(x)) %>% 
+	apply(.,1,function(x) sum(x)) %>% mutate(df, clim.sd = ., logArea = log(Area), sqrtArea = sqrt(Area), cols = "#000000", loc = "interior")
+
+#Model to add line to plot.
+remove = c("08MH006","08MH076","08MH090","08MH056","08LG016","08LG048","08MH103","08MH001","08MH016","08MH029")
+df$cols[which(df$Station.ID%in%remove)] = "#969696"
+df$loc[which(df$Station.ID%in%remove)] = "coastal"
+ggplot(df,aes(clim.sd, slope, color = cols)) + geom_point()
+
+mod1 = lm(clim.sd~logArea, data = df, subset = which(Station.ID%in%remove==F))
+mod2 = lm(clim.sd~logArea, data = df, subset = which(Station.ID%in%remove==T))
+plot(clim.sd~logArea, data = df)
+abline(mod1)
+abline(mod2, lty = 2)
+
+sim = data.frame(logArea = log(seq(100,250000,by = 100)))
+preds1 = predict.lm(object = mod1, newdata = sim, se.fit = T)
+preds2 = predict.lm(object = mod2, newdata = sim, se.fit = T)
+preds = data.frame(area = exp(sim$logArea), preds1 = preds1$fit, se1 = preds1$se, preds2 = preds2$fit, se2 = preds2$se)
+
+pdf("Fig2_ClimatePortfolio.pdf", width = 7.5, height = 5.25)
+
+LW = 1; ALS = 1.1; LL = 3.1
+par(oma = c(4,0,0,0), mar = c(0,5,0.2,0), family = "serif", bg = "white",fg = "white", mgp = c(2,0.7,0))
+plot(clim.sd~sqrtArea, data = df, xlab = "", ylab = "", axes = F, xlim = c(0,sqrt(250000)), ylim = c(0,5))
+par(fg = "black")
+axis(1, lwd = LW, cex.axis = ALS, outer = T, hadj = 1,
+		 labels = c("0", "10000", "40000", "90000", "160000", "250000"), at = c(0, 100, 200, 300, 400, 500))
+mtext(expression("Area (km"^2 * ")"), side = 1, line = LL, 
+			cex = ALS)
[…]

### 2017-03-07 tmb-example.R

@@ -39,31 +39,42 @@ rep
 
 
 # Replicate with Fraser Data.
-library(tidyverse)
+library(tidyverse);library(attenPlot)
+load("03_Data_Annual.RData")
 load("out_doy2.RData")
 load("out_med.RData")
 load("out_max.RData")
 load("out_min.RData")
 
-x = out_med
+remove = c("08MH006","08MH076","08MH090","08MH056","08LG016","08LG048","08MH103","08MH001","08MH016","08MH029")
+x = out_doy2
 
-y = x$Area$real_slopes%>% select(8:ncol(.)) %>% apply(.,2,function(x) zero_one(x)) %>% 
-	apply(.,1,function(x) sum(x)) %>% mutate(x$Area$real_slopes, clim.sd = .) %>%  gather(., climate, index, emt.sd,ext.sd,map.sd,mat.sd,pas.sd,clim.sd)
+y = x$Area$real_slopes %>% select(8:ncol(.)) %>% apply(.,2,function(x) zero_one(x)) %>% 
+	apply(.,1,function(x) sum(x)) %>% mutate(x$Area$real_slopes, clim.sd = .)
+y = Y.Data %>% select(Station.ID, Lat, Long) %>% distinct() %>% left_join(y,., by = "Station.ID")
+
+ggplot(y, aes(log(y$Area), y$clim.sd, color = Lat, label=Station.ID)) + 
+	geom_point() +
+	scale_colour_gradient(low="blue", high = "red") + 
+	geom_text(hjust = 0, nudge_x = 0.05)
+
+y = x$Area$real_slopes
+y = y %>% select(8:ncol(.)) %>% apply(.,2,function(x) zero_one(x)) %>% 
+	apply(.,1,function(x) sum(x)) %>% mutate(y, clim.sd = .) %>%
+	gather(., climate, index, emt.sd,ext.sd,map.sd,mat.sd,pas.sd,clim.sd)
 
 ggplot(y, aes(index, slope)) +
 	geom_point() +
 	facet_wrap(~climate, scales = "free_x")
 
-#Function for scaling values between 0 and 1.
-zero_one = function(x) ((x-min(x)))/(diff(range(x)))
 #Function for standardizing values.
 std = function(x) (x/sd(x))
 
-u = x$Area$real_slopes$slope
[…]

### 2017-03-07 attenPlot/R/arrange_data.R
Attenuation Plot Library
Add functions for building attenuation and density plots.

@@ -1,0 +1,17 @@
+#' Simplify data and add climate index values.
+#'
+#' This function allows you to summarize climate data into a single index values across variables. It requires the zero_one() function.
+#' @param df Requires a list with nested response variables with a further nested dataframe of model output. This dataframe must contain 7 initial columns of model results with all subsequent columns containing climate data.
+#' @param response Requires a character object referencing one of the nested response variables in df.
+#' @keywords climate index
+#' @export
+#' @examples
+#' arrange.data(df = out_doy2, response = "Area")
+#
+arrange.data = function(df, response){
+	tp = df[[grep(response, names(df))]]$real_slopes
+	tp = tp %>% select(8:ncol(tp)) %>% apply(.,2,function(x) zero_one(x)) %>% 
+		apply(.,1,function(x) sum(x)) %>% mutate(tp, std.clim = .)
+	names(tp)[grep(response,names(tp))] = "resp"
+	tp %>% mutate(sqrt.resp = sqrt(resp)) %>% mutate(se.upper = slope+se, se.lower = slope-se) %>% arrange(resp)
+}

### 2017-03-07 attenPlot/R/axis.density.R

@@ -1,0 +1,28 @@
+#' Density Plot Axis' and Labels.
+#'
+#' This function allows you to draw the density plot axis' and labels.
+#' @param type Requires a character vector of "int" if plotting intercept data. Anything less (e.g., "exp") will default to plotting variance exponent data.
+#' @param ALS Requires a numeric that defines the axis line size.
+#' @param LL Requires a numeric that defines the margin line where the x-axis label will be drawn.
+#' @param LW Requires a numeric that defines the line width.
+#' @param annual Requires a T/F as to whether the data are annually summarized (T) or monthly summarized (F). Defaults to False.
+#' @keywords axis, labels
+#' @export
+#' @examples
+#' axis.density(type = "exp", ALS = 1, LL = 1)
+#
+axis.density.draw = function(type, ALS, LL, LW, annual = F){
+	if(type == "int"){
+		label = round(exp(c(-0.008,-0.004,0,0.004,0.008,0.012,0.016)*10)*100-100,0)
+		axis(1, lwd = 0.5, cex.axis = ALS, outer = T, hadj = 1, labels = label, at = c(-0.008,-0.004,0,0.004,0.008,0.012,0.016))
+		mtext(expression("Intercept | %Change"%.%"Decade"^-1), side = 1, line = LL, cex = ALS)
+	} else{
+		if(annual == T){
+			mtext(expression("Var. Exp. Param. ("*delta*")"), side = 1, line = LL, cex = ALS - 0.2) # x-axis label.	
+			axis(1, lwd = LW, tck = -0.04, cex.axis = ALS, outer = T)
+		} else {
+			axis(1, lwd = 0.5, cex.axis = ALS, outer = T, hadj = 0.5, labels = c("-12","-8","-4","0","4"), at = c(-12,-8,-4,0,4), cex.axis = ALS)
+			mtext(expression("Var. Exp. Param. ("*delta*")"), side = 1, line = LL, cex = ALS)	
+		}
+	}
+}

### 2017-03-07 attenPlot/R/axis_Draw.R

@@ -1,0 +1,23 @@
+#' Draw x and y axis.
+#'
+#' This function allows you to draw the x and y axis'.
+#' @param LW Requires a numeric that defines the line width.
+#' @param ALS Requires a numeric that defines the axis line size.
+#' @param ALY Requires a character object that defines the y-axis label.
+#' @param LL Requires a numeric that defines the margin line where the x-axis label will be drawn.
+#' @param YLim Requires a numeric vector for the location of tick mark labels. These values are output from axis.Fun().
+#' @param YLab Require a character vector fo the labels of tick mark labels. These values are output from axis.Fun().
+#' @param xAxis Requires a boolean T/F and determines whether the x-axis is drawn. xAxis defaults to False.
+#' @keywords draw axis'
+#' @export
+#' @examples
+#' axis.Draw(LW = 1, ALS = 1, ALY = "Y-Axis", LL = 1, YLim = ylimits, YLab = labels, xAxis = T)
+axis.Draw = function(LW, ALS, ALY, LL, YLim, YLab, xAxis = F){
+	if(xAxis == T){
+		axis(1, lwd = LW, cex.axis = ALS, outer = T, hadj = 1,
+				 labels = c("0","10000","40000","90000","160000","250000"), at = c(0,100,200,300,400,500))
+		mtext(expression("Area (km"^2*")"), side = 1, line = LL, cex = ALS) # x-axis label.
+	} else axis(1, lwd = LW, tick = TRUE, labels = FALSE, tck = 0)
+	axis(2, lwd = LW, cex.axis = ALS, at = YLim, labels = YLab, las = 1)
+	mtext(ALY, side=2, line=LL, cex = ALS, las = 0)  #y-axis label.
+}

### 2017-03-07 attenPlot/R/axis_Fun.R

@@ -1,0 +1,23 @@
+#' Y-Axis Labels
+#'
+#' Manually or automatically define readable y-axis labels for logit transformed day of year (DOY) trend variables or log transformed trend variables.
+#' @param df Requires a list with a nested dataframe of model results.
+#' @param yaxis.DOY Requires a boolean T/F. This argument refers to whether or not the y-axis is logit transformed DOY or not. It defaults to False.
+#' @param ylimit Allows the user to override estimates and force the bounds of the y-axis.
+#' @keywords y-axis
+#' @export
+#' @examples
+#' axis.Fun(df, yaxis.DOY = F, ylimit = c(-0.04,0.04))
+axis.Fun = function(df, yaxis.DOY=F, ylimit=NULL){
+	if(yaxis.DOY == T) {
+		ylimits <<- pretty(c(min(df$slope), max(df$slope)))
+		label <<- round((ylimits/4*38*365)/3.8,0)
+	} else {
+		ylimits <<- pretty(c(min(df$se.lower), max(df$se.upper)))
+		label <<- round(exp(ylimits*10)*100-100,0)
+	}
+	if(is.null(ylimit)==F){
+		ylimits <<- pretty(ylimit)
+		label <<- round(exp(ylimits*10)*100-100,0)
+	}
+}

### 2017-03-07 attenPlot/R/color_gradient.R

@@ -1,0 +1,22 @@
+#' Color ramp function.
+#'
+#' This function allows you to create a scaled color ramp for a vector of data.
+#' @param df Requires a vector of data.
+#' @param pal Requires a character object designating the color pallet to apply. Here we require palletes provided by RColorBrewer and default ot a Green Blue gradient pallet.
+#' @param shade Requires a character object between 0 and 100. The lower the value the greater the transparency. Default is "99".
+#' @keywords color
+#' @export
+#' @examples
+#' color.gradient(vec, pal = "GnBu", shade = "99")
+color.gradient = function(vec, pal = "GnBu", shade = "99"){
+	library(RColorBrewer)
+	#vector scaled 0-1 for function
+	col = zero_one(vec)
+	#color ramp function
+	gradient = brewer.pal(9, pal)[3:9]
+	FUN = colorRamp(gradient, bias=1)
+	#apply function
+	cols = FUN(col)
+	cols = rgb(cols, maxColorValue=256)
+	if(as.numeric(shade)<100) paste(cols, shade, sep = "")
+}

### 2017-03-07 attenPlot/R/gather_density_data.R

@@ -1,0 +1,25 @@
+#' Gather and Arrange Density Plot Data
+#'
+#' Gather and arrange observed and simulated variance and intercept parameter estimates.
+#' @param df Requires a list with nested response variables with a further nested dataframe of model output.
+#' @param response Requires a character object referencing one of the nested response variables in df.
+#' @param type Requires a character vector of "int" if plotting intercept data. Anything less (e.g., "exp") will default to plotting variance exponent data.
+#' @keywords density
+#' @export
+#' @examples
+#' gather.density.data(df = out_doy2, response = "Area", type = "int")
+gather.density.data = function(df, response, type){
+	tp = df[[grep(response, names(df))]]
+	Rvarexp = tp$real_varexp$varexp 
+	Svarexp = tp$sim_varexp$varexp
+	Rint = tp$real_varexp$intercept
+	Sint = tp$sim_varexp$intercept
+	
+	if(type == "int"){
+		d <<- density(Sint)
+		var.real <<- Rint}
+	else{d <<- density(Svarexp)
+	var.real <<- Rvarexp}
+	varR <<- Rvarexp
+	varS <<- Svarexp
+}

### 2017-03-07 attenPlot/R/line_sims.R

@@ -1,0 +1,25 @@
+#' Draw Simulation Lines
+#'
+#' This function allows you to add simulation lines to plotting region.
+#' @param df A dataframe containing a column of variance exponent parameters.
+#' @param col A character object defining the colour of the lines.
+#' @param LW A numeric object defining the width of each variance line.
+#' @param shade Requires a character object between 0 and 100. The lower the value the greater the transparency. Default is "10".
+#' @param transp A on/off (T/F) switch for making the lines transparent.
+#' @param intercept Requires a numeric object with the observed intercept.
+#' @param slope Requires a numeric object with the observed slope.
+#' @param resp.sqrt Requires a numeric object with the observed response variables. This variable (resp.sqrt) was added by arrange.data().
+#' @param sigma Requires a numeric object with the observed sigma.
+#' @keywords simlines
+#' @export
+#' @examples
+#' line.sims(df, col = "#000000", LW = 1, shade = "10", transp = T, intercep = 1, slope = 1, resp.sqrt = df$resp.sqrt, sigma = 2)
+line.sims = function(df, col, LW, shade = "10", transp = T, intercept, slope, resp.sqrt, sigma){
+	apply(df,1,function(y){
+		if(transp == T) colur = paste(col,shade,sep="")
+		else colur = col
+		upper = intercept + slope*(resp.sqrt - mean(resp.sqrt)) + 1.96 * sqrt(sigma^2 * exp(2*(resp.sqrt/1e3)*y[1]))
+		lines(resp.sqrt, upper, col = colur, lwd = LW)
+		lower = intercept + slope*(resp.sqrt - mean(resp.sqrt)) - 1.96 * sqrt(sigma^2 * exp(2*(resp.sqrt/1e3)*y[1]))
+		lines(resp.sqrt, lower, col = colur, lwd = LW)
+	})}

### 2017-03-07 attenPlot/R/plygnPrptns.R

@@ -1,0 +1,36 @@
+#' Plot Density Plots
+#'
+#' Plot proportional plots of the variance parameter depicting the simulated values greather than and less than the observed.
+#' @param df Requires a dataframe with model simulation results. Must have a column of "var".
+#' @param type Requires a character vector of "int" if plotting intercept data. Anything less (e.g., "exp") will default to plotting variance exponent data.
+#' @param varR Observed numeric value of the variance exponent. Output generated from gather_density_data().
+#' @param varS Simulated numeric values of the variance exponent. Output generated from gather_density_data().
+#' @param annual Requires a T/F as to whether the data are annually summarized (T) or monthly summarized (F). Defaults to False.
+#' @param col1 Polygon color (character HEX) of simulated variance exponent parameters that are less extreme than the observed.
+#' @param col1 Line color (character HEX) of real variance exponent parameter.
+#' @param col3 Polygon color (character HEX) of simulated variance exponent parameters that are more extreme than the observed.
+#' @param col4 Outline color  (character HEX) of simulated variance exponent parameters that are more extreme than the observed.
+#' @param shade Requires a character object between 0 and 100. The lower the value the greater the transparency. Default is "99".
+#' @keywords Plot, Density
+#' @export
+#' @examples
+#' plygnPrptns(df, type = "exp", varR = varR, varS = varS, col1 = "#F2AD00", col3 = "#F98400", col4 = "#FF0000")
+plygnPrptns = function(df, type, varR, varS, annual = F, col1="#F2AD00", col2 = "#67A9CF", col3="#F98400", col4="#FF0000", shade = "99"){
+	if(type != "int"){
+		Upper = filter(df, var>varR) %>% bind_rows(., data.frame(var = varR, dens = 0))
+		Lower = filter(df, var<varR) %>% bind_rows(., data.frame(var = varR, dens = 0))
+		if(is.null(Upper)==F & nrow(Upper)>1)	polygon(Upper$var,Upper$dens, 
+																									col=paste(col1,shade,sep=""),
+																									border=paste(col1,shade,sep=""))
+		if(is.null(Upper)==F & nrow(Lower)>1) polygon(Lower$var,Lower$dens, 
+																									col=paste(col3,shade,sep=""),
+																									border=paste(col4,shade,sep=""))
+		prop = round(sum(varR<varS)/length(varS),2)
[…]

### 2017-03-07 attenPlot/R/points_SE.R

@@ -1,0 +1,18 @@
+#' Add Points and Standard Errors
+#'
+#' This function allows you to add points and SE lines to plotting region.
+#' @param df Requires a dataframe with slope (slope), standard error (se) and sqrt reponse variable (sqrt.resp) columns. The sqrt.resp variable column should be produced by the arrange.data() function.
+#' @param lcol Requires a character object defining the color of the lines. The default is grey.
+#' @param cols Requires a character vector of either 1 or the same length as the number of plot points. These values should be output from the color.gradient() function.
+#' @param LW Requires a numeric value that defines the line width of the SE segments.
+#' @param shade Requires a character object between 0 and 100. The lower the value the greater the transparency. Default is "99".
+#' @keywords points, SE-lines
+#' @export
+#' @examples
+#' points.SE(df, lcol = "#969696", cols = cols, LW = 1, shade = "99")
+pointsSE = function(df, lcol ="#969696", cols, LW, shade = "99"){
+	segments(df$sqrt.resp, df$slope+df$se, df$sqrt.resp, df$slope-df$se,
+					 col = paste(lcol, shade, sep = ""), lwd = LW)
+	segments(0,0,500,0, lty = 2, lwd = LW)
+	points(df$sqrt.resp, df$slope, col = cols, pch = 16, cex = 1.5)
+}

### 2017-03-07 attenPlot/R/screenDivide.R

@@ -1,0 +1,19 @@
+#' Break Up Canvas
+#'
+#' This function allows you to break up the screen canvas into two columns of twelve for density plots and a column of 4 rows for representitive seasonal funnel plots.
+#' @param empty No arguments required.
+#' @keywords scale
+#' @export
+#' @examples
+#' zero_one()
+monthlyScreen = function(){layout(cbind(c(1:12),c(1:12),c(1:12),c(1:12),c(13:24),c(13:24),c(13:24),c(13:24),c(rep(25,3),rep(26,3),rep(27,3),rep(28,3)),c(rep(25,3),rep(26,3),rep(27,3),rep(28,3)),c(rep(25,3),rep(26,3),rep(27,3),rep(28,3)),c(rep(25,3),rep(26,3),rep(27,3),rep(28,3)),c(rep(25,3),rep(26,3),rep(27,3),rep(28,3))))}
+
+#' Break Up Canvas
+#'
+#' This function allows you to break up the screen canvas into 2 columns of 2 for funnel plots and a column of 4 rows for density plots.
+#' @param empty No arguments required.
+#' @keywords scale
+#' @export
+#' @examples
+#' zero_one()
+annualScreen = function(){layout(rbind(c(1,1,3,3,2),c(1,1,3,3,4),c(5,5,7,7,6),c(5,5,7,7,8)))}

### 2017-03-07 attenPlot/R/zero_one.R

@@ -1,0 +1,9 @@
+#' 0 to 1 scaling function.
+#'
+#' This function allows you to scale data between 0 and 1.
+#' @param x Requires a vector of data.
+#' @keywords scale
+#' @export
+#' @examples
+#' zero_one()
+zero_one = function(x) ((x-min(x)))/(diff(range(x)))

### 2017-03-07 Code_06 Annual_Funnel_Plot.R
Retool plot code to incorporate new attenPlot library.

@@ -2,116 +2,36 @@
 ################## Plot Annual Flow Trends Against Area and Display Climate Index Data ##################
 #########################################################################################################
 
-#Function for scaling values between 0 and 1.
-zero_one = function(x) ((x-min(x)))/(diff(range(x)))
+library(attenPlot) #Funnel and Density plot function library.
 
-#Function for simplifying data and adding climate index values.
-arrange.data = function(df, response){
-	tp = df[[grep(response, names(df))]]$real_slopes
-	tp = tp %>% select(8:ncol(tp)) %>% apply(.,2,function(x) zero_one(x)) %>% 
-		apply(.,1,function(x) sum(x)) %>% mutate(tp, std.clim = .)
-	names(tp)[grep(response,names(tp))] = "resp"
-	tp %>% mutate(sqrt.resp = sqrt(resp)) %>% mutate(se.upper = slope+se, se.lower = slope-se) %>% arrange(resp)
-}
-
-#Add colour gradient to climate index.
-color.gradient = function(df, pal = "GnBu", shade = "99"){
-	#vector scaled 0-1 for function
-	col = zero_one(df$std.clim)
-	#color ramp function
-	RdYlBu = brewer.pal(9, pal)[3:9]
-	FUN = colorRamp(RdYlBu, bias=1)
-	#apply function
-	cols = FUN(col)
-	cols = rgb(cols, maxColorValue=256)
-	paste(cols, shade, sep = "")
-}
-
-#Axis label determination
-axis.Fun = function(df, yaxis.DOY=F){
-	if(yaxis.DOY == T) {
-		ylimits <<- pretty(c(min(df$slope), max(df$slope)))
-		label <<- round((ylimits/4*38*365)/3.8,0)
-	} else {
-		ylimits <<- pretty(c(min(df$se.lower), max(df$se.upper)))
-		label <<- round(exp(ylimits*10)*100-100,0)
-	}
-}
-
-#Draw x and y axis.
-axis.Draw = function(LW, ALS, ALX, ALY, LL, YLim, YLab, xAxis = T){
-	if(xAxis == T){
-		axis(1, lwd = LW, cex.axis = ALS, outer = T, hadj = 1,
[…]

### 2017-03-07 Code_07 Monthly_Plot.R

@@ -1,135 +1,10 @@
-###########################################################################################################
-######################################## Sub-Functions ####################################################
-###########################################################################################################
+#########################################################################################################
+################## Plot Monthly Flow Trends Against Area and Display Climate Index Data #################
+#########################################################################################################
 
+library(attenPlot)#Funnel and Density plot function library.
 
-############################################ Funnel Plot Functions ###############################################
-
-#Function for scaling values between 0 and 1.
-zero_one = function(x) ((x-min(x)))/(diff(range(x)))
-
-#Function for simplifying data and adding climate index values.
-arrange.data = function(df, response){
-	tp = df[[grep(response, names(df))]]$real_slopes
-	tp = tp %>% select(8:ncol(tp)) %>% apply(.,2,function(x) zero_one(x)) %>% 
-		apply(.,1,function(x) sum(x)) %>% mutate(tp, std.clim = .)
-	names(tp)[grep(response,names(tp))] = "resp"
-	tp %>% mutate(sqrt.resp = sqrt(resp)) %>% mutate(se.upper = slope+se, se.lower = slope-se) %>% arrange(resp)
-}
-
-#Add colour gradient to climate index.
-color.gradient = function(df, pal = "GnBu", shade = "99"){
-	#vector scaled 0-1 for function
-	col = zero_one(df$std.clim)
-	#color ramp function
-	RdYlBu = brewer.pal(9, pal)[3:9]
-	FUN = colorRamp(RdYlBu, bias=1)
-	#apply function
-	cols = FUN(col)
-	cols = rgb(cols, maxColorValue=256)
-	paste(cols, shade, sep = "")
-}
-
-#Axis label determination
-axis.Fun = function(df, yaxis.DOY=F, ylimit=NULL){
-	if(yaxis.DOY == T) {
-		ylimits <<- pretty(c(min(df$slope), max(df$slope)))
-		label <<- round((ylimits/4*38*365)/3.8,0)
[…]

### 2017-03-11 drafts/AGU_Journal/ms_Geo_Let_Sup.tex
Fixed Climate Data Indices
Corrected zonal stats problem and re-incorporated climate data into
“out” datasets. Updated code to reflect working models for flow
attenuation.

\documentclass[draft]{agujournal}
\usepackage{gensymb}
\usepackage{textcomp}
\usepackage{upgreek}
\journalname{Geophysical Research Letters}
\begin{document}
\supportinginfo{River networks dampen long-term hydrological signals of climate change}	
\authors{K. A. Chezik\affil{1},
	S. C. Anderson\affil{2},
	J. W. Moore\affil{1}}

\affiliation{1}{Earth to Ocean Research Group, Department of Biological Sciences, Simon Fraser University, 8888 University Dr., Burnaby, British Columbia V5A 1S6, Canada}
\affiliation{2}{School of Aquatic and Fishery Sciences, University of Washington, Box 455020, Seattle, WA 98195, USA}

\correspondingauthor{K. A. Chezik}{kchezik@sfu.ca}
\section*{Contents}

\begin{enumerate}
\item Figures S1 to S4
\end{enumerate}

\section*{Introduction}
The following supplementary material includes 4 additional figures. Figures S1 and S2 are similar to figure 4 in the main text where monthly trends in flow are shown for the basin along with trend attenuation results. In these supplementary figures we highlight minimum and median flow trends rather than maximum flow trends. Figures S3 and S4 are example plots demonstrating a single null-model iteration for the median flow metric.

\section{Supplementary Figures}
\begin{figure}[h]
	\centerline{\includegraphics[height=5.2in]{FigS1_Min-Monthly-Density-Funnel.pdf}}
	\caption{Monthly minimum flow trend attenuation within the Fraser River basin. (\textbf{Left}) Fraser River's basin-wide minimum-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulation. Observed values falling further from the center of the density distribution suggest greater evidence for changes in minimum flow and a greater shift in magnitude. (\textbf{Center}) Observed monthly Fraser River minimum-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates. Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed. (\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}) for four seasonally representative months. These reflect months in the prior columns and describe the variation in percent change per decade of minimum flow among sites. Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
	\label{S1}
\end{figure}

\begin{figure}[h]
	\centerline{\includegraphics[height=5.2in]{FigS2_Median-Monthly-Density-Funnel.pdf}}
	\caption{Monthly median flow trend attenuation within the Fraser River basin. (\textbf{Left}) Fraser River's basin-wide median-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulation. Observed values falling further from the center of the density distribution suggest greater evidence for changes in median flow and a greater shift in magnitude. (\textbf{Center}) Observed monthly Fraser River median-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates. Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed. (\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}) for four seasonally representative months. These reflect months in the prior columns and describe the variation in percent change per decade of median flow among sites. Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
	\label{S2}
\end{figure}

\begin{sidewaysfigure}
	\includegraphics[width=7in]{FigS3_site_simulation.pdf}}
	\caption{Example site specific flow simulations. Annual median-flow simulations (yellow) and observations (blue) for all 55 sites considered in this study found within the Fraser River basin. Simulations at each site were parameterized using the observed data but no trend was imposed. Fifty-five site simulations equate to one basin-wide simulation. Each flow metric was simulated 1000 times basin-wide per response variable. See supplemental figure~\ref{S4} for resultant basin-wide simulation.}
	\label{S3}
\end{sidewaysfigure}

\begin{figure}
	\centerline{\includegraphics[height=4in]{FigS4_basin_simulation.pdf}}
	\caption{Example flow simulation results. Trend estimates $\pm$SE of simulated null-model data (yellow) and the observed data upon which simulations were based (blue), plotted against watershed area. Curved lines represent the estimated variance function ($\hat{\updelta}$) given the simulated data. This is a single example result of the 1000 null-model simulations produced from the data in supplemental figure 3.}
	\label{S4}
\end{figure}

\end{document}

### 2017-03-11 Code_02 ClimatePortfolio.R

@@ -78,9 +78,10 @@ for(i in dir()){
 		
 		#Collect analysis results from clusters and compile into a single object.
 		df4 = collect(df3)
+		browser()
 		df5 = df4 %>% spread(key = measure, value = slope)
-		df %>% select(id1, latitude, longitude, elevation) %>% 
-			left_join(., df5) %>% 
+		df6 = df %>% select(id1, latitude, longitude, elevation) %>% 
+			left_join(.,df5,by="id1") %>% distinct()%>% 
 			write_csv(paste("../trends_split/",i,sep = ""))
 	})
 	rm(df,df1,df2,df3,df4,df5)

### 2017-03-11 Code_03_Post_Analysis_Null_Test.R

@@ -15,28 +15,27 @@ AnnualVars = vars[which(nchar(vars) == 3)]
 MonthlyVars = vars[which(nchar(vars) > 3 & nchar(vars)<7)]
 #Annual
 for(i in AnnualVars){
-	cDat = read.table(paste(i,"-stats.txt",sep = ""), header = T, as.is = T) %>% select(file, std)
-	names(cDat) = c("Station.ID", paste(i,".sd", sep = ""))
+	labels = c("Station.ID","count","min","mean","max",paste(i,".sd",sep = ""),"P25","P75")
+	cDat = read_csv(paste(i,"-stats.csv", sep = ""), col_names =  labels) %>% select(1,6)
 	Y.Data = dplyr::left_join(Y.Data, cDat, by = "Station.ID")
 }
-Y.Data = Y.Data %>% ungroup(); rm(cDat, AnnualVars, i, logit, vars)
+Y.Data = Y.Data %>% ungroup(); rm(cDat, AnnualVars, i, logit, vars, labels)
 
 #Monthly
 vars = unique(str_extract(string = MonthlyVars, pattern = "([^0-9]){3,}"))
 M.Data = lapply(vars,function(x){
-	clim = MonthlyVars[grep(pattern = x, MonthlyVars)] %>% paste(.,"-stats.txt", sep = "") %>%
+	clim = MonthlyVars[grep(pattern = x, MonthlyVars)] %>% paste(.,"-stats.csv", sep = "") %>%
 		lapply(., function(y){
-			read_table(y) %>% select(file, std) %>% mutate(nMonth = as.numeric(str_extract(string = y, pattern = "([0-9]{2})")))
+			labels = c("Station.ID","count","min","mean","max",paste(x,".sd", sep = ""),"P25","P75")
+			read_csv(y, col_names = labels) %>% select(1,6) %>% mutate(nMonth = as.numeric(str_extract(string = y, pattern = "([0-9]{2})")))
 		}) %>% bind_rows()
-		
-		names(clim) = c("Station.ID",paste(x,".sd", sep = ""),"nMonth")
 		M.Data <<- left_join(M.Data, clim, by = c("Station.ID","nMonth"))
 })[[length(vars)]] %>% ungroup()
 rm(vars, MonthlyVars)
 
-remove = c("08MH006","08MH076","08MH090","08MH056","08LG016","08LG048","08MH103","08MH001","08MH016","08MH029")
-M.Data = filter(M.Data, Station.ID%in%remove == F)
[…]

### 2017-03-11 Code_06 Annual_Funnel_Plot.R

@@ -43,7 +43,7 @@ funnel.wrapper = function(dat, response, yaxis.DOY = F, xAxis = T, pane, axis.Ly
 	
 	########################################### Plot Raw Data ##############################################
 	#Set initial plotting parameters
-	par(mar = c(0,5,0.2,0), family = "serif", bg = "white",fg = "white", mgp = c(2,0.7,0))
+	par(mar = c(0,5,0.2,0), family = "serif", bg = "white",fg = "white", mgp = c(2,0.7,0), las = 1)
 	#Determine axis labels.
 	axis.Fun(flow.trends, yaxis.DOY)
 	#Plot initial blank region to be filled.
@@ -92,7 +92,7 @@ load("out_min.RData")
 load("out_max.RData")
 load("out_med.RData")
 
-#pdf("Fig3_Annual-Funnel.pdf", width = 11, height = 8.5)
+pdf("Fig3_Annual-Funnel.pdf", width = 11, height = 8.5)
 annualScreen()
 par(oma = c(4,0,0,0))
 funnel.wrapper(out_doy2, "Area", pane = "a", xAxis = F, yaxis.DOY = T,  
@@ -103,4 +103,4 @@ funnel.wrapper(out_max, "Area", pane = "c",
 							 axis.Ly = expression("Maximum-Flow | %Change"%.%"Decade"^-1))
 funnel.wrapper(out_med, "Area", pane = "d", d.axis = T,
 							 axis.Ly = expression("Median-Flow | %Change"%.%"Decade"^-1))
-#dev.off()
+dev.off()

### 2017-03-11 Code_06 AttenStats.R

@@ -1,4 +1,8 @@
 #Attenuation Stats.
+load("out_doy2.RData")
+load("out_med.RData")
+load("out_min.RData")
+load("out_max.RData")
 
 #This function finds the predicted attenuation values at any two catchment areas and returns the percent decrease in predicted range values when moving from small catchments to a large catchment.
 pred.F = function(x, Area){
@@ -7,11 +11,11 @@ pred.F = function(x, Area){
 	percent.reduced = (1-sqrt(diff(c(prediction.1[2],prediction.2[2]))^2/diff(c(prediction.1[1],prediction.2[1]))^2))*100
 	data.frame(percent.reduced)}
 
-pred.F(out_max,range(out_max$real_slopes$Area))
-pred.F(out_min,range(out_max$real_slopes$Area))
-pred.F(out_med,range(out_max$real_slopes$Area))
-pred.F(out_doy2,range(out_max$real_slopes$Area))
-pred.F(out_med,c(5000,60000))
+pred.F(out_max$Area,range(out_max$Area$real_slopes$Area))
+pred.F(out_min$Area,range(out_max$Area$real_slopes$Area))
+pred.F(out_med$Area,range(out_max$Area$real_slopes$Area))
+pred.F(out_doy2$Area,range(out_max$Area$real_slopes$Area))
+pred.F(out_med$Area,c(5000,60000))
 
 #This function produces lower and upper range estimates for trend values among smaller catchments (Small_Area) and then at a larger catchment (Large_Area). Values are returned as the percent change per decade.
 pred.F2 = function(x){
@@ -22,9 +26,9 @@ pred.F2 = function(x){
 	lower.est = round((exp(range(lower)*10)-1)*100,2)
 	data.frame(upper.est, lower.est, row.names = c("Small_Area", "Large_Area"))
 	}
-pred.F2(out_max)
-pred.F2(out_min)
-pred.F2(out_med)
+pred.F2(out_max$Area)
+pred.F2(out_min$Area)
+pred.F2(out_med$Area)
 
 ##############################################################
 #     Determine the ratio of observed to null attenuation    #
@@ -38,12 +42,12 @@ ratio_atten = function(varexp_real, varexp_null, area1, area2){
 	#obs / null
 }
 
-Area_Small = min(out_med$real_slopes$Area)
+Area_Small = min(out_med$Area$real_slopes$Area)
 #Area_Small = 5000
-Area_Large = max(out_med$real_slopes$Area)
[…]

### 2017-03-11 Code_08 Area_ClimatePortfolioPlot.R

@@ -1,31 +1,36 @@
-load("out_doy2.RData")
-library(tidyverse); library(attenPlot)
-df = out_doy2$Area$real_slopes
-df = df %>% select(8:ncol(.)) %>% apply(.,2,function(x) zero_one(x)) %>% 
-	apply(.,1,function(x) sum(x)) %>% mutate(df, clim.sd = ., logArea = log(Area), sqrtArea = sqrt(Area), cols = "#000000", loc = "interior")
+load("out_med.RData")
+library(MASS);library(tidyverse); library(attenPlot);
+df = out_med$Area$real_slopes
+df = df %>% mutate(logArea = log(Area), sqrtArea = sqrt(Area), cols = "#000000", loc = "interior")
 
 #Model to add line to plot.
-remove = c("08MH006","08MH076","08MH090","08MH056","08LG016","08LG048","08MH103","08MH001","08MH016","08MH029")
+remove = c("08MH006","08MH076")
+#"08MH090","08MH056","08LG016","08LG048","08MH103","08MH001","08MH016","08MH029"
 df$cols[which(df$Station.ID%in%remove)] = "#969696"
 df$loc[which(df$Station.ID%in%remove)] = "coastal"
-ggplot(df,aes(clim.sd, slope, color = cols)) + geom_point()
+ggplot(df,aes(std.clim, slope, color = cols)) + geom_point()
 
-mod1 = lm(clim.sd~logArea, data = df, subset = which(Station.ID%in%remove==F))
-mod2 = lm(clim.sd~logArea, data = df, subset = which(Station.ID%in%remove==T))
-plot(clim.sd~logArea, data = df)
-abline(mod1)
-abline(mod2, lty = 2)
+#Look at each climate variable and the index against area.
+gather(data = df, key = "climateIndex", value = "IndexValue", emt.sd, ext.sd, map.sd, mat.sd, pas.sd, std.clim) %>% 
+	ggplot(., aes(logArea, IndexValue, label = Station.ID)) + geom_text(check_overlap = T) + geom_point(aes(color = cols)) + facet_wrap(~climateIndex, scales = "free_y") + theme(legend.position="none")
 
-sim = data.frame(logArea = log(seq(100,250000,by = 100)))
-preds1 = predict.lm(object = mod1, newdata = sim, se.fit = T)
[…]

### 2017-03-15 drafts/PNAS/ms_pnas.tex
Monthly Color Gradient / Fist MS PNAS Draft
Made the monthly color gradients scale across all climate index values
rather than just within a month. Also adding PNAS draft. Also, adding a
bit of code that plays with the effect of N on SD when SD is expected
to grow with N. Finally, added a bit of code that shows how climate
index values become less strong and homogenous across the Fraser basin
during seasonal transitions.

--- 2017-01-11 AGU_Journal/ms_Geo_Let.tex
@@
-\documentclass[linenumbers,draft]{AGUJournal}
-\usepackage{gensymb}
-\usepackage{textcomp}
-\usepackage{upgreek}
-\usepackage{natbib}
-\usepackage[parfill]{parskip}
-\journalname{Geophysical Research Letters}
-\begin{document}
+\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+% Use the lineno option to display guide line numbers if required.
+% Note that the use of elements such as single-column equations
+% may affect the guide line number alignment.
+
+\templatetype{pnasresearcharticle} % Choose template 
+% {pnasresearcharticle} = Template for a two-column research article
+% {pnasmathematics} = Template for a one-column mathematics article
+% {pnasinvited} = Template for a PNAS invited submission
+
 \title{River networks dampen long-term hydrological signals of climate change}
-\authors{K. A. Chezik\affil{1},
-	S. C. Anderson\affil{2},
-J. W. Moore\affil{1}}
 
-\affiliation{1}{Earth to Ocean Research Group, Department of Biological Sciences, Simon Fraser University, 8888 University Dr., Burnaby, British Columbia V5A 1S6, Canada}
-\affiliation{2}{School of Aquatic and Fishery Sciences, University of Washington, Box 455020, Seattle, WA 98195, USA}
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a,1]{Kyle A. Chezik}
+\author[b,1]{Sean C. Anderson} 
+\author[a,1]{Jonathan W. Moore}
 
-\correspondingauthor{K. A. Chezik}{kchezik@sfu.ca}
+\affil[a]{Earth to Ocean Research Group, Department of Biological Sciences, Simon Fraser University, 8888 University Dr., Burnaby, British Columbia V5A 1S6, Canada}
+\affil[b]{School of Aquatic and Fishery Sciencies, University of Washington, Box 455020, Seattle, WA 98195, USA}
 
-\begin{keypoints}
-\item River networks act as diversified portfolios of climate; by integrating varied climate signals the network substantially dampens downstream climate-driven hydrologic trends.
-\item River network climate dampening varied seasonally in the Fraser River basin, declining sharply with the spring freshet, suggesting that basin-wide events homogenize the climate portfolio, thereby degrading the network dampening effect.
-\item Rivers that integrate a diverse climate portfolio offer climate change relief to riverine biota that is many times greater than homogenized systems.
-Leveraging network dampening properties by promoting land use practices that maintain or enhance climate portfolio diversity may be a powerful tool in mitigating climate change impacts.
-\end{keypoints}
+% Please give the surname of the lead author for the running footer
+\leadauthor{Chezik}
+
+% Please add here a significance statement to explain the relevance of your work
+\significancestatement{Many ecosystems posses stabilizing processes that provide a natural defense against the impacts of climate change.
+The branching tree structure of river networks, provide a framework for aggregating the varied signals of climate change within a watershed and a possible mechanism for dampening long-term climate driven trends.
+Varied landscape attributes within a watershed produce a mosaic or portfolio of climate which in turn results in a variety of climate change manifestations.
+Here we demonstrate that large free-flowing river networks dampen local long-term flow trends by integrating a large and varied climate portfolio across a diverse and complex landscape.
+Integrating the benefits of these ecosystems into climate change mitigation stategies is essential in reducing the increasingly extreme impacts of climate change.}
+
+% Please include corresponding author, author contribution and author declaration information
+\authorcontributions{J.W.
+Moore conceived the study and analysis, S.C.
+Anderson formalized the analysis, K.A.
+Chezik gathered and analyzed the data and wrote the paper.
+All authors contributed in all aspects to varying degrees.}
+\authordeclaration{The authors declare no conflict of interest.}
+\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: kchezik@sfu.ca}
+
+% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
+\keywords{rivers networks $|$ climate portfolio $|$ climate change $|$ trend attenuation $|$ climate dampening}
 
 \begin{abstract}
-River networks may dampen local hydrologic climate change by aggregating a diverse climate portfolio with variable upstream climate signals.
-Here we examine this hypothesis using flow trend estimates (1970--2007) at 55 hydrometric gauge stations within the Fraser River watershed in British Columbia, Canada.
-The Fraser River dampened variability in downstream long-term median discharge by 95\% from between -6 and 18\% per-decade among small watersheds to between 0.8 and 1.6\% per-decade near the confluence.
-A null model approach indicated that dampening of annual median flow trends within the Fraser River was typically 5.4 times greater than would be expected of a watershed exhibiting a consistent basin-wide climate response.
+	Climate change is dramatically altering temperate rivers --- snow-melt is shifting ever earlier \citep{Rauscher:2008]} floods are becoming more frequent \citep{Hirabayashi:2013}, and low flows more persistent \citep{McCabe:2014}.
+Here we describe a novel pathway by which river networks dampen local hydrologic climate change through the aggregation of upstream climate portfolio assets.
+We examine this hypothesis using flow and climate trend estimates (1970--2007), calculated at 55 hydrometric gauge stations and their contributing watersheds', within the vast Fraser River basin in British Columbia, Canada.
+Using a null hypothesis framework we compared our observed attenuation of river flow trends as a function of increasing area and climate trend diversity, with null-simulated estimates to gauge the likelihood and strength of our observations.
+We found the Fraser River dampened variability in downstream long-term discharge by >91\% across all flow trend response variables.
+For instance, median flow trends ranged from between -6 and 19\% per-decade among small watersheds, attenuating to between 0.8 and 1.6\% per-decade near the confluence.
+Our null-model approach indicated that dampening of annual median flow trends within the Fraser River was typically 3.3 times greater than would be expected of a watershed exhibiting a consistent basin-wide climate response.
 Although climate dampening strength varied seasonally, our findings indicate that large free-flowing rivers offer a powerful and largely unappreciated process of climate change mitigation.
@@
 
-\section{Introduction} 
-As billions are spent adapting to climate change by such means as building dikes or improving stormwater drainage \citep{Narain:2011}, certain habitats and processes provide a natural defense system that mitigates climate change impacts \citep{Jones:2012}.
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+
+\begin{document}
+
+% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
+% You should only change this length when you've finalised the article contents.
+\verticaladjustment{-2pt}
+
+\maketitle
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
+
+% If your first paragraph (i.e.
+with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
+If this is the case, add \parshape=0 to the end of the list environment.
+\dropcap{A}s billions are spent adapting to climate change by such means as building dikes or improving stormwater drainage \citep{Narain:2011}, certain habitats and processes provide a natural defense system that mitigates climate change impacts \citep{Jones:2012}.
 For example, coastal habitats such as oyster reefs, mangrove forests, and eelgrass beds shield 67\% of the United States coastline and 1.4 million people from sea level rise and storms \citep{Arkema:2013}.
@@
 The value of climate buffering habitats stands only to increase as average global temperatures climb.
-Recent international agreements in Paris aim to limit global temperature rise to 1.5\degree{C} \citep{Hulme:2016}, a goal that concedes a minimum doubling of observed warming \citep{Hartmann:2013}; therefore, an urgent challenge and opportunity is discovering systems that naturally buffer against oncoming impacts of climate change.
+The Paris Agreement of 2016 aims to limit global temperature rise to 1.5\degree{C} \citep{Hulme:2016}, a goal that concedes a minimum doubling of observed warming \citep{Hartmann:2013}; therefore, an urgent challenge and opportunity is discovering, maintaining and restoring systems that naturally buffer against the oncoming impacts of climate change.
 
-River systems may offer an unappreciated defense against some aspects of climate change.
-Climate change is altering global precipitation patterns \citep{Donat:2016} and subsequently river flow regimes \citep{Hartmann:2013,Palmer:2009}.
+In the face of changing global precipitation patterns \citep{Donat:2016}, river networks may offer an unappreciated defense against shifting flow regimes under climate change\citep{Hartmann:2013,Palmer:2009}.
 Earlier snowmelt \citep{Rauscher:2008}, reduced snow pack \citep{McCabe:2014}, shifts from snow to rain, and changes in the annual distribution of precipitation are all impacting stream flow \citep{Hartmann:2013}, resulting in flashier flows with increased frequency of flooding \citep{Hirabayashi:2013}, longer periods of drought and critically low flows for wildlife and agriculture \citep{Melillo:2014}.
-However, we hypothesize that large free-flowing river networks may naturally dampen these signals of climate change by integrating across different manifestations of climate.
+However, we hypothesize that large free-flowing river networks may naturally dampen these signals of climate change by integrating across varied landscapes and different manifestations of climate.
 Processes that aggregate across asynchronous components tend to dampen the aggregate, a process known as the portfolio effect \citep{Doak:1998}.
-Climate is expressed differently and asynchronously over the landscape \citep{George:2015} and on account of their branching architecture and directional flow, river networks directionally integrate these varied climate manifestations \citep{Peterson:2013}, potentially dampening the local climatic response of sub-catchments.
+Climate is expressed differently and asynchronously over the landscape \citep{George:2015} and on account of their branching architecture and directional flow, river networks integrate these varied climate manifestations \citep{Peterson:2013}, potentially dampening the local climatic response of sub-catchments.
 Large free-flowing rivers may have diverse climate portfolios, with sub-catchments acting as assets in their portfolio, thereby dampening local climate impacts.
@@
 
-Here we examine whether rivers attenuate local long-term climate change through analyses of hydrometric data from one of the largest free-flowing rivers in North America.
-Located in British Columbia, Canada, the Fraser River drains an area approximately the size of the United Kingdom (\texttildelow 217,000 km\textsuperscript{2}) and in a rare combination is fairly well monitored while also having no dams on its main stem \citep{Vorosmarty:2010} (Figure~\ref{figone}).
+Here we consider whether river networks attenuate local long-term climate change through analysis of hydrometric data from one of the largest free-flowing rivers in North America.
+Located in British Columbia, Canada, the Fraser River drains an area approximately the size of the United Kingdom (\texttildelow 217,000 km\textsuperscript{2}) and in a rare combination is fairly well monitored while also having no dams on its main stem \citep{Vorosmarty:2010} (Fig.~\ref{fig:1}).
 This watershed drains interior high plateau, coastal mountains, and the Canadian Rockies, and thus integrates a diverse mosaic of landscapes, weather, and climate.
@@
 \begin{figure}[h]
-	\centerline{\includegraphics[height=4in]{Fig1_Map2.pdf}}
-	\caption{Flow gauge sites within the Fraser River basin (dark grey) in British Columbia Canada (light grey).
-Flow gauge sites (yellow dots) are scaled by the size of the contributing area.
+\centering
+\includegraphics[width=17.8cm]{Fig1_Map.pdf}
+	\caption{Climate trends in mean annual temperature (MAT) and mean annual precipitation (MAP) within the Fraser River basin overlayed on a digital elevation model of British Columbia, Canada.
+Flow gauge sites (dots) are scaled by the size of the contributing area.
 The map is projected in Albers BC UTMs to provide an equal area depiction of the region but labels are expressed in WGS84 latitude and longitude.}
-	\label{figone}
+\label{fig:1}
 \end{figure}
 
-\section{Data and Methods}
-\subsection{Discharge Data and Flow Metrics}
-To assess signals of climate change in hydrology within the Fraser River basin we downloaded daily flow data from Environment Canada's HYDAT database.
-Within HYDAT we selected hydrometric gauge stations using three criteria: (i) the station was within the Fraser River basin, (ii) the station's flow data were not observed to be dam influenced, and (iii) the station collected data in each month between 1970 and 2007 where no month was missing more than 5 days of data and at least 35 complete years were present.
-We selected this 38-year range because this period maximized the number of gauge stations in the Fraser River basin (\textit{n}=55) operating concurrently over a long enough time period in which climate trends may be observable.
-We interpolated missing data using a linear model in log-space, but of the 763,345 days of data only 23 days were interpolated.
-Because dams control flow and stabilize flow trends, we removed sites where daily flow exhibited unusually high autocorrelation and low variability relative to other sites of similar size, and were observed downstream of, and atypically close to, a dam (\textit{n}=3).
-The downstream distance of gauge stations from dams was determined using GIS spatial layers available in the \textit{BC Data Catalog}, maintained by the provincial government of British Columbia.
-
+\section*{Detecting Climate Change in the Fraser River Basin}
+We focused on changes in hydrology as our riverine manifestation of climate signaling.
 As climate change alters precipitation patterns from snow to rain in temperate regions, rivers are generally predicted to have lower low-flows and higher high-flows, as well as earlier snowmelt \citep{Nijssen:2001}.
 These manifestations of rising temperatures are impacting species \citep{Xenopoulos:2006} and people \citep{Hirabayashi:2013}.
-We summarized our daily discharge data into annual and monthly flow metrics that capture these transitions.
-We used maximum- and minimum-flow estimates to capture trends in extreme discharge and median-flow estimates to capture general annual and seasonal trends.
+Extracting annual and monthly metrics of flow from Fraser River basin gauge station records, we used rolling average maximum- and minimum-flow estimates to capture trends in extreme discharge and rolling average median-flow estimates to capture general annual and seasonal trends.
 Because the timing of flow is particularly important to the phenology of many organisms, we also calculated the day-of-year in which half the annual flow was reached.
 
-\subsection{River Discharge Trend Analysis}
-For a river network to act as a portfolio of climate and dampen the effects of climate change, expressions of climate and their trends over time need to vary throughout the network (Figure~\ref{figtwo}).
-We estimated annual and monthly flow trends for each response variable at each gauge station using a generalized least squares (GLS) model with an AR1 autocorrelation function:
+To quantify changes in flow we fit generalized least squares models with an autoregressive error structure for each of our flow metrics between 1970 and 2007 for each site (\textit{Materials and Methods}).
+Long-term changes in flow metrics varied substantially across the watershed.
+For instance, annual maximum flow decreased in 40 sites but increased in 15 over the last four decades (-26 to 10\% change$\cdot$decade\textsuperscript{-1}, std.
+dev.
+= 6\%).
+Annual median flow showed similar variability but a greater tendency towards rising flow trends with 49 increasing and only six decreasing (-5 to 25\% change$\cdot$decade\textsuperscript{-1} std.
+dev.
+= 5\%).
+Thus, within the Fraser River basin, there are a variety of long-term trends in flow, suggesting a diversity of climate responses and landscape effects.
+
+For a river network to act as a portfolio of climate and dampen the effects of climate change, expressions of climate and their trends over time need to vary throughout the network.
+Using the Western North America Climate tool (ClimateWNA), we quantified the diversity in climate trends within the Fraser River basin over the years concurrent with our hydrology trend estimates \citep{Wang:2016} (Fig.~\ref{fig:1}) (\textit{Materials and Methods}).
+We then summarized the climate portfolio of each flow gauge by calculating the standard deviation in climate trends within their catchments.
+Temperature and precipitation are likely to drive changing hydrology, therefore we included variables in our climate index that capture changes in extreme and mean temperature, as well as changes in precipitation volume and physical state.
+The standard deviations of these climate variables were scaled between 0 and 1 and summed into a general climate index for each catchment.
+Following our hypothesis, this climate portfolio index increases with area (Fig.~\ref{fig:2}), rapidly asymptoting in catchments greater than 40,000 km\textsuperscript{2}.
+
+\begin{figure}%[tbhp]
+\centering
+\includegraphics[width=.8\linewidth]{Fig_ClimPort.pdf}
+	\caption{Climate portfolio index as a function of catchment area.
+The mean trend line (blue) was estimated using robust linear regression and standard error estimates (grey) were calculated using the weighted standard deviation estimate.}
+\label{fig:2}
+\end{figure}
+
+\section*{Quantifying the River Network Portfolio Effect}
+Climate portfolio diversity is largely a function of area (Fig.~\ref{fig:2}).
+Larger areas contain greater landscape complexity and thus a greater variety of climatic expressions.
+The influence of any individual sub-basin on downstream flow trends is also a function of area.
+Larger catchments simply aggregate more water such that their climate portfolio will carry more weight on downstream dampening.
+The interplay between climate portfolio and catchment area on the impact of downstream flow trend dampening makes area a more robust response variable than simply the climate portfolio index.
+Therefore, to test whether networks dampen climate signals we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model (\textit{Materials and Methods}).
+To quantify attenuation, we modeled the change in trend variability with the change in catchment area using an exponential variance function \citep{Pinheiro:2006}.
+The variance function allowed us to predict the range of flow trend values that would be expected as watershed area increased.
+A decreasing range of flow trends with increasing watershed area indicates a dampening network effect.
+
+In the observed data, small watersheds exhibited greater variability around their trend estimates than large watersheds, likely because of greater short-term variation in small catchments \ref{fig:3} \citep{Moore:2015}.
+This relationship between watershed size and trend certainty may pull small watershed trend estimates away from zero, thereby creating the appearance of decreasing trend variability among sites as watershed size increases.
+Using a null-hypothesis framework (\textit{Materials and Methods}) we simulated time-series for each site with no underlying trend.
+These simulations were generated by a random walk process with the observed standard deviation and autocorrelation parameters ($\hat{\upsigma}$, $\hat{\upphi}$) at our sites.
+Using the same generalized least squares model applied to the observed data, we estimated trends for 1000 simulated time series at each site for each response variable (e.g., Figure S3).
+This simulation process, a form of parametric bootstrapping, created distributions of null-expectations with which we compared our observed results.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=17.8]{Fig_AnnFun.pdf}
+	\caption{Annual flow trend attenuation within the Fraser River basin.
+(\textbf{Left}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large).
+Blue lines represent observed attenuation; orange and red represent simulated attenuation that is weaker and stronger than observed.
+(\textbf{Right}) Density plots show null model simulated variance exponents ($\hat{\updelta}$) and the proportion on either side of the observed variance exponent (blue).
+Flow metrics include long-term flow-timing shifts (change per decade in day-of-year (DOY) to half annual flow), where decreasing trends suggest more annual flow is occurring earlier in the year (a), and the percent change per decade in minimum (b), maximum (c) and median (d) annual flow.
+Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
+\label{fig:3}
+\end{figure}
+
+\section*{Results}
+\subsection{Climate Portfolios and Flow Trend Dampening.} Long-term changes in hydrology were less variable in larger catchments, as predicted, demonstrating between 92 and 96\% dampening of flow trends moving from headwater catchments to the network's confluence.
+In other words, the largest catchments were approximately 10 times more stable in their climate response than the smallest catchments.
+For instance, trends ranged between a 6\% reduction and 19\% increase in median flow per decade among small watersheds, while large watersheds ranged between 0.8 and 1.6\% change per decade.
+Similar attenuation was seen in flow timing with a 94\% reduction in the day-of-year to half-annual-flow trends such that small watersheds were reaching half their annual flow between 8 days later and 25 days earlier while large watersheds have only shifted 2 to 4 days earlier between 1970 and 2007.
+Our null model approach illustrates that support for a river network portfolio effect is greater than 90\% for all four annual flow trend metrics and as high as 98\% for median-annual-flow trends (Fig.~\ref{fig:3}).
+Trend attenuation was 3.0 (0.7--9.4 this and hereafter are 90\% intervals), 4.8 (1.1--13.8), 3.4 (0.9--9.0) and 5.3(1.4--14.8) times greater than the null model for annual-flow timing, and minimum, maximum and median annual-flow respectively.
+These statistics provide strong support for the hypothesis that river network portfolio effects contribute to climate dampening in large rivers.
+
+\subsection{Seasonally Shifting Flow Trend Dampening}
+Regional flow-trends and attenuation strength varied by season.
+Analysis of monthly trends revealed that winter flows are getting higher and summer flows are getting lower over time, but that these trends have been more variable in smaller catchments (Fig.~\ref{fig:4}, \textit{SI Appendix}, Figs.
+S~\ref{fig:S1}, S~\ref{fig:S2}).
+For example, winter maximum flows in small catchments ranged from nearly no change to dramatic 47\% increases per decade with decreases only as large as 9\%.
+Despite these extreme locations, on average the basin has only experienced moderate winter maximum flow increases of 5-9\%.
+Summer flows exhibited the opposite, with decreases in high flows over time.
+However, there was weak evidence of attenuation in the spring (Fig.~\ref{fig:4}, \textit{SI Appendix},Figs.
+S~\ref{fig:S1}, S~\ref{fig:S2}).
+Overall, climate portfolios tended to be greater during stable periods of the year and decrease during seasonal transitions (\textit{SI Appendix}, Fig.~\ref{S7}).
+In the spring the likelihood of network-driven attenuation of maximum flow trends was as low as 25\% (Fig.~\ref{fig:4}).
+The deterioration of the network's attenuation strength was coincident with the spring freshet, when snowmelt drives high flows across the basin.
+Synchronization events such as the freshet may subvert river network dampening mechanisms by homogenizing the region's response to seasonally driven climate shifts.
+
+\begin{figure}%[tbhp]
+\centering
+\includegraphics[width=.8\linewidth]{Figure4.pdf}
+	\caption{Monthly maximum flow trend attenuation within the Fraser River basin.
+(\textbf{Left}) Fraser River's basin-wide maximum-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulations.
+Observed values falling further from the center of the density distribution suggest greater evidence for changes in maximum flow and a greater shift in magnitude.
+(\textbf{Center}) Observed monthly Fraser River maximum-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates.
+Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed.
+(\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large), for four seasonally representative months.
+These reflect months in the prior columns and describe the variation in percent change per decade of maximum flow among sites.
+Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
+\label{fig:4}
+\end{figure}
+
+Despite considerable dampening relative to smaller sub-basins, the furthest downstream reaches of the Fraser River are still exhibiting small but detectable signals of climate change.
+Late fall, winter and early spring flows are generally increasing while late spring, summer and early fall flows are decreasing (Fig.~\ref{fig:4}, \textit{SI Appendix} Figs.
+S~\ref{fig:S1}, S~\ref{fig:S2}).
+These findings generally support previous studies demonstrating earlier spring runoff \citep{Dery:2012}, decreasing summer and fall flows \citep{Stahl:2006} and increased winter flows \citep{Healey:2011}.
+Together, this body of work indicates basin-wide effects of climate change, where shifts in winter precipitation from snow to rain is resulting in increased winter runoff and decreased snow pack contributions to summer flows.
+Our work suggests that because the basin is responding heterogeneously to the shifting climate and integrating climate trends, the downstream impacts of climate change on flow are greatly reduced.
+In contrast to previously studied types of climate mitigation where habitats modify climate drivers (e.g., carbon storage) or physically absorb climate change impacts (e.g., mangroves), network climate dampening smooths out extreme climatic trends thereby reducing the impact of local extremes.
+
+\section*{Discussion}
+
+Climate change is already causing economic and conservation challenges for river systems worldwide\citep{Palmer:2009}.
+For example, sockeye salmon in the Fraser River have seen as much as 80\% pre-spawn mortality in years with later run-off and elevated temperatures, leading managers to close the fishery in 2013 and 2015.
+In the absence of climate dampening the impacts of these temperature extremes could have been dramatically worse.
+As climate change progresses, river hydrology will continue to shift, further stressing riverine ecosystems and subsequently demanding responsive management.
+However, river networks provide an underappreciated defense against these climate-change impacts as they are uniquely organized to leverage locally filtered expressions of climate into stability.
+A simple product of form and gravity, stability emerges as river networks integrate landscape heterogeneity, creating a downstream portfolio of climate that smooths local extremes and tempers long-term flow trends.
+Therefore, larger rivers should have flow regimes that are less sensitive to local climate trends.
+
+Our study provides insight into the spatial scaling of climate change.
+We focused on a vast watershed with a remarkably diverse climate portfolio, yet we would still expect that smaller rivers could dampen some degree of climate variability.
+For instance, based on our data, we predict that rivers draining a catchment of 60,000 km\textsuperscript{2} would have 66\% less variability in the rate of change in hydrology than a smaller river draining ~5000 km\textsuperscript{2}.
+While this dampening is less than the 10-fold decrease observed in the entire 220,000 km\textsuperscript{2} Fraser River, smaller watersheds can offer a defense against the impacts of climate change.
+
+There have been extensive efforts to predict the hydrologic sensitivity of rivers due to climate change through downscaled climate projections \citep{Nijssen:2001}.
+Here we demonstrate that large rivers may have an inherent ability to absorb local climate change.
+But just as the destruction of coastal habitats degrades their natural capacity to mitigate sea-level rise or storm events \citep{Arkema:2013}, the management of river basins likely impacts the ability for river networks to attenuate climate trends.
+For example, dams synchronize the flow regimes of rivers \citep{Poff:2007} and logging increases the magnitude and frequency of extreme events such as high and low flows through increased runoff rates \citep{Zhang:2014}.
+These anthropogenic activities may magnify the impacts of climatic shifts such as the transition from snow- to rain-dominated precipitation in some basins.
+The climate cost of these watershed activities should be considered in environmental decision making.
+Climate change is a global challenge; here we suggest large rivers dampen local change by leveraging landscape diversity into a portfolio of climate, an important tool in the climate-mitigation toolbox.
+
+\subsubsection*{SI Information}
+
+\subsubsection*{SI Figures}
+
+\begin{figure}%[tbhp]
+\centering
+\includegraphics[width=.8\linewidth]{FigureS1.pdf}
+	\caption{Monthly minimum flow trend attenuation within the Fraser River basin.
+(\textbf{Left}) Fraser River's basin-wide minimum-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulation.
+Observed values falling further from the center of the density distribution suggest greater evidence for changes in minimum flow and a greater shift in magnitude.
+(\textbf{Center}) Observed monthly Fraser River minimum-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates.
+Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed.
+(\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large), for four seasonally representative months.
+These reflect months in the prior columns and describe the variation in percent change per decade of minimum flow among sites.
+Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
+\label{fig:S1}
+\end{figure}
+
+\begin{figure}%[tbhp]
+\centering
+\includegraphics[width=.8\linewidth]{FigureS2.pdf}
+	\caption{Monthly median flow trend attenuation within the Fraser River basin.
+(\textbf{Left}) Fraser River's basin-wide median-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulation.
+Observed values falling further from the center of the density distribution suggest greater evidence for changes in median flow and a greater shift in magnitude.
+(\textbf{Center}) Observed monthly Fraser River median-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates.
+Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed.
+(\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large),for four seasonally representative months.
+These reflect months in the prior columns and describe the variation in percent change per decade of median flow among sites.
+Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
+\label{fig:S2}
+\end{figure}
+
+\matmethods{
+\subsection*{Materials and Methods}
+\subsection{Data.} We downloaded daily flow data from Environment Canada's HYDAT database.
+Within HYDAT we selected hydro-metric gauge stations using three criteria: (i) the station was within the Fraser River basin, (ii) the station's flow data were not observed to be dam influenced, and (iii) the station collected data in each month between 1970 and 2007 and no month was missing more than 5 days data.
+We selected this 38-year range because this period maximized the number of gauge stations in the Fraser River basin (*n*=55) operating concurrently over a long enough time period in which climate trends may be observable.
+We interpolated missing data using a linear model in log-space, but of the 13,879 days of data only 23 days were estimated.
+These daily data were smoothed using a five day rolling average before summarizing into annual and monthly response variables.
+This smoothing technique was used to reduce the influence of erroneous and unusually extreme values.
+Because dams control flow and stabilize flow trends, we removed sites where daily flow exhibited unusually high autocorrelation and low variability relative to other sites of similar size, and were observed downstream of, and atypically close to, a dam (*n*=3).
+The downstream distance of gauge stations from dams was determined using GIS spatial layers available in the *BC Data Catalog*, maintained by the provincial government of British Columbia.
+Finally, we summarized the daily data into annual and monthly maximum, median and minimum flow estimates and the day-of-year to half-annual-flow.
+
+\subsection{River discharge trend analysis.} We estimated annual and monthly flow trends for each response variable at each gauge station using a generalized least squares (GLS) model with an AR1 autocorrelation function:
 \begin{linenomath*}
@@
 
-\begin{figure}[h]
-	\centerline{\includegraphics[height=3in]{Fig2_Flow-Time.pdf}}
-	\caption{Annual median flow at 55 sites within the Fraser River basin.
-Flow rates are scaled at each site by the root mean square in order to make flow rates comparable among sites.
-Lines are loess smoothers with a smoothing span of 0.25 predicted each month between 1970 and 2007.
-Line and point size are scaled to each flow gauge site's watershed area.
-The data show inter-annual variability within sites with varying magnitudes and directions among sites, illustrating variable long-term change in hydrology across the basin.}
-	\label{figtwo}
-\end{figure}
-
-\subsection{River Network Trend Attenuation Analysis}
-We hypothesized that larger catchments within the Fraser River watershed with more diverse climate portfolios would have dampened long-term hydrological change when compared to their smaller sub-catchments.
-Therefore, to test whether networks dampen climate signals we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model:
+\subsection{Basin trend attenuation analysis.}
+To measure network driven dampening of climate signalling, we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model:
 \begin{linenomath*}
@@
 This variance function allowed us to predict the range of flow trend values that would be expected as watershed area increased.
-A decreasing range of flow trends with increasing watershed area would indicate a climate-dampening network effect.
 
-\subsection{River Discharge Trend Simulation}
+\subsection{River Attnuation Null Model Simulations}
 In the observed data, small watersheds exhibited greater variability around their trend estimates than large watersheds, likely because of greater short-term variation in small catchments \citep{Moore:2015}.
@@
 
-\begin{figure}[h]
-	\centerline{\includegraphics[height=4.5in]{Fig3_Annual-Funnel.pdf}}
-	\caption{Annual flow trend attenuation within the Fraser River basin.
-(\textbf{Left}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}).
-Blue represents observed attenuation; orange and red represent simulated attenuation that is weaker and stronger than observed.
-(\textbf{Right}) Density plots show null model simulated variance exponents ($\hat{\updelta}$) and the proportion on either side of the observed variance exponent (blue).
-Flow metrics include long-term flow-timing shifts (change per decade in day-of-year (DOY) to half annual flow), where decreasing trends suggest more annual flow is occurring earlier in the year (a), and the percent change per decade in minimum (b), maximum (c) and median (d) annual flow.
-Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
-	\label{figthree}
-\end{figure}
-
-\subsection{Dampening Summary Statistics}
-We calculated attenuation certainty as the percentage of simulated $\hat{\updelta}$ that were less than the observed $\hat{\updelta}$.
-We calculated attenuation strength as the ratio between the standard deviation at the smallest and largest watersheds as defined by the exponential variance function (i.e.,\\$\sqrt{f(\mathrm{area}_1)} / \sqrt{f(\mathrm{area}_2)}$ ).
-To compare the degree of attenuation observed with the null-simulated attenuation, we compared this ratio with the same ratio calculated from the null-simulated data (i.e., $\left( \sqrt{f(\mathrm{area}_1)} / \sqrt{f(\mathrm{area}_2)} \right) / \left( \sqrt{f(\mathrm{area}_{\mathrm{null} 1})} / \sqrt{f(\mathrm{area}_{\mathrm{null} 2})} \right)$ ).
+\subsection{Dampening summary statistics.} We calculated attenuation certainty as the percentage of simulated $\hat{\updelta}$ that were less than the observed $\hat{\updelta}$.
+We calculated attenuation strength as the ratio between the standard deviation at the smallest and largest watersheds as defined by the exponential variance function (i.e., $\sqrt{f(\mathrm{area}_1)} / \sqrt{f(\mathrm{area}_2)}$).
+To compare the degree of attenuation observed with the null-simulated attenuation, we compared this ratio with the same ratio calculated from the null-simulated data (i.e., $\left( \sqrt{f(\mathrm{area}_1)} / \sqrt{f(\mathrm{area}_2)} \right) / \left( \sqrt{f(\mathrm{area}_{\mathrm{null} 1})} / \sqrt{f(\mathrm{area}_{\mathrm{null} 2})} \right)$).
 Finally, to estimate how the day-of-year to half-annual-flow was changing, we estimated the steepest point of each site's logistic curve and multiplied this slope value by 365 to revert from our 0--1 scale to our original annual range of 0--365.
 To get a decadal rate we multiplied these results by 3.8 (number of decades in 38 years).
-We estimated the steepest part of the logistic curve using the \textit{divide by four} rule \citep{Gelman:2008}, where dividing the coefficient by four equals the first derivative (slope) of the logistic curve at its steepest point.
+We estimated the steepest part of the logistic curve using the *divide by four* rule[@Gelman:2008] (dividing the coefficient by four equals the first derivitive of the logistic curve at its steepest point).
 
-\section{Results}
-\subsection{Hydrologic Trend Diversity}
-Long-term changes in flow metrics varied substantially across the watershed (Figure~\ref{figthree}).
-For instance, annual maximum flow decreased in 38 sites but increased in 17 over the last four decades (-28 to 15\% change$\cdot$decade\textsuperscript{-1}, std.
-dev.
-= 6\%).
-Annual median flow showed similar variability but a greater tendency towards rising flow trends with 49 increasing and only six decreasing (-5 to 25\% change$\cdot$decade\textsuperscript{-1} std.
-dev.
-= 5\%).
-Thus, within the Fraser River basin, there are a variety of long-term trends in flow, suggesting a diversity of climate responses.
+\subsection*{Climate} In order to develop a climate index for each flow-gauge catchment we needed to estimate climate trends across the Fraser River basin and then spatially summarise them by flow-gauge catchment.
+Using Whitebox Geospatial Analysis Tools, a freely available and open source geospatial analysis software \citep{}, we delineated the Fraser River basin and the catchment areas of each flow-gauge station.
+Digital elevation models provided by the provincial government of British Columbia and state goverment of Washinton were pre-processed with a breaching algorithm and stream burned to facilitate proper flow path and accumulation models.
+We then used Climate WNA to estimate historic climate values on an evenly spaced 1 km\textsuperscript{2} grid across the Fraser Basin for each year and month in our study.
+Climate variables included the mean annual temperature (MAT), and precipitation (MAP), extreme minimum (EMT) and maximum (EXT) temperature and precipitation as snow (PAS).
+Climate variable trends were calculated at each grid point using \ref{eq1}.
 
-\subsection{Annual Hydrologic Trend Dampening}
-Long-term changes in hydrology were less variable in larger catchments, as predicted, demonstrating between 85 and 92\% dampening of flow trends moving from headwater catchments to the network's confluence.
-In other words, the largest catchments were approximately 10 times more stable in their climate response than the smallest catchments.
-For instance, trends ranged between a 6\% reduction and 19\% increase in median flow per decade among small watersheds, while large watersheds ranged between 0.8 and 1.6\% change per decade.
-Similar attenuation was seen in flow timing with a 94\% reduction in the day-of-year to half-annual-flow trends such that small watersheds were reaching half their annual flow between 8 days later and 25 days earlier while large watersheds have only shifted 2 to 4 days earlier between 1970 and 2007.
+The standard deviation in climate trends represent the strength of each basins' climate portfolio.
+Climate trend standard deviation estimates were calculated using the `zonal\_stats` function in the \textit{rasterstats} python module.
+To demonstrate how this climate index scales with watershed area we regressed climate portoflio index values onto watershed area (log(km\texsuperscript{2})).
+Two outlier catchments had high leverage climate portfolio index values.
+Rather than remove these points, we used robust linear regression to downweight their influence using Huber's method of M-estimation.
 
-A null model approach illustrated that support for a river network portfolio effect was greater than 90\% for three of four annual flow trend metrics and as high as 98\% for median-annual-flow trends (Figure~\ref{figthree}).
-Trend attenuation was 3.1 (0.7--9.1 this and hereafter are 90\% intervals), 1.6 (0.3--4.6), 3.4 (1.0--8.5) and 5.4(1.5--14.6) times greater than the null model for annual-flow timing, and minimum, maximum and median annual-flow respectively.
-These statistics provide strong support for the hypothesis that river network portfolio effects contribute to climate dampening in large rivers.
+\showmatmethods % Display the Materials and Methods section
 
-\subsection{Seasonal Hydrologic Trend Dampening}
-Regional flow-trends and attenuation strength varied by season.
-Analysis of monthly trends revealed that winter flows have been getting higher and summer flows have been getting lower over time, but that these trends have been more variable in smaller catchments (Figure~\ref{figfour}, S1, S2).
-For example, winter maximum flows in small catchments ranged from nearly no change to dramatic 44\% increases per decade with decreases only as large as 15\%.
-Despite these extreme locations, on average the basin has only experienced moderate winter maximum flow increases of 5--10\%.
-Summer flows exhibited the opposite, with decreases in high flows over time.
-However, there was weak evidence of attenuation in the spring (Figure~\ref{figfour}, S1, S2).
-For example, in the spring the likelihood of network-driven attenuation of maximum flow trends was as low as 24\% (Figure~\ref{figfour}).
-The deterioration of the network's attenuation strength was coincident with the spring freshet, when snowmelt drives high flows across the basin.
-Synchronization events such as the freshet may subvert river network dampening mechanisms by homogenizing the region's response to climate shifts.
-
-\begin{figure}[htp]
-	\centerline{\includegraphics[height=5.2in]{Fig4_Max-Monthly-Density-Funnel.pdf}}
-	\caption{Monthly maximum flow trend attenuation within the Fraser River basin.
-(\textbf{Left}) Fraser River's basin-wide maximum-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulations.
-Observed values falling further from the center of the density distribution suggest greater evidence for changes in maximum flow and a greater shift in magnitude.
-(\textbf{Center}) Observed monthly Fraser River maximum-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates.
-Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed.
-(\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}) for four seasonally representative months.These reflect months in the prior columns and describe the variation in percent change per decade of maximum flow among sites.
-Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
-	\label{figfour}
-\end{figure}
-
-\section{Discussion}
-Climate change is already causing economic and conservation challenges for river systems worldwide \citep{Palmer:2009}.
-For example, sockeye salmon in the Fraser River have seen as much as 80\% pre-spawn mortality in years with later run-off and elevated temperatures, leading managers to reduce harvest rates in these years.
-In the absence of climate dampening the impacts of these temperature extremes could have been dramatically worse.
-As climate change progresses, river hydrology will continue to shift, further stressing riverine ecosystems and subsequently demanding responsive management.
-However, river networks provide an underappreciated defense against these climate-change impacts as they are uniquely organized to leverage locally-filtered expressions of climate into stability.
-A simple product of form and gravity, stability emerges as river networks integrate landscape heterogeneity, creating a downstream portfolio of climate that smooths local extremes and tempers long-term flow trends.
-Therefore, larger rivers should have flow regimes that are less sensitive to local climate trends.
-
-Despite these dampening processes, river networks are still vulnerable to climate change.
-In the Fraser River, the furthest downstream reaches are still exhibiting small but detectable signals of hydrological shifts.
-Late fall, winter and early spring flows have generally increased while late spring, summer and early fall flows have decreased (Figure~\ref{figfour}, S1, S2).
-These findings generally support previous studies demonstrating earlier spring runoff  \citep{Dery:2012}, decreasing summer and fall flows  \citep{Stahl:2006} and increased winter flows  \citep{Healey:2011}.
-Together, this body of work indicates basin-wide effects of climate change, where shifts in winter precipitation from snow to rain is resulting in increased winter runoff and decreased snow pack contributions to summer flows.
-Our work suggests that because the basin is responding heterogeneously to the shifting climate and integrating climate trends, the downstream impacts of climate change on flow have been greatly reduced.
-In contrast to previously studied types of climate mitigation where habitats modify climate drivers (e.g., carbon storage) or physically absorb climate change impacts (e.g., mangroves), network climate dampening smooths out extreme climatic trends thereby reducing but not eliminating the impact of local extremes.
-
-Our study provides insight into the spatial scaling of climate change.
-We focused on a vast watershed with a remarkably diverse climate portfolio, yet we would still expect that smaller rivers could dampen some degree of climate variability.
-For instance, based on our data, we predict that rivers draining a catchment of 60,000 km\textsuperscript{2} would have 66\% less variability in the rate of change in hydrology than a smaller river draining ~5000 km\textsuperscript{2}.
-While this dampening is less than the 10-fold decrease observed in the entire 220,000 km\textsuperscript{2} Fraser River, smaller watersheds can offer a defense against the impacts of climate change.
-
-There have been extensive efforts to predict the hydrologic sensitivity of rivers due to climate change through downscaled climate projections \citep{Nijssen:2001}.
-Here we demonstrate that large rivers may have an inherent ability to absorb local climate change.
-But just as the destruction of coastal habitats degrades their natural capacity to mitigate sea-level rise or storm events \citep{Arkema:2013}, the management of river basins likely impacts the ability for river networks to attenuate climate trends.
-For example, dams synchronize the flow regimes of rivers  \citep{Poff:2007} and logging increases the magnitude and frequency of extreme events such as high and low flows through increased runoff rates \citep{Zhang:2014}.
-These anthropogenic activities may magnify the impacts of climatic shifts such as the transition from snow- to rain-dominated precipitation in some basins.
-The climate cost of these watershed activities should be considered in environmental decision making.
-Climate change is a global challenge; here we suggest large rivers dampen local change by leveraging landscape diversity into a portfolio of climate, an important tool in the climate-mitigation toolbox.
-
-\acknowledgments
-All the data and R scripts used in this study are freely available on github at https://github.com/kchezik/River-Network-Flow-Trends.
-This study could not have been done without the decades of flow data collected by those at Environment Canada.
+\acknow{This study could not have been done without the decades of flow data collected by those at Environment Canada.
 K.A.
@@
 Smith Conservation Research Fellowship.
+\showacknow % Display the acknowledgments section
 
-\begin{thebibliography}{30}
-\providecommand{\natexlab}[1]{#1}
-\expandafter\ifx\csname urlstyle\endcsname\relax
-  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
-  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
-  \urlstyle{rm}\Url}\fi
+% \pnasbreak splits and balances the columns before the references.
+% If you see unexpected formatting errors, try commenting out this line
+% as it can run into problems with floats and footnotes on the final page.
+\pnasbreak
+
+% Bibliography
+\bibliography{pnas-sample}
 
@@
 
-\end{thebibliography}
-
-%\bibliography{/Users/kylechezik/Documents/lit}
-
-\end{document}
+\bibitem{Wang:2016}
+T.~Wang, A.~Hamann, D.~Spittlehouse, and C.~Carroll.
+\newblock Locally downscaled and spatially customizable climate data for
+  historical and future periods for north america.
+\newblock {\em PLoS ONE}, 11(6):e0156720, 2016.

### 2017-03-15 Code_03_Post_Analysis_Null_Test.R

@@ -33,10 +33,6 @@ M.Data = lapply(vars,function(x){
 })[[length(vars)]] %>% ungroup()
 rm(vars, MonthlyVars)
 
-#remove = c("08MH006","08MH076")
-#M.Data = filter(M.Data, Station.ID%in%remove == F)
-#Y.Data = filter(Y.Data, Station.ID%in%remove == F)
-
 setwd("~/sfuvault/Simon_Fraser_University/PhD_Research/Projects/River-Network-Flow-Trends")
 # this function fits slopes to real data
 fit_slopes <- function(flow.dat, response, pred.var, vars) {
@@ -87,7 +83,8 @@ fit_var <- function(slopes, ind, clim, var2 = F) {
 	w = clim/sd(clim)
 	if(var2 == T){
 		tryCatch({
-			m <- gls(scaled_slope~x, weights = varComb(varExp(form = ~v), varExp(form = ~w)),
+			#m <- gls(scaled_slope~x, weights = varComb(varExp(form = ~v), varExp(form = ~w)),
+			m <- gls(scaled_slope~x, weights = varComb(varExp(form= ~exp(clim)),varExp(form= ~sqrt(ind)/1e3)),
 							 control = glsControl(maxIter = 1000L, msMaxIter = 1000L))
 			varexpA <- m$model[[1]][[1]][[1]]
 			varexpC <- m$model[[1]][[2]][[1]]
@@ -103,6 +100,7 @@ fit_var <- function(slopes, ind, clim, var2 = F) {
 	if(var2 == F){
 		tryCatch({
 			m <- gls(scaled_slope~x, weights = varExp(form= ~sqrt(ind)/1e3),
+			#m <- gls(scaled_slope~x, weights = varExp(form= ~exp(clim)),
 							 control = glsControl(maxIter = 1000L, msMaxIter = 1000L))
 			varexp <- m$model[[1]][[1]]
 			sigma <- m$sigma * scale_factor
@@ -157,7 +155,7 @@ vars = c("Area","emt.sd","ext.sd","map.sd","mat.sd","pas.sd")
 out_doy2 <- iter_null_sim(Y.Data, "DOY2.logit", "Area", F, vars, 1000L, T)
 out_min <- iter_null_sim(Y.Data, "log(min.log.sd)", "Area", F, vars, 1000L, T)
 out_max <- iter_null_sim(Y.Data, "log(max.log.sd)", "Area", F, vars, 1000L, T)
[…]

### 2017-03-15 Code_06 AttenStats.R

@@ -3,6 +3,7 @@ load("out_doy2.RData")
 load("out_med.RData")
 load("out_min.RData")
 load("out_max.RData")
+library(tidyverse)
 
 #This function finds the predicted attenuation values at any two catchment areas and returns the percent decrease in predicted range values when moving from small catchments to a large catchment.
 pred.F = function(x, Area){
@@ -46,8 +47,8 @@ Area_Small = min(out_med$Area$real_slopes$Area)
 #Area_Small = 5000
 Area_Large = max(out_med$Area$real_slopes$Area)
 #Area_Large = 60000
-expReal = out_max$Area$real_varexp$varexp
-expNull = out_max$Area$sim_varexp$varexp
+expReal = out_med$Area$real_varexp$varexp
+expNull = out_med$Area$sim_varexp$varexp
 atten_compare = ratio_atten(expReal,expNull,Area_Small,Area_Large)
 
 #Uncomment the last line of the ratio_atten function before plotting. Also, nice to comment out the print function and the plotting line for computational speed.  
@@ -77,7 +78,7 @@ Y.Data = plyr::ddply(Y.Data,"Station.ID",plyr::mutate, med.log.sd = scale(Median
 #function to predict doy to half annual flow given slope, intercept and year.
 logit.pred = function(slope, intercept, year.center){data.frame(pred.points = plogis(intercept+slope*year.center))}
 #expand the slope/intercept data so each site has a row for each year.
-new.dat = expand.grid(Station.ID = out_doy2$real_slopes$Station.ID, year.center = sort(unique(Y.Data$Year.Center))) %>% inner_join(out_doy2$real_slopes)
+new.dat = expand.grid(Station.ID = out_doy2$Area$real_slopes$Station.ID, year.center = sort(unique(Y.Data$Year.Center))) %>% inner_join(out_doy2$Area$real_slopes)
 #apply the logit.pred function to the new.dat dataframe.
 for.plot = plyr::mdply(select(new.dat, intercept, slope, year.center), logit.pred)
 #add the station.id to the predicted data.
@@ -88,10 +89,10 @@ ggplot(for.plot, aes(year.center, pred.points*365, group = station.id))+
 	theme_classic()
 
 #isolate the doy gls coefficients.
-varexp = out_doy2$real_varexp$varexp
[…]

### 2017-03-15 Code_07 Monthly_Plot.R

@@ -47,7 +47,7 @@ density_plot = function(data, response, type, xlimit){
 }
 
 #Funnel Plot Wrapper.
-funnel.plots = function(dat, response, yaxis.DOY = F, d.xaxis = F, pane, funnel.y = "", axis.Ly = "", xAxis = F, ylimit = NULL, iter = 200){
+funnel.plots = function(dat, response, yaxis.DOY = F, d.xaxis = F, pane, funnel.y = "", axis.Ly = "", xAxis = F, ylimit = NULL, iter = 200, monthly = F, min = NA, max = NA){
 	library(tidyverse);library(wesanderson)
 	
 	############################################ Prepare Data ###############################################
@@ -60,12 +60,12 @@ funnel.plots = function(dat, response, yaxis.DOY = F, d.xaxis = F, pane, funnel.
 	#Cleanup and arrange observed flow and climate trend data.
 	flow.trends = arrange.data(dat,response)
 	#Create a colour gradient for the climate index values.
-	cols = color.gradient(flow.trends$std.clim)
+	cols = color.gradient(flow.trends$std.clim, monthly = monthly, min = min, max = max)
 	
 	########################################### Plot Raw Data ##############################################
 	
 	#Set initial plotting parameters
-	par(mar = c(0,5,0.2,0), family = "serif", bg = "white",fg = "white")
+	par(mar = c(0,5,0.2,0), family = "serif", bg = "white",fg = "white", las = 1)
 	#Determine axis labels.
 	axis.Fun(flow.trends, yaxis.DOY, ylimit)
 	#Plot initial blank region to be filled.
@@ -107,15 +107,25 @@ funnel.plots = function(dat, response, yaxis.DOY = F, d.xaxis = F, pane, funnel.
 	}
 }
 
+MaxMinClim = function(dat){
+	tp = range(unlist(lapply(dat,function(x){
+		range(x$Area$real_slopes$std.clim)
+	})))
+	minC <<- tp[1]
+	maxC <<- tp[2]
+}
+
+
 load("max_month.RData")
+MaxMinClim(out_max_month)
 #pdf("Fig4_Max-Monthly-Density-Funnel.pdf", width = 11, height = 8.5)
 monthlyScreen()
[…]

### 2017-03-15 Code_08 Area_ClimatePortfolioPlot.R

@@ -8,7 +8,7 @@ remove = c("08MH006","08MH076")
 #"08MH090","08MH056","08LG016","08LG048","08MH103","08MH001","08MH016","08MH029"
 df$cols[which(df$Station.ID%in%remove)] = "#969696"
 df$loc[which(df$Station.ID%in%remove)] = "coastal"
-ggplot(df,aes(std.clim, slope, color = cols)) + geom_point()
+ggplot(df,aes(exp(std.clim), slope, color = cols)) + geom_point()
 
 #Look at each climate variable and the index against area.
 gather(data = df, key = "climateIndex", value = "IndexValue", emt.sd, ext.sd, map.sd, mat.sd, pas.sd, std.clim) %>% 

### 2017-03-15 Code_08 Supplementary_Plots.R

@@ -88,3 +88,21 @@ ggplot(final, aes(sqrt(Area), slope, colour = label)) +
 ggsave("FigS4_basin_simulation.pdf", width = 11, height = 8.5)
 
 
+#S7 Plot the Climate Portfolio over each month 
+load("out_max_month.RData")
+dat = plyr::ldply(out_max_month, function(x){
+	clim = x$Area$real_slopes$std.clim
+})
+names(dat)[2:56] = out_max_month$`1`$Area$real_slopes$Station.ID
+
+df = gather(dat, key = "Station.ID",value = "value", 2:56)
+df = df %>% group_by(Station.ID) %>% mutate(value.adj = zero_one(value))
+df1 = df %>% mutate(year = 0)
+df2 = df %>% mutate(year = 12)
+df3 = df %>% mutate(year = 24)
+df = bind_rows(df1,df2,df3)
+df = df %>% mutate(time = nMonth + year)
+
+ggplot(df, aes(time, value.adj, color = Station.ID)) + geom_smooth(se = F, span = 1, n = 36) + labs(x = "Month", y = "Cliamte Portfolio") + scale_x_continuous(breaks = c(13:24), limits = c(13,24), labels = as.character(1:12)) + theme_minimal() + theme(legend.position = "none")
+
+ggsave("FigS7_Seasonal_Climate_Portfolio.pdf", width = 11, height = 8.5)

### 2017-03-15 attenPlot/R/arrange_data.R

@@ -10,8 +10,8 @@
 #
 arrange.data = function(df, response){
 	tp = df[[grep(response, names(df))]]$real_slopes
-	tp = tp %>% select(8:ncol(tp)) %>% apply(.,2,function(x) zero_one(x)) %>% 
-		apply(.,1,function(x) sum(x)) %>% mutate(tp, std.clim = .)
+	#tp = tp %>% select(8:ncol(tp)) %>% apply(.,2,function(x) zero_one(x)) %>% 
+	#	apply(.,1,function(x) sum(x)) %>% mutate(tp, std.clim = .)
 	names(tp)[grep(response,names(tp))] = "resp"
 	tp %>% mutate(sqrt.resp = sqrt(resp)) %>% mutate(se.upper = slope+se, se.lower = slope-se) %>% arrange(resp)
 }

### 2017-03-15 attenPlot/R/color_gradient.R

@@ -8,10 +8,12 @@
 #' @export
 #' @examples
 #' color.gradient(vec, pal = "GnBu", shade = "99")
-color.gradient = function(vec, pal = "GnBu", shade = "99"){
+color.gradient = function(vec, pal = "GnBu", shade = "99",monthly = F, min = NA, max = NA){
 	library(RColorBrewer)
 	#vector scaled 0-1 for function
-	col = zero_one(vec)
+	if(monthly == T){
+		col = ((vec-min))/(diff(c(min,max)))
+	} else col = zero_one(vec)
 	#color ramp function
 	gradient = brewer.pal(9, pal)[3:9]
 	FUN = colorRamp(gradient, bias=1)

### 2017-03-15 attenPlot/R/line_sims.R

@@ -23,3 +23,17 @@ line.sims = function(df, col, LW, shade = "10", transp = T, intercept, slope, re
 		lower = intercept + slope*(resp.sqrt - mean(resp.sqrt)) - 1.96 * sqrt(sigma^2 * exp(2*(resp.sqrt/1e3)*y[1]))
 		lines(resp.sqrt, lower, col = colur, lwd = LW)
 	})}
+
+# line.sims = function(df, col, LW, shade = "10", transp = T, intercept, slope, resp.sqrt, sigma, clim){
+# 	apply(df,1,function(y){
+# 		if(transp == T) colur = paste(col,shade,sep="")
+# 		else colur = col
+# 		v = resp.sqrt/sd(resp.sqrt)
+# 		w = clim/sd(clim)
+# 		upper = intercept + slope*(resp.sqrt - mean(resp.sqrt)) + 1.96 * sqrt(sigma^2 * exp(2*(w*v)*y[1]))
+# 		#upper = predict(loess(upper~resp.sqrt, span = 1))
+# 		lines(resp.sqrt, upper, col = colur, lwd = LW)
+# 		lower = intercept + slope*(resp.sqrt - mean(resp.sqrt)) - 1.96 * sqrt(sigma^2 * exp(2*(w*v)*y[1]))
+# 		#lower = predict(loess(lower~resp.sqrt, span = 1))
+# 		lines(resp.sqrt, lower, col = colur, lwd = LW)
+# 	})}

### 2017-03-15 test.R

@@ -1,0 +1,11 @@
+library(attenPlot); library(tidyverse)
+n = exp(seq(log(35),log(250000), by = 0.2))
+index.sd = -1.3575 + log(n)*0.3869
+plot(index.sd~sqrt(n))
+df = expand.grid(n = n, sd.samp = c(1:1000)) %>% left_join(.,data.frame(n = n, index.sd = index.sd))
+
+df = df %>% mutate(sd.sim = apply(., 1, function(x){
+		sd(rnorm(n = x[["n"]], mean = 10000, sd = x[["index.sd"]]))
+	}))
+
+ggplot(df,aes(log(n),sd.sim)) + geom_point() + geom_abline(intercept = 1) + geom_line(aes(log(n),index.sd), color = "blue")

### 2017-03-17 drafts/PNAS/Chezik_PNAS_MS.tex
Updated PNAS paper. Merged old drafts.

--- 2017-03-15 drafts/PNAS/ms_pnas.tex
@@
 
+\usepackage{upgreek}
+
 \templatetype{pnasresearcharticle} % Choose template
-% {pnasresearcharticle} = Template for a two-column research article
-% {pnasmathematics} = Template for a one-column mathematics article
-% {pnasinvited} = Template for a PNAS invited submission
 
@@
 Here we demonstrate that large free-flowing river networks dampen local long-term flow trends by integrating a large and varied climate portfolio across a diverse and complex landscape.
-Integrating the benefits of these ecosystems into climate change mitigation stategies is essential in reducing the increasingly extreme impacts of climate change.}
+Integrating the benefits of these ecosystems into climate change mitigation strategies is essential in reducing the increasingly extreme impacts of climate change.}
 
@@
 \begin{abstract}
-	Climate change is dramatically altering temperate rivers --- snow-melt is shifting ever earlier \citep{Rauscher:2008]} floods are becoming more frequent \citep{Hirabayashi:2013}, and low flows more persistent \citep{McCabe:2014}.
+	Climate change is dramatically altering temperate rivers --- snow-melt is shifting ever earlier \cite{Rauscher:2008} floods are becoming more frequent \cite{Hirabayashi:2013}, and low flows more persistent \cite{McCabe:2014}.
 Here we describe a novel pathway by which river networks dampen local hydrologic climate change through the aggregation of upstream climate portfolio assets.
@@
 If this is the case, add \parshape=0 to the end of the list environment.
-\dropcap{A}s billions are spent adapting to climate change by such means as building dikes or improving stormwater drainage \citep{Narain:2011}, certain habitats and processes provide a natural defense system that mitigates climate change impacts \citep{Jones:2012}.
-For example, coastal habitats such as oyster reefs, mangrove forests, and eelgrass beds shield 67\% of the United States coastline and 1.4 million people from sea level rise and storms \citep{Arkema:2013}.
-Protection and enhancement of these natural systems has become a key component of many climate mitigation strategies \citep{Guerry:2015}, as they return multiple cost-effective services in contrast to engineered alternatives \citep{Jones:2012}.
-For instance, revitalization of the Mississippi delta not only improves storm protection, but also increases production and sustainability of fisheries and wildlife resources, protects the water supply, and reduces the impact of wastewater effluent \citep{Conservation:1998}.
+\dropcap{A}s billions are spent adapting to climate change by such means as building dikes or improving stormwater drainage \cite{Narain:2011}, certain habitats and processes provide a natural defense system that mitigates climate change impacts \cite{Jones:2012}.
+For example, coastal habitats such as oyster reefs, mangrove forests, and eelgrass beds shield 67\% of the United States coastline and 1.4 million people from sea level rise and storms \cite{Arkema:2013}.
+Protection and enhancement of these natural systems has become a key component of many climate mitigation strategies \cite{Guerry:2015}, as they return multiple cost-effective services in contrast to engineered alternatives \cite{Jones:2012}.
+For instance, revitalization of the Mississippi delta not only improves storm protection, but also increases production and sustainability of fisheries and wildlife resources, protects the water supply, and reduces the impact of wastewater effluent \cite{Conservation:1998}.
 The value of climate buffering habitats stands only to increase as average global temperatures climb.
-The Paris Agreement of 2016 aims to limit global temperature rise to 1.5\degree{C} \citep{Hulme:2016}, a goal that concedes a minimum doubling of observed warming \citep{Hartmann:2013}; therefore, an urgent challenge and opportunity is discovering, maintaining and restoring systems that naturally buffer against the oncoming impacts of climate change.
+The Paris Agreement of 2016 aims to limit global temperature rise to 1.5\degree{C} \cite{Hulme:2016}, a goal that concedes a minimum doubling of observed warming \cite{Hartmann:2013}; therefore, an urgent challenge and opportunity is discovering, maintaining and restoring systems that naturally buffer against the oncoming impacts of climate change.
 
-In the face of changing global precipitation patterns \citep{Donat:2016}, river networks may offer an unappreciated defense against shifting flow regimes under climate change\citep{Hartmann:2013,Palmer:2009}.
-Earlier snowmelt \citep{Rauscher:2008}, reduced snow pack \citep{McCabe:2014}, shifts from snow to rain, and changes in the annual distribution of precipitation are all impacting stream flow \citep{Hartmann:2013}, resulting in flashier flows with increased frequency of flooding \citep{Hirabayashi:2013}, longer periods of drought and critically low flows for wildlife and agriculture \citep{Melillo:2014}.
+In the face of changing global precipitation patterns \cite{Donat:2016}, river networks may offer an unappreciated defense against shifting flow regimes under climate change\cite{Hartmann:2013,Palmer:2009}.
+Earlier snowmelt \cite{Rauscher:2008}, reduced snow pack \cite{McCabe:2014}, shifts from snow to rain, and changes in the annual distribution of precipitation are all impacting stream flow \cite{Hartmann:2013}, resulting in flashier flows with increased frequency of flooding \cite{Hirabayashi:2013}, longer periods of drought and critically low flows for wildlife and agriculture \cite{Melillo:2014}.
 However, we hypothesize that large free-flowing river networks may naturally dampen these signals of climate change by integrating across varied landscapes and different manifestations of climate.
-Processes that aggregate across asynchronous components tend to dampen the aggregate, a process known as the portfolio effect \citep{Doak:1998}.
-Climate is expressed differently and asynchronously over the landscape \citep{George:2015} and on account of their branching architecture and directional flow, river networks integrate these varied climate manifestations \citep{Peterson:2013}, potentially dampening the local climatic response of sub-catchments.
+Processes that aggregate across asynchronous components tend to dampen the aggregate, a process known as the portfolio effect \cite{Doak:1998}.
+Climate is expressed differently and asynchronously over the landscape \cite{George:2015} and on account of their branching architecture and directional flow, river networks integrate these varied climate manifestations \cite{Peterson:2013}, potentially dampening the local climatic response of sub-catchments.
 Large free-flowing rivers may have diverse climate portfolios, with sub-catchments acting as assets in their portfolio, thereby dampening local climate impacts.
-Portfolio effects in rivers are already known to dampen short-term variation \citep{Moore:2015,Yeakel:2014} but attenuation of long-term climate driven shifts have yet to be examined.
+Portfolio effects in rivers are already known to dampen short-term variation \cite{Moore:2015,Yeakel:2014} but attenuation of long-term climate driven shifts have yet to be examined.
 
 Here we consider whether river networks attenuate local long-term climate change through analysis of hydrometric data from one of the largest free-flowing rivers in North America.
-Located in British Columbia, Canada, the Fraser River drains an area approximately the size of the United Kingdom (\texttildelow 217,000 km\textsuperscript{2}) and in a rare combination is fairly well monitored while also having no dams on its main stem \citep{Vorosmarty:2010} (Fig.~\ref{fig:1}).
+Located in British Columbia, Canada, the Fraser River drains an area approximately the size of the United Kingdom ($\sim$217,000 km\textsuperscript{2}) and in a rare combination is fairly well monitored while also having no dams on its main stem \cite{Vorosmarty:2010} (Fig.~\ref{fig:1}).
 This watershed drains interior high plateau, coastal mountains, and the Canadian Rockies, and thus integrates a diverse mosaic of landscapes, weather, and climate.
-As with other mid-latitude rivers \citep{Bindoff:2013}, the Fraser River has expressed increasing trends in discharge volume and variability over the last several decades \citep{Dery:2012,Morrison:2002}.
+As with other mid-latitude rivers \cite{Bindoff:2013}, the Fraser River has expressed increasing trends in discharge volume and variability over the last several decades \cite{Dery:2012,Morrison:2002}.
 The river's discharge is monitored by hydrometric gauge stations throughout the watershed.
@@
 
-\begin{figure}[h]
+\begin{figure*}[h]
 \centering
@@
 \label{fig:1}
-\end{figure}
+\end{figure*}
 
@@
 We focused on changes in hydrology as our riverine manifestation of climate signaling.
-As climate change alters precipitation patterns from snow to rain in temperate regions, rivers are generally predicted to have lower low-flows and higher high-flows, as well as earlier snowmelt \citep{Nijssen:2001}.
-These manifestations of rising temperatures are impacting species \citep{Xenopoulos:2006} and people \citep{Hirabayashi:2013}.
+As climate change alters precipitation patterns from snow to rain in temperate regions, rivers are generally predicted to have lower low-flows and higher high-flows, as well as earlier snowmelt \cite{Nijssen:2001}.
+These manifestations of rising temperatures are impacting species \cite{Xenopoulos:2006} and people \cite{Hirabayashi:2013}.
 Extracting annual and monthly metrics of flow from Fraser River basin gauge station records, we used rolling average maximum- and minimum-flow estimates to capture trends in extreme discharge and rolling average median-flow estimates to capture general annual and seasonal trends.
@@
 For a river network to act as a portfolio of climate and dampen the effects of climate change, expressions of climate and their trends over time need to vary throughout the network.
-Using the Western North America Climate tool (ClimateWNA), we quantified the diversity in climate trends within the Fraser River basin over the years concurrent with our hydrology trend estimates \citep{Wang:2016} (Fig.~\ref{fig:1}) (\textit{Materials and Methods}).
+Using the Western North America Climate tool (ClimateWNA) \cite{Wang:2016}, we quantified the diversity in climate trends within the Fraser River basin over the years concurrent with our hydrology trend estimates (Fig.~\ref{fig:1}) (\textit{Materials and Methods}).
 We then summarized the climate portfolio of each flow gauge by calculating the standard deviation in climate trends within their catchments.
@@
 The standard deviations of these climate variables were scaled between 0 and 1 and summed into a general climate index for each catchment.
-Following our hypothesis, this climate portfolio index increases with area (Fig.~\ref{fig:2}), rapidly asymptoting in catchments greater than 40,000 km\textsuperscript{2}.
+Following our hypothesis, this climate portfolio index increases with area (Fig.~\ref{fig:2}), rapidly asymptoting in catchments greater than 40,000~km\textsuperscript{2}.
 
-\begin{figure}%[tbhp]
+\begin{figure}[h]
 \centering
-\includegraphics[width=.8\linewidth]{Fig_ClimPort.pdf}
+\includegraphics[width=.8\linewidth]{Fig2_ClimPort.pdf}
 	\caption{Climate portfolio index as a function of catchment area.
@@
 Therefore, to test whether networks dampen climate signals we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model (\textit{Materials and Methods}).
-To quantify attenuation, we modeled the change in trend variability with the change in catchment area using an exponential variance function \citep{Pinheiro:2006}.
+To quantify attenuation, we modeled the change in trend variability with the change in catchment area using an exponential variance function \cite{Pinheiro:2000}.
 The variance function allowed us to predict the range of flow trend values that would be expected as watershed area increased.
@@
 
-In the observed data, small watersheds exhibited greater variability around their trend estimates than large watersheds, likely because of greater short-term variation in small catchments \ref{fig:3} \citep{Moore:2015}.
+In the observed data, small watersheds exhibited greater variability around their trend estimates than large watersheds, likely because of greater short-term variation in small catchments (Fig.~\ref{fig:3}) \cite{Moore:2015}.
 This relationship between watershed size and trend certainty may pull small watershed trend estimates away from zero, thereby creating the appearance of decreasing trend variability among sites as watershed size increases.
@@
 These simulations were generated by a random walk process with the observed standard deviation and autocorrelation parameters ($\hat{\upsigma}$, $\hat{\upphi}$) at our sites.
-Using the same generalized least squares model applied to the observed data, we estimated trends for 1000 simulated time series at each site for each response variable (e.g., Figure S3).
+Using the same generalized least squares model applied to the observed data, we estimated trends for 1000 simulated time series at each site for each response variable (e.g., Fig.~S3).
 This simulation process, a form of parametric bootstrapping, created distributions of null-expectations with which we compared our observed results.
 
-\begin{figure}[h]
+\begin{figure*}[h]
 \centering
-\includegraphics[width=17.8]{Fig_AnnFun.pdf}
+\includegraphics[width=17.8cm]{Fig3_AnnFun.pdf}
 	\caption{Annual flow trend attenuation within the Fraser River basin.
@@
 \label{fig:3}
-\end{figure}
+\end{figure*}
 
 \section*{Results}
-\subsection{Climate Portfolios and Flow Trend Dampening.} Long-term changes in hydrology were less variable in larger catchments, as predicted, demonstrating between 92 and 96\% dampening of flow trends moving from headwater catchments to the network's confluence.
+\subsection*{Climate Portfolios and Flow Trend Dampening} Long-term changes in hydrology were less variable in larger catchments, as predicted, demonstrating between 92 and 96\% dampening of flow trends moving from headwater catchments to the network's confluence.
 In other words, the largest catchments were approximately 10 times more stable in their climate response than the smallest catchments.
@@
 Similar attenuation was seen in flow timing with a 94\% reduction in the day-of-year to half-annual-flow trends such that small watersheds were reaching half their annual flow between 8 days later and 25 days earlier while large watersheds have only shifted 2 to 4 days earlier between 1970 and 2007.
+
 Our null model approach illustrates that support for a river network portfolio effect is greater than 90\% for all four annual flow trend metrics and as high as 98\% for median-annual-flow trends (Fig.~\ref{fig:3}).
@@
 
-\subsection{Seasonally Shifting Flow Trend Dampening}
-Regional flow-trends and attenuation strength varied by season.
-Analysis of monthly trends revealed that winter flows are getting higher and summer flows are getting lower over time, but that these trends have been more variable in smaller catchments (Fig.~\ref{fig:4}, \textit{SI Appendix}, Figs.
-S~\ref{fig:S1}, S~\ref{fig:S2}).
+\subsection*{Seasonally Shifting Flow Trend Dampening} Regional flow-trends and attenuation strength varied by season.
+Analysis of monthly trends revealed that winter flows are getting higher and summer flows are getting lower over time, but that these trends have been more variable in smaller catchments (Fig.~\ref{fig:4}, \textit{SI Appendix}, Figs.~S1, S2).
 For example, winter maximum flows in small catchments ranged from nearly no change to dramatic 47\% increases per decade with decreases only as large as 9\%.
@@
 Summer flows exhibited the opposite, with decreases in high flows over time.
-However, there was weak evidence of attenuation in the spring (Fig.~\ref{fig:4}, \textit{SI Appendix},Figs.
-S~\ref{fig:S1}, S~\ref{fig:S2}).
-Overall, climate portfolios tended to be greater during stable periods of the year and decrease during seasonal transitions (\textit{SI Appendix}, Fig.~\ref{S7}).
+However, there was weak evidence of attenuation in the spring (Fig.~\ref{fig:4}, \textit{SI Appendix},Figs.~S1,S2).
+Overall, climate portfolios tended to be greater during stable periods of the year and decrease during seasonal transitions (\textit{SI Appendix}, Fig.~S3).
 In the spring the likelihood of network-driven attenuation of maximum flow trends was as low as 25\% (Fig.~\ref{fig:4}).
@@
 
-\begin{figure}%[tbhp]
+\begin{figure*}[h]
 \centering
-\includegraphics[width=.8\linewidth]{Figure4.pdf}
+\includegraphics[width=17.8cm]{Fig4_MaxMonth.pdf}
 	\caption{Monthly maximum flow trend attenuation within the Fraser River basin.
@@
 \label{fig:4}
-\end{figure}
+\end{figure*}
 
 Despite considerable dampening relative to smaller sub-basins, the furthest downstream reaches of the Fraser River are still exhibiting small but detectable signals of climate change.
-Late fall, winter and early spring flows are generally increasing while late spring, summer and early fall flows are decreasing (Fig.~\ref{fig:4}, \textit{SI Appendix} Figs.
-S~\ref{fig:S1}, S~\ref{fig:S2}).
-These findings generally support previous studies demonstrating earlier spring runoff \citep{Dery:2012}, decreasing summer and fall flows \citep{Stahl:2006} and increased winter flows \citep{Healey:2011}.
+Late fall, winter and early spring flows are generally increasing while late spring, summer and early fall flows are decreasing (Fig.~\ref{fig:4}, \textit{SI Appendix} Figs.~S1,S2).
+These findings generally support previous studies demonstrating earlier spring runoff \cite{Dery:2012}, decreasing summer and fall flows \cite{Stahl:2006} and increased winter flows \cite{Healey:2011}.
 Together, this body of work indicates basin-wide effects of climate change, where shifts in winter precipitation from snow to rain is resulting in increased winter runoff and decreased snow pack contributions to summer flows.
@@
 \section*{Discussion}
-
-Climate change is already causing economic and conservation challenges for river systems worldwide\citep{Palmer:2009}.
+Climate change is already causing economic and conservation challenges for river systems worldwide\cite{Palmer:2009}.
 For example, sockeye salmon in the Fraser River have seen as much as 80\% pre-spawn mortality in years with later run-off and elevated temperatures, leading managers to close the fishery in 2013 and 2015.
@@
 
-There have been extensive efforts to predict the hydrologic sensitivity of rivers due to climate change through downscaled climate projections \citep{Nijssen:2001}.
+There have been extensive efforts to predict the hydrologic sensitivity of rivers due to climate change through downscaled climate projections \cite{Nijssen:2001}.
 Here we demonstrate that large rivers may have an inherent ability to absorb local climate change.
-But just as the destruction of coastal habitats degrades their natural capacity to mitigate sea-level rise or storm events \citep{Arkema:2013}, the management of river basins likely impacts the ability for river networks to attenuate climate trends.
-For example, dams synchronize the flow regimes of rivers \citep{Poff:2007} and logging increases the magnitude and frequency of extreme events such as high and low flows through increased runoff rates \citep{Zhang:2014}.
+But just as the destruction of coastal habitats degrades their natural capacity to mitigate sea-level rise or storm events \cite{Arkema:2013}, the management of river basins likely impacts the ability for river networks to attenuate climate trends.
+For example, dams synchronize the flow regimes of rivers \cite{Poff:2007} and logging increases the magnitude and frequency of extreme events such as high and low flows through increased runoff rates \cite{Zhang:2014}.
 These anthropogenic activities may magnify the impacts of climatic shifts such as the transition from snow- to rain-dominated precipitation in some basins.
@@
 
-\subsubsection*{SI Information}
-
-\subsubsection*{SI Figures}
-
-\begin{figure}%[tbhp]
-\centering
-\includegraphics[width=.8\linewidth]{FigureS1.pdf}
-	\caption{Monthly minimum flow trend attenuation within the Fraser River basin.
-(\textbf{Left}) Fraser River's basin-wide minimum-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulation.
-Observed values falling further from the center of the density distribution suggest greater evidence for changes in minimum flow and a greater shift in magnitude.
-(\textbf{Center}) Observed monthly Fraser River minimum-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates.
-Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed.
-(\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large), for four seasonally representative months.
-These reflect months in the prior columns and describe the variation in percent change per decade of minimum flow among sites.
-Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
-\label{fig:S1}
-\end{figure}
-
-\begin{figure}%[tbhp]
-\centering
-\includegraphics[width=.8\linewidth]{FigureS2.pdf}
-	\caption{Monthly median flow trend attenuation within the Fraser River basin.
-(\textbf{Left}) Fraser River's basin-wide median-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulation.
-Observed values falling further from the center of the density distribution suggest greater evidence for changes in median flow and a greater shift in magnitude.
-(\textbf{Center}) Observed monthly Fraser River median-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates.
-Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed.
-(\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large),for four seasonally representative months.
-These reflect months in the prior columns and describe the variation in percent change per decade of median flow among sites.
-Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
-\label{fig:S2}
-\end{figure}
-
 \matmethods{
-\subsection*{Materials and Methods}
-\subsection{Data.} We downloaded daily flow data from Environment Canada's HYDAT database.
-Within HYDAT we selected hydro-metric gauge stations using three criteria: (i) the station was within the Fraser River basin, (ii) the station's flow data were not observed to be dam influenced, and (iii) the station collected data in each month between 1970 and 2007 and no month was missing more than 5 days data.
-We selected this 38-year range because this period maximized the number of gauge stations in the Fraser River basin (*n*=55) operating concurrently over a long enough time period in which climate trends may be observable.
+\subsection*{Data} We downloaded daily flow data from Environment Canada's HYDAT database.
+Within HYDAT we selected Fraser River basin hydro-metric gauge stations that were not observed to be dam influenced and which collected data in each month between 1970 and 2007 where no month was missing more than 5 days data.
+We selected this 38-year range because this period maximized the number of gauge stations in the Fraser River basin (\textit{n}=55) operating concurrently over a long enough time period in which climate trends may be observable.
 We interpolated missing data using a linear model in log-space, but of the 13,879 days of data only 23 days were estimated.
@@
 This smoothing technique was used to reduce the influence of erroneous and unusually extreme values.
-Because dams control flow and stabilize flow trends, we removed sites where daily flow exhibited unusually high autocorrelation and low variability relative to other sites of similar size, and were observed downstream of, and atypically close to, a dam (*n*=3).
-The downstream distance of gauge stations from dams was determined using GIS spatial layers available in the *BC Data Catalog*, maintained by the provincial government of British Columbia.
+Because dams control flow and stabilize flow trends, we removed sites where daily flow exhibited unusually high autocorrelation and low variability relative to other sites of similar size, and were observed downstream of, and atypically close to, a dam (\textit{n}=3).
+The downstream distance of gauge stations from dams was determined using GIS spatial layers available in the \textit{BC Data Catalog}, maintained by the provincial government of British Columbia.
 Finally, we summarized the daily data into annual and monthly maximum, median and minimum flow estimates and the day-of-year to half-annual-flow.
 
-\subsection{River discharge trend analysis.} We estimated annual and monthly flow trends for each response variable at each gauge station using a generalized least squares (GLS) model with an AR1 autocorrelation function:
-\begin{linenomath*}
+\subsection*{River discharge trend analysis} We estimated annual and monthly flow trends for each response variable at each gauge station using a generalized least squares (GLS) model with an AR1 autocorrelation function:
 \begin{equation}
@@
 \end{equation}
-\end{linenomath*}
 where $\mathrm{flow}_{s,t}$ represents a flow metric [log(maximum), log(minimum), log(median), logit(day-of-year to half-annual-flow)] at each site ($s$) and time point ($t$),  and $\mathrm{year}_{s,t}$ represents the time in years (e.g., 1970, 1971, \ldots, 2007) with 1988 subtracted to approximately center the predictor.
@@
 
-\subsection{Basin trend attenuation analysis.}
-To measure network driven dampening of climate signalling, we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model:
-\begin{linenomath*}
+\subsection*{Basin trend attenuation analysis} To measure network driven dampening of climate signalling, we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model:
 \begin{equation}
@@
 \end{equation}
-\end{linenomath*}
 where $\hat{b}_{s}$ represents a flow trend at a given site ($s$) and $\mathrm{area}_{s}$ represents site $s$'s watershed area.
 Fitted $d$ and $c$ parameters represent the mean effect of $\sqrt{\mathrm{watershed area}}$ on flow trends and the mean basin-wide flow trend at a theoretical watershed area of zero, respectively.
-To quantify attenuation, we modeled the change in trend variability with the change in catchment area using an exponential variance function \citep[p.~211]{Pinheiro:2006}:
-\begin{linenomath*}
+To quantify attenuation, we modeled the change in trend variability with the change in catchment area using an exponential variance function \cite[p.~211]{Pinheiro:2000}:
 \begin{equation}
-	f(\mathrm{area}_{s})=\sigma_b^2 \exp(2\updelta\sqrt{\mathrm{area}_{s}}) \label{eq3},
+	f(\mathrm{area}_{s}) = \sigma_b^2 \exp(2\updelta\sqrt{\mathrm{area}_{s}}) \label{eq3},
 \end{equation}
-\end{linenomath*} 
 where the variance of the estimated error ($\eta$) changes exponentially with increasing $\sqrt{\mathrm{area}}$.
@@
 
-\subsection{River Attnuation Null Model Simulations}
-In the observed data, small watersheds exhibited greater variability around their trend estimates than large watersheds, likely because of greater short-term variation in small catchments \citep{Moore:2015}.
+\subsection*{River Attnuation Null Model Simulations} In the observed data, small watersheds exhibited greater variability around their trend estimates than large watersheds, likely because of greater short-term variation in small catchments \cite{Moore:2015}.
 This relationship between watershed size and trend certainty may pull small watershed trend estimates away from zero, thereby creating the appearance of decreasing trend variability among sites as watershed size increases.
@@
 These simulations were generated by a random walk process with the observed standard deviation and autocorrelation parameters ($\hat{\upsigma}$, $\hat{\upphi}$) at our sites.
-Using the same GLS model and AR1 correlation structure as applied to the observed data (eq.~\ref{eq1}), we estimated trends for 1000 simulated time series at each site for each response variable (e.g., Figure S3).
+Using the same GLS model and AR1 correlation structure as applied to the observed data (eq.
+\ref{eq1}), we estimated trends for 1000 simulated time series at each site for each response variable (e.g., Fig.~S4).
 This simulation process, a form of parametric bootstrapping, created distributions of null-expectations with which we compared our observed results.
 
-We then applied equation~\ref{eq2} to each of the 1000 basin-wide simulations, resulting in 1 observed variance exponent parameter ($\hat{\updelta}$) and 1000 simulated $\hat{\updelta}$ for each flow metric.
-By comparing our observed attenuation with our basin-wide null-model simulations we addressed the potential that more variable flows in smaller catchments contribute to the observed pattern of flow trends as a function of watershed area (Figure~\ref{figthree}).
+We then applied equation \ref{eq2} to each of the 1000 basin-wide simulations (e.g.
+Fig.~S5), resulting in 1 observed variance exponent parameter ($\hat{\updelta}$) and 1000 simulated $\hat{\updelta}$ for each flow metric.
+By comparing our observed attenuation with our basin-wide null-model simulations we addressed the potential that more variable flows in smaller catchments contribute to the observed pattern of flow trends as a function of watershed area (Fig.~\ref{fig:3}).
 This null-model approach asks how likely our observation is due to a sampling effect versus a network portfolio effect.
 
-\subsection{Dampening summary statistics.} We calculated attenuation certainty as the percentage of simulated $\hat{\updelta}$ that were less than the observed $\hat{\updelta}$.
+\subsection*{Dampening summary statistics} We calculated attenuation certainty as the percentage of simulated $\hat{\updelta}$ that were less than the observed $\hat{\updelta}$.
 We calculated attenuation strength as the ratio between the standard deviation at the smallest and largest watersheds as defined by the exponential variance function (i.e., $\sqrt{f(\mathrm{area}_1)} / \sqrt{f(\mathrm{area}_2)}$).
@@
 \subsection*{Climate} In order to develop a climate index for each flow-gauge catchment we needed to estimate climate trends across the Fraser River basin and then spatially summarise them by flow-gauge catchment.
-Using Whitebox Geospatial Analysis Tools, a freely available and open source geospatial analysis software \citep{}, we delineated the Fraser River basin and the catchment areas of each flow-gauge station.
-Digital elevation models provided by the provincial government of British Columbia and state goverment of Washinton were pre-processed with a breaching algorithm and stream burned to facilitate proper flow path and accumulation models.
+Using Whitebox Geospatial Analysis Tools, a freely available and open source geospatial analysis software \cite{Lindsay:2016}, we delineated the Fraser River basin and the catchment areas of each flow-gauge station.
+Digital elevation models provided by the provincial government of British Columbia, CA and state goverment of Washinton, USA were pre-processed with a breaching algorithm and stream burned to facilitate proper flow path and accumulation models.
 We then used Climate WNA to estimate historic climate values on an evenly spaced 1 km\textsuperscript{2} grid across the Fraser Basin for each year and month in our study.
 Climate variables included the mean annual temperature (MAT), and precipitation (MAP), extreme minimum (EMT) and maximum (EXT) temperature and precipitation as snow (PAS).
-Climate variable trends were calculated at each grid point using \ref{eq1}.
+Climate variable trends were calculated at each grid point using a modified eq.~\ref{eq1}, where time points were no longer considered and our response variable becomes $\mathrm{climate}_{s}$.
 
-The standard deviation in climate trends represent the strength of each basins' climate portfolio.
-Climate trend standard deviation estimates were calculated using the `zonal\_stats` function in the \textit{rasterstats} python module.
-To demonstrate how this climate index scales with watershed area we regressed climate portoflio index values onto watershed area (log(km\texsuperscript{2})).
-Two outlier catchments had high leverage climate portfolio index values.
-Rather than remove these points, we used robust linear regression to downweight their influence using Huber's method of M-estimation.
+\subsection*{Code and Data Repository} All flow data munging and analysis was done using R v3.3.2 (\url{www.r-project.org/}).
+Spatial analysis of watersheds were done using GDAL/OGR v1.11.5, and Whitebox GAT v3.4 (\url{www.uoguelph.ca/~hydrogeo/Whitebox/}) with visualization help from QGIS v2.18.2 (\url{www.qgis.org/}).
+Climate data were gathered using ClimateWNA v5.4 (\url{www.climatewna.com}) and summary statitics were calculated using R and the \textit{rasterstats} module (\url{https://github.com/perrygeo/python-rasterstats.git}) in Python v2.7.
+All code can be found on \url{www.github.com/} at \url{github.com/kchezik/River-Network-Flow-Trends.git}.
+Raw HYDAT data can be obtained via Environment Canada (\url{http://collaboration.cmc.ec.gc.ca/cmc/hydrometrics/www/}) or a simplified version can be found on GitHub.
+}
 
-\showmatmethods % Display the Materials and Methods section
+\showmatmethods{} % Display the Materials and Methods section
 
@@
 Anderson was supported by the David H.
-Smith Conservation Research Fellowship.
-\showacknow % Display the acknowledgments section
+Smith Conservation Research Fellowship.}
+
+\showacknow{} % Display the acknowledgments section
 
@@
 % Bibliography
-\bibliography{pnas-sample}
+\bibliography{ms}
 
-\bibitem[{\textit{Arkema et~al.}(2013)\textit{Arkema, Guannel, Verutes, Wood,
-  Guerry, Ruckelshaus, Kareiva, Lacayo, and Silver}}]{Arkema:2013}
-Arkema, K.~K., G.~Guannel, G.~Verutes, S.~A. Wood, A.~Guerry, M.~Ruckelshaus,
-  P.~Kareiva, M.~Lacayo, and J.~M. Silver (2013), Coastal habitats shield
-  people and property from sea-level rise and storms, \textit{Nature Climate
-  Change}, \textit{3}(10), 913--918.
-
-\bibitem[{\textit{Bindoff et~al.}(2013)\textit{Bindoff, Stott, AchutaRao,
-  Gillett, Gutzler, Hansingo, Hegerl, Hu, Jain, Mokhov, Overland, Perlwitz,
-  Sebbari, and Zhang}}]{Bindoff:2013}
-Bindoff, N.~L., P.~A. Stott, M.~R. AchutaRao, N.~P. Gillett, D.~Gutzler,
-  K.~Hansingo, G.~C. Hegerl, Y.~Hu, S.~Jain, I.~I. Mokhov, J.~Overland,
-  J.~Perlwitz, R.~Sebbari, and X.~Zhang (2013), Detection and attribution of
-  climate change: from global to regional, \textit{Tech. rep.},
-  Intergovernmental Panel on Climate Change.
-
-\bibitem[{\textit{D{\'e}ry et~al.}(2012)\textit{D{\'e}ry,
-  Hern{\'a}ndez-Henr{\'\i}quez, Owens, Parkes, and Petticrew}}]{Dery:2012}
-D{\'e}ry, S.~J., M.~A. Hern{\'a}ndez-Henr{\'\i}quez, P.~N. Owens, M.~W. Parkes,
-  and E.~L. Petticrew (2012), A century of hydrological variability and trends
-  in the {F}raser {R}iver basin, \textit{Environmental Research Letters},
-  \textit{7}(2), 024,019.
-
-\bibitem[{\textit{Doak et~al.}(1998)\textit{Doak, Bigger, Harding, Marvier,
-  O'malley, and Thomson}}]{Doak:1998}
-Doak, D.~F., D.~Bigger, E.~Harding, M.~Marvier, R.~O'malley, and D.~Thomson
-  (1998), The statistical inevitability of stability-diversity relationships in
-  community ecology, \textit{The American Naturalist}, \textit{151}(3),
-  264--276 %@ 0003--0147.
-
-\bibitem[{\textit{Donat et~al.}(2016)\textit{Donat, Lowry, Alexander, O'Gorman,
-  and Maher}}]{Donat:2016}
-Donat, M.~G., A.~L. Lowry, L.~V. Alexander, P.~A. O'Gorman, and N.~Maher
-  (2016), More extreme precipitation in the world's dry and wet regions,
-  \textit{Nature Climate Change}, \textit{6}(5), 508--513,
-  \doi{10.1038/nclimate2941}.
-
-\bibitem[{\textit{Gelman and Hill}(2008)}]{Gelman:2008}
-Gelman, A., and J.~Hill (2008), Data analysis using regression and multilevel
-  hierarchical models, \doi{10.1080/02664760801949043}.
-
-\bibitem[{\textit{George et~al.}(2015)\textit{George, Thompson~III, and
-  Faaborg}}]{George:2015}
-George, A.~D., F.~R. Thompson~III, and J.~Faaborg (2015), Using {L}i{DAR} and
-  remote microclimate loggers to downscale near-surface air temperatures for
-  site-level studies, \textit{Remote Sensing Letters}, \textit{6}(12),
-  924--932.
-
-\bibitem[{\textit{Guerry et~al.}(2015)\textit{Guerry, Polasky, Lubchenco,
-  Chaplin-Kramer, Daily, Griffin, Ruckelshaus, Bateman, Duraiappah, Elmqvist,
-  Feldman, Folke, Hoekstra, Kareiva, Keeler, Li, McKenzie, Ouyang, Reyers,
-  Ricketts, Rockstr{\"o}m, Tallis, and Vira}}]{Guerry:2015}
-Guerry, A.~D., S.~Polasky, J.~Lubchenco, R.~Chaplin-Kramer, G.~C. Daily,
-  R.~Griffin, M.~Ruckelshaus, I.~J. Bateman, A.~Duraiappah, T.~Elmqvist, M.~W.
-  Feldman, C.~Folke, J.~Hoekstra, P.~M. Kareiva, B.~L. Keeler, S.~Li,
-  E.~McKenzie, Z.~Ouyang, B.~Reyers, T.~H. Ricketts, J.~Rockstr{\"o}m,
-  H.~Tallis, and B.~Vira (2015), Natural capital and ecosystem services
-  informing decisions: From promise to practice, \textit{Proceedings of the
-  National Academy of Sciences}, \textit{112}(24), 7348--7355,
-  \doi{10.1073/pnas.1503751112}.
-
-\bibitem[{\textit{Hartmann et~al.}(2013)\textit{Hartmann, Klein, Rusticucci,
-  Alexander, Bronnimann, Charabi, Dentener, Dlugokencky, Easterling, Kaplan,
-  Soden, Thorne, Wild, and Zhai}}]{Hartmann:2013}
-Hartmann, D.~L., A.~M.~G. Klein, M.~Rusticucci, L.~V. Alexander, S.~Bronnimann,
-  Y.~Charabi, F.~J. Dentener, D.~R. Dlugokencky, A.~Easterling, B.~J. Kaplan,
-  B.~J. Soden, P.~W. Thorne, M.~Wild, and P.~M. Zhai (2013), Climate change
-  2013: The physical science basis., \textit{Tech. rep.}, Intergovernmental
-  Panel on Climate Change.
-
-\bibitem[{\textit{Healey and Bradford}(2011)}]{Healey:2011}
-Healey, M., and M.~Bradford (2011), The cumulative impacts of climate change on
-  {F}raser {R}iver sockeye salmon (\textit{{O}ncorhynchus nerka}) and
-  implications for management, \textit{Can. J. Fish. Aquat. Sci.},
-  \textit{68}(4), 718--737.
-
-\bibitem[{\textit{Hirabayashi et~al.}(2013)\textit{Hirabayashi, Mahendran,
-  Koirala, Konoshima, Yamazaki, Watanabe, Kim, and Kanae}}]{Hirabayashi:2013}
-Hirabayashi, Y., R.~Mahendran, S.~Koirala, L.~Konoshima, D.~Yamazaki,
-  S.~Watanabe, H.~Kim, and S.~Kanae (2013), Global flood risk under climate
-  change, \textit{Nature Climate Change}, \textit{3}, 816--821.
-
-\bibitem[{\textit{Hulme}(2016)}]{Hulme:2016}
-Hulme, M. (2016), 1.5$\degree${C} and climate research after the {P}aris
-  {A}greement, \textit{Nature Climate Change}, \textit{6}(3), 222--224.
-
-\bibitem[{\textit{Jones et~al.}(2012)\textit{Jones, Hole, and
-  Zavaleta}}]{Jones:2012}
-Jones, H.~P., D.~G. Hole, and E.~S. Zavaleta (2012), Harnessing nature to help
-  people adapt to climate change, \textit{Nature Climate Change},
-  \textit{2}(7), 504--509.
-
-\bibitem[{\textit{McCabe and Wolock}(2014)}]{McCabe:2014}
-McCabe, G.~J., and D.~M. Wolock (2014), Spatial and temporal patterns in
-  conterminous {U}nited {S}tates streamflow characteristics,
-  \textit{Geophysical Research Letters}, \textit{41}(19), 6889--6897.
-
-\bibitem[{\textit{Melillo et~al.}(2014)\textit{Melillo, Richmond, and
-  Yohe}}]{Melillo:2014}
-Melillo, J.~M., T.~Richmond, and G.~W. Yohe (2014), Climate change impacts in
-  the {U}nited {S}tates: the third national climate assessment, \textit{US
-  Global change research program}, \textit{841}.
-
-\bibitem[{\textit{Moore et~al.}(2015)\textit{Moore, Beakes, Nesbitt, Yeakel,
-  Patterson, Thompson, Phillis, Braun, Favaro, Scott et~al.}}]{Moore:2015}
-Moore, J.~W., M.~P. Beakes, H.~K. Nesbitt, J.~D. Yeakel, D.~A. Patterson, L.~A.
-  Thompson, C.~Phillis, D.~C. Braun, C.~Favaro, D.~C. Scott, et~al. (2014),
-  Emergent stability in a large free-flowing watershed, \textit{Ecology}, 
-  \textit{96}(2), 340--347.
-
-\bibitem[{\textit{Morrison et~al.}(2002)\textit{Morrison, Quick, and
-  Foreman}}]{Morrison:2002}
-Morrison, J., M.~C. Quick, and M.~G. Foreman (2002), Climate change in the
-  {F}raser {R}iver watershed: flow and temperature projections, \textit{Journal
-  of Hydrology}, \textit{263}(1), 230--244.
-
-\bibitem[{\textit{Narain et~al.}(2011)\textit{Narain, Margulis, and
-  Essam}}]{Narain:2011}
-Narain, U., S.~Margulis, and T.~Essam (2011), Estimating costs of adaptation to
-  climate change, \textit{Climate Policy}, \textit{11}(3), 1001--1019.
-
-\bibitem[{\textit{Nijssen et~al.}(2001)\textit{Nijssen, O'Donnell, Hamlet, and
-  Lettenmaier}}]{Nijssen:2001}
-Nijssen, B., G.~M. O'Donnell, A.~F. Hamlet, and D.~P. Lettenmaier (2001),
-  Hydrologic sensitivity of global rivers to climate change, \textit{Climatic
-  Change}, \textit{50}(1-2), 143--175.
-
-\bibitem[{\textit{Palmer et~al.}(2009)\textit{Palmer, Lettenmaier, Poff,
-  Postel, Richter, and Warner}}]{Palmer:2009}
-Palmer, M.~A., D.~P. Lettenmaier, N.~L. Poff, S.~L. Postel, B.~Richter, and
-  R.~Warner (2009), Climate change and river ecosystems: protection and
-  adaptation options, \textit{Environmental Management}, \textit{44}(6),
-  1053--1068.
-
-\bibitem[{\textit{Peterson et~al.}(2013)\textit{Peterson, Ver~Hoef, Isaak,
-  Falke, Fortin, Jordan, McNyset, Monestiez, Ruesch, Sengupta
-  et~al.}}]{Peterson:2013}
-Peterson, E.~E., J.~M. Ver~Hoef, D.~J. Isaak, J.~A. Falke, M.-J. Fortin, C.~E.
-  Jordan, K.~McNyset, P.~Monestiez, A.~S. Ruesch, A.~Sengupta, et~al. (2013),
-  Modelling dendritic ecological networks in space: an integrated network
-  perspective, \textit{Ecology Letters}, \textit{16}(5), 707--719.
-
-\bibitem[{\textit{Pinheiro and Bates}(2006)}]{Pinheiro:2006}
-Pinheiro, J., and D.~Bates (2006), Mixed-effects models in {S} and {S-PLUS},
-  \textit{Springer Science \& Business Media}.
-
-\bibitem[{\textit{Poff et~al.}(2007)\textit{Poff, Olden, Merritt, and
-  Pepin}}]{Poff:2007}
-Poff, L.~N., J.~D. Olden, D.~M. Merritt, and D.~M. Pepin (2007), Homogenization
-  of regional river dynamics by dams and global biodiversity implications,
-  \textit{Proceedings of the National Academy of Sciences}, \textit{104}(14),
-  5732--5737, \doi{10.1073/pnas.0609812104}.
-
-\bibitem[{\textit{Rauscher et~al.}(2008)\textit{Rauscher, Pal, Diffenbaugh, and
-  Benedetti}}]{Rauscher:2008}
-Rauscher, S.~A., J.~S. Pal, N.~S. Diffenbaugh, and M.~M. Benedetti (2008),
-  Future changes in snowmelt-driven runoff timing over the western {US},
-  \textit{Geophysical Research Letters}, \textit{35}(16).
-
-\bibitem[{\textit{\relax{Louisiana Coastal Wetlands Conservation and
-  Restoration Task Force and the Wetlands Conservation and Restoration
-  Authority}}(1998)}]{Conservation:1998}
-\relax{Louisiana Coastal Wetlands Conservation and Restoration Task Force and
-  the Wetlands Conservation and Restoration Authority} (1998), Coast 2050:
-  {T}oward a sustainable coastal louisiana, \textit{Tech. rep.}, Louisiana
-  Department of Natural Resources.
-
-\bibitem[{\textit{Stahl and Moore}(2006)}]{Stahl:2006}
-Stahl, K., and R.~Moore (2006), Influence of watershed glacier coverage on
-  summer streamflow in {B}ritish {C}olumbia, {C}anada, \textit{Water Resources
-  Research}, \textit{42}(6).
-
-\bibitem[{\textit{V{\"o}r{\"o}smarty et~al.}(2010)\textit{V{\"o}r{\"o}smarty,
-  McIntyre, Gessner, Dudgeon, Prusevich, Green, Glidden, Bunn, Sullivan,
-  Liermann et~al.}}]{Vorosmarty:2010}
-V{\"o}r{\"o}smarty, C.~J., P.~McIntyre, M.~O. Gessner, D.~Dudgeon,
-  A.~Prusevich, P.~Green, S.~Glidden, S.~E. Bunn, C.~A. Sullivan, C.~R.
-  Liermann, et~al. (2010), Global threats to human water security and river
-  biodiversity, \textit{Nature}, \textit{467}(7315), 555--561.
-
-\bibitem[{\textit{Xenopoulos and Lodge}(2006)}]{Xenopoulos:2006}
-Xenopoulos, M.~A., and D.~M. Lodge (2006), Going with the flow: using
-  species-discharge relationships to forecast losses in fish biodiversity,
-  \textit{Ecology}, \textit{87}(8), 1907--1914.
-
-\bibitem[{\textit{Yeakel et~al.}(2014)\textit{Yeakel, Moore, Guimaraes, and
-  Aguiar}}]{Yeakel:2014}
-Yeakel, J. D., J. W.~Moore, P. R.~Guimaraes, and M. A. M. de~Aguiar (2014), Synchronisation and
-  stability in river metapopulation networks, \textit{Ecology Letters},
-  \textit{17}(3), 273--283.
-
-\bibitem[{\textit{Zhang and Wei}(2014)}]{Zhang:2014}
-Zhang, M., and X.~Wei (2014), Alteration of flow regimes caused by large-scale
-  forest disturbance: a case study from a large watershed in the interior of
-  {B}ritish {C}olumbia, {C}anada, \textit{Ecohydrology}, \textit{7}(2),
-  544--556, \doi{10.1002/eco.1374}.
-
-\bibitem{Wang:2016}
-T.~Wang, A.~Hamann, D.~Spittlehouse, and C.~Carroll.
-\newblock Locally downscaled and spatially customizable climate data for
-  historical and future periods for north america.
-\newblock {\em PLoS ONE}, 11(6):e0156720, 2016.
+\end{document}

### 2017-03-17 drafts/PNAS/Chezik_PNAS_SI.tex
Updated PNAS paper. Merged old drafts.

--- 2017-01-11 AGU_Journal/ms_Geo_Let.tex
@@
-\documentclass[linenumbers,draft]{AGUJournal}
-\usepackage{gensymb}
-\usepackage{textcomp}
+\documentclass[9pt]{pnas-new}
+
 \usepackage{upgreek}
-\usepackage{natbib}
-\usepackage[parfill]{parskip}
-\journalname{Geophysical Research Letters}
+
+\templatetype{pnasresearcharticle}
+
+\title{River networks dampen long-term hydrological signals of climate change}
+
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a,1]{Kyle A. Chezik}
+\author[b,1]{Sean C. Anderson} 
+\author[a,1]{Jonathan W. Moore}
+
+\affil[a]{Earth to Ocean Research Group, Department of Biological Sciences, Simon Fraser University, 8888 University Dr., Burnaby, British Columbia V5A 1S6, Canada}
+\affil[b]{School of Aquatic and Fishery Sciencies, University of Washington, Box 455020, Seattle, WA 98195, USA}
+
+% Please give the surname of the lead author for the running footer
+\leadauthor{Chezik}
+
 \begin{document}
-\title{River networks dampen long-term hydrological signals of climate change}
-\authors{K. A. Chezik\affil{1},
-	S. C. Anderson\affil{2},
-J. W. Moore\affil{1}}
 
-\affiliation{1}{Earth to Ocean Research Group, Department of Biological Sciences, Simon Fraser University, 8888 University Dr., Burnaby, British Columbia V5A 1S6, Canada}
-\affiliation{2}{School of Aquatic and Fishery Sciences, University of Washington, Box 455020, Seattle, WA 98195, USA}
+\maketitle
 
-\correspondingauthor{K. A. Chezik}{kchezik@sfu.ca}
+\section*{Supplementary Information}
+The following supplementary material includes 5 additional figures.
+Figures \ref{fig:S1} and \ref{fig:S2} are similar to figure 4 in the main text where monthly trends in flow are shown for the basin along with trend attenuation results.
+In these supplementary figures we highlight minimum and median flow trends rather than maximum flow trends.
+Figure \ref{fig:S3} demonstrates the effect of the seasonal transition of spring and fall on the climate portfolio index values within and among Fraser River sub-basins.
+Figures \ref{fig:S4} and \ref{fig:S5} are example plots demonstrating a single null-model iteration for the median flow metric.
 
-\begin{keypoints}
-\item River networks act as diversified portfolios of climate; by integrating varied climate signals the network substantially dampens downstream climate-driven hydrologic trends.
-\item River network climate dampening varied seasonally in the Fraser River basin, declining sharply with the spring freshet, suggesting that basin-wide events homogenize the climate portfolio, thereby degrading the network dampening effect.
-\item Rivers that integrate a diverse climate portfolio offer climate change relief to riverine biota that is many times greater than homogenized systems.
-Leveraging network dampening properties by promoting land use practices that maintain or enhance climate portfolio diversity may be a powerful tool in mitigating climate change impacts.
-\end{keypoints}
+\subsection*{Supplementary Information Figures}
 
-\begin{abstract}
-River networks may dampen local hydrologic climate change by aggregating a diverse climate portfolio with variable upstream climate signals.
-Here we examine this hypothesis using flow trend estimates (1970--2007) at 55 hydrometric gauge stations within the Fraser River watershed in British Columbia, Canada.
-The Fraser River dampened variability in downstream long-term median discharge by 95\% from between -6 and 18\% per-decade among small watersheds to between 0.8 and 1.6\% per-decade near the confluence.
-A null model approach indicated that dampening of annual median flow trends within the Fraser River was typically 5.4 times greater than would be expected of a watershed exhibiting a consistent basin-wide climate response.
-Although climate dampening strength varied seasonally, our findings indicate that large free-flowing rivers offer a powerful and largely unappreciated process of climate change mitigation.
-River networks that integrate a diverse climate portfolio can dampen local extremes and offer climate relief to riverine biota.
-\end{abstract}
+\begin{figure*}[b]
+\centering
+\includegraphics[width=17.8cm]{FigS1_MinMonth.pdf}
+	\caption{Monthly minimum flow trend attenuation within the Fraser River basin.
+(\textbf{Left}) Fraser River's basin-wide minimum-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulations.
+Observed values falling further from the center of the density distribution suggest greater evidence for changes in minimum flow and a greater shift in magnitude.
+(\textbf{Center}) Observed monthly Fraser River minimum-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates.
+Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed.
+(\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large), for four seasonally representative months.
+These reflect months in the prior columns and describe the variation in percent change per decade of minimum flow among sites.
+Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
+\label{fig:S1}
+\end{figure*}
 
-\section{Introduction} 
-As billions are spent adapting to climate change by such means as building dikes or improving stormwater drainage \citep{Narain:2011}, certain habitats and processes provide a natural defense system that mitigates climate change impacts \citep{Jones:2012}.
-For example, coastal habitats such as oyster reefs, mangrove forests, and eelgrass beds shield 67\% of the United States coastline and 1.4 million people from sea level rise and storms \citep{Arkema:2013}.
-Protection and enhancement of these natural systems has become a key component of many climate mitigation strategies \citep{Guerry:2015}, as they return multiple cost-effective services in contrast to engineered alternatives \citep{Jones:2012}.
-For instance, revitalization of the Mississippi delta not only improves storm protection, but also increases production and sustainability of fisheries and wildlife resources, protects the water supply, and reduces the impact of wastewater effluent \citep{Conservation:1998}.
-The value of climate buffering habitats stands only to increase as average global temperatures climb.
-Recent international agreements in Paris aim to limit global temperature rise to 1.5\degree{C} \citep{Hulme:2016}, a goal that concedes a minimum doubling of observed warming \citep{Hartmann:2013}; therefore, an urgent challenge and opportunity is discovering systems that naturally buffer against oncoming impacts of climate change.
+\begin{figure*}
+\centering
+\includegraphics[width=17.8cm]{FigS2_MedMonth.pdf}
+	\caption{Monthly median flow trend attenuation within the Fraser River basin.
+(\textbf{Left}) Fraser River's basin-wide median-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulations.
+Observed values falling further from the center of the density distribution suggest greater evidence for changes in median flow and a greater shift in magnitude.
+(\textbf{Center}) Observed monthly Fraser River median-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates.
+Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed.
+(\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large),for four seasonally representative months.
+These reflect months in the prior columns and describe the variation in percent change per decade of median flow among sites.
+Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
+\label{fig:S2}
+\end{figure*}
 
-River systems may offer an unappreciated defense against some aspects of climate change.
-Climate change is altering global precipitation patterns \citep{Donat:2016} and subsequently river flow regimes \citep{Hartmann:2013,Palmer:2009}.
-Earlier snowmelt \citep{Rauscher:2008}, reduced snow pack \citep{McCabe:2014}, shifts from snow to rain, and changes in the annual distribution of precipitation are all impacting stream flow \citep{Hartmann:2013}, resulting in flashier flows with increased frequency of flooding \citep{Hirabayashi:2013}, longer periods of drought and critically low flows for wildlife and agriculture \citep{Melillo:2014}.
-However, we hypothesize that large free-flowing river networks may naturally dampen these signals of climate change by integrating across different manifestations of climate.
-Processes that aggregate across asynchronous components tend to dampen the aggregate, a process known as the portfolio effect \citep{Doak:1998}.
-Climate is expressed differently and asynchronously over the landscape \citep{George:2015} and on account of their branching architecture and directional flow, river networks directionally integrate these varied climate manifestations \citep{Peterson:2013}, potentially dampening the local climatic response of sub-catchments.
-Large free-flowing rivers may have diverse climate portfolios, with sub-catchments acting as assets in their portfolio, thereby dampening local climate impacts.
-Portfolio effects in rivers are already known to dampen short-term variation \citep{Moore:2015,Yeakel:2014} but attenuation of long-term climate driven shifts have yet to be examined.
+\begin{figure*}
+\centering
+\includegraphics[width=11.4cm]{FigS3_SeasonClimPort.pdf}
+	\caption{Climate portfolio index by month at 55 sites within  the Fraser River basin.
+Lines are loess smoothers with a smoothing span of 1 to reduce noise and highlight general season shifts in climate portfolio among and across Fraser River sub-basins.}
+	\label{fig:S3}
+\end{figure*}
 
-Here we examine whether rivers attenuate local long-term climate change through analyses of hydrometric data from one of the largest free-flowing rivers in North America.
-Located in British Columbia, Canada, the Fraser River drains an area approximately the size of the United Kingdom (\texttildelow 217,000 km\textsuperscript{2}) and in a rare combination is fairly well monitored while also having no dams on its main stem \citep{Vorosmarty:2010} (Figure~\ref{figone}).
-This watershed drains interior high plateau, coastal mountains, and the Canadian Rockies, and thus integrates a diverse mosaic of landscapes, weather, and climate.
-As with other mid-latitude rivers \citep{Bindoff:2013}, the Fraser River has expressed increasing trends in discharge volume and variability over the last several decades \citep{Dery:2012,Morrison:2002}.
-The river's discharge is monitored by hydrometric gauge stations throughout the watershed.
-We analyzed 38 years of data collected concurrently at 55 of these stations and applied a novel analytical approach to examine whether this large river exhibits signals of long-term climate dampening.
+\begin{figure*}
+	\includegraphics[width=17.8cm]{FigS4_SiteSim.pdf}
+	\caption{Example site specific flow simulations.
+Annual median-flow simulations (yellow) and observations (blue) for all 55 sites considered in this study found within the Fraser River basin.
+Simulations at each site were parameterized using the observed data but no trend was imposed.
+Fifty-five site simulations equate to one basin-wide simulation.
+Each flow metric was simulated 1000 times basin-wide per response variable.
+See supplemental figure \ref{fig:S5} for resultant basin-wide simulation.}
+	\label{fig:S4}
+\end{figure*}
 
-\begin{figure}[h]
-	\centerline{\includegraphics[height=4in]{Fig1_Map2.pdf}}
-	\caption{Flow gauge sites within the Fraser River basin (dark grey) in British Columbia Canada (light grey).
-Flow gauge sites (yellow dots) are scaled by the size of the contributing area.
-The map is projected in Albers BC UTMs to provide an equal area depiction of the region but labels are expressed in WGS84 latitude and longitude.}
-	\label{figone}
-\end{figure}
-
-\section{Data and Methods}
-\subsection{Discharge Data and Flow Metrics}
-To assess signals of climate change in hydrology within the Fraser River basin we downloaded daily flow data from Environment Canada's HYDAT database.
-Within HYDAT we selected hydrometric gauge stations using three criteria: (i) the station was within the Fraser River basin, (ii) the station's flow data were not observed to be dam influenced, and (iii) the station collected data in each month between 1970 and 2007 where no month was missing more than 5 days of data and at least 35 complete years were present.
-We selected this 38-year range because this period maximized the number of gauge stations in the Fraser River basin (\textit{n}=55) operating concurrently over a long enough time period in which climate trends may be observable.
-We interpolated missing data using a linear model in log-space, but of the 763,345 days of data only 23 days were interpolated.
-Because dams control flow and stabilize flow trends, we removed sites where daily flow exhibited unusually high autocorrelation and low variability relative to other sites of similar size, and were observed downstream of, and atypically close to, a dam (\textit{n}=3).
-The downstream distance of gauge stations from dams was determined using GIS spatial layers available in the \textit{BC Data Catalog}, maintained by the provincial government of British Columbia.
-
-As climate change alters precipitation patterns from snow to rain in temperate regions, rivers are generally predicted to have lower low-flows and higher high-flows, as well as earlier snowmelt \citep{Nijssen:2001}.
-These manifestations of rising temperatures are impacting species \citep{Xenopoulos:2006} and people \citep{Hirabayashi:2013}.
-We summarized our daily discharge data into annual and monthly flow metrics that capture these transitions.
-We used maximum- and minimum-flow estimates to capture trends in extreme discharge and median-flow estimates to capture general annual and seasonal trends.
-Because the timing of flow is particularly important to the phenology of many organisms, we also calculated the day-of-year in which half the annual flow was reached.
-
-\subsection{River Discharge Trend Analysis}
-For a river network to act as a portfolio of climate and dampen the effects of climate change, expressions of climate and their trends over time need to vary throughout the network (Figure~\ref{figtwo}).
-We estimated annual and monthly flow trends for each response variable at each gauge station using a generalized least squares (GLS) model with an AR1 autocorrelation function:
-\begin{linenomath*}
-\begin{equation}
-  \mathrm{flow}_{s,t} = a_s + b_s \mathrm{year}_{s,t} + \epsilon_{s,t}, \quad 
-  \epsilon_{s,t} \sim \mathcal{N}(\upphi \epsilon_{s,t-1}, \sigma_\mathrm{flow}^2) \label{eq1},
-\end{equation}
-\end{linenomath*}
-where $\mathrm{flow}_{s,t}$ represents a flow metric [log(maximum), log(minimum), log(median), logit(day-of-year to half-annual-flow)] at each site ($s$) and time point ($t$),  and $\mathrm{year}_{s,t}$ represents the time in years (e.g., 1970, 1971, \ldots, 2007) with 1988 subtracted to approximately center the predictor.
-Parameters $b_{s}$ (slope) and $a_{s}$ (intercept) represent the estimated mean effect of time on $\mathrm{flow}_{s,t}$ and the estimate of $\mathrm{flow}_{s,t}$ at time zero (i.e., 1988) respectively.
-Error ($\epsilon_{s,t}$) of the current time step for a given site was allowed to be correlated with that of the previous time step by $\upphi$ and was assumed to be normally distributed with a variance of $\sigma^{2}$.
-To ensure our model estimates and simulations remained within the calendar year we logit transformed our flow-timing response variable (i.e., day-of-year to half-annual-flow) after scaling the data between 0 and 1 (i.e., dividing by 365).
-
-\begin{figure}[h]
-	\centerline{\includegraphics[height=3in]{Fig2_Flow-Time.pdf}}
-	\caption{Annual median flow at 55 sites within the Fraser River basin.
-Flow rates are scaled at each site by the root mean square in order to make flow rates comparable among sites.
-Lines are loess smoothers with a smoothing span of 0.25 predicted each month between 1970 and 2007.
-Line and point size are scaled to each flow gauge site's watershed area.
-The data show inter-annual variability within sites with varying magnitudes and directions among sites, illustrating variable long-term change in hydrology across the basin.}
-	\label{figtwo}
-\end{figure}
-
-\subsection{River Network Trend Attenuation Analysis}
-We hypothesized that larger catchments within the Fraser River watershed with more diverse climate portfolios would have dampened long-term hydrological change when compared to their smaller sub-catchments.
-Therefore, to test whether networks dampen climate signals we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model:
-\begin{linenomath*}
-\begin{equation}
-	\hat{b}_{s} = c + d\sqrt{\mathrm{area}_{s}} + \eta_{s}, \quad
-  \eta_{s} \sim \mathcal{N}(0, f(\mathrm{area}_{s})) \label{eq2},
-\end{equation}
-\end{linenomath*}
-where $\hat{b}_{s}$ represents a flow trend at a given site ($s$) and $\mathrm{area}_{s}$ represents site $s$'s watershed area.
-Fitted $d$ and $c$ parameters represent the mean effect of $\sqrt{\mathrm{watershed area}}$ on flow trends and the mean basin-wide flow trend at a theoretical watershed area of zero, respectively.
-To quantify attenuation, we modeled the change in trend variability with the change in catchment area using an exponential variance function \citep[p.~211]{Pinheiro:2006}:
-\begin{linenomath*}
-\begin{equation}
-	f(\mathrm{area}_{s})=\sigma_b^2 \exp(2\updelta\sqrt{\mathrm{area}_{s}}) \label{eq3},
-\end{equation}
-\end{linenomath*} 
-where the variance of the estimated error ($\eta$) changes exponentially with increasing $\sqrt{\mathrm{area}}$.
-This variance function allowed us to predict the range of flow trend values that would be expected as watershed area increased.
-A decreasing range of flow trends with increasing watershed area would indicate a climate-dampening network effect.
-
-\subsection{River Discharge Trend Simulation}
-In the observed data, small watersheds exhibited greater variability around their trend estimates than large watersheds, likely because of greater short-term variation in small catchments \citep{Moore:2015}.
-This relationship between watershed size and trend certainty may pull small watershed trend estimates away from zero, thereby creating the appearance of decreasing trend variability among sites as watershed size increases.
-Using a null-hypothesis framework we simulated time-series for each site with no underlying trend.
-These simulations were generated by a random walk process with the observed standard deviation and autocorrelation parameters ($\hat{\upsigma}$, $\hat{\upphi}$) at our sites.
-Using the same GLS model and AR1 correlation structure as applied to the observed data (eq.~\ref{eq1}), we estimated trends for 1000 simulated time series at each site for each response variable (e.g., Figure S3).
-This simulation process, a form of parametric bootstrapping, created distributions of null-expectations with which we compared our observed results.
-
-We then applied equation~\ref{eq2} to each of the 1000 basin-wide simulations, resulting in 1 observed variance exponent parameter ($\hat{\updelta}$) and 1000 simulated $\hat{\updelta}$ for each flow metric.
-By comparing our observed attenuation with our basin-wide null-model simulations we addressed the potential that more variable flows in smaller catchments contribute to the observed pattern of flow trends as a function of watershed area (Figure~\ref{figthree}).
-This null-model approach asks how likely our observation is due to a sampling effect versus a network portfolio effect.
-
-\begin{figure}[h]
-	\centerline{\includegraphics[height=4.5in]{Fig3_Annual-Funnel.pdf}}
-	\caption{Annual flow trend attenuation within the Fraser River basin.
-(\textbf{Left}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}).
-Blue represents observed attenuation; orange and red represent simulated attenuation that is weaker and stronger than observed.
-(\textbf{Right}) Density plots show null model simulated variance exponents ($\hat{\updelta}$) and the proportion on either side of the observed variance exponent (blue).
-Flow metrics include long-term flow-timing shifts (change per decade in day-of-year (DOY) to half annual flow), where decreasing trends suggest more annual flow is occurring earlier in the year (a), and the percent change per decade in minimum (b), maximum (c) and median (d) annual flow.
-Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
-	\label{figthree}
-\end{figure}
-
-\subsection{Dampening Summary Statistics}
-We calculated attenuation certainty as the percentage of simulated $\hat{\updelta}$ that were less than the observed $\hat{\updelta}$.
-We calculated attenuation strength as the ratio between the standard deviation at the smallest and largest watersheds as defined by the exponential variance function (i.e.,\\$\sqrt{f(\mathrm{area}_1)} / \sqrt{f(\mathrm{area}_2)}$ ).
-To compare the degree of attenuation observed with the null-simulated attenuation, we compared this ratio with the same ratio calculated from the null-simulated data (i.e., $\left( \sqrt{f(\mathrm{area}_1)} / \sqrt{f(\mathrm{area}_2)} \right) / \left( \sqrt{f(\mathrm{area}_{\mathrm{null} 1})} / \sqrt{f(\mathrm{area}_{\mathrm{null} 2})} \right)$ ).
-Finally, to estimate how the day-of-year to half-annual-flow was changing, we estimated the steepest point of each site's logistic curve and multiplied this slope value by 365 to revert from our 0--1 scale to our original annual range of 0--365.
-To get a decadal rate we multiplied these results by 3.8 (number of decades in 38 years).
-We estimated the steepest part of the logistic curve using the \textit{divide by four} rule \citep{Gelman:2008}, where dividing the coefficient by four equals the first derivative (slope) of the logistic curve at its steepest point.
-
-\section{Results}
-\subsection{Hydrologic Trend Diversity}
-Long-term changes in flow metrics varied substantially across the watershed (Figure~\ref{figthree}).
-For instance, annual maximum flow decreased in 38 sites but increased in 17 over the last four decades (-28 to 15\% change$\cdot$decade\textsuperscript{-1}, std.
-dev.
-= 6\%).
-Annual median flow showed similar variability but a greater tendency towards rising flow trends with 49 increasing and only six decreasing (-5 to 25\% change$\cdot$decade\textsuperscript{-1} std.
-dev.
-= 5\%).
-Thus, within the Fraser River basin, there are a variety of long-term trends in flow, suggesting a diversity of climate responses.
-
-\subsection{Annual Hydrologic Trend Dampening}
-Long-term changes in hydrology were less variable in larger catchments, as predicted, demonstrating between 85 and 92\% dampening of flow trends moving from headwater catchments to the network's confluence.
-In other words, the largest catchments were approximately 10 times more stable in their climate response than the smallest catchments.
-For instance, trends ranged between a 6\% reduction and 19\% increase in median flow per decade among small watersheds, while large watersheds ranged between 0.8 and 1.6\% change per decade.
-Similar attenuation was seen in flow timing with a 94\% reduction in the day-of-year to half-annual-flow trends such that small watersheds were reaching half their annual flow between 8 days later and 25 days earlier while large watersheds have only shifted 2 to 4 days earlier between 1970 and 2007.
-
-A null model approach illustrated that support for a river network portfolio effect was greater than 90\% for three of four annual flow trend metrics and as high as 98\% for median-annual-flow trends (Figure~\ref{figthree}).
-Trend attenuation was 3.1 (0.7--9.1 this and hereafter are 90\% intervals), 1.6 (0.3--4.6), 3.4 (1.0--8.5) and 5.4(1.5--14.6) times greater than the null model for annual-flow timing, and minimum, maximum and median annual-flow respectively.
-These statistics provide strong support for the hypothesis that river network portfolio effects contribute to climate dampening in large rivers.
-
-\subsection{Seasonal Hydrologic Trend Dampening}
-Regional flow-trends and attenuation strength varied by season.
-Analysis of monthly trends revealed that winter flows have been getting higher and summer flows have been getting lower over time, but that these trends have been more variable in smaller catchments (Figure~\ref{figfour}, S1, S2).
-For example, winter maximum flows in small catchments ranged from nearly no change to dramatic 44\% increases per decade with decreases only as large as 15\%.
-Despite these extreme locations, on average the basin has only experienced moderate winter maximum flow increases of 5--10\%.
-Summer flows exhibited the opposite, with decreases in high flows over time.
-However, there was weak evidence of attenuation in the spring (Figure~\ref{figfour}, S1, S2).
-For example, in the spring the likelihood of network-driven attenuation of maximum flow trends was as low as 24\% (Figure~\ref{figfour}).
-The deterioration of the network's attenuation strength was coincident with the spring freshet, when snowmelt drives high flows across the basin.
-Synchronization events such as the freshet may subvert river network dampening mechanisms by homogenizing the region's response to climate shifts.
-
-\begin{figure}[htp]
-	\centerline{\includegraphics[height=5.2in]{Fig4_Max-Monthly-Density-Funnel.pdf}}
-	\caption{Monthly maximum flow trend attenuation within the Fraser River basin.
-(\textbf{Left}) Fraser River's basin-wide maximum-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulations.
-Observed values falling further from the center of the density distribution suggest greater evidence for changes in maximum flow and a greater shift in magnitude.
-(\textbf{Center}) Observed monthly Fraser River maximum-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates.
-Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed.
-(\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}) for four seasonally representative months.These reflect months in the prior columns and describe the variation in percent change per decade of maximum flow among sites.
-Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
-	\label{figfour}
-\end{figure}
-
-\section{Discussion}
-Climate change is already causing economic and conservation challenges for river systems worldwide \citep{Palmer:2009}.
-For example, sockeye salmon in the Fraser River have seen as much as 80\% pre-spawn mortality in years with later run-off and elevated temperatures, leading managers to reduce harvest rates in these years.
-In the absence of climate dampening the impacts of these temperature extremes could have been dramatically worse.
-As climate change progresses, river hydrology will continue to shift, further stressing riverine ecosystems and subsequently demanding responsive management.
-However, river networks provide an underappreciated defense against these climate-change impacts as they are uniquely organized to leverage locally-filtered expressions of climate into stability.
-A simple product of form and gravity, stability emerges as river networks integrate landscape heterogeneity, creating a downstream portfolio of climate that smooths local extremes and tempers long-term flow trends.
-Therefore, larger rivers should have flow regimes that are less sensitive to local climate trends.
-
-Despite these dampening processes, river networks are still vulnerable to climate change.
-In the Fraser River, the furthest downstream reaches are still exhibiting small but detectable signals of hydrological shifts.
-Late fall, winter and early spring flows have generally increased while late spring, summer and early fall flows have decreased (Figure~\ref{figfour}, S1, S2).
-These findings generally support previous studies demonstrating earlier spring runoff  \citep{Dery:2012}, decreasing summer and fall flows  \citep{Stahl:2006} and increased winter flows  \citep{Healey:2011}.
-Together, this body of work indicates basin-wide effects of climate change, where shifts in winter precipitation from snow to rain is resulting in increased winter runoff and decreased snow pack contributions to summer flows.
-Our work suggests that because the basin is responding heterogeneously to the shifting climate and integrating climate trends, the downstream impacts of climate change on flow have been greatly reduced.
-In contrast to previously studied types of climate mitigation where habitats modify climate drivers (e.g., carbon storage) or physically absorb climate change impacts (e.g., mangroves), network climate dampening smooths out extreme climatic trends thereby reducing but not eliminating the impact of local extremes.
-
-Our study provides insight into the spatial scaling of climate change.
-We focused on a vast watershed with a remarkably diverse climate portfolio, yet we would still expect that smaller rivers could dampen some degree of climate variability.
-For instance, based on our data, we predict that rivers draining a catchment of 60,000 km\textsuperscript{2} would have 66\% less variability in the rate of change in hydrology than a smaller river draining ~5000 km\textsuperscript{2}.
-While this dampening is less than the 10-fold decrease observed in the entire 220,000 km\textsuperscript{2} Fraser River, smaller watersheds can offer a defense against the impacts of climate change.
-
-There have been extensive efforts to predict the hydrologic sensitivity of rivers due to climate change through downscaled climate projections \citep{Nijssen:2001}.
-Here we demonstrate that large rivers may have an inherent ability to absorb local climate change.
-But just as the destruction of coastal habitats degrades their natural capacity to mitigate sea-level rise or storm events \citep{Arkema:2013}, the management of river basins likely impacts the ability for river networks to attenuate climate trends.
-For example, dams synchronize the flow regimes of rivers  \citep{Poff:2007} and logging increases the magnitude and frequency of extreme events such as high and low flows through increased runoff rates \citep{Zhang:2014}.
-These anthropogenic activities may magnify the impacts of climatic shifts such as the transition from snow- to rain-dominated precipitation in some basins.
-The climate cost of these watershed activities should be considered in environmental decision making.
-Climate change is a global challenge; here we suggest large rivers dampen local change by leveraging landscape diversity into a portfolio of climate, an important tool in the climate-mitigation toolbox.
-
-\acknowledgments
-All the data and R scripts used in this study are freely available on github at https://github.com/kchezik/River-Network-Flow-Trends.
-This study could not have been done without the decades of flow data collected by those at Environment Canada.
-K.A.
-Chezik and J.W.
-Moore were supported by the Liber Ero Chair of Coastal Science and Management and Simon Fraser University.
-S.C.
-Anderson was supported by the David H.
-Smith Conservation Research Fellowship.
-
-\begin{thebibliography}{30}
-\providecommand{\natexlab}[1]{#1}
-\expandafter\ifx\csname urlstyle\endcsname\relax
-  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
-  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
-  \urlstyle{rm}\Url}\fi
-
-\bibitem[{\textit{Arkema et~al.}(2013)\textit{Arkema, Guannel, Verutes, Wood,
-  Guerry, Ruckelshaus, Kareiva, Lacayo, and Silver}}]{Arkema:2013}
-Arkema, K.~K., G.~Guannel, G.~Verutes, S.~A. Wood, A.~Guerry, M.~Ruckelshaus,
-  P.~Kareiva, M.~Lacayo, and J.~M. Silver (2013), Coastal habitats shield
-  people and property from sea-level rise and storms, \textit{Nature Climate
-  Change}, \textit{3}(10), 913--918.
-
-\bibitem[{\textit{Bindoff et~al.}(2013)\textit{Bindoff, Stott, AchutaRao,
-  Gillett, Gutzler, Hansingo, Hegerl, Hu, Jain, Mokhov, Overland, Perlwitz,
-  Sebbari, and Zhang}}]{Bindoff:2013}
-Bindoff, N.~L., P.~A. Stott, M.~R. AchutaRao, N.~P. Gillett, D.~Gutzler,
-  K.~Hansingo, G.~C. Hegerl, Y.~Hu, S.~Jain, I.~I. Mokhov, J.~Overland,
-  J.~Perlwitz, R.~Sebbari, and X.~Zhang (2013), Detection and attribution of
-  climate change: from global to regional, \textit{Tech. rep.},
-  Intergovernmental Panel on Climate Change.
-
-\bibitem[{\textit{D{\'e}ry et~al.}(2012)\textit{D{\'e}ry,
-  Hern{\'a}ndez-Henr{\'\i}quez, Owens, Parkes, and Petticrew}}]{Dery:2012}
-D{\'e}ry, S.~J., M.~A. Hern{\'a}ndez-Henr{\'\i}quez, P.~N. Owens, M.~W. Parkes,
-  and E.~L. Petticrew (2012), A century of hydrological variability and trends
-  in the {F}raser {R}iver basin, \textit{Environmental Research Letters},
-  \textit{7}(2), 024,019.
-
-\bibitem[{\textit{Doak et~al.}(1998)\textit{Doak, Bigger, Harding, Marvier,
-  O'malley, and Thomson}}]{Doak:1998}
-Doak, D.~F., D.~Bigger, E.~Harding, M.~Marvier, R.~O'malley, and D.~Thomson
-  (1998), The statistical inevitability of stability-diversity relationships in
-  community ecology, \textit{The American Naturalist}, \textit{151}(3),
-  264--276 %@ 0003--0147.
-
-\bibitem[{\textit{Donat et~al.}(2016)\textit{Donat, Lowry, Alexander, O'Gorman,
-  and Maher}}]{Donat:2016}
-Donat, M.~G., A.~L. Lowry, L.~V. Alexander, P.~A. O'Gorman, and N.~Maher
-  (2016), More extreme precipitation in the world's dry and wet regions,
-  \textit{Nature Climate Change}, \textit{6}(5), 508--513,
-  \doi{10.1038/nclimate2941}.
-
-\bibitem[{\textit{Gelman and Hill}(2008)}]{Gelman:2008}
-Gelman, A., and J.~Hill (2008), Data analysis using regression and multilevel
-  hierarchical models, \doi{10.1080/02664760801949043}.
-
-\bibitem[{\textit{George et~al.}(2015)\textit{George, Thompson~III, and
-  Faaborg}}]{George:2015}
-George, A.~D., F.~R. Thompson~III, and J.~Faaborg (2015), Using {L}i{DAR} and
-  remote microclimate loggers to downscale near-surface air temperatures for
-  site-level studies, \textit{Remote Sensing Letters}, \textit{6}(12),
-  924--932.
-
-\bibitem[{\textit{Guerry et~al.}(2015)\textit{Guerry, Polasky, Lubchenco,
-  Chaplin-Kramer, Daily, Griffin, Ruckelshaus, Bateman, Duraiappah, Elmqvist,
-  Feldman, Folke, Hoekstra, Kareiva, Keeler, Li, McKenzie, Ouyang, Reyers,
-  Ricketts, Rockstr{\"o}m, Tallis, and Vira}}]{Guerry:2015}
-Guerry, A.~D., S.~Polasky, J.~Lubchenco, R.~Chaplin-Kramer, G.~C. Daily,
-  R.~Griffin, M.~Ruckelshaus, I.~J. Bateman, A.~Duraiappah, T.~Elmqvist, M.~W.
-  Feldman, C.~Folke, J.~Hoekstra, P.~M. Kareiva, B.~L. Keeler, S.~Li,
-  E.~McKenzie, Z.~Ouyang, B.~Reyers, T.~H. Ricketts, J.~Rockstr{\"o}m,
-  H.~Tallis, and B.~Vira (2015), Natural capital and ecosystem services
-  informing decisions: From promise to practice, \textit{Proceedings of the
-  National Academy of Sciences}, \textit{112}(24), 7348--7355,
-  \doi{10.1073/pnas.1503751112}.
-
-\bibitem[{\textit{Hartmann et~al.}(2013)\textit{Hartmann, Klein, Rusticucci,
-  Alexander, Bronnimann, Charabi, Dentener, Dlugokencky, Easterling, Kaplan,
-  Soden, Thorne, Wild, and Zhai}}]{Hartmann:2013}
-Hartmann, D.~L., A.~M.~G. Klein, M.~Rusticucci, L.~V. Alexander, S.~Bronnimann,
-  Y.~Charabi, F.~J. Dentener, D.~R. Dlugokencky, A.~Easterling, B.~J. Kaplan,
-  B.~J. Soden, P.~W. Thorne, M.~Wild, and P.~M. Zhai (2013), Climate change
-  2013: The physical science basis., \textit{Tech. rep.}, Intergovernmental
-  Panel on Climate Change.
-
-\bibitem[{\textit{Healey and Bradford}(2011)}]{Healey:2011}
-Healey, M., and M.~Bradford (2011), The cumulative impacts of climate change on
-  {F}raser {R}iver sockeye salmon (\textit{{O}ncorhynchus nerka}) and
-  implications for management, \textit{Can. J. Fish. Aquat. Sci.},
-  \textit{68}(4), 718--737.
-
-\bibitem[{\textit{Hirabayashi et~al.}(2013)\textit{Hirabayashi, Mahendran,
-  Koirala, Konoshima, Yamazaki, Watanabe, Kim, and Kanae}}]{Hirabayashi:2013}
-Hirabayashi, Y., R.~Mahendran, S.~Koirala, L.~Konoshima, D.~Yamazaki,
-  S.~Watanabe, H.~Kim, and S.~Kanae (2013), Global flood risk under climate
-  change, \textit{Nature Climate Change}, \textit{3}, 816--821.
-
-\bibitem[{\textit{Hulme}(2016)}]{Hulme:2016}
-Hulme, M. (2016), 1.5$\degree${C} and climate research after the {P}aris
-  {A}greement, \textit{Nature Climate Change}, \textit{6}(3), 222--224.
-
-\bibitem[{\textit{Jones et~al.}(2012)\textit{Jones, Hole, and
-  Zavaleta}}]{Jones:2012}
-Jones, H.~P., D.~G. Hole, and E.~S. Zavaleta (2012), Harnessing nature to help
-  people adapt to climate change, \textit{Nature Climate Change},
-  \textit{2}(7), 504--509.
-
-\bibitem[{\textit{McCabe and Wolock}(2014)}]{McCabe:2014}
-McCabe, G.~J., and D.~M. Wolock (2014), Spatial and temporal patterns in
-  conterminous {U}nited {S}tates streamflow characteristics,
-  \textit{Geophysical Research Letters}, \textit{41}(19), 6889--6897.
-
-\bibitem[{\textit{Melillo et~al.}(2014)\textit{Melillo, Richmond, and
-  Yohe}}]{Melillo:2014}
-Melillo, J.~M., T.~Richmond, and G.~W. Yohe (2014), Climate change impacts in
-  the {U}nited {S}tates: the third national climate assessment, \textit{US
-  Global change research program}, \textit{841}.
-
-\bibitem[{\textit{Moore et~al.}(2015)\textit{Moore, Beakes, Nesbitt, Yeakel,
-  Patterson, Thompson, Phillis, Braun, Favaro, Scott et~al.}}]{Moore:2015}
-Moore, J.~W., M.~P. Beakes, H.~K. Nesbitt, J.~D. Yeakel, D.~A. Patterson, L.~A.
-  Thompson, C.~Phillis, D.~C. Braun, C.~Favaro, D.~C. Scott, et~al. (2014),
-  Emergent stability in a large free-flowing watershed, \textit{Ecology}, 
-  \textit{96}(2), 340--347.
-
-\bibitem[{\textit{Morrison et~al.}(2002)\textit{Morrison, Quick, and
-  Foreman}}]{Morrison:2002}
-Morrison, J., M.~C. Quick, and M.~G. Foreman (2002), Climate change in the
-  {F}raser {R}iver watershed: flow and temperature projections, \textit{Journal
-  of Hydrology}, \textit{263}(1), 230--244.
-
-\bibitem[{\textit{Narain et~al.}(2011)\textit{Narain, Margulis, and
-  Essam}}]{Narain:2011}
-Narain, U., S.~Margulis, and T.~Essam (2011), Estimating costs of adaptation to
-  climate change, \textit{Climate Policy}, \textit{11}(3), 1001--1019.
-
-\bibitem[{\textit{Nijssen et~al.}(2001)\textit{Nijssen, O'Donnell, Hamlet, and
-  Lettenmaier}}]{Nijssen:2001}
-Nijssen, B., G.~M. O'Donnell, A.~F. Hamlet, and D.~P. Lettenmaier (2001),
-  Hydrologic sensitivity of global rivers to climate change, \textit{Climatic
-  Change}, \textit{50}(1-2), 143--175.
-
-\bibitem[{\textit{Palmer et~al.}(2009)\textit{Palmer, Lettenmaier, Poff,
-  Postel, Richter, and Warner}}]{Palmer:2009}
-Palmer, M.~A., D.~P. Lettenmaier, N.~L. Poff, S.~L. Postel, B.~Richter, and
-  R.~Warner (2009), Climate change and river ecosystems: protection and
-  adaptation options, \textit{Environmental Management}, \textit{44}(6),
-  1053--1068.
-
-\bibitem[{\textit{Peterson et~al.}(2013)\textit{Peterson, Ver~Hoef, Isaak,
-  Falke, Fortin, Jordan, McNyset, Monestiez, Ruesch, Sengupta
-  et~al.}}]{Peterson:2013}
-Peterson, E.~E., J.~M. Ver~Hoef, D.~J. Isaak, J.~A. Falke, M.-J. Fortin, C.~E.
-  Jordan, K.~McNyset, P.~Monestiez, A.~S. Ruesch, A.~Sengupta, et~al. (2013),
-  Modelling dendritic ecological networks in space: an integrated network
-  perspective, \textit{Ecology Letters}, \textit{16}(5), 707--719.
-
-\bibitem[{\textit{Pinheiro and Bates}(2006)}]{Pinheiro:2006}
-Pinheiro, J., and D.~Bates (2006), Mixed-effects models in {S} and {S-PLUS},
-  \textit{Springer Science \& Business Media}.
-
-\bibitem[{\textit{Poff et~al.}(2007)\textit{Poff, Olden, Merritt, and
-  Pepin}}]{Poff:2007}
-Poff, L.~N., J.~D. Olden, D.~M. Merritt, and D.~M. Pepin (2007), Homogenization
-  of regional river dynamics by dams and global biodiversity implications,
-  \textit{Proceedings of the National Academy of Sciences}, \textit{104}(14),
-  5732--5737, \doi{10.1073/pnas.0609812104}.
-
-\bibitem[{\textit{Rauscher et~al.}(2008)\textit{Rauscher, Pal, Diffenbaugh, and
-  Benedetti}}]{Rauscher:2008}
-Rauscher, S.~A., J.~S. Pal, N.~S. Diffenbaugh, and M.~M. Benedetti (2008),
-  Future changes in snowmelt-driven runoff timing over the western {US},
-  \textit{Geophysical Research Letters}, \textit{35}(16).
-
-\bibitem[{\textit{\relax{Louisiana Coastal Wetlands Conservation and
-  Restoration Task Force and the Wetlands Conservation and Restoration
-  Authority}}(1998)}]{Conservation:1998}
-\relax{Louisiana Coastal Wetlands Conservation and Restoration Task Force and
-  the Wetlands Conservation and Restoration Authority} (1998), Coast 2050:
-  {T}oward a sustainable coastal louisiana, \textit{Tech. rep.}, Louisiana
-  Department of Natural Resources.
-
-\bibitem[{\textit{Stahl and Moore}(2006)}]{Stahl:2006}
-Stahl, K., and R.~Moore (2006), Influence of watershed glacier coverage on
-  summer streamflow in {B}ritish {C}olumbia, {C}anada, \textit{Water Resources
-  Research}, \textit{42}(6).
-
-\bibitem[{\textit{V{\"o}r{\"o}smarty et~al.}(2010)\textit{V{\"o}r{\"o}smarty,
-  McIntyre, Gessner, Dudgeon, Prusevich, Green, Glidden, Bunn, Sullivan,
-  Liermann et~al.}}]{Vorosmarty:2010}
-V{\"o}r{\"o}smarty, C.~J., P.~McIntyre, M.~O. Gessner, D.~Dudgeon,
-  A.~Prusevich, P.~Green, S.~Glidden, S.~E. Bunn, C.~A. Sullivan, C.~R.
-  Liermann, et~al. (2010), Global threats to human water security and river
-  biodiversity, \textit{Nature}, \textit{467}(7315), 555--561.
-
-\bibitem[{\textit{Xenopoulos and Lodge}(2006)}]{Xenopoulos:2006}
-Xenopoulos, M.~A., and D.~M. Lodge (2006), Going with the flow: using
-  species-discharge relationships to forecast losses in fish biodiversity,
-  \textit{Ecology}, \textit{87}(8), 1907--1914.
-
-\bibitem[{\textit{Yeakel et~al.}(2014)\textit{Yeakel, Moore, Guimaraes, and
-  Aguiar}}]{Yeakel:2014}
-Yeakel, J. D., J. W.~Moore, P. R.~Guimaraes, and M. A. M. de~Aguiar (2014), Synchronisation and
-  stability in river metapopulation networks, \textit{Ecology Letters},
-  \textit{17}(3), 273--283.
-
-\bibitem[{\textit{Zhang and Wei}(2014)}]{Zhang:2014}
-Zhang, M., and X.~Wei (2014), Alteration of flow regimes caused by large-scale
-  forest disturbance: a case study from a large watershed in the interior of
-  {B}ritish {C}olumbia, {C}anada, \textit{Ecohydrology}, \textit{7}(2),
-  544--556, \doi{10.1002/eco.1374}.
-
-\end{thebibliography}
-
-%\bibliography{/Users/kylechezik/Documents/lit}
+\begin{figure*}
+\centering
+\includegraphics[width=11.4cm]{FigS5_BasinSim.pdf}
+	\caption{Example flow simulation results.
+Trend estimates $\pm$SE of simulated null-model data (yellow) and the observed data upon which simulations were based (blue), plotted against watershed area.
+Curved lines represent the estimated variance function ($\hat{\updelta}$) given the simulated data.
+This is a single example result of the 1000 null-model simulations produced from the data in supplemental figure \ref{fig:S3}.}
+	\label{fig:S5}
+\end{figure*}
 

### 2017-04-08 drafts/PNAS/Chezik_PNAS_MS.tex
Added Sean and Jon comments to first draft.

@@
 \author[a,1]{Kyle A. Chezik}
-\author[b,1]{Sean C. Anderson} 
+\author[b]{Sean C. Anderson} 
 \author[a,1]{Jonathan W. Moore}
@@
 % Please add here a significance statement to explain the relevance of your work
-\significancestatement{Many ecosystems posses stabilizing processes that provide a natural defense against the impacts of climate change.
-The branching tree structure of river networks, provide a framework for aggregating the varied signals of climate change within a watershed and a possible mechanism for dampening long-term climate driven trends.
-Varied landscape attributes within a watershed produce a mosaic or portfolio of climate which in turn results in a variety of climate change manifestations.
-Here we demonstrate that large free-flowing river networks dampen local long-term flow trends by integrating a large and varied climate portfolio across a diverse and complex landscape.
-Integrating the benefits of these ecosystems into climate change mitigation strategies is essential in reducing the increasingly extreme impacts of climate change.}
+\significancestatement{Many ecosystems posses processes that provide a natural defense against the impacts of climate change.
+Here we demonstrate that large free-flowing river networks dampen local long-term flow trends by integrating varied climate responses across a diverse and complex landscape.
+We show that larger catchments exhibit greater climate trend diversity and that this climate diversity dampens the local extremes of climate change.
+Thus, rivers may possess a previously under-appreciated capacity for climate relief that can be incorporated into climate change mitigation strategies.}
 
@@
 \authorcontributions{J.W.
-Moore conceived the study and analysis, S.C.
-Anderson formalized the analysis, K.A.
-Chezik gathered and analyzed the data and wrote the paper.
-All authors contributed in all aspects to varying degrees.}
+Moore conceived the project, K.A.
+Chezik, S.C.
+Anderson and J.W.
+Moore, designed the study, K.A.
+Chezik gathered and processed the data, K.A.
+Chezik and S.C.
+Anderson analyzed the data and K.A.
+Chezik wrote the paper with input from all authors.}
 \authordeclaration{The authors declare no conflict of interest.}
-\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: kchezik@sfu.ca}
+\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: kchezik@sfu.ca or jwmoore@sfu.ca}
 
 % Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
-\keywords{rivers networks $|$ climate portfolio $|$ climate change $|$ trend attenuation $|$ climate dampening}
+\keywords{dendritic networks $|$ climate portfolio $|$ flow trends $|$ trend attenuation $|$ climate dampening}
 
 \begin{abstract}
-	Climate change is dramatically altering temperate rivers --- snow-melt is shifting ever earlier \cite{Rauscher:2008} floods are becoming more frequent \cite{Hirabayashi:2013}, and low flows more persistent \cite{McCabe:2014}.
+	Climate change is dramatically altering temperate rivers --- snow-melt is shifting ever earlier,  floods are becoming more frequent and low flows more persistent.
 Here we describe a novel pathway by which river networks dampen local hydrologic climate change through the aggregation of upstream climate portfolio assets.
@@
 Although climate dampening strength varied seasonally, our findings indicate that large free-flowing rivers offer a powerful and largely unappreciated process of climate change mitigation.
-River networks that integrate a diverse climate portfolio can dampen local extremes and offer climate relief to riverine biota.
+River networks that integrate a diverse climate portfolio can dampen local extremes and offer climate change relief to riverine biota.
 \end{abstract}
@@
 Here we consider whether river networks attenuate local long-term climate change through analysis of hydrometric data from one of the largest free-flowing rivers in North America.
-Located in British Columbia, Canada, the Fraser River drains an area approximately the size of the United Kingdom ($\sim$217,000 km\textsuperscript{2}) and in a rare combination is fairly well monitored while also having no dams on its main stem \cite{Vorosmarty:2010} (Fig.~\ref{fig:1}).
+Located in British Columbia, Canada, the Fraser River drains an area approximately the size of the United Kingdom ($\sim$217,000 km\textsuperscript{2}) and in a rare combination is fairly well-monitored while also having no dams on its main stem \cite{Vorosmarty:2010} (Fig.~\ref{fig:1}).
 This watershed drains interior high plateau, coastal mountains, and the Canadian Rockies, and thus integrates a diverse mosaic of landscapes, weather, and climate.
-As with other mid-latitude rivers \cite{Bindoff:2013}, the Fraser River has expressed increasing trends in discharge volume and variability over the last several decades \cite{Dery:2012,Morrison:2002}.
-The river's discharge is monitored by hydrometric gauge stations throughout the watershed.
+As with other mid-latitude rivers \cite{Bindoff:2013}, the Fraser River's discharge volume and variability have increased over the last several decades \cite{Dery:2012,Morrison:2002}.
+The River's discharge is monitored by hydrometric gauge stations throughout the watershed.
 We analyzed 38 years of data collected concurrently at 55 of these stations and applied a novel analytical approach to examine whether this large river exhibits signals of long-term climate dampening.
@@
 \section*{Detecting Climate Change in the Fraser River Basin}
-We focused on changes in hydrology as our riverine manifestation of climate signaling.
-As climate change alters precipitation patterns from snow to rain in temperate regions, rivers are generally predicted to have lower low-flows and higher high-flows, as well as earlier snowmelt \cite{Nijssen:2001}.
+We focused on changes in hydrology as our riverine manifestation of climate change.
+As climate change shifts precipitation patterns from snow to rain in temperate regions, rivers are generally predicted to have lower low-flows and higher high-flows, as well as earlier snowmelt \cite{Nijssen:2001}.
 These manifestations of rising temperatures are impacting species \cite{Xenopoulos:2006} and people \cite{Hirabayashi:2013}.
-Extracting annual and monthly metrics of flow from Fraser River basin gauge station records, we used rolling average maximum- and minimum-flow estimates to capture trends in extreme discharge and rolling average median-flow estimates to capture general annual and seasonal trends.
+Extracting annual and monthly metrics of flow from Fraser River basin gauge station records, we used five day rolling average maximum- and minimum-flow estimates to capture trends in extreme discharge and rolling average median-flow estimates to capture general annual and seasonal trends.
 Because the timing of flow is particularly important to the phenology of many organisms, we also calculated the day-of-year in which half the annual flow was reached.
@@
 Long-term changes in flow metrics varied substantially across the watershed.
-For instance, annual maximum flow decreased in 40 sites but increased in 15 over the last four decades (-26 to 10\% change$\cdot$decade\textsuperscript{-1}, std.
-dev.
-= 6\%).
-Annual median flow showed similar variability but a greater tendency towards rising flow trends with 49 increasing and only six decreasing (-5 to 25\% change$\cdot$decade\textsuperscript{-1} std.
-dev.
-= 5\%).
+For instance, annual maximum flow decreased in 40 sites but increased in 15 over the last four decades (-26 to 10\% change$\cdot$decade\textsuperscript{-1}, SD = 6\%).
+Annual median flow showed similar variability but a greater tendency towards rising flow trends with 49 increasing and only six decreasing (-5 to 25\% change$\cdot$decade\textsuperscript{-1} SD = 5\%).
 Thus, within the Fraser River basin, there are a variety of long-term trends in flow, suggesting a diversity of climate responses and landscape effects.
@@
 Using the Western North America Climate tool (ClimateWNA) \cite{Wang:2016}, we quantified the diversity in climate trends within the Fraser River basin over the years concurrent with our hydrology trend estimates (Fig.~\ref{fig:1}) (\textit{Materials and Methods}).
-We then summarized the climate portfolio of each flow gauge by calculating the standard deviation in climate trends within their catchments.
-Temperature and precipitation are likely to drive changing hydrology, therefore we included variables in our climate index that capture changes in extreme and mean temperature, as well as changes in precipitation volume and physical state.
-The standard deviations of these climate variables were scaled between 0 and 1 and summed into a general climate index for each catchment.
-Following our hypothesis, this climate portfolio index increases with area (Fig.~\ref{fig:2}), rapidly asymptoting in catchments greater than 40,000~km\textsuperscript{2}.
+The rate of warming varies across the watershed (from increases of 0.03 to 0.04\degree{C}$\cdot$year\textsuperscript{-1}).Further, there have been increases in precipitation towards the more northeastern portion.
+
+This spatial variation in climate change means that catchments of different size integrate different amounts of climate variability.
+We summarized the climate variability integrated by each flow gauge by calculating the standard deviation in climate trends within their catchments.
+Temperature and precipitation are likely to drive changing hydrology, therefore we included variables in our climate index that capture changes in extreme and mean temperature, as well as changes in precipitation volume and physical state (snow vs.
+rain).
+The standard deviations of these climate variables were scaled between 0 and 1 and summed as a general climate index for each catchment.
+Following our hypothesis, this climate variability index increases with area (Fig.~\ref{fig:2}), asymptoting in catchments greater than 40,000~km\textsuperscript{2}.Thus, rivers that drain larger catchments integrated more climate variability.
+Larger areas contain greater landscape complexity and thus a greater variety of climatic expressions.
 
@@
 \includegraphics[width=.8\linewidth]{Fig2_ClimPort.pdf}
-	\caption{Climate portfolio index as a function of catchment area.
+	\caption{Climate variability index as a function of catchment area.
 The mean trend line (blue) was estimated using robust linear regression and standard error estimates (grey) were calculated using the weighted standard deviation estimate.}
@@
 \section*{Quantifying the River Network Portfolio Effect}
-Climate portfolio diversity is largely a function of area (Fig.~\ref{fig:2}).
-Larger areas contain greater landscape complexity and thus a greater variety of climatic expressions.
-The influence of any individual sub-basin on downstream flow trends is also a function of area.
-Larger catchments simply aggregate more water such that their climate portfolio will carry more weight on downstream dampening.
-The interplay between climate portfolio and catchment area on the impact of downstream flow trend dampening makes area a more robust response variable than simply the climate portfolio index.
-Therefore, to test whether networks dampen climate signals we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model (\textit{Materials and Methods}).
+We hypothesize that flow gauge sites with larger climate porfolios will exhibit greater long-term flow trend stability than sites with smaller climate portfolios.
+We used catchment area as a proxy for the climate portfolio diversity of a given site, given that larger catchments have more variable climate portfolios (Fig.~\ref{fig:2}) as well as aggregate more water and carry more weight on downstream dampening.
+Therefore, to test whether networks dampen climate signals we regressed site-specific estimates of hydrologic trend (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model (\textit{Materials and Methods}).
 To quantify attenuation, we modeled the change in trend variability with the change in catchment area using an exponential variance function \cite{Pinheiro:2000}.
@@
 In the observed data, small watersheds exhibited greater variability around their trend estimates than large watersheds, likely because of greater short-term variation in small catchments (Fig.~\ref{fig:3}) \cite{Moore:2015}.
-This relationship between watershed size and trend certainty may pull small watershed trend estimates away from zero, thereby creating the appearance of decreasing trend variability among sites as watershed size increases.
+This relationship between watershed size and trend certainty could pull small watershed trend estimates away from zero, thereby creating the appearance of decreasing trend variability among sites as watershed size increases.
 Using a null-hypothesis framework (\textit{Materials and Methods}) we simulated time-series for each site with no underlying trend.
-These simulations were generated by a random walk process with the observed standard deviation and autocorrelation parameters ($\hat{\upsigma}$, $\hat{\upphi}$) at our sites.
+These simulations were generated by a random-walk process with the observed standard deviation and autocorrelation parameters ($\hat{\upsigma}$, $\hat{\upphi}$) at our sites.
 Using the same generalized least squares model applied to the observed data, we estimated trends for 1000 simulated time series at each site for each response variable (e.g., Fig.~S3).
@@
 Our null model approach illustrates that support for a river network portfolio effect is greater than 90\% for all four annual flow trend metrics and as high as 98\% for median-annual-flow trends (Fig.~\ref{fig:3}).
-Trend attenuation was 3.0 (0.7--9.4 this and hereafter are 90\% intervals), 4.8 (1.1--13.8), 3.4 (0.9--9.0) and 5.3(1.4--14.8) times greater than the null model for annual-flow timing, and minimum, maximum and median annual-flow respectively.
+Trend attenuation was; 3.0 (0.7--9.4 this and hereafter are 90\% intervals), 4.8 (1.1--13.8), 3.4 (0.9--9.0) and 5.3(1.4--14.8) times greater than the null model for annual-flow timing, and minimum, maximum and median annual-flow, respectively.
 These statistics provide strong support for the hypothesis that river network portfolio effects contribute to climate dampening in large rivers.
@@
 Summer flows exhibited the opposite, with decreases in high flows over time.
-However, there was weak evidence of attenuation in the spring (Fig.~\ref{fig:4}, \textit{SI Appendix},Figs.~S1,S2).
-Overall, climate portfolios tended to be greater during stable periods of the year and decrease during seasonal transitions (\textit{SI Appendix}, Fig.~S3).
-In the spring the likelihood of network-driven attenuation of maximum flow trends was as low as 25\% (Fig.~\ref{fig:4}).
+However, there was weak evidence of attenuation in the spring (Fig.~\ref{fig:4}, \textit{SI Appendix}, Figs.~S1,S2).
+Overall, climate portfolios tended to have greater attenuation during stable periods of the year and decreased during seasonal transitions (\textit{SI Appendix}, Fig.~S3).
+In the spring, the likelihood of network-driven attenuation of maximum flow trends was as low as 25\% (Fig.~\ref{fig:4}).
 The deterioration of the network's attenuation strength was coincident with the spring freshet, when snowmelt drives high flows across the basin.
@@
 These findings generally support previous studies demonstrating earlier spring runoff \cite{Dery:2012}, decreasing summer and fall flows \cite{Stahl:2006} and increased winter flows \cite{Healey:2011}.
-Together, this body of work indicates basin-wide effects of climate change, where shifts in winter precipitation from snow to rain is resulting in increased winter runoff and decreased snow pack contributions to summer flows.
+In this large free-flowing river, climate change is shifting winter precipitation from snow to rain, increasing winter runoff, and decreasing snow pack contributions to summer flows.
 Our work suggests that because the basin is responding heterogeneously to the shifting climate and integrating climate trends, the downstream impacts of climate change on flow are greatly reduced.
-In contrast to previously studied types of climate mitigation where habitats modify climate drivers (e.g., carbon storage) or physically absorb climate change impacts (e.g., mangroves), network climate dampening smooths out extreme climatic trends thereby reducing the impact of local extremes.
 
 \section*{Discussion}
-Climate change is already causing economic and conservation challenges for river systems worldwide\cite{Palmer:2009}.
-For example, sockeye salmon in the Fraser River have seen as much as 80\% pre-spawn mortality in years with later run-off and elevated temperatures, leading managers to close the fishery in 2013 and 2015.
-In the absence of climate dampening the impacts of these temperature extremes could have been dramatically worse.
+Climate change is already causing economic and conservation challenges for river systems worldwide\cite{Palmer:2009,Pecl:2017}.
+For example, sockeye salmon in the Fraser River have seen as much as 80\% pre-spawn mortality in years with later run-off and elevated temperatures, leading managers to reduce or close the fishery (e.g., 2013 and 2015).
+In the absence of climate dampening the impacts of these climate extremes could have been dramatically worse.
 As climate change progresses, river hydrology will continue to shift, further stressing riverine ecosystems and subsequently demanding responsive management.
-However, river networks provide an underappreciated defense against these climate-change impacts as they are uniquely organized to leverage locally filtered expressions of climate into stability.
-A simple product of form and gravity, stability emerges as river networks integrate landscape heterogeneity, creating a downstream portfolio of climate that smooths local extremes and tempers long-term flow trends.
+However, river networks provide an underappreciated defense against these climate-change impacts as they are uniquely organized to leverage locally-filtered expressions of climate into stability.
+A simple product of network form and gravity, stability emerges as river networks integrate landscape heterogeneity, creating a downstream portfolio of climate that smooths local extremes and tempers long-term flow trends.
 Therefore, larger rivers should have flow regimes that are less sensitive to local climate trends.
 
-Our study provides insight into the spatial scaling of climate change.
-We focused on a vast watershed with a remarkably diverse climate portfolio, yet we would still expect that smaller rivers could dampen some degree of climate variability.
+Our study provides insight into the spatial scaling of a river networks dampening capacity.
+Although we focused on a vast watershed with a remarkably diverse climate portfolio, we would still expect that smaller rivers will dampen climate variability.
 For instance, based on our data, we predict that rivers draining a catchment of 60,000 km\textsuperscript{2} would have 66\% less variability in the rate of change in hydrology than a smaller river draining ~5000 km\textsuperscript{2}.
@@
 
-There have been extensive efforts to predict the hydrologic sensitivity of rivers due to climate change through downscaled climate projections \cite{Nijssen:2001}.
-Here we demonstrate that large rivers may have an inherent ability to absorb local climate change.
+In contrast to previously studied types of climate mitigation where habitats modify climate drivers (e.g., carbon storage) or physically absorb climate change impacts (e.g., mangroves), network climate dampening smooths out extreme climatic trends thereby reducing the impact of local extremes.
 But just as the destruction of coastal habitats degrades their natural capacity to mitigate sea-level rise or storm events \cite{Arkema:2013}, the management of river basins likely impacts the ability for river networks to attenuate climate trends.
@@
 We interpolated missing data using a linear model in log-space, but of the 13,879 days of data only 23 days were estimated.
-These daily data were smoothed using a five day rolling average before summarizing into annual and monthly response variables.
+These daily data were smoothed using a five-day rolling average before summarizing into annual and monthly response variables.
 This smoothing technique was used to reduce the influence of erroneous and unusually extreme values.
@@
 
-\subsection*{Basin trend attenuation analysis} To measure network driven dampening of climate signalling, we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model:
+\subsection*{Basin trend attenuation analysis} To measure network-driven dampening of climate signalling, we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model:
 \begin{equation}
@@
 
-\subsection*{River Attnuation Null Model Simulations} In the observed data, small watersheds exhibited greater variability around their trend estimates than large watersheds, likely because of greater short-term variation in small catchments \cite{Moore:2015}.
+\subsection*{River Attenuation Null Model Simulations} In the observed data, small watersheds exhibited greater variability around their trend estimates than large watersheds, likely because of greater short-term variation in small catchments \cite{Moore:2015}.
 This relationship between watershed size and trend certainty may pull small watershed trend estimates away from zero, thereby creating the appearance of decreasing trend variability among sites as watershed size increases.

### 2017-04-11 drafts/PNAS/Chezik_PNAS_MS.tex
Organized
Removed Lingering Defunct Files and Cleaned Up For Reproducibility.

@@
 \usepackage{upgreek}
+\usepackage{gensymb}
 
@@
 The value of climate buffering habitats stands only to increase as average global temperatures climb.
-The Paris Agreement of 2016 aims to limit global temperature rise to 1.5\degree{C} \cite{Hulme:2016}, a goal that concedes a minimum doubling of observed warming \cite{Hartmann:2013}; therefore, an urgent challenge and opportunity is discovering, maintaining and restoring systems that naturally buffer against the oncoming impacts of climate change.
+The Paris Agreement of 2016 aims to limit global temperature rise to 1.5\celsius~\cite{Hulme:2016}, a goal that concedes a minimum doubling of observed warming \cite{Hartmann:2013}; therefore, an urgent challenge and opportunity is discovering, maintaining and restoring systems that naturally buffer against the oncoming impacts of climate change.
 
-In the face of changing global precipitation patterns \cite{Donat:2016}, river networks may offer an unappreciated defense against shifting flow regimes under climate change\cite{Hartmann:2013,Palmer:2009}.
+In the face of changing global precipitation patterns \cite{Donat:2016}, river networks may offer an unappreciated defense against shifting flow regimes under climate change \cite{Hartmann:2013,Palmer:2009}.
 Earlier snowmelt \cite{Rauscher:2008}, reduced snow pack \cite{McCabe:2014}, shifts from snow to rain, and changes in the annual distribution of precipitation are all impacting stream flow \cite{Hartmann:2013}, resulting in flashier flows with increased frequency of flooding \cite{Hirabayashi:2013}, longer periods of drought and critically low flows for wildlife and agriculture \cite{Melillo:2014}.
@@
 Processes that aggregate across asynchronous components tend to dampen the aggregate, a process known as the portfolio effect \cite{Doak:1998}.
-Climate is expressed differently and asynchronously over the landscape \cite{George:2015} and on account of their branching architecture and directional flow, river networks integrate these varied climate manifestations \cite{Peterson:2013}, potentially dampening the local climatic response of sub-catchments.
+Climate is expressed differently and asynchronously over the landscape \cite[e.g.][]{George:2015} and on account of their branching architecture and directional flow, river networks integrate these varied climate manifestations \cite{Peterson:2013}, potentially dampening the local climatic response of sub-catchments.
 Large free-flowing rivers may have diverse climate portfolios, with sub-catchments acting as assets in their portfolio, thereby dampening local climate impacts.
 Portfolio effects in rivers are already known to dampen short-term variation \cite{Moore:2015,Yeakel:2014} but attenuation of long-term climate driven shifts have yet to be examined.
+
+\begin{figure*}[h]
+\centering
+\includegraphics[width=17.7cm]{Fig1_Map.pdf}
+	\caption{Climate trends in mean annual temperature (MAT) and mean annual precipitation (MAP) within the Fraser River basin overlayed on a digital elevation model of British Columbia, Canada.
+Flow gauge sites (dots) are scaled by the size of the contributing area.
+The map is projected in Albers BC UTMs to provide an equal area depiction of the region but labels are expressed in WGS84 latitude and longitude.}
+\label{fig:1}
+\end{figure*}
 
@@
 We analyzed 38 years of data collected concurrently at 55 of these stations and applied a novel analytical approach to examine whether this large river exhibits signals of long-term climate dampening.
-
-\begin{figure*}[h]
-\centering
-\includegraphics[width=17.8cm]{Fig1_Map.pdf}
-	\caption{Climate trends in mean annual temperature (MAT) and mean annual precipitation (MAP) within the Fraser River basin overlayed on a digital elevation model of British Columbia, Canada.
-Flow gauge sites (dots) are scaled by the size of the contributing area.
-The map is projected in Albers BC UTMs to provide an equal area depiction of the region but labels are expressed in WGS84 latitude and longitude.}
-\label{fig:1}
-\end{figure*}
 
@@
 Using the Western North America Climate tool (ClimateWNA) \cite{Wang:2016}, we quantified the diversity in climate trends within the Fraser River basin over the years concurrent with our hydrology trend estimates (Fig.~\ref{fig:1}) (\textit{Materials and Methods}).
-The rate of warming varies across the watershed (from increases of 0.03 to 0.04\degree{C}$\cdot$year\textsuperscript{-1}).Further, there have been increases in precipitation towards the more northeastern portion.
+The rate of warming varies across the watershed (from increases of 0.03 to 0.04\celsius$\cdot$year\textsuperscript{-1}).
+Furthermore, there is an observed decrease in preciptiation along the Canadian Rockies transitioning to increasingly wetter trends along the coast.
 
@@
 The standard deviations of these climate variables were scaled between 0 and 1 and summed as a general climate index for each catchment.
-Following our hypothesis, this climate variability index increases with area (Fig.~\ref{fig:2}), asymptoting in catchments greater than 40,000~km\textsuperscript{2}.Thus, rivers that drain larger catchments integrated more climate variability.
+Following our hypothesis, this climate variability index increases with area (Fig.~\ref{fig:2}), asymptoting in catchments greater than 40,000~km\textsuperscript{2}.
+Thus, rivers that drain larger catchments integrated more climate variability.
 Larger areas contain greater landscape complexity and thus a greater variety of climatic expressions.
@@
 \centering
-\includegraphics[width=.8\linewidth]{Fig2_ClimPort.pdf}
+\includegraphics[width=8.69cm]{Fig2_ClimPort.pdf}
 	\caption{Climate variability index as a function of catchment area.
-The mean trend line (blue) was estimated using robust linear regression and standard error estimates (grey) were calculated using the weighted standard deviation estimate.}
+The mean trend line (blue) was estimated using robust linear regression and standard error estimates (SE) (2$\cdot$SE = grey) were calculated using the weighted standard deviation estimate.}
 \label{fig:2}
@@
 This relationship between watershed size and trend certainty could pull small watershed trend estimates away from zero, thereby creating the appearance of decreasing trend variability among sites as watershed size increases.
-Using a null-hypothesis framework (\textit{Materials and Methods}) we simulated time-series for each site with no underlying trend.
-These simulations were generated by a random-walk process with the observed standard deviation and autocorrelation parameters ($\hat{\upsigma}$, $\hat{\upphi}$) at our sites.
-Using the same generalized least squares model applied to the observed data, we estimated trends for 1000 simulated time series at each site for each response variable (e.g., Fig.~S3).
+Using a null-hypothesis framework (\textit{Materials and Methods}) we simulated time-series for each site with the observed standard deviation and autocorrelation parameters ($\hat{\sigma}$, $\hat{\upphi}$) but with no underlying trend.
+Using the same generalized least squares model applied to the observed data, we estimated trends for 1000 simulated time series at each site for each response variable.
 This simulation process, a form of parametric bootstrapping, created distributions of null-expectations with which we compared our observed results.
@@
 \centering
-\includegraphics[width=17.8cm]{Fig3_AnnFun.pdf}
+\includegraphics[width=17.8cm, height=10.8cm]{Fig3_AnnFun.pdf}
 	\caption{Annual flow trend attenuation within the Fraser River basin.
@@
 However, there was weak evidence of attenuation in the spring (Fig.~\ref{fig:4}, \textit{SI Appendix}, Figs.~S1,S2).
-Overall, climate portfolios tended to have greater attenuation during stable periods of the year and decreased during seasonal transitions (\textit{SI Appendix}, Fig.~S3).
+Overall, climate portfolios tended to have greater attenuation during stable periods of the year and decreased during seasonal transitions.
 In the spring, the likelihood of network-driven attenuation of maximum flow trends was as low as 25\% (Fig.~\ref{fig:4}).
@@
 
-\begin{figure*}[h]
+\begin{SCfigure*}[\sidecaptionrelwidth][h]
 \centering
-\includegraphics[width=17.8cm]{Fig4_MaxMonth.pdf}
+\includegraphics[width=11.4cm, height=8.5cm]{Fig4_MaxMonth.pdf}
 	\caption{Monthly maximum flow trend attenuation within the Fraser River basin.
@@
 \label{fig:4}
-\end{figure*}
+\end{SCfigure*}
 
@@
 \section*{Discussion}
-Climate change is already causing economic and conservation challenges for river systems worldwide\cite{Palmer:2009,Pecl:2017}.
+Climate change is already causing economic and conservation challenges for river systems worldwide \cite{Palmer:2009,Pecl:2017}.
 For example, sockeye salmon in the Fraser River have seen as much as 80\% pre-spawn mortality in years with later run-off and elevated temperatures, leading managers to reduce or close the fishery (e.g., 2013 and 2015).
-In the absence of climate dampening the impacts of these climate extremes could have been dramatically worse.
+In the absence of climate dampening, the impacts of these climate extremes could have been considerably worse.
 As climate change progresses, river hydrology will continue to shift, further stressing riverine ecosystems and subsequently demanding responsive management.
@@
 We selected this 38-year range because this period maximized the number of gauge stations in the Fraser River basin (\textit{n}=55) operating concurrently over a long enough time period in which climate trends may be observable.
-We interpolated missing data using a linear model in log-space, but of the 13,879 days of data only 23 days were estimated.
+We linearly interpolated missing data in log-space, but of the 13,879 days of data only 23 days were estimated.
 These daily data were smoothed using a five-day rolling average before summarizing into annual and monthly response variables.
 This smoothing technique was used to reduce the influence of erroneous and unusually extreme values.
-Because dams control flow and stabilize flow trends, we removed sites where daily flow exhibited unusually high autocorrelation and low variability relative to other sites of similar size, and were observed downstream of, and atypically close to, a dam (\textit{n}=3).
+Because dams control flow and stabilize flow trends, we removed sites where daily flow exhibited unusually high autocorrelation and low variability relative to other sites of similar size, and sites that were observed downstream of, and atypically close to, a dam (\textit{n}=3).
 The downstream distance of gauge stations from dams was determined using GIS spatial layers available in the \textit{BC Data Catalog}, maintained by the provincial government of British Columbia.
@@
 where $\mathrm{flow}_{s,t}$ represents a flow metric [log(maximum), log(minimum), log(median), logit(day-of-year to half-annual-flow)] at each site ($s$) and time point ($t$),  and $\mathrm{year}_{s,t}$ represents the time in years (e.g., 1970, 1971, \ldots, 2007) with 1988 subtracted to approximately center the predictor.
-Parameters $b_{s}$ (slope) and $a_{s}$ (intercept) represent the estimated mean effect of time on $\mathrm{flow}_{s,t}$ and the estimate of $\mathrm{flow}_{s,t}$ at time zero (i.e., 1988) respectively.
+Parameters $b_{s}$ (slope) and $a_{s}$ (intercept) represent the estimated mean effect of time on $\mathrm{flow}_{s,t}$ and the estimate of $\mathrm{flow}_{s,t}$ at time zero (i.e., 1988), respectively.
 Error ($\epsilon_{s,t}$) of the current time step for a given site was allowed to be correlated with that of the previous time step by $\upphi$ and was assumed to be normally distributed with a variance of $\sigma^{2}$.
@@
 where $\hat{b}_{s}$ represents a flow trend at a given site ($s$) and $\mathrm{area}_{s}$ represents site $s$'s watershed area.
-Fitted $d$ and $c$ parameters represent the mean effect of $\sqrt{\mathrm{watershed area}}$ on flow trends and the mean basin-wide flow trend at a theoretical watershed area of zero, respectively.
+Fitted $d$ and $c$ parameters represent the mean effect of watershed area ($\sqrt{\mathrm{area}}$), on flow trends and the mean basin-wide flow trend at a theoretical watershed area of zero, respectively.
 To quantify attenuation, we modeled the change in trend variability with the change in catchment area using an exponential variance function \cite[p.~211]{Pinheiro:2000}:
@@
 This relationship between watershed size and trend certainty may pull small watershed trend estimates away from zero, thereby creating the appearance of decreasing trend variability among sites as watershed size increases.
-Using a null-hypothesis framework we simulated time-series for each site with no underlying trend.
-These simulations were generated by a random walk process with the observed standard deviation and autocorrelation parameters ($\hat{\upsigma}$, $\hat{\upphi}$) at our sites.
+Using a null-hypothesis framework we simulated time-series with no underlying trend from each sites' fitted model, using the observed standard deviation and autocorrelation parameters ($\hat{\sigma}$, $\hat{\upphi}$).
 Using the same GLS model and AR1 correlation structure as applied to the observed data (eq.
-\ref{eq1}), we estimated trends for 1000 simulated time series at each site for each response variable (e.g., Fig.~S4).
+\ref{eq1}), we estimated trends for 1000 simulated time series at each site for each response variable (e.g., Fig.~S3).
 This simulation process, a form of parametric bootstrapping, created distributions of null-expectations with which we compared our observed results.
@@
 We then applied equation \ref{eq2} to each of the 1000 basin-wide simulations (e.g.
-Fig.~S5), resulting in 1 observed variance exponent parameter ($\hat{\updelta}$) and 1000 simulated $\hat{\updelta}$ for each flow metric.
+Fig.~S4), resulting in 1 observed variance exponent parameter ($\hat{\updelta}$) and 1000 simulated $\hat{\updelta}$ for each flow metric.
 By comparing our observed attenuation with our basin-wide null-model simulations we addressed the potential that more variable flows in smaller catchments contribute to the observed pattern of flow trends as a function of watershed area (Fig.~\ref{fig:3}).
@@
 \subsection*{Dampening summary statistics} We calculated attenuation certainty as the percentage of simulated $\hat{\updelta}$ that were less than the observed $\hat{\updelta}$.
-We calculated attenuation strength as the ratio between the standard deviation at the smallest and largest watersheds as defined by the exponential variance function (i.e., $\sqrt{f(\mathrm{area}_1)} / \sqrt{f(\mathrm{area}_2)}$).
-To compare the degree of attenuation observed with the null-simulated attenuation, we compared this ratio with the same ratio calculated from the null-simulated data (i.e., $\left( \sqrt{f(\mathrm{area}_1)} / \sqrt{f(\mathrm{area}_2)} \right) / \left( \sqrt{f(\mathrm{area}_{\mathrm{null} 1})} / \sqrt{f(\mathrm{area}_{\mathrm{null} 2})} \right)$).
+We calculated attenuation strength as the ratio between the standard deviation at the smallest and largest watersheds as defined by the exponential variance function (i.e., $\mathrm{AS}=\sqrt{f(\mathrm{area}_{s=1})} / \sqrt{f(\mathrm{area}_{s=2})}$).
+To compare the degree of attenuation observed with the null-simulated attenuation, we compared this ratio with the same ratio calculated from the null-simulated data (i.e., $\mathrm{AS} / \mathrm{AS}_{NULL}$).
 Finally, to estimate how the day-of-year to half-annual-flow was changing, we estimated the steepest point of each site's logistic curve and multiplied this slope value by 365 to revert from our 0--1 scale to our original annual range of 0--365.
 To get a decadal rate we multiplied these results by 3.8 (number of decades in 38 years).
-We estimated the steepest part of the logistic curve using the *divide by four* rule[@Gelman:2008] (dividing the coefficient by four equals the first derivitive of the logistic curve at its steepest point).
+We estimated the steepest part of the logistic curve using the \textit{divide by four} rule \cite{Gelman:2008} (dividing the coefficient by four equals the first derivitive of the logistic curve at its steepest point).
 
@@
 Climate variables included the mean annual temperature (MAT), and precipitation (MAP), extreme minimum (EMT) and maximum (EXT) temperature and precipitation as snow (PAS).
-Climate variable trends were calculated at each grid point using a modified eq.~\ref{eq1}, where time points were no longer considered and our response variable becomes $\mathrm{climate}_{s}$.
+Climate variable trends were calculated at each grid point using eq.~\ref{eq1} but where our response variable becomes $\mathrm{climate}_{s,t}$.
 
-\subsection*{Code and Data Repository} All flow data munging and analysis was done using R v3.3.2 (\url{www.r-project.org/}).
-Spatial analysis of watersheds were done using GDAL/OGR v1.11.5, and Whitebox GAT v3.4 (\url{www.uoguelph.ca/~hydrogeo/Whitebox/}) with visualization help from QGIS v2.18.2 (\url{www.qgis.org/}).
-Climate data were gathered using ClimateWNA v5.4 (\url{www.climatewna.com}) and summary statitics were calculated using R and the \textit{rasterstats} module (\url{https://github.com/perrygeo/python-rasterstats.git}) in Python v2.7.
-All code can be found on \url{www.github.com/} at \url{github.com/kchezik/River-Network-Flow-Trends.git}.
+\subsection*{Code and Data Repository} All flow data analysis was done using R v3.3.2 (\url{www.r-project.org/}).
+Spatial analysis of watersheds was done using GDAL/OGR v1.11.5, and Whitebox GAT v3.4 \cite{Lindsay:2016} with visualization help from QGIS v2.18.2 (\url{www.qgis.org/}).
+Climate data were gathered using ClimateWNA v5.4 \cite{Wang:2016} and summary statitics were calculated using R and the \textit{rasterstats} module (\url{https://github.com/perrygeo/python-rasterstats.git}) in Python v2.7.
+All code can be found on \url{www.github.com/} at \url{http://github.com/kchezik/River-Network-Flow-Trends.git}.
 Raw HYDAT data can be obtained via Environment Canada (\url{http://collaboration.cmc.ec.gc.ca/cmc/hydrometrics/www/}) or a simplified version can be found on GitHub.

### 2017-04-11 drafts/PNAS/Chezik_PNAS_SI.tex
Organized
Removed Lingering Defunct Files and Cleaned Up For Reproducibility.

@@
 \section*{Supplementary Information}
-The following supplementary material includes 5 additional figures.
+The following supplementary material includes 4 additional figures.
 Figures \ref{fig:S1} and \ref{fig:S2} are similar to figure 4 in the main text where monthly trends in flow are shown for the basin along with trend attenuation results.
-In these supplementary figures we highlight minimum and median flow trends rather than maximum flow trends.
-Figure \ref{fig:S3} demonstrates the effect of the seasonal transition of spring and fall on the climate portfolio index values within and among Fraser River sub-basins.
-Figures \ref{fig:S4} and \ref{fig:S5} are example plots demonstrating a single null-model iteration for the median flow metric.
+These supplementary figures highlight minimum and median seasonal flow trends rather than maximum seasonal flow trends.
+Figures \ref{fig:S3} and \ref{fig:S4} are example plots demonstrating a single null-model iteration for the median flow metric.
 
@@
 \begin{figure*}
-\centering
-\includegraphics[width=11.4cm]{FigS3_SeasonClimPort.pdf}
-	\caption{Climate portfolio index by month at 55 sites within  the Fraser River basin.
-Lines are loess smoothers with a smoothing span of 1 to reduce noise and highlight general season shifts in climate portfolio among and across Fraser River sub-basins.}
-	\label{fig:S3}
-\end{figure*}
-
-\begin{figure*}
-	\includegraphics[width=17.8cm]{FigS4_SiteSim.pdf}
+	\includegraphics[width=17.8cm]{FigS3_SiteSim.pdf}
 	\caption{Example site specific flow simulations.
@@
 See supplemental figure \ref{fig:S5} for resultant basin-wide simulation.}
-	\label{fig:S4}
+	\label{fig:S3}
 \end{figure*}
@@
 \centering
-\includegraphics[width=11.4cm]{FigS5_BasinSim.pdf}
+\includegraphics[width=11.4cm]{FigS4_BasinSim.pdf}
 	\caption{Example flow simulation results.
@@
 This is a single example result of the 1000 null-model simulations produced from the data in supplemental figure \ref{fig:S3}.}
-	\label{fig:S5}
+	\label{fig:S4}
 \end{figure*}

### 2017-04-11 Code_03 Null-Sim_ARIMA-Mod.R

@@ -1,66 +1,0 @@
-# test gls model and arima simulation
-# prove that they are interchangeable
-# answer: yes.
-#
-# gls() with arima residuals is interchangeable with fitted arima model
-# with same order and drift
-
-set.seed(1)
-library(forecast) # for Arima()
-library(dplyr)
-
-out <- plyr::rdply(200, function() {
-
-  y <- as.numeric(arima.sim(n = 60L, model = list(model = c(1, 0, 0),
-        ar = 0.3), sd = 0.2))
-
-  m1_arima <- Arima(y, order = c(1L, 0L, 0L), include.drift = TRUE)
-  phi_arima <- coef(m1_arima)[[1]]
-  intercept_arima <- coef(m1_arima)[[2]]
-  drift_arima <- coef(m1_arima)[[3]]
-  sigma_arima <- sqrt(m1_arima$sigma2)
-
-  m1 <- gls(y~seq_along(y), correlation = corAR1())
-  phi <- coef(m1$model[[1]], unconstrained = FALSE)[[1]]
-  sigma <- m1$sigma
-  intercept <- coef(m1)[[1]]
-  slope <- coef(m1)[[2]]
-
-  ysim <- arima.sim(n = length(y), model = list(order = c(1, 0, 0),
-      ar = phi), sd = sigma) %>% as.numeric()
-
-  m2 <- gls(ysim~seq_along(ysim), correlation = corAR1())
-  phi_sim <- coef(m2$model[[1]], unconstrained = FALSE)[[1]]
-  sigma_sim <- m2$sigma
-  intercept_sim <- coef(m2)[[1]]
-
-  data.frame(phi, sigma, phi_sim, sigma_sim, phi_arima, intercept_arima,
-    sigma_arima, drift_arima, slope, intercept)
-})
-
-out <- out %>% mutate(phi_re = (phi_sim - phi) / phi,
-  sigma_re = (sigma_sim - sigma) / sigma)
-
-par(mfrow = c(3, 2))
-hist(out$phi)
-abline(v = 0.3, col = "red")
-hist(out$sigma)
-abline(v = 0.2, col = "red")
-hist(out$phi_sim)
-abline(v = 0.3, col = "red")
-hist(out$sigma_sim)
-abline(v = 0.2, col = "red")
-hist(out$phi_re)
-abline(v = 0, col = "red")
-hist(out$sigma_re)
-abline(v = 0, col = "red")
-
-par(mfrow = c(2, 2))
[…]

### 2017-04-11 Code_03_Post_Analysis_Null_Test.R

@@ -1,254 +1,0 @@
-library(tidyverse); library(stringr);library(attenPlot)
-#Load/Organize Annual Data
-load("03_Data_Annual.RData")
-logit = function(p){log(p/(1-p))}
-Y.Data = Y.Data %>% group_by(Station.ID) %>% mutate(med.log.sd = scale(Median.F, center = F), max.log.sd = scale(Max.F, center = F), min.log.sd = scale(Min.F, center = F), DOY2.logit = logit(DOY2/365), Year.Center = Year-1988)
-#Load/Orgnize Montly Data
-load("03_Data_Monthly.RData")
-M.Data = M.Data %>% group_by(Station.ID) %>% mutate(med.log.sd = scale(median.M.flow, center = F), max.log.sd = scale(max.M.flow, center = F), min.log.sd = scale(min.M.flow, center = F), Year.Center = Year-1988)
-#Scaling the data by site has no impact on the results, rather it only stands to reduce the likelyhood of computational issues during model bootstraping.
-
-#Incorporate the Climate Data
-setwd("~/sfuvault/Simon_Fraser_University/PhD_Research/Projects/Data/Original_Data/GIS_Data/BC_Climate_Data/ClimateTrends/zonal_stats")
-vars = unlist(lapply(strsplit(dir(), "-"), function(x) x[1]))
-AnnualVars = vars[which(nchar(vars) == 3)]
-MonthlyVars = vars[which(nchar(vars) > 3 & nchar(vars)<7)]
-#Annual
-for(i in AnnualVars){
-	labels = c("Station.ID","count","min","mean","max",paste(i,".sd",sep = ""),"P25","P75")
-	cDat = read_csv(paste(i,"-stats.csv", sep = ""), col_names =  labels) %>% select(1,6)
-	Y.Data = dplyr::left_join(Y.Data, cDat, by = "Station.ID")
-}
-Y.Data = Y.Data %>% ungroup(); rm(cDat, AnnualVars, i, logit, vars, labels)
-
-#Monthly
-vars = unique(str_extract(string = MonthlyVars, pattern = "([^0-9]){3,}"))
-M.Data = lapply(vars,function(x){
-	clim = MonthlyVars[grep(pattern = x, MonthlyVars)] %>% paste(.,"-stats.csv", sep = "") %>%
-		lapply(., function(y){
[…]

### 2017-04-11 Code_03b Post_Analysis_Null_Test.R

@@ -1,0 +1,254 @@
+library(tidyverse); library(stringr);library(attenPlot)
+#Load/Organize Annual Data
+load("03_Data_Annual.RData")
+logit = function(p){log(p/(1-p))}
+Y.Data = Y.Data %>% group_by(Station.ID) %>% mutate(med.log.sd = scale(Median.F, center = F), max.log.sd = scale(Max.F, center = F), min.log.sd = scale(Min.F, center = F), DOY2.logit = logit(DOY2/365), Year.Center = Year-1988)
+#Load/Orgnize Montly Data
+load("03_Data_Monthly.RData")
+M.Data = M.Data %>% group_by(Station.ID) %>% mutate(med.log.sd = scale(median.M.flow, center = F), max.log.sd = scale(max.M.flow, center = F), min.log.sd = scale(min.M.flow, center = F), Year.Center = Year-1988)
+#Scaling the data by site has no impact on the results, rather it only stands to reduce the likelyhood of computational issues during model bootstraping.
+
+#Incorporate the Climate Data
+setwd("~/sfuvault/Simon_Fraser_University/PhD_Research/Projects/Data/Original_Data/GIS_Data/BC_Climate_Data/ClimateTrends/zonal_stats")
+vars = unlist(lapply(strsplit(dir(), "-"), function(x) x[1]))
+AnnualVars = vars[which(nchar(vars) == 3)]
+MonthlyVars = vars[which(nchar(vars) > 3 & nchar(vars)<7)]
+#Annual
+for(i in AnnualVars){
+	labels = c("Station.ID","count","min","mean","max",paste(i,".sd",sep = ""),"P25","P75")
+	cDat = read_csv(paste(i,"-stats.csv", sep = ""), col_names =  labels) %>% select(1,6)
+	Y.Data = dplyr::left_join(Y.Data, cDat, by = "Station.ID")
+}
+Y.Data = Y.Data %>% ungroup(); rm(cDat, AnnualVars, i, logit, vars, labels)
+
+#Monthly
+vars = unique(str_extract(string = MonthlyVars, pattern = "([^0-9]){3,}"))
+M.Data = lapply(vars,function(x){
+	clim = MonthlyVars[grep(pattern = x, MonthlyVars)] %>% paste(.,"-stats.csv", sep = "") %>%
+		lapply(., function(y){
[…]

### 2017-04-11 Code_04 Annual_Funnel_Plot.R

@@ -1,0 +1,106 @@
+#########################################################################################################
+################## Plot Annual Flow Trends Against Area and Display Climate Index Data ##################
+#########################################################################################################
+
+library(attenPlot) #Funnel and Density plot function library.
+
+#Density Plot Wrapper
+densityP = function(df, response, type = "exp", d.axis=F, LW=L.Width, ALS=Axis.Lab.Size, LL,	
+										shade = "99", col1, col2, col3, col4, vline, pane){
+	
+	gather.density.data(df, response, type)
+	d_dat = dplyr::tbl_df(data.frame(var = d$x, dens = d$y))
+	
+	par(mar = c(0,0.3,1.5,0.5), mgp = c(2,0.60,0))
+	plot(dens~var, data = d_dat, type = "n", xlab = "", ylab = "", las = 1,
+			 bty = "n", yaxt = "n", xaxt = "n", axes = F, xaxs = "i", xlim = c(-12,5))
+	
+	if(d.axis == TRUE){
+		axis.density.draw(type = type, ALS = ALS+0.2, LL = LL, LW = LW, annual = T)
+	} else axis(1, lwd = LW, labels = F, tck = 0)
+	
+	plygnPrptns(d_dat, type, annual = T, varR, varS)
+	
+	for(i in 1:2) abline(v = varR, col = paste(col2, shade, sep=""), lwd = vline)
+	text(4, max(d_dat$dens)-0.01, labels = pane, pos = 2, cex = ALS+0.2)
+}
+
+
+#Funnel Plot Wrapper.
+funnel.wrapper = function(dat, response, yaxis.DOY = F, xAxis = T, pane, axis.Ly=NULL, d.axis = F, iter=1000){
+	
+	############################################ Prepare Data ###############################################
+	#Control Panel
+	L.Width = 0.5; Axis.Lab.Size = 0.7; LabLoc = 2.1; vLine = 2.5
+	#Colour Control Panel
+	col1="#F2AD00"; col2="#67A9CF"; col3="#F98400"; col4="#FF0000"; lcol="#969696"
+	#Libraries to load.
+	library(tidyverse);library(wesanderson);library(RColorBrewer)
+	#Cleanup and arrange observed flow and climate trend data.
+	flow.trends = arrange.data(dat,response)
+	#Create a colour gradient for the climate index values.
+	cols = color.gradient(flow.trends$std.clim)
+	
[…]

### 2017-04-11 Code_04 Flow_Time_Plot.R

@@ -1,77 +1,0 @@
-#library(dplyr)
-load("03_Data_Annual.RData")
-
-Y.Data = Y.Data %>%
-	group_by(Station.ID) %>%
-	mutate(Med.sc = as.vector(scale(Median.F, center = F)),
-				 Min.sc = as.vector(scale(Min.F, center = F)),
-				 Max.sc = as.vector(scale(Max.F, center = F)))
-
-Flow.Time = function(time, response, weight, group, alpha = 0.25, seq.inc = 0.083, Te.cex = 0.7, xlab= "time", ylab ="response"){
-	#vector scaled 0-1 for function 
-	col = ((weight-min(weight))/(diff(range(weight))))^0.3
-	
-	#color ramp function
-	library(RColorBrewer)
-	Blues = brewer.pal(9, "Blues")#; RdYlBu = RdYlBu[-c(5,6,7)]
-	FUN = colorRamp(Blues[c(2:7)], bias=1)
-	
-	#apply function
-	cols = FUN(col)
-	cols = rgb(cols, maxColorValue=256)
-	cols = paste(cols, "99", sep = "")
-	
-	#Determination of point/line size and legend point data.
-	Area.width = ((((weight-min(weight))/(diff(range(weight))))^0.3)+0.01)*8 #Line width.
-	Area.size = (((weight-min(weight))/(diff(range(weight))))^0.3)*3 #Point size.
-	Area.values = c(500,1500,10000,100000,200000) #Area values of interest for the legends.
-	Legend.Area.size = (((Area.values-min(weight))/(diff(range(weight))))^0.3)*3 
-	#Equivalent size given legend point size values.
-	Legend.Area.size.lines = (((Area.values-min(weight))/(diff(range(weight))))^0.3)*8
-	#Equivalent size given legend line size values.
-	Legend.x = seq(1999,2006,length.out = 5) #x values for line/point legend.
-	Legend.y = rep(max(pretty(response)), 5) #y values for line/point legend.
-	Legend.x.col = seq(1999,2006,length.out = 20000) #x values for color legend.
-	Legend.y.col = rep(max(pretty(response)), 20000) #y values for color legend.
-	
-	#color legend point data scaled from 0-1.
-	legend.col = (Legend.x.col-min(Legend.x.col))/(diff(range(Legend.x.col)))
[…]

### 2017-04-11 Code_05 Monthly_Plot.R

@@ -1,0 +1,150 @@
+#########################################################################################################
+################## Plot Monthly Flow Trends Against Area and Display Climate Index Data #################
+#########################################################################################################
+
+library(attenPlot)#Funnel and Density plot function library.
+
+#Density Plot Wrapper
+density_plot = function(data, response, type, xlimit){
+	
+	############################################ Prepare Data ###############################################
+	#Libraries to load.
+	library(tidyverse);library(wesanderson);library(RColorBrewer);library(lubridate)
+	#Control Panel
+	L.Width = 0.5; Axis.Lab.Size = 0.7; Lab.Loc = 2.1; vLine = 2
+	#Colour Control Panel
+	col1 = if_else(type == "int","#000000","#F2AD00"); col2= if_else(type=="int","#525252","#67A9CF")
+	col3="#F98400"; col4="#FF0000"; lcol="#969696"; shade = "99"
+	#Determine if we are plotting intercept (int) or exponent (exp) data and set boundaries accordingly.
+	if(type == "int") par(oma = c(3,1.5,0.2,0.2), mar = c(0,0,0.1,0.3), mgp = c(3,0.5,0), family = "serif")
+	else par(mar = c(0,0.3,0.1,0), family = "serif")
+	#Create list of month labels to iterate through.
+	Months = as.character(lubridate::month(as.numeric(names(data)), label = T)); count = 1
+	
+	plyr::l_ply(data,function(x){
+		######## Gather and arrange observed and simulated variance and intercept parameter estimates. ########
+		gather.density.data(x,response, type)
+		d_dat = dplyr::tbl_df(data.frame(var = d$x, dens = d$y))
+		
+		########################################### Plot Raw Data ##############################################
+		plot(dens~var, data = d_dat, type="l", xlab = "", ylab = "", las = 1,
+				 bty = "n", yaxt = "n", xaxt = "n", xaxs = "i", xlim = xlimit, col = col1,
+				 lwd = L.Width)
[…]

### 2017-04-11 Code_06 Annual_Funnel_Plot.R

@@ -1,106 +1,0 @@
-#########################################################################################################
-################## Plot Annual Flow Trends Against Area and Display Climate Index Data ##################
-#########################################################################################################
-
-library(attenPlot) #Funnel and Density plot function library.
-
-#Density Plot Wrapper
-densityP = function(df, response, type = "exp", d.axis=F, LW=L.Width, ALS=Axis.Lab.Size, LL,	
-										shade = "99", col1, col2, col3, col4, vline, pane){
-	
-	gather.density.data(df, response, type)
-	d_dat = dplyr::tbl_df(data.frame(var = d$x, dens = d$y))
-	
-	par(mar = c(0,0.3,1.5,0.5), mgp = c(2,0.60,0))
-	plot(dens~var, data = d_dat, type = "n", xlab = "", ylab = "", las = 1,
-			 bty = "n", yaxt = "n", xaxt = "n", axes = F, xaxs = "i", xlim = c(-12,5))
-	
-	if(d.axis == TRUE){
-		axis.density.draw(type = type, ALS = ALS, LL = LL, LW = LW, annual = T)
-	} else axis(1, lwd = LW, labels = F, tck = 0)
-	
-	plygnPrptns(d_dat, type, annual = T, varR, varS)
-	
-	for(i in 1:2) abline(v = varR, col = paste(col2, shade, sep=""), lwd = vline)
-	text(4, max(d_dat$dens), labels = pane, pos = 2, cex = ALS)
-}
-
-
-#Funnel Plot Wrapper.
-funnel.wrapper = function(dat, response, yaxis.DOY = F, xAxis = T, pane, axis.Ly=NULL, d.axis = F, iter=1000){
-	
-	############################################ Prepare Data ###############################################
-	#Control Panel
-	L.Width = 1; Axis.Lab.Size = 1.1; LabLoc = 3.1; vLine = 4
-	#Colour Control Panel
-	col1="#F2AD00"; col2="#67A9CF"; col3="#F98400"; col4="#FF0000"; lcol="#969696"
-	#Libraries to load.
-	library(tidyverse);library(wesanderson);library(RColorBrewer)
-	#Cleanup and arrange observed flow and climate trend data.
-	flow.trends = arrange.data(dat,response)
-	#Create a colour gradient for the climate index values.
-	cols = color.gradient(flow.trends$std.clim)
-	
[…]

### 2017-04-11 Code_07 Area_ClimatePortfolioPlot.R

@@ -1,0 +1,50 @@
+load("out_med.RData")
+library(MASS);library(tidyverse); library(attenPlot);
+df = out_med$Area$real_slopes
+df = df %>% mutate(logArea = log(Area), sqrtArea = sqrt(Area), cols = "#000000", loc = "interior")
+
+#Model to add line to plot.
+remove = c("08MH006","08MH076")
+#"08MH090","08MH056","08LG016","08LG048","08MH103","08MH001","08MH016","08MH029"
+df$cols[which(df$Station.ID%in%remove)] = "#969696"
+df$loc[which(df$Station.ID%in%remove)] = "coastal"
+ggplot(df,aes(exp(std.clim), slope, color = cols)) + geom_point()
+
+#Look at each climate variable and the index against area.
+gather(data = df, key = "climateIndex", value = "IndexValue", emt.sd, ext.sd, map.sd, mat.sd, pas.sd, std.clim) %>% 
+	ggplot(., aes(logArea, IndexValue, label = Station.ID)) + geom_text(check_overlap = T) + geom_point(aes(color = cols)) + facet_wrap(~climateIndex, scales = "free_y") + theme(legend.position="none")
+
+#Model with weighted regresssion to downweight outliers/leverage points.
+mod1 = rlm(std.clim~logArea, data = df)
+se = summary(mod1)[[4]][4]
+int = summary(mod1)[[4]][1]
+slope = summary(mod1)[[4]][2]
+
+sim = log(seq(100,250000,by = 100))
+preds = int + sim*slope
+se_upper = preds + (2*se)
+se_lower = preds - (2*se)
+preds = data.frame(area = exp(sim), preds = preds, se_upper = se_upper, se_lower = se_lower)
+#par(mar = c(0,0.3,1.5,0.5), mgp = c(2,0.60,0))
+wdth = 8.7/2.54; hght = wdth*0.6
+pdf("Fig2_ClimPort.pdf", width = wdth, height = hght)
+
+LW = 1; ALS = 0.7; LL = 1.2
+par(oma = c(2,0,0,0), mar = c(0,1.8,0.2,0), family = "serif", bg = "white",fg = "white", mgp = c(2,0.2,0))
+plot(std.clim~sqrtArea, data = df, xlab = "", ylab = "", axes = F, xlim = c(0,sqrt(250000)), ylim = c(0,4))
+par(fg = "black")
[…]

### 2017-04-11 Code_07 Monthly_Plot.R

@@ -1,151 +1,0 @@
-#########################################################################################################
-################## Plot Monthly Flow Trends Against Area and Display Climate Index Data #################
-#########################################################################################################
-
-library(attenPlot)#Funnel and Density plot function library.
-
-#Density Plot Wrapper
-density_plot = function(data, response, type, xlimit){
-	
-	############################################ Prepare Data ###############################################
-	#Libraries to load.
-	library(tidyverse);library(wesanderson);library(RColorBrewer);library(lubridate)
-	#Control Panel
-	L.Width = 1; Axis.Lab.Size = 1.1; Lab.Loc = 3.1; vLine = 4
-	#Colour Control Panel
-	col1 = if_else(type == "int","#000000","#F2AD00"); col2= if_else(type=="int","#525252","#67A9CF")
-	col3="#F98400"; col4="#FF0000"; lcol="#969696"; shade = "99"
-	#Determine if we are plotting intercept (int) or exponent (exp) data and set boundaries accordingly.
-	if(type == "int") par(oma = c(4,3.5,0.2,0.2), mar = c(0,0,0.2,0.5), family = "serif")
-	else par(mar = c(0,0.5,0.2,0), family = "serif")
-	#Create list of month labels to iterate through.
-	Months = as.character(lubridate::month(as.numeric(names(data)), label = T)); count = 1
-	
-	
-	plyr::l_ply(data,function(x){
-		######## Gather and arrange observed and simulated variance and intercept parameter estimates. ########
-		gather.density.data(x,response, type)
-		d_dat = dplyr::tbl_df(data.frame(var = d$x, dens = d$y))
-		
-		########################################### Plot Raw Data ##############################################
-		plot(dens~var, data = d_dat, type="l", xlab = "", ylab = "", las = 1,
-				 bty = "n", yaxt = "n", xaxt = "n", xaxs = "i", xlim = xlimit, col = col1,
-				 lwd = L.Width)
-		#Add month label if plotting intercept (i.e., far left on page) data.
[…]

### 2017-04-11 Code_08 Area_ClimatePortfolioPlot.R

@@ -1,49 +1,0 @@
-load("out_med.RData")
-library(MASS);library(tidyverse); library(attenPlot);
-df = out_med$Area$real_slopes
-df = df %>% mutate(logArea = log(Area), sqrtArea = sqrt(Area), cols = "#000000", loc = "interior")
-
-#Model to add line to plot.
-remove = c("08MH006","08MH076")
-#"08MH090","08MH056","08LG016","08LG048","08MH103","08MH001","08MH016","08MH029"
-df$cols[which(df$Station.ID%in%remove)] = "#969696"
-df$loc[which(df$Station.ID%in%remove)] = "coastal"
-ggplot(df,aes(exp(std.clim), slope, color = cols)) + geom_point()
-
-#Look at each climate variable and the index against area.
-gather(data = df, key = "climateIndex", value = "IndexValue", emt.sd, ext.sd, map.sd, mat.sd, pas.sd, std.clim) %>% 
-	ggplot(., aes(logArea, IndexValue, label = Station.ID)) + geom_text(check_overlap = T) + geom_point(aes(color = cols)) + facet_wrap(~climateIndex, scales = "free_y") + theme(legend.position="none")
-
-#Model with weighted regresssion to downweight outliers/leverage points.
-mod1 = rlm(std.clim~logArea, data = df)
-se = summary(mod1)[[4]][4]
-int = summary(mod1)[[4]][1]
-slope = summary(mod1)[[4]][2]
-
-sim = log(seq(100,250000,by = 100))
-preds = int + sim*slope
-se_upper = preds + (2*se)
-se_lower = preds - (2*se)
-preds = data.frame(area = exp(sim), preds = preds, se_upper = se_upper, se_lower = se_lower)
-
-pdf("Fig2_ClimatePortfolio.pdf", width = 7.5, height = 5.25)
-
-LW = 1; ALS = 1.1; LL = 3.1
-par(oma = c(4,0,0,0), mar = c(0,5,0.2,0), family = "serif", bg = "white",fg = "white", mgp = c(2,0.7,0))
-plot(std.clim~sqrtArea, data = df, xlab = "", ylab = "", axes = F, xlim = c(0,sqrt(250000)), ylim = c(0,4))
-par(fg = "black")
-axis(1, lwd = LW, cex.axis = ALS, outer = T, hadj = 1,
[…]

### 2017-04-11 Code_08 Supplementary_Plots.R

@@ -1,12 +1,12 @@
 #supplementary plots
-library(dplyr)
+library(tidyverse);library(attenPlot)
 load("out_med.RData")
 load("03_Data_Annual.RData")
 logit = function(p){log(p/(1-p))}
 Y.Data = plyr::ddply(Y.Data,"Station.ID",plyr::mutate, med.log.sd = scale(Median.F, center = F), max.log.sd = scale(Max.F, center = F), min.log.sd = scale(Min.F, center = F), DOY2.logit = logit(DOY2/365), Year.Center = Year-1988)
 
 
-dat_sim = out_med$example_ts; names(dat_sim)[2] = "Year.Center"
+dat_sim = out_med$Area$example_ts; names(dat_sim)[2] = "Year.Center"
 dat = dplyr::left_join(dat_sim, Y.Data, by = c("Station.ID", "Year.Center"))
 Adjuster = Y.Data %>%
 	group_by(Station.ID) %>%
@@ -25,7 +25,7 @@ ggplot(dat, aes(Year, med_sim_unscale)) +
 	theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm")) +
 	labs(x = "Year", y = expression("Simulated & Observed Annual Median-Flow (m"^3%.%"sec"^-1*")"%.%"year"^-1))
 
-ggsave("FigS3_site_simulation.pdf", width = 11, height = 8.5)
+ggsave("FigS3_SiteSim.pdf", width = 11, height = 8.5)
 
 #clean up NA's in simulated data where no observed data existed.
 dat$Area = apply(dat, 1, function(x){
@@ -58,13 +58,18 @@ m = gls(slope~sqrt(Area), data = sim_out, weights = varExp(form= ~sqrt(Area)/1e3
 							control = glsControl(maxIter = 1000L, msMaxIter = 1000L))
 varexp <- m$model[[1]][[1]]; sigma <- m$sigma; intercept <- coef(m)[[1]]; slope = m$coefficients[[2]]; sim_varexp = data.frame(varexp, sigma, intercept, slope)
 
-var_upper = out_med$real_varexp$intercept + out_med$real_varexp$slope*(sqrt(sim_out$Area) - mean(sqrt(sim_out$Area))) + 1.96 * sqrt(out_med$real_varexp$sigma^2 * exp(2*(sqrt(sim_out$Area)/1e3)*sim_varexp$varexp))
[…]

### 2017-04-11 Shiny_App1/app.R

@@ -1,109 +1,0 @@
-library(shiny) #Call shiny library
-
-#######################################
-#### Initialize the User Interface ####
-#######################################
-
-ui <- fluidPage(
-	includeCSS("style.css"),
-	fluidRow(
-		column(2),
-		column(4,tags$h4("Fraser River Annual Flow", align = "center"), style = "height: 70px; border-right: 1px solid #EEEEEE; top: 25px; position: relative"),
-		column(4, sliderInput(inputId = "num",
-													label = "Smoother Period (Years)",
-													value = 10, min = 3, max = 15, ticks = F)
-					 , align = "center", offset = 1, style = "font-family: 'Raleway'; postion: relative; top: 30px"),
-		column(2)
-	),
-	tags$hr(),
-	fluidRow(
-		column(12, plotOutput("plot"), align = "center")
-	)
-)
-
-#Load Data
-load("app_data.RData")
-
-#Set initial parameters
-time = Y.Data$Year; response = Y.Data$Med.sc; weight = Y.Data$Area; group = Y.Data$Station.ID; xlab = "Year"; ylab = expression("Scaled Median Flow (m"^3%.%"sec"^-1*")"%.%"year"^-1); seq.inc = 0.1; Te.cex = 1.3
-
-##################################
-#### Colour and Scale of Data ####
-##################################
-
-#vector scaled 0-1 for function 
-col = ((weight-min(weight))/(diff(range(weight))))^0.3
-
-#color ramp function
-library(RColorBrewer)
-Blues = brewer.pal(9, "Blues")#; RdYlBu = RdYlBu[-c(5,6,7)]
-FUN = colorRamp(Blues[c(2:7)], bias=1)
-
-#apply function
-cols = FUN(col)
-cols = rgb(cols, maxColorValue=256)
-cols = paste(cols, "99", sep = "")
-
-#####################################
-#### Legend Placement and Colour ####
-#####################################
-
-#Determination of point/line size and legend point data.
-Area.width = ((((weight-min(weight))/(diff(range(weight))))^0.3)+0.01)*8 #Line width.
-Area.size = (((weight-min(weight))/(diff(range(weight))))^0.3)*3 #Point size.
-Area.values = c(500,1500,10000,100000,200000) #Area values of interest for the legends.
[…]

### 2017-04-11 attenPlot/R/axis_Draw.R

@@ -12,12 +12,12 @@
 #' @export
 #' @examples
 #' axis.Draw(LW = 1, ALS = 1, ALY = "Y-Axis", LL = 1, YLim = ylimits, YLab = labels, xAxis = T)
-axis.Draw = function(LW, ALS, ALY, LL, YLim, YLab, xAxis = F){
+axis.Draw = function(LW, ALS, ALY, LL, YLim, YLab, xAxis = F, ALabS=ALS){
 	if(xAxis == T){
-		axis(1, lwd = LW, cex.axis = ALS, outer = T, hadj = 1,
+		axis(1, lwd = LW, cex.axis = ALS+0.2, outer = T, hadj = 1,
 				 labels = c("0","10000","40000","90000","160000","250000"), at = c(0,100,200,300,400,500))
-		mtext(expression("Area (km"^2*")"), side = 1, line = LL, cex = ALS) # x-axis label.
+		mtext(expression("Area (km"^2*")"), side = 1, line = LL, cex = ALabS) # x-axis label.
 	} else axis(1, lwd = LW, tick = TRUE, labels = FALSE, tck = 0)
-	axis(2, lwd = LW, cex.axis = ALS, at = YLim, labels = YLab, las = 1)
+	axis(2, lwd = LW, cex.axis = ALS+0.2, at = YLim, labels = YLab, las = 1)
 	mtext(ALY, side=2, line=LL, cex = ALS, las = 0)  #y-axis label.
 }

### 2017-04-11 attenPlot/R/plygnPrptns.R

@@ -28,9 +28,9 @@ plygnPrptns = function(df, type, varR, varS, annual = F, col1="#F2AD00", col2 = 
 		prop = round(sum(varR<varS)/length(varS),2)
 		if(annual == T) {
 			text(2,max(df$dens)/2,
-					 labels = as.character(prop), cex = 1.1)
-			text(-8,max(df$dens)/2,
-					 labels = as.character(round(1-prop,2)), cex = 1.1)
-		} else mtext(prop, side = 3, adj = 0.05, line = -1.3, cex = 0.75)
+					 labels = as.character(prop), cex = 0.8)
+			text(-8.25,max(df$dens)/2,
+					 labels = as.character(round(1-prop,2)), cex = 0.8)
+		} else mtext(prop, side = 3, adj = 0.05, line = -1.3, cex = 0.5)
 	}
 }

### 2017-04-11 attenPlot/R/points_SE.R

@@ -10,9 +10,9 @@
 #' @export
 #' @examples
 #' points.SE(df, lcol = "#969696", cols = cols, LW = 1, shade = "99")
-pointsSE = function(df, lcol ="#969696", cols, LW, shade = "99"){
+pointsSE = function(df, lcol ="#969696", cols, LW, shade = "99", ptsize = 1){
 	segments(df$sqrt.resp, df$slope+df$se, df$sqrt.resp, df$slope-df$se,
 					 col = paste(lcol, shade, sep = ""), lwd = LW)
 	segments(0,0,500,0, lty = 2, lwd = LW)
-	points(df$sqrt.resp, df$slope, col = cols, pch = 16, cex = 1.5)
+	points(df$sqrt.resp, df$slope, col = cols, pch = 16, cex = ptsize)
 }

### 2017-04-11 attenPlot/R/screenDivide.R

@@ -6,7 +6,7 @@
 #' @export
 #' @examples
 #' zero_one()
-monthlyScreen = function(){layout(cbind(c(1:12),c(1:12),c(1:12),c(1:12),c(13:24),c(13:24),c(13:24),c(13:24),c(rep(25,3),rep(26,3),rep(27,3),rep(28,3)),c(rep(25,3),rep(26,3),rep(27,3),rep(28,3)),c(rep(25,3),rep(26,3),rep(27,3),rep(28,3)),c(rep(25,3),rep(26,3),rep(27,3),rep(28,3)),c(rep(25,3),rep(26,3),rep(27,3),rep(28,3))))}
+monthlyScreen = function(IntExp = 4, Atten = 5){layout(cbind(matrix(data = c(1:12), nrow = 12, ncol = IntExp),matrix(data = c(13:24), nrow = 12, ncol = IntExp),matrix(c(rep(25,3),rep(26,3),rep(27,3),rep(28,3)), nrow = 12, ncol = Atten)))}
 
 #' Break Up Canvas
 #'

### 2017-04-11 test.R

@@ -1,11 +1,0 @@
-library(attenPlot); library(tidyverse)
-n = exp(seq(log(35),log(250000), by = 0.2))
-index.sd = -1.3575 + log(n)*0.3869
-plot(index.sd~sqrt(n))
-df = expand.grid(n = n, sd.samp = c(1:1000)) %>% left_join(.,data.frame(n = n, index.sd = index.sd))
-
-df = df %>% mutate(sd.sim = apply(., 1, function(x){
-		sd(rnorm(n = x[["n"]], mean = 10000, sd = x[["index.sd"]]))
-	}))
-
-ggplot(df,aes(log(n),sd.sim)) + geom_point() + geom_abline(intercept = 1) + geom_line(aes(log(n),index.sd), color = "blue")

### 2017-04-11 tmb-example.R

@@ -1,109 +1,0 @@
-library(TMB)
-compile("varianceFunction.cpp")
-dyn.load(dynlib("varianceFunction"))
-
-set.seed(1234)
-b0 <- 0.5
-b1 <- 0.3
-b2 <- 0.1
-sigma0 <- 0.1
-sigma1 <- -0.2
-sigma2 <- 0.3
-N <- 2000
-s1 <- sample(seq(1, 10), N, replace = TRUE)
-s2 <- sample(seq(1, 10), N, replace = TRUE)
-
-# This is the way that nlme::varExp() has the function set up:
-# var * exp(2 * t * v)
-# This one is set up a little bit differently but has the same general shape:
-u <- rnorm(N, b0 + b1 * s1 + b2 * s2, sqrt(exp(sigma0 + sigma1 * s1 + sigma2 * s2)))
-# matching it exactly:
-# u <- rnorm(N, b0 + b1 * s, sqrt(sigma0^2 * exp(2 * sigma1 * s)))
-# Not using this one right now because I'd have to tweak the model matrix below 
-# and I'm out of time. 
-plot(s1, u)
-plot(s2, u)
-
-mm <- model.matrix(~ s1 + s2)
-
-obj <- MakeADFun(data = list(x_ij = mm, y_i = u),
-  parameters = list(b_j = c(0, 0, 0), sigma_j = c(0, 0, 0)),
-  DLL = "varianceFunction")
-
-opt <- nlminb(start=obj$env$last.par.best, objective=obj$fn, gradient=obj$gr)
-opt$par
-# If you need standard errors and therefore confidence intervals... exp(sigma1 + 1.96 * SE)... :
-rep <- sdreport(obj)
-rep
-
-
-
-# Replicate with Fraser Data.
-library(tidyverse);library(attenPlot)
-load("03_Data_Annual.RData")
-load("out_doy2.RData")
-load("out_med.RData")
-load("out_max.RData")
-load("out_min.RData")
-
-remove = c("08MH006","08MH076","08MH090","08MH056","08LG016","08LG048","08MH103","08MH001","08MH016","08MH029")
-x = out_doy2
-
-y = x$Area$real_slopes %>% select(8:ncol(.)) %>% apply(.,2,function(x) zero_one(x)) %>% 
-	apply(.,1,function(x) sum(x)) %>% mutate(x$Area$real_slopes, clim.sd = .)
[…]

### 2017-04-11 varianceFunction.cpp

@@ -1,28 +1,0 @@
-#include <TMB.hpp>
-
-template<class Type>
-Type objective_function<Type>::operator() () {
-  // data:
-  DATA_MATRIX(x_ij);
-  DATA_VECTOR(y_i);
-  
-  // parameters:
-  PARAMETER_VECTOR(b_j)
-  PARAMETER_VECTOR(sigma_j);
-  
-  int n_data = y_i.size(); // get number of data points to loop over
-  
-  // Linear predictor
-  vector<Type> linear_predictor_i(n_data);
-  vector<Type> linear_predictor_sigma_i(n_data);
-  linear_predictor_i = x_ij*b_j;
-  linear_predictor_sigma_i = sqrt(exp(x_ij*sigma_j));
-  
-  Type nll = 0.0; // initialize negative log likelihood
-  
-  for(int i = 0; i < n_data; i++){
-    nll -= dnorm(y_i(i), linear_predictor_i(i) , linear_predictor_sigma_i(i), true);
-  }
-  
-  return nll;
-}

### 2017-04-13 drafts/PNAS/Chezik_PNAS_MS.tex
Corrected spelling errors in the manuscript and added Pages files to gitignore.

@@
 The rate of warming varies across the watershed (from increases of 0.03 to 0.04\celsius$\cdot$year\textsuperscript{-1}).
-Furthermore, there is an observed decrease in preciptiation along the Canadian Rockies transitioning to increasingly wetter trends along the coast.
+Furthermore, there is an observed decrease in precipitation along the Canadian Rockies transitioning to increasingly wetter trends along the coast.
 
@@
 \section*{Quantifying the River Network Portfolio Effect}
-We hypothesize that flow gauge sites with larger climate porfolios will exhibit greater long-term flow trend stability than sites with smaller climate portfolios.
+We hypothesize that flow gauge sites with larger climate portfolios will exhibit greater long-term flow trend stability than sites with smaller climate portfolios.
 We used catchment area as a proxy for the climate portfolio diversity of a given site, given that larger catchments have more variable climate portfolios (Fig.~\ref{fig:2}) as well as aggregate more water and carry more weight on downstream dampening.
-Therefore, to test whether networks dampen climate signals we regressed site-specific estimates of hydrologic trend (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model (\textit{Materials and Methods}).
-To quantify attenuation, we modeled the change in trend variability with the change in catchment area using an exponential variance function \cite{Pinheiro:2000}.
+Therefore, to test whether networks dampen climate signals we regressed site-specific estimates of hydrologic trend (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model (\textit{Materials and Methods}, eq.~\ref{eq2}).
+To quantify attenuation, we modeled the change in trend variability with the change in catchment area using an exponential variance function \cite{Pinheiro:2000}, (eq.~\ref{eq3}).
 The variance function allowed us to predict the range of flow trend values that would be expected as watershed area increased.
@@
 
-\subsection*{Basin trend attenuation analysis} To measure network-driven dampening of climate signalling, we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model:
+\subsection*{Basin trend attenuation analysis} To measure network-driven dampening of climate signaling, we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model:
 \begin{equation}

### 2017-04-18 drafts/PNAS/Chezik_PNAS_SI.tex
Final Pre-Submission Edits

@@
 The following supplementary material includes 4 additional figures.
-Figures \ref{fig:S1} and \ref{fig:S2} are similar to figure 4 in the main text where monthly trends in flow are shown for the basin along with trend attenuation results.
+Figures \ref{fig:S1} and \ref{fig:S2} are similar to figure 4 in the main text where Fraser River basin monthly flow trends are shown along with trend attenuation results.
 These supplementary figures highlight minimum and median seasonal flow trends rather than maximum seasonal flow trends.

### 2017-05-30 Code_03b Post_Analysis_Null_Test.R
Analysis of changes in forest cover on flow trends. Subsequent supplementary figures.

@@ -33,22 +33,63 @@ M.Data = lapply(vars,function(x){
 })[[length(vars)]] %>% ungroup()
 rm(vars, MonthlyVars)
 
+#Incorporate Land cover data.
+setwd("~/sfuvault/Simon_Fraser_University/PhD_Research/Projects/Data/Original_Data/GIS_Data/BC_Climate_Data/ClimateTrends/forest_stats")
+#Gather Together all the sites timber harvest results.
+df = dir() %>% 
+	lapply(.,function(x){
+		year = as.numeric(substr(x,1,4))
+		read_csv(x,col_names = c("Station.ID","count")) %>% 
+			mutate(Year = year)
+	}) %>% 
+	bind_rows()
+#Convert the 100m X 100m raster count data to area of timber harvest in km^2.
+df = df %>% mutate(harvest = count*(0.1*0.1))
+#Create a rolling average column to account for recent cuts.
+roll = plyr::ddply(df, "Station.ID", function(x){
+	c = 1; FiveHarvest = NULL
+	for(i in 1970:2007){
+		FiveHarvest[c] = x %>% filter(Year <= i, Year >= i-5) %>% .$harvest %>% sum(.)
+		c = c + 1
+	}
+	data.frame(FiveHarvest, Year = c(1970:2007))
+})
+df = df %>% filter(Year >= 1970) %>% left_join(., roll, by = c("Station.ID","Year"))
+
+#
+M.Data = df %>% select(-count) %>% left_join(M.Data, ., by = c("Station.ID","Year"))
+Y.Data = df %>% select(-count) %>% left_join(Y.Data, ., by = c("Station.ID","Year"))
+M.Data = M.Data %>% mutate(pHarvest = harvest/Area, p5Harvest = FiveHarvest/Area)
+Y.Data = Y.Data %>% mutate(pHarvest = harvest/Area, p5Harvest = FiveHarvest/Area)
+M.Data = M.Data %>% group_by(Station.ID) %>% mutate(p5H_Center = scale(p5Harvest, scale=F))
+Y.Data = Y.Data %>% group_by(Station.ID) %>% mutate(p5H_Center = scale(p5Harvest, scale=F))
+rm(df)
+
 setwd("~/sfuvault/Simon_Fraser_University/PhD_Research/Projects/River-Network-Flow-Trends")
+save(Y.Data, file = "05_MonthDat_ClimForest.RData")
[…]

### 2017-05-30 drafts/AGU_Journal/submission 2/ms_Geo_Let.tex
GPRL MS, SI and BIB

%% To submit your paper:
\documentclass[draft,linenumbers]{AGUJournal}
\drafttrue

%Packages
\usepackage{upgreek}
\usepackage{gensymb}

%% For final version.
% \documentclass{AGUJournal}

\journalname{Geophysical Research Letters}

\begin{document}

%% ------------------------------------------------------------------------ %%
%  Title
%% ------------------------------------------------------------------------ %%

\title{River networks dampen long-term hydrological signals of climate change}

%% ------------------------------------------------------------------------ %%
%  AUTHORS AND AFFILIATIONS
%% ------------------------------------------------------------------------ %%

\authors{K. A. Chezik\affil{1}, S. C. Anderson\affil{2}, J. W. Moore\affil{1}}

\affiliation{1}{Earth to Ocean Research Group, Department of Biological Sciences, Simon Fraser University, 8888 University Dr., Burnaby, British Columbia V5A 1S6, Canada}
\affiliation{2}{School of Aquatic and Fishery Sciences, University of Washington, Box 455020, Seattle, WA 98195, USA}

%% Corresponding Author:

\correspondingauthor{Kyle A. Chezik}{kchezik@sfu.ca}

%% Keypoints, final entry on title page.

\begin{keypoints}
\item River networks dampen long-term flow trends by integrating varied climate responses across complex landscapes.
\item Large sub-catchments integrate greater climate trend diversity, dampening the local extremes of climate change.
\item Rivers may possess an under-appreciated capacity, and provide a tool, for mitigating the impacts of climate change.
\end{keypoints}

%% ------------------------------------------------------------------------ %%
%  ABSTRACT
%% ------------------------------------------------------------------------ %%

\begin{abstract}
	River networks may dampen local hydrologic climate change through the aggregation of upstream climate portfolio assets. Here we examine this hypothesis using flow and climate trend estimates (1970--2007) at 55 hydrometric gauge stations and across their contributing watersheds' within the Fraser River basin in British Columbia, Canada. Using a null hypothesis framework we compared our observed attenuation of river flow trends as a function of increasing area and climate trend diversity, with null-simulated estimates to gauge the likelihood and strength of our observations. We found the Fraser River reduced variability in downstream long-term discharge by $>$91\%, with $>$3.1 times the attenuation than would be expected under null-simulation. Although the strength of dampening varied seasonally, our findings indicate that large free-flowing rivers offer a powerful and largely unappreciated process of climate change mitigation. River networks that integrate a diverse climate portfolio can dampen local extremes and offer climate change relief to riverine biota.
\end{abstract}

%% ------------------------------------------------------------------------ %%
\section{Introduction}

As billions are spent adapting to climate change by such means as building dikes or improving stormwater drainage \citep{Narain:2011}, certain habitats and processes provide a natural defense system that mitigates climate change impacts \citep{Jones:2012}. For example, coastal habitats such as oyster reefs, mangrove forests, and eelgrass beds shield 67\% of the United States coastline and 1.4 million people from sea level rise and storms \citep{Arkema:2013}. Protection and enhancement of these natural systems has become a key component of many climate mitigation strategies \citep{Guerry:2015}, as they return multiple cost-effective services in contrast to engineered alternatives \citep{Jones:2012}. For instance, revitalization of the Mississippi delta not only improves storm protection, but also increases production and sustainability of fisheries and wildlife resources, protects the water supply, and reduces the impact of wastewater effluent \citep{LDNR:1998}. The value of climate-buffering habitats stands only to increase as average global temperatures climb. The Paris Agreement of 2016 aims to limit global temperature rise to 1.5\celsius~\citep{Hulme:2016}, a goal that concedes a minimum doubling of observed warming \citep{Hartmann:2013}; therefore, an urgent challenge and opportunity is discovering, maintaining and restoring systems that naturally buffer against the oncoming impacts of climate change.

	In the face of changing global precipitation patterns \citep{Donat:2016}, river networks may offer an unappreciated defense against shifting flow regimes under climate change \citep{Hartmann:2013,Palmer:2009}. Earlier snowmelt \citep{Rauscher:2008}, reduced snow pack \citep{McCabe:2014}, shifts from snow to rain, and changes in the annual distribution of precipitation are all impacting stream flow \citep{Hartmann:2013}, resulting in flashier flows with increased frequency of flooding \citep{Hirabayashi:2013}, longer periods of drought and critically low flows for wildlife and agriculture \citep{Melillo:2014}. However, we hypothesize that large free-flowing river networks may naturally dampen these signals of climate change by integrating across varied landscapes and different manifestations of climate. Processes that aggregate across asynchronous components tend to dampen the aggregate, a process known as the portfolio effect \citep{Doak:1998}. Climate is expressed differently and asynchronously over the landscape \citep [e.g.,][]{Wang:2012} and on account of their branching architecture and directional flow, river networks integrate these varied climate manifestations \citep{Peterson:2013}, potentially dampening the local climatic response of sub-catchments. Large free-flowing rivers may have diverse climate portfolios, with sub-catchments acting as assets in their portfolio, thereby dampening local climate impacts. Portfolio effects in rivers are already known to dampen short-term variation \citep{Moore:2015,Yeakel:2014} but attenuation of long-term climate driven shifts have yet to be examined.

\begin{figure}[h]
\centering
\includegraphics[width=35pc]{Fig1_Map.pdf}
	\caption{Climate trends (1970 - 2007) in mean annual temperature (MAT) and mean annual precipitation (MAP) within the Fraser River basin overlayed on a digital elevation model of British Columbia, Canada. Flow gauge sites (dots) are scaled by the size of the contributing area. The map is projected in Albers BC UTMs to provide an equal area depiction of the region but labels are expressed in WGS84 latitude and longitude.}
\label{fig:1}
\end{figure}

Here we consider whether river networks attenuate local long-term climate change through analysis of hydrometric data from one of the largest free-flowing rivers in North America. Located in British Columbia, Canada, the Fraser River drains an area approximately the size of the United Kingdom ($\sim$217,000 km\textsuperscript{2}) and in a rare combination is fairly well-monitored while also having no dams on its main stem \citep{Vorosmarty:2010} (Figure~\ref{fig:1}). This watershed drains interior high plateau, coastal mountains, and the Canadian Rockies, and thus integrates a diverse mosaic of landscapes, weather, and climate. As with other mid-latitude rivers \citep{Bindoff:2013}, the Fraser River's discharge volume and variability have increased over the last several decades \citep{Dery:2012,Morrison:2002}. Following projected pattens of climate change, the Fraser River basin is increasingly shifting from a snow dominated to rain-snow hybrid system \citep{Kang:2016}. Using 38 years of flow data collected concurrently at 55 hydrometric gauge stations throughout the watershed, we use a novel analytical approach to examine whether this large river exhibits signals of long-term climate dampening.

\section{Materials and Methods}

\subsection{Discharge Data and Flow Metrics}

	To assess signals of climate change in hydrology within the Fraser River basin we downloaded daily flow data from Environment Canada's HYDAT database. Within HYDAT we selected Fraser River basin hydro-metric gauge stations that were not observed to be dam influenced and which collected data in each month between 1970 and 2007; where no month was missing more than 5 days data and at least 35 complete years were present. We selected this 38-year range because this period maximized the number of gauge stations in the Fraser River basin (\textit{n}=55) operating concurrently over a long enough time period in which climate trends may be observable. We linearly interpolated missing data in log-space, but of the 746,544 days of data only 23 days were interpolated. These daily data were smoothed using a five-day rolling average before summarizing into annual and monthly response variables. This smoothing technique was used to reduce the influence of erroneous and unusually extreme values. Because dams control flow and stabilize flow trends, we removed sites where daily flow exhibited unusually high autocorrelation and low variability relative to other sites of similar size, and sites that were observed downstream of, and atypically close to, a dam (\textit{n}=3). The downstream distance of gauge stations from dams was determined using GIS spatial layers available in the \textit{BC Data Catalog}, maintained by the provincial government of British Columbia.

	As climate change shifts precipitation patterns from snow to rain in temperate regions, rivers are generally predicted to have lower low-flows and higher high-flows, as well as earlier snowmelt \citep{Nijssen:2001}. These manifestations of rising temperatures are impacting species \citep{Xenopoulos:2006} and people \citep{Hirabayashi:2013}. We summarized our daily rolling average discharge data into annual and monthly flow metrics that capture these transitions. We used maximum- and minimum-flow estimates to capture trends in extreme discharge and median-flow estimates to capture general annual and seasonal trends. Because the timing of flow is particularly important to the phenology of many organisms, we also calculated the day-of-year in which half the annual flow was reached.

\subsection{River discharge trend analysis}

To test for a river network portfolio effect, we need to detect varied and asynchronous, long-term changes in flow throughout the watershed. To do this we estimated annual and monthly flow trends for each response variable at each gauge station using a generalized least squares (GLS) model with an AR1 autocorrelation function:
\begin{linenomath*}
\begin{equation}
  \mathrm{Q}_{s,t} = a_s + b_s \mathrm{Y}_{s,t} + \epsilon_{s,t}, \quad 
  \epsilon_{s,t} \sim \mathcal{N}(\upphi \epsilon_{s,t-1}, \sigma_\mathrm{Q}^2) \label{eq1},
\end{equation}
\end{linenomath*}
where $\mathrm{Q}_{s,t}$ represents a flow metric [log(maximum), log(minimum), log(median), logit(day-of-year to half-annual-flow)] at each site ($s$) and time point ($t$), and $\mathrm{Y}_{s,t}$ represents the time in years (e.g., 1970, 1971, \ldots, 2007) with 1988 subtracted to approximately center the predictor. Flow metrics were scaled by site to facilitate cross-site comparisons. Parameters $b_{s}$ (slope) and $a_{s}$ (intercept) represent the estimated mean effect of time on $\mathrm{Q}_{s,t}$ and the estimate of $\mathrm{Q}_{s,t}$ at time zero (i.e., 1988), respectively. Error ($\epsilon_{s,t}$) of the current time step for a given site was allowed to be correlated with that of the previous time step by $\upphi$ and was assumed to be normally distributed with a variance of $\sigma_\mathrm{Q}^{2}$. To ensure our model estimates and simulations remained within the calendar year we logit transformed our flow-timing response variable (i.e., day-of-year to half-annual-flow) after scaling the data between 0 and 1 (i.e., dividing by 365).

\subsection{Quantifying Climate Variability}

For a river network to act as a portfolio of climate and dampen the effects of climate change, expressions of climate and their trends need to vary throughout the network. In order to quantify climate trend diversity we developed a climate index for each flow-gauge catchment. Using the Western North America Climate tool (ClimateWNA) \citep{Wang:2016}, we estimated historic climate values on an evenly spaced 1 km\textsuperscript{2} grid across the Fraser River basin for each year and month in our study. Temperature and precipitation are likely to drive changing hydrology, therefore we included variables in our climate index that capture changes in extreme and mean temperature, as well as changes in precipitation volume and physical state (snow vs. rain). Climate variables included the mean annual temperature (MAT), and precipitation (MAP), extreme minimum (EMT) and maximum (EXT) temperature, and precipitation as snow (PAS). Climate variable trends were calculated at each grid point using a GLS model similar to eq.~\ref{eq1};
\begin{linenomath*}
\begin{equation}
	\mathrm{C}_{s,t} = \alpha_s + \beta_s \mathrm{Y}_{s,t} + \epsilon_{s,t}, \quad 
  \epsilon_{s,t} \sim \mathcal{N}(\upphi \epsilon_{s,t-1}, \sigma_\mathrm{C}^2) \label{eq2},
\end{equation}
\end{linenomath*}
where $\mathrm{C}_{s,t}$ represents a climate variable and $\mathrm{Y}_{s,t}$ represents the time in years (e.g., 1970, 1971, \ldots, 2007). Parameters $\beta_{s}$ (slope) and $\alpha_{s}$ (intercept) represent the estimated mean effect of time on $\mathrm{C}_{s,t}$ and the estimate of $\mathrm{climate}_{s,t}$ at time zero, respectively. Autocorrelated errors were incorporated as in eq.~\ref{eq1}.

	Spatial variation in climate change would suggest catchments of different size integrate different amounts of climate variability. In order to calculate a climate index for each gauge station we delineated flow-gauge catchments and summarized climate trends within those catchments. Using Whitebox Geospatial Analysis Tools \citep{Lindsay:2016}, we delineated the Fraser River basin and the catchment areas of each flow-gauge station. Digital elevation models were provided by the provincial government of British Columbia, Canada and state government of Washington, USA and were pre-processed with a breaching algorithm and stream burned to facilitate proper flow path and accumulation models \citep{Woodrow:2016}. We summarized the climate variability integrated at each flow gauge by calculating the standard deviation in climate trends within their catchments. The standard deviations of these climate variables were scaled between 0 and 1 and summed as a general climate index for each catchment.

\subsection{Quantifying the River Network Portfolio Effect}

We hypothesize that flow gauge sites with larger climate portfolios will exhibit greater long-term flow trend stability than sites with smaller climate portfolios. We used catchment area as a proxy for the climate portfolio diversity of a given site, given that larger catchments have more variable climate portfolios (Figure~\ref{fig:2}) as well as aggregate more water and carry more weight on downstream dampening. Therefore, to test whether networks dampen climate signals, we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model:
\begin{linenomath*}
\begin{equation}
	\hat{b}_{s} = c + d\sqrt{\mathrm{A}_{s}} + \eta_{s}, \quad
  \eta_{s} \sim \mathcal{N}(0, f(\mathrm{A}_{s})) \label{eq3},
\end{equation}
\end{linenomath*}
where $\hat{b}_{s}$ represents a flow trend at a given site ($s$) and $\mathrm{A}_{s}$ represents site $s$'s watershed area. We adjusted $\sqrt{\mathrm{A}_{s}}$ by subtracting the mean $\sqrt{\mathrm{A}}$, thereby centering our predictor. Fitted $d$ and $c$ parameters represent the mean effect of watershed area ($\sqrt{\mathrm{A}}$), on flow trends and the mean flow trend for a $\sim$5,700 km\textsuperscript{2} watershed, respectively. Captured in our climate portfolio proxy variable ($\mathrm{A}$), is the interaction between climate and landuse. Undoubtedly, timber harvest, the dominant source of land cover change in the Fraser River basin, scales with sub-basin area and contributes to changes in flow. Although exploratory linear models at the site level do suggest an effect of timber harvest on long-term flow trends for some gauge stations (Figure~S1, S2), when information is shared across sites in a global model, climate effects dominate (Figure~S3). Thus we did not attempt to separate the contributions of land cover change, climate change and their interaction on long-term flow trends.

To quantify attenuation, we modeled the change in trend variability with the change in catchment area using an exponential variance function \citep[p.~211]{Pinheiro:2000}: 
\begin{linenomath*}
\begin{equation}
	f(\mathrm{A}_{s}) = \sigma_b^2 \exp(2\updelta\sqrt{\mathrm{A}_{s}}) \label{eq4},
\end{equation}
\end{linenomath*}
where the variance of the estimated error ($\eta_{s}$) changes exponentially with increasing $\sqrt{\mathrm{A}}$. This variance function allowed us to predict the range of flow trend values that would be expected as watershed area increased. A decreasing range of flow trends with increasing watershed area indicates a dampening network effect.

\subsection{River Attenuation Null Model Simulations}

In the observed data, small watersheds exhibited greater variability around their trend estimates than large watersheds, likely because of greater short-term variation in small catchments \citep{Moore:2015}. This relationship between watershed size and trend certainty may pull small watershed trend estimates away from zero, thereby creating the appearance of decreasing trend variability among sites as watershed size increases. Using a null-hypothesis framework we simulated time-series with no underlying trend from each sites' fitted model, using the estimated standard deviation and autocorrelation parameters ($\hat{\sigma}_{Q}$, $\hat{\upphi}$). Using the same GLS model and AR1 correlation structure as applied to the observed data (eq. \ref{eq1}), we estimated trends for 1000 simulated time series at each site for each response variable (e.g., Figure~S4). This simulation process, a form of parametric bootstrapping, created distributions of null-expectations with which we compared our observed results.

We then applied equation \ref{eq3} to each of the 1000 basin-wide simulations (e.g. Figure~S5), resulting in 1 observed variance exponent parameter ($\hat{\updelta}$) and 1000 simulated $\hat{\updelta}$ for each flow metric. By comparing our observed attenuation with our basin-wide null-model simulations we addressed the potential that more variable flows in smaller catchments contribute to the observed pattern of flow trends as a function of watershed area (Figure~\ref{fig:3}). This null-model approach asks how likely our observation is due to a sampling effect versus a network portfolio effect.

\subsection{Dampening summary statistics}

We calculated attenuation certainty as the percentage of simulated $\hat{\updelta}$ that were less than the observed $\hat{\updelta}$. We calculated attenuation strength ($\mathrm{AS}$) as the ratio between the standard deviation at the smallest and largest watersheds as defined by the exponential variance function (i.e., $\mathrm{AS}=\sqrt{f(\mathrm{A}_{s=1})} / \sqrt{f(\mathrm{A}_{s=2})}$). To compare the degree of attenuation observed with the null-simulated attenuation, we compared this ratio with the same ratio calculated from the null-simulated data (i.e., $\mathrm{AS} / \mathrm{AS_{NULL}}$). Finally, to estimate how the day-of-year to half-annual-flow was changing, we estimated the steepest point of each site's logistic curve and multiplied this slope value by 365 to revert from our 0--1 scale to our original annual range of 0--365. To get a decadal rate we multiplied these results by 3.8 (number of decades in 38 years). We estimated the steepest part of the logistic curve using the \textit{divide by four} rule \citep{Gelman:2008} (dividing the coefficient by four equals the first derivative of the logistic curve at its steepest point).

\section{Results}

\subsection{Hydrologic Trend Diversity}

Long-term changes in flow metrics varied substantially across the watershed. For instance, annual maximum flow decreased in 40 sites but increased in 15 over the last four decades (-26 to 10\% change$\cdot$decade\textsuperscript{-1}, SD = 6\%). Annual median flow showed similar variability but a greater tendency towards rising flow trends with 49 increasing and only six decreasing (-5 to 25\% change$\cdot$decade\textsuperscript{-1} SD = 5\%). Thus, within the Fraser River basin, there are a variety of long-term trends in flow, suggesting a diversity of climate responses and landscape effects.

\subsection{Climate Trend Diversity}

Our index of climate variability increased with area (Figure~\ref{fig:2}), asymptoting in catchments larger than 40,000~km\textsuperscript{2}. Thus, rivers that drain larger catchments integrate a greater diversity of shifting climates. These findings suggest that larger areas contain greater landscape complexity resulting in a greater variety of climatic expressions that respond to changes in global climate differently.

\begin{figure}[h]
\centering
\includegraphics[width=20pc]{Fig2_ClimPort.pdf}
	\caption{Climate variability index as a function of catchment area. The mean trend line (black) was estimated using robust linear regression and standard error estimates (SE) (2$\cdot$SE = grey) were calculated using the weighted standard deviation estimate.}
\label{fig:2}
\end{figure}

\subsection{Climate Portfolios and Flow Trend Dampening}

Long-term changes in hydrology were less variable in larger catchments, as predicted, demonstrating between 92 and 96\% dampening of flow trends moving from headwater catchments to the network's confluence. In other words, the largest catchments were approximately 10 times more stable in their climate response than the smallest catchments. For instance, trends ranged between a 6\% reduction and 19\% increase in median flow per decade among small watersheds, while large watersheds ranged between 0.8 and 1.6\% change per decade. Similar attenuation was seen in flow timing with a 94\% reduction in the day-of-year to half-annual-flow trends such that small watersheds were reaching half their annual flow between 8 days later and 25 days earlier while large watersheds have only shifted 2 to 4 days earlier between 1970 and 2007.

\begin{figure}[h]
\centering
\includegraphics[width=35pc]{Fig4_AnnFun.pdf}
	\caption{Annual flow trend attenuation within the Fraser River basin. (\textbf{Left}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large). Blue lines represent observed attenuation; orange and red represent simulated attenuation that is weaker and stronger than observed. (\textbf{Right}) Density plots show null model simulated variance exponents ($\hat{\updelta}$) and the proportion on either side of the observed variance exponent (blue). Flow metrics include long-term flow-timing shifts (change per decade in day-of-year (DOY) to half annual flow), where decreasing trends suggest more annual flow is occurring earlier in the year (a), and the percent change per decade in minimum (b), maximum (c) and median (d) annual flow. Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
\label{fig:3}
\end{figure}

Our null model approach illustrates that support for a river network portfolio effect is greater than 90\% for all four annual flow trend metrics and as high as 98\% for median-annual-flow trends (Figure~\ref{fig:3}). Trend attenuation was; 3.1 (0.6--9.8 this and hereafter are 90\% intervals), 4.7 (1.1--14.7), 3.2 (0.9--9.0) and 5.2 (1.4--14.4) times greater than the null model for annual-flow timing, and minimum, maximum and median annual-flow, respectively. These statistics provide strong support for the hypothesis that river network portfolio effects contribute to climate dampening in large rivers.

\subsection{Seasonally Shifting Flow Trend Dampening}

Regional flow-trends and attenuation strength varied by season. Analysis of monthly trends revealed that winter flows are getting higher and summer flows are getting lower over time, but that these trends have been more variable in smaller catchments (Figure~\ref{fig:4}, Figures~S6, S7). For example, winter maximum flows in small catchments ranged from nearly no change to dramatic 47\% increases per decade with decreases only as large as 9\%. Despite these extreme locations, on average the basin has only experienced moderate winter maximum flow increases of 5-9\%. Summer flows exhibited the opposite, with decreases in high flows over time. However, there was weak evidence of attenuation in the spring (Figure~\ref{fig:4}, Figures.~S6, S7). Overall, climate portfolios tended to have greater attenuation during stable periods of the year and decreased during seasonal transitions. In the spring, the likelihood of network-driven attenuation of maximum flow trends was as low as 24\% (Figure~\ref{fig:4}). The deterioration of the network's attenuation strength was coincident with the spring freshet, when snowmelt drives high flows across the basin. Synchronization events such as the freshet may subvert river network dampening mechanisms by homogenizing the region's response to seasonally driven climate shifts.

\begin{figure}[h]
\centering
\includegraphics[width=30pc]{Fig5_MaxMonth.pdf}
	\caption{Monthly maximum flow trend attenuation within the Fraser River basin. (\textbf{Left}) Fraser River's basin-wide maximum-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulations. Observed values falling further from the center of the density distribution suggest greater evidence for changes in maximum flow and a greater shift in magnitude. (\textbf{Center}) Observed monthly Fraser River maximum-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates. Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed. (\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large), for four seasonally representative months. These reflect months in the prior columns and describe the variation in percent change per decade of maximum flow among sites. Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
\label{fig:4}
\end{figure}

\section{Discussion}

Our results suggest that river networks mitigate and dampen long-term hydrologic trends by integrating a diverse climate portfolio over a heterogeneous landscape. By quantifying the diversity of climate and flow trends within many sub-basins of the Fraser River watershed (Figures~\ref{fig:1}, \ref{fig:2}) we demonstrate as much as a 10 fold decrease in flow trend variability with increasing watershed area (i.e., climate portfolio) (Figures~\ref{fig:3}, \ref{fig:4}). Similar to other temperate rivers \citep{Rauscher:2008} and consistent with recent studies \citep [e.g.,][]{Kang:2016, Kang:2014, Shrestha:2012}, we show the Fraser River basin is exhibiting climate driven shifts with increasing flows over winter transitioning to decreasing flows over the summer (Figure~\ref{fig:4}). These findings are consistent with a general body of work that suggests climate change is transitioning previously snow driven systems to rain, resulting in a redistribution of water towards winter flows \citep{Bindoff:2013}. As climate change progresses, river hydrology will continue to shift, further stressing riverine ecosystems and subsequently demanding responsive management. However, our work suggests that river networks provide an underappreciated defense against these climate-change impacts as they are uniquely organized to leverage locally-filtered expressions of climate into stability. A simple product of network form and gravity, stability emerges as river networks integrate landscape heterogeneity, creating a downstream portfolio of climate that smooths local extremes and tempers long-term flow trends. Therefore, larger rivers should have flow regimes that are less sensitive to local climate trends.

The Fraser River basin has seen changes in landcover between 1970 and 2007, which have impacted river discharge \citep[e.g.,][]{Zhang:2014}. Over this period, timber harvest has been the dominant contributor to changes in forest cover but increasingly natural events are having a greater impact (Figure~S8). For instance, abnormally hot, dry weather in 2003 led to fires that destroyed an unprecedented 2,600 km\textsuperscript{2} of forest largely within the Fraser River basin \citep{Filmon:2003}. Beginning in the early 2000's, western mountain pine beetle outbreaks began consuming large swaths of forest, impacting 20\% of British Columbia by 2013 \citep{schnorbus:2010}. These natural events are facilitated by climate change and are becoming increasingly common and extreme \citep{Melillo:2014, Maness:2013}. Pine beetle outbreaks are often subsequently heavily logged and timber removal can occur in anticipation of fires, resulting in harvest practices responding dynamically to climate change. Analyses indicated that while timber harvest rates did have some local impacts on hydrological change, harvest did not have substantial impacts on flows in a global model (Figure S3); thus we focus on long-term hydrological change and make the simplifying assumption that different forest harvest rates are subsumed by long-term climate change. Indeed, timber harvest was relatively consistent over the last 38 years across the vast majority of our 55 study catchments, with the exception of four relatively small, heavily fire and pine beetle impacted sites (Figure S8). Therefore, our study captures the interwoven direct and indirect impacts of climate change by considering changes in precipitation but also latent affects through climate filtering of altered landscapes due to climate change.

Our study provides insight into the spatial scaling of a river network's dampening capacity. Although we focused on a vast watershed with a remarkably diverse climate portfolio, we would still expect that smaller rivers or watersheds with less topographic complexity, will dampen climate variability. For instance, based on our data, we predict that rivers draining a catchment of 60,000 km\textsuperscript{2} would have 66\% less variability in the rate of change in hydrology than a smaller river draining ~5000 km\textsuperscript{2}. While this dampening is less than the 10-fold decrease observed in the entire 217,000 km\textsuperscript{2} Fraser River watershed, smaller watersheds can offer a defense against the impacts of climate change.

\section{Conclusion}

Climate change is causing economic and conservation challenges for river systems worldwide \citep{Palmer:2009,Pecl:2017}. In the Fraser River, sockeye salmon have seen as much as 80\% pre-spawn mortality in years with later run-off and elevated temperatures, leading managers to reduce or close the fishery (e.g., 2013 and 2015). In the absence of climate dampening, the impacts of these climate extremes could have been considerably worse. In contrast to previously studied types of climate mitigation where habitats modify climate drivers (e.g., carbon storage) or physically absorb climate change impacts (e.g., mangroves), river network climate dampening smooths out extreme climatic trends thereby reducing the impact of local extremes. But just as the destruction of coastal habitats degrades their natural capacity to mitigate sea-level rise or storm events \citep{Arkema:2013}, the management of river basins likely impacts the ability for river networks to attenuate climate trends. For example, dams synchronize the flow regimes of rivers \citep{Poff:2007} and logging increases the magnitude and frequency of extreme events such as high and low flows through increased runoff rates \citep{Zhang:2014}. These anthropogenic activities may magnify the impacts of climatic shifts such as the transition from snow- to rain-dominated precipitation in some basins. The climate cost of these watershed activities should be considered in environmental decision making. Climate change is a global challenge; here we suggest large rivers dampen local change by leveraging landscape diversity into a portfolio of climate, an important tool in the climate-mitigation toolbox.

%% ------------------------------------------------------------------------ %%

%  ACKNOWLEDGMENTS
\acknowledgments
This paper was possible only through the extensive use of many open source, freely available tools. All flow data analysis was done using R v3.3.2 (\url{www.r-project.org/}) and all R code can be found at \url{http://github.com/kchezik/River-Network-Flow-Trends.git}. Spatial analysis of watersheds was done using GDAL/OGR v1.11.5, and Whitebox GAT v3.4 \citep{Lindsay:2016} with visualization help from QGIS v2.18.2 (\url{www.qgis.org/}). Climate data were gathered using ClimateWNA v5.4 \citep{Wang:2016} and summary statitics were calculated using R and the \textit{rasterstats} module (\url{https://github.com/perrygeo/python-rasterstats.git}) in Python v2.7. Data processing was facilitated greatly by GNU Parallel \citep{Tange2011a}. This study could not have been done without the decades of flow data collected by those at Environment Canada. Raw HYDAT data can be obtained via Environment Canada or a simplified version can be found on GitHub. K.A. Chezik and J.W. Moore were supported by the Liber Ero Chair of Coastal Science and Management and Simon Fraser University. S.C. Anderson was supported by the David H. Smith Conservation Research Fellowship.

%% ------------------------------------------------------------------------ %%

%% Citations

%\bibliography{ms.bib}

\begin{thebibliography}{38}
\providecommand{\natexlab}[1]{#1}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi

\bibitem[{\textit{Arkema et~al.}(2013)\textit{Arkema, Guannel, Verutes, Wood,
  Guerry, Ruckelshaus, Kareiva, Lacayo, and Silver}}]{Arkema:2013}
Arkema, K.~K., G.~Guannel, G.~Verutes, S.~A. Wood, A.~Guerry, M.~Ruckelshaus,
  P.~Kareiva, M.~Lacayo, and J.~M. Silver (2013), Coastal habitats shield
  people and property from sea-level rise and storms, \textit{Nat Clim Chang},
  \textit{3}(10), 913--918, \doi{10.1038/nclimate1944}.

\bibitem[{\textit{Bindoff et~al.}(2013)\textit{Bindoff, Stott, {AchutaRao},
  Allen, Gillett, Gutzler, Hansingo, Hegerl, Hu, Jain, Mokhov, Overland,
  Perlwitz, Sebbari, and Zhang}}]{Bindoff:2013}
Bindoff, N.~L., P.~A. Stott, M.~{AchutaRao}, M.~R. Allen, N.~Gillett,
  D.~Gutzler, K.~Hansingo, G.~Hegerl, Y.~Hu, S.~Jain, I.~Mokhov, J.~Overland,
  J.~Perlwitz, R.~Sebbari, and X.~Zhang (2013), \textit{{Detection and
  Attribution of Climate Change: from Global to Regional}. In: {Climate Change
  2013}: The Physical Science Basis.~Contribution of Working Group I to the
  Fifth Assessment Report of the {Intergovernmental Panel on Climate Change}
  {[Stocker, T.F., D. Qin, G.-K. Plattner, M. Tignor, S.K Allen, J. Boschung,
  A. Nauels, Y. Xia, V. Bex and P.M. Midgley (eds.)].}}, Cambridge University
  Press.

\bibitem[{\textit{D{\'e}ry et~al.}(2012)\textit{D{\'e}ry,
  {Hern{\'a}ndez-Henr{\'\i}quez}, Owens, Parkes, and Petticrew}}]{Dery:2012}
D{\'e}ry, S.~J., M.~A. {Hern{\'a}ndez-Henr{\'\i}quez}, P.~N. Owens, M.~W.
  Parkes, and E.~L. Petticrew (2012), A century of hydrological variability and
  trends in the {Fraser River} basin, \textit{Environ Res Lett}, \textit{7}(2),
  \doi{10.1088/1748-9326/7/2/024019}.

\bibitem[{\textit{Do~Hyuk~Kang et~al.}(2016)\textit{Do~Hyuk~Kang, Shi,
  ul~Islam, and D{\'e}ry}}]{Kang:2016}
Do~Hyuk~Kang, H.~G., X.~Shi, S.~ul~Islam, and S.~J. D{\'e}ry (2016), Impacts of
		a rapidly declining mountain snowpack on streamflow timing in {C}anada{'}s
  {Fraser River} basin, \textit{Scientific reports}, \textit{6}.

\bibitem[{\textit{Doak et~al.}(1998)\textit{Doak, Bigger, Harding, Marvier,
  {RE}, and Thomson}}]{Doak:1998}
Doak, D., D.~Bigger, E.~Harding, M.~Marvier, O.~{RE}, and D.~Thomson (1998),
		The statistical inevitability of stability{-}diversity relationships in
  community ecology, \textit{Am Nat}, \textit{151}(3), 264--276,
  \doi{10.1086/286117}.

\bibitem[{\textit{Donat et~al.}(2016)\textit{Donat, Lowry, Alexander,
  {O'Gorman}, and Maher}}]{Donat:2016}
Donat, M.~G., A.~L. Lowry, L.~V. Alexander, P.~A. {O'Gorman}, and N.~Maher
  (2016), More extreme precipitation in the world's dry and wet regions,
  \textit{Nat Clim Chang}, \textit{6}(5), 508--513, \doi{10.1038/nclimate2941}.

\bibitem[{\textit{Filmon}(2003)}]{Filmon:2003}
Filmon, G. (0), Firestorm 2003.

\bibitem[{\textit{Gelman and Hill}(2008)}]{Gelman:2008}
Gelman, A., and J.~Hill (2008), Data analysis using regression and multilevel
  hierarchical models, \doi{10.1080/02664760801949043}.

\bibitem[{\textit{Guerry et~al.}(2015)\textit{Guerry, Polasky, Lubchenco,
  Rebecca, Daily, Griffin, Ruckelshaus, Bateman, Duraiappah, Elmqvist, Feldman,
  Folke, Hoekstra, Kareiva, Keeler, Li, Emily, Ouyang, Reyers, Ricketts,
  Rockstr{\"o}m, Tallis, and Vira}}]{Guerry:2015}
Guerry, A.~D., S.~Polasky, J.~Lubchenco, C.~Rebecca, G.~C. Daily, R.~Griffin,
  M.~Ruckelshaus, I.~J. Bateman, A.~Duraiappah, T.~Elmqvist, M.~W. Feldman,
  C.~Folke, J.~Hoekstra, P.~M. Kareiva, B.~L. Keeler, S.~Li, M.~Emily,
  Z.~Ouyang, B.~Reyers, T.~H. Ricketts, J.~Rockstr{\"o}m, H.~Tallis, and
  B.~Vira (2015), Natural capital and ecosystem services informing decisions:
  From promise to practice, \textit{Proc Natl Acad Sci}, \textit{112}(24),
  7348--7355, \doi{10.1073/pnas.1503751112}.

\bibitem[{\textit{Hartmann et~al.}(2013)\textit{Hartmann, Klein~Tank,
  Rusticucci, Alexander, Br\diaeresis{o}nnimann, Charabi, Dentener,
  Dlugokencky, Easterling, Kaplan, Soden, Thorne, Wild, and
  Zhai}}]{Hartmann:2013}
Hartmann, D., A.~Klein~Tank, M.~Rusticucci, L.~Alexander,
  S.~Br\diaeresis{o}nnimann, Y.~Charabi, F.~Dentener, E.~Dlugokencky,
  D.~Easterling, A.~Kaplan, B.~Soden, P.~Thorne, M.~Wild, and P.~Zhai (2013),
  \textit{Observations: {Atmosphere and Surface}. In: {Climate Change 2013}:
  The Physical Science Basis.~Contribution of Working Group I to the Fifth
  Assessment Report of the {Intergovernmental Panel on Climate Change}
  {[Stocker, T.F., D. Qin, G.-K. Plattner, M. Tignor, S.K Allen, J. Boschung,
  A. Nauels, Y. Xia, V. Bex and P.M. Midgley (eds.)].}}, Cambridge University
  Press.

\bibitem[{\textit{Hirabayashi et~al.}(2013)\textit{Hirabayashi, Mahendran,
  Koirala, Konoshima, Yamazaki, Watanabe, Kim, and Kanae}}]{Hirabayashi:2013}
Hirabayashi, Y., R.~Mahendran, S.~Koirala, L.~Konoshima, D.~Yamazaki,
  S.~Watanabe, H.~Kim, and S.~Kanae (2013), Global flood risk under climate
  change, \textit{Nat Clim Chang}, \textit{3}(9), 816--821,
  \doi{10.1038/nclimate1911}.

\bibitem[{\textit{Hulme}(2016)}]{Hulme:2016}
Hulme, M. (2016), 1.5\celsius~ and climate research after the {Paris
  Agreement}, \textit{Nat Clim Chang}, \textit{6}(3), 222--224,
  \doi{10.1038/nclimate2939}.

\bibitem[{\textit{Jones et~al.}(2012)\textit{Jones, Hole, and
  Zavaleta}}]{Jones:2012}
Jones, H.~P., D.~G. Hole, and E.~S. Zavaleta (2012), Harnessing nature to help
  people adapt to climate change, \textit{Nat Clim Chang}, \textit{2}(7),
  504--509, \doi{10.1038/nclimate1463}.

\bibitem[{\textit{Kang et~al.}(2014)\textit{Kang, Shi, Gao, and
  D{\'e}ry}}]{Kang:2014}
Kang, D., X.~Shi, H.~Gao, and S.~D{\'e}ry (2014), On the changing contribution
  of snow to the hydrology of the {F}raser {R}iver basin, {C}anada.

\bibitem[{\textit{LDNR}(1998)}]{LDNR:1998}
LDNR (1998), Coast 2050: Toward a sustainable coastal {L}ouisiana.

\bibitem[{\textit{Lindsay}(2016)}]{Lindsay:2016}
Lindsay, J. (2016), Whitebox {GAT:} a case study in geomorphometric analysis,
  \textit{Comp Geosci}, \textit{95}, 75--84.

\bibitem[{\textit{Maness et~al.}(2013)\textit{Maness, Kushner, and
  Fung}}]{Maness:2013}
Maness, H., P.~J. Kushner, and I.~Fung (2013), Summertime climate response to
  mountain pine beetle disturbance in {B}ritish {C}olumbia, \textit{Nature
  Geoscience}, \textit{6}(1), 65--70.

\bibitem[{\textit{{McCabe} and Wolock}(2014)}]{McCabe:2014}
{McCabe}, G.~J., and D.~M. Wolock (2014), Spatial and temporal patterns in
  conterminous {U}nited {S}tates streamflow characteristics, \textit{Geophys Res
  Lett}, \textit{41}(19), 6889--6897, \doi{10.1002/2014GL061980}.

\bibitem[{\textit{Melillo et~al.}(2014)\textit{Melillo, Richmond, and
  Yohe}}]{Melillo:2014}
Melillo, J., T.~Richmond, and G.~W. Yohe (2014), \textit{Climate change impacts
  in the {U}nited {S}tates: The Third National Climate Assessment.}, 841 pp.,
  {U.S.} Government Printing Office.

\bibitem[{\textit{Moore et~al.}(2015)\textit{Moore, Beakes, Nesbitt, Yeakel,
  Patterson, Thompson, Phillis, Braun, Favaro, Scott, Charmaine, and
  Atlas}}]{Moore:2015}
Moore, J.~W., M.~P. Beakes, H.~K. Nesbitt, J.~D. Yeakel, D.~A. Patterson, L.~A.
  Thompson, C.~C. Phillis, D.~C. Braun, C.~Favaro, D.~Scott, C.~Charmaine, and
  W.~I. Atlas (2015), Emergent stability in a large, free{-}flowing watershed,
  \textit{Ecology}, \textit{96}(2), 340--347,
  \doi{10.1890/14-0326.1}.

\bibitem[{\textit{Morrison et~al.}(2002)\textit{Morrison, Quick, and
  Foreman}}]{Morrison:2002}
Morrison, J., M.~Quick, and M.~Foreman (2002), Climate change in the {Fraser
  River} watershed: flow and temperature projections, \textit{J Hydrol},
  \textit{263}, 230--244.

\bibitem[{\textit{Narain et~al.}(2011)\textit{Narain, Margulis, and
  Essam}}]{Narain:2011}
Narain, U., S.~Margulis, and T.~Essam (2011), Estimating costs of adaptation to
  climate change, \textit{Climate Policy}, \textit{11}(3), 1001--1019,
  \doi{10.1080/14693062.2011.582387}.

\bibitem[{\textit{Nijssen et~al.}(2001)\textit{Nijssen, {O'Donnell}, Hamlet,
  and Lettenmaier}}]{Nijssen:2001}
Nijssen, B., G.~{O'Donnell}, A.~Hamlet, and D.~Lettenmaier (2001), Hydrologic
  sensitivity of global rivers to climate change, \textit{Clim Change},
  \textit{50}, 143--175.

\bibitem[{\textit{Palmer et~al.}(2009)\textit{Palmer, Lettenmaier, Poff,
  Postel, Richter, and Warner}}]{Palmer:2009}
Palmer, M.~A., D.~P. Lettenmaier, L.~N. Poff, S.~L. Postel, B.~Richter, and
  R.~Warner (2009), Climate change and river ecosystems: Protection and
  adaptation options, \textit{Environ Manage}, \textit{44}(6), 1053--1068,
  \doi{10.1007/s00267-009-9329-1}.

\bibitem[{\textit{Pecl et~al.}(2017)\textit{Pecl, Ara{\'u}jo, Bell, Blanchard,
  Bonebrake, Chen, Clark, Colwell, Danielsen, Eveng{\aa}rd, Falconi, Ferrier,
  Frusher, Garcia, Griffis, Hobday, Janion-Scheepers, Jarzyna, Jennings,
  Lenoir, Linnetved, Martin, McCormack, McDonald, Mitchell, Mustonen, Pandolfi,
  Pettorelli, Popova, Robinson, Scheffers, Shaw, Sorte, Strugnell, Sunday,
  Tuanmu, Verg{\'e}s, Villanueva, Wernberg, Wapstra, and Williams}}]{Pecl:2017}
Pecl, G.~T., M.~B. Ara{\'u}jo, J.~D. Bell, J.~Blanchard, T.~C. Bonebrake, I.-C.
  Chen, T.~D. Clark, R.~K. Colwell, F.~Danielsen, B.~Eveng{\aa}rd, L.~Falconi,
  S.~Ferrier, S.~Frusher, R.~A. Garcia, R.~B. Griffis, A.~J. Hobday,
  C.~Janion-Scheepers, M.~A. Jarzyna, S.~Jennings, J.~Lenoir, H.~I. Linnetved,
  V.~Y. Martin, P.~C. McCormack, J.~McDonald, N.~J. Mitchell, T.~Mustonen,
  J.~M. Pandolfi, N.~Pettorelli, E.~Popova, S.~A. Robinson, B.~R. Scheffers,
  J.~D. Shaw, C.~J.~B. Sorte, J.~M. Strugnell, J.~M. Sunday, M.-N. Tuanmu,
  A.~Verg{\'e}s, C.~Villanueva, T.~Wernberg, E.~Wapstra, and S.~E. Williams
  (2017), Biodiversity redistribution under climate change: Impacts on
  ecosystems and human well-being, \textit{Science}, \textit{355}(6332),
  \doi{10.1126/science.aai9214}.

\bibitem[{\textit{Peterson et~al.}(2013)\textit{Peterson, Hoef, Isaak, Falke,
  Fortin, Jordan, Kristina, Monestiez, Ruesch, Sengupta, Som, Steel, Theobald,
  Torgersen, and Wenger}}]{Peterson:2013}
Peterson, E.~E., J.~M. Hoef, D.~J. Isaak, J.~A. Falke, M.~Fortin, C.~E. Jordan,
  M.~Kristina, P.~Monestiez, A.~S. Ruesch, A.~Sengupta, N.~Som, A.~E. Steel,
  D.~M. Theobald, C.~E. Torgersen, and S.~J. Wenger (2013), Modelling dendritic
  ecological networks in space: an integrated network perspective, \textit{Ecol
  Lett}, \textit{16}(5), 707--719, \doi{10.1111/ele.12084}.

\bibitem[{\textit{Pinheiro and Bates}(2000)}]{Pinheiro:2000}
Pinheiro, J., and D.~Bates (2000), \textit{{Mixed-Effects} Models in S and
  {S-PLUS}}, Springer.

\bibitem[{\textit{Poff et~al.}(2007)\textit{Poff, Olden, Merritt, and
  Pepin}}]{Poff:2007}
Poff, L.~N., J.~D. Olden, D.~M. Merritt, and D.~M. Pepin (2007), Homogenization
  of regional river dynamics by dams and global biodiversity implications,
  \textit{Proc Natl Acad Sci}, \textit{104}(14), 5732--5737,
  \doi{10.1073/pnas.0609812104}.

\bibitem[{\textit{Rauscher et~al.}(2008)\textit{Rauscher, Pal, Diffenbaugh, and
  Benedetti}}]{Rauscher:2008}
Rauscher, S.~A., J.~S. Pal, N.~S. Diffenbaugh, and M.~M. Benedetti (2008),
  Future changes in snowmelt{-}driven runoff timing over the western {US},
  \textit{Geophys Res Lett}, \textit{35}(16), \doi{10.1029/2008GL034424}.

\bibitem[{\textit{Schnorbus et~al.}(2010)\textit{Schnorbus, Werner, and
  Bennett}}]{schnorbus:2010}
Schnorbus, M., A.~Werner, and K.~Bennett (2010), \textit{Quantifying the water
  resource impacts of mountain pine beetle and associated salvage harvest
  operations across a range of watershed scales: Hydrologic modeling of the
  {F}raser {R}iver Basin}, vol. 423, Pacific Forestry Centre.

\bibitem[{\textit{Shrestha et~al.}(2012)\textit{Shrestha, Schnorbus, Werner,
  and Berland}}]{Shrestha:2012}
Shrestha, R.~R., M.~A. Schnorbus, A.~T. Werner, and A.~J. Berland (2012),
  Modelling spatial and temporal variability of hydrologic impacts of climate
  change in the {F}raser {R}iver basin, {B}ritish {C}olumbia, {C}anada, \textit{Hydrol
  Process}, \textit{26}(12), 1840--1860, \doi{10.1002/hyp.9283}.

\bibitem[{\textit{Tange}(2011)}]{Tange2011a}
Tange, O. (2011), Gnu parallel - the command-line power tool, \textit{;login:
  The USENIX Magazine}, \textit{36}(1), 42--47,
  \doi{http://dx.doi.org/10.5281/zenodo.16303}.

\bibitem[{\textit{V{\"o}r{\"o}smarty et~al.}(2010)\textit{V{\"o}r{\"o}smarty,
  {PB}, Gessner, Dudgeon, Prusevich, Green, Glidden, Bunn, Sullivan, and
  Liermann}}]{Vorosmarty:2010}
V{\"o}r{\"o}smarty, C., M.~{PB}, M.~Gessner, D.~Dudgeon, A.~Prusevich,
  P.~Green, S.~Glidden, S.~Bunn, C.~Sullivan, and D.~P. Liermann, {CR} (2010),
  Global threats to human water security and river biodiversity,
  \textit{Nature}, \textit{467}, 555--561, \doi{10.1038/nature09440}.

\bibitem[{\textit{Wang et~al.}(2012)\textit{Wang, Hamann, Spittlehouse, and
  Murdock}}]{Wang:2012}
Wang, T., A.~Hamann, D.~L. Spittlehouse, and T.~Q. Murdock (2012),
  {ClimateWNA}---high-resolution spatial climate data for {W}estern {N}orth {A}merica,
  \textit{Journal of Applied Meteorology and Climatology}, \textit{51}(1),
  16--29, \doi{10.1175/JAMC-D-11-043.1}.

\bibitem[{\textit{Wang et~al.}(2016)\textit{Wang, Hamann, Spittlehouse, and
  Carroll}}]{Wang:2016}
Wang, T., A.~Hamann, D.~Spittlehouse, and C.~Carroll (2016), Locally downscaled
  and spatially customizable climate data for historical and future periods for
  {North America}, \textit{PLoS One}, \textit{11}(6), e0156,720,
  \doi{10.1371/journal.pone.0156720}.

\bibitem[{\textit{Woodrow et~al.}(2016)\textit{Woodrow, Lindsay, and
  Berg}}]{Woodrow:2016}
Woodrow, K., J.~B. Lindsay, and A.~A. Berg (2016), Evaluating {DEM}
  conditioning techniques, elevation source data, and grid resolution for
  field-scale hydrological parameter extraction, \textit{540}, 1022--1029.

\bibitem[{\textit{Xenopoulos and Lodge}(2006)}]{Xenopoulos:2006}
Xenopoulos, M.~A., and D.~M. Lodge (2006), Going with the flow: using
  species-discharge relationships to forecast losses in fish biodiversity.,
  \textit{Ecology}, \textit{87}(8), 1907--14.

\bibitem[{\textit{Yeakel et~al.}(2014)\textit{Yeakel, Moore, Guimar{\~a}es, and
  Aguiar}}]{Yeakel:2014}
Yeakel, J., J.~Moore, P.~Guimar{\~a}es, and M.~Aguiar (2014), Synchronisation
  and stability in river metapopulation networks, \textit{Ecol Lett},
  \textit{17}(3), 273--283, \doi{10.1111/ele.12228}.

\bibitem[{\textit{Zhang and Wei}(2014)}]{Zhang:2014}
Zhang, M., and X.~Wei (2014), Alteration of flow regimes caused by
  large{-}scale forest disturbance: a case study from a large watershed in the
  interior of {British Columbia, Canada}, \textit{Ecohydrology}, \textit{7}(2),
  544--556, \doi{10.1002/eco.1374}.

\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Track Changes:
% To add words, \added{<word added>}
% To delete words, \deleted{<word deleted>}
% To replace words, \replace{<word to be replaced>}{<replacement word>}
% To explain why change was made: \explain{<explanation>} This will put
% a comment into the right margin.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% At the end of the document, use \listofchanges, which will list the
% changes and the page and line number where the change was made.

% When final version, \listofchanges will not produce anything,
% \added{<word or words>} word will be printed, \deleted{<word or words} will take away the word,
% \replaced{<delete this word>}{<replace with this word>} will print only the replacement word.
%  In the final version, \explain will not print anything.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%
\listofchanges
%%%

\end{document}

### 2017-05-30 drafts/AGU_Journal/submission 2/ms_Geo_Let_SI.tex
GPRL MS, SI and BIB

\documentclass[draft,grl]{agutexSI}

\usepackage{upgreek}
\usepackage{gensymb}
\usepackage{rotating}

%  Uncomment the following command to include .eps files
  \usepackage[pdftex]{graphicx}
%
%  Uncomment the following command to allow illustrations to print
%   when using Draft:
  \setkeys{Gin}{draft=false}

%% ------------------------------------------------------------------------ %%
%  ENTER PREAMBLE
%% ------------------------------------------------------------------------ %%

% Author names in capital letters:
\authorrunninghead{CHEZIK ET AL.}

% Shorter version of title entered in capital letters:
\titlerunninghead{RIVER NETWORKS DAMPEN CLIMATE CHANGE}

%Corresponding author mailing address and e-mail address:
\authoraddr{Corresponding author: K. A. Chezik,
Department of Biological Sciences, Simon Fraser University, 
Burnaby, BC V5A1S6, Canada.
(kchezik@sfu.ca)}

\begin{document}

%% ------------------------------------------------------------------------ %%
%  TITLE
%% ------------------------------------------------------------------------ %%

%\includegraphics{agu_pubart-white_reduced.eps}

\title{Supporting Information for RIVER NETWORKS DAMPEN LONG-TERM HYRDOLOGICAL SIGNALS OF CLIMATE CHANGE.}
%DOI: 10.1002/%insert paper number here%

%% ------------------------------------------------------------------------ %%
%  AUTHORS AND AFFILIATIONS
%% ------------------------------------------------------------------------ %%

\authors{K. A. Chezik,\altaffilmark{1} S. C. Anderson,\altaffilmark{2} J. W. Moore,\altaffilmark{1}}

\altaffiltext{1}{Earth to Ocean Research Group, Department of Biological Sciences, Simon Fraser University, 8888 University Dr., Burnaby, British Columbia V5A 1S6, Canada}
\altaffiltext{2}{School of Aquatic and Fishery Sciences, University of Washington, Box 455020, Seattle, WA 98195, USA}

%% ------------------------------------------------------------------------ %%
%  BEGIN ARTICLE
%% ------------------------------------------------------------------------ %%

\begin{article}

%% ------------------------------------------------------------------------ %%
%  TEXT
%% ------------------------------------------------------------------------ %%

\noindent\textbf{Contents of this file}

\begin{enumerate}
\item Figures S1 to S8
\end{enumerate}

\clearpage

\noindent\textbf{Introduction}

	The following supplementary material includes 8 additional figures. Figures~\ref{fig:S1}, \ref{fig:S2}, \ref{fig:S3} and \ref{fig:S8} explore the impact of timber harvest on flow in the Fraser River basin in British Columbia, CA. These figures aim to demonstrate the complicated and unclear relationship between flow trends and forest cover change. Figures \ref{fig:S6} and \ref{fig:S7} are similar to figure 4 in the main text where monthly flow trends in the Fraser River basin are shown along with trend attenuation results. These supplementary figures highlight minimum and median seasonal flow trends rather than maximum seasonal flow trends. Figures \ref{fig:S4} and \ref{fig:S5} are example plots demonstrating a single null-model iteration for the median flow metric.

%% ------------------------------------------------------------------------ %%
%  END ARTICLE
%% ------------------------------------------------------------------------ %%
\end{article}
\clearpage

\begin{sidewaysfigure}[h]
	\centering
	\setfigurenum{S1}
	\noindent\includegraphics[width=40pc]{FigS1_SiteLogModels.pdf}
	\caption{Median annual flow plotted at an annual time step between 1970 and 2007 for 55 flow gauge stations within the Fraser River basin. Points are colored by the percent of the total watershed area harvested in the previous five years. Grey (blue: sites with no harvest) lines indicate the trend in flow over time using a linear model. Light grey around each trend line represents the 95\% confidence interval.}
	\label{fig:S1}
\end{sidewaysfigure}

\begin{sidewaysfigure}[h]
	\centering
	\setfigurenum{S2}
	\noindent\includegraphics[width=40pc]{FigS2_SiteMonthLogModels.pdf}
	\caption{The response of monthly median flow trends (1970-2007) to logging in 55 sub-basins (colors) of the Fraser River watershed. Timber harvest is quantified as the percent of watershed area cut in the previous five years. Flow trends represent the mean change in median flow over time after being scaled by site and log transformed to facilitate cross site comparsion. Lines, estimated using a linear model, depict the change in flow trends with increasing timber harvest for each flow gauge station.}
	\label{fig:S2}
\end{sidewaysfigure}

\begin{figure}[h]
	\centering
	\setfigurenum{S3}
	\noindent\includegraphics[width=35pc]{FigS3_GlobalLogModel.pdf}
	\caption{The global response of median annual flow trends in the Fraser River basin to timber harvest between 1970 and 2007. Points are colored by flow gauge station. Flow trends represent the mean change in the median flow over time after being scaled by site and log transformed for cross site comparison. The black line represents the fixed effect estimate of a generalized linear mixed model with a Gamma distribution and log link function, where random intercepts were allowed to vary by flow gauge station with a single shared slope. This model describes the typical impact of harvest on median flow trends across the Fraser River basin.}
	\label{fig:S3}
\end{figure}

\begin{sidewaysfigure}[h]
	\centering
	\setfigurenum{S4}
	\noindent\includegraphics[width=40pc]{FigS4_SiteSim.pdf}
	\caption{Example site specific flow simulations. Annual median-flow simulations (yellow) and observations (blue) for all 55 sites considered in this study found within the Fraser River basin. Simulations at each site were parameterized using the observed data but no trend was imposed. Fifty-five site simulations equate to one basin-wide simulation. Each flow metric was simulated 1000 times basin-wide per response variable. See supplemental figure \ref{fig:S5} for resultant basin-wide simulation.}
	\label{fig:S4}
\end{sidewaysfigure}

\begin{figure}[h]
	\centering
	\setfigurenum{S5}
	\noindent\includegraphics[width=35pc]{FigS5_BasinSim.pdf}
	\caption{Example flow simulation results. Trend estimates (dots) $\pm$SE (vertical lines) of simulated null-model data (yellow) and the observed data upon which simulations were based (blue), plotted against watershed area. Curved lines represent the estimated variance function given the simulated data. This is a single example result of the 1000 null-model simulations produced from the data in supplemental figure \ref{fig:S4}.}
	\label{fig:S5}
\end{figure}

\begin{figure}[h]
	\centering
	\setfigurenum{S6}
	\noindent\includegraphics[width=35pc]{FigS6_MinMonth.pdf}
	\caption{Monthly minimum flow trend attenuation within the Fraser River basin. (\textbf{Left}) Fraser River's basin-wide minimum-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulations. Observed values falling further from the center of the density distribution suggest greater evidence for changes in minimum flow and a greater shift in magnitude. (\textbf{Center}) Observed monthly Fraser River minimum-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates. Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed. (\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large), for four seasonally representative months. These reflect months in the prior columns and describe the variation in percent change per decade of minimum flow among sites. Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
	\label{fig:S6}
\end{figure}

\begin{figure}[h]
	\centering
	\setfigurenum{S7}
	\noindent\includegraphics[width=35pc]{FigS7_MedMonth.pdf}
	\caption{Monthly median flow trend attenuation within the Fraser River basin. (\textbf{Left}) Fraser River's basin-wide median-flow trend estimates (i.e., intercept = vertical grey lines) by month with density distributions of null-model simulations. Observed values falling further from the center of the density distribution suggest greater evidence for changes in median flow and a greater shift in magnitude. (\textbf{Center}) Observed monthly Fraser River median-flow variance exponent ($\hat{\updelta}$, blue) and associated density distribution of simulated $\hat{\updelta}$ estimates. Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed. (\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large),for four seasonally representative months. These reflect months in the prior columns and describe the variation in percent change per decade of median flow among sites. Simulated lines ignore variance in the intercept and slope to focus visually on attenuation.}
	\label{fig:S7}
\end{figure}

\begin{sidewaysfigure}[h]
	\centering
	\setfigurenum{S8}
	\noindent\includegraphics[width=40pc]{FigS8_HarvestYear.pdf}
	\caption{Timber harvest between 1970 and 2007 within 55 sub-basins of the Fraser River watershed in British Columbia, CA. Timber harvest is quantified as the percent of each sub-basins area harvested in the previous five years. Lines are loess smoothers indicating general trends over time for each sub-basin and are colored by sub-basin drainage area.}
	\label{fig:S8}
\end{sidewaysfigure}

\end{document}

### 2017-05-30 Code_07 Area_ClimatePortfolioPlot.R
Cleaned up colors on Climate Portfolio plot

@@ -30,7 +30,7 @@ wdth = 8.7/2.54; hght = wdth*0.6
 pdf("Fig2_ClimPort.pdf", width = wdth, height = hght)
 
 LW = 1; ALS = 0.7; LL = 1.2
-par(oma = c(2,0,0,0), mar = c(0,1.8,0.2,0), family = "serif", bg = "white",fg = "white", mgp = c(2,0.2,0))
+par(oma = c(2,0,0,0), mar = c(0,1.8,0.2,0), family = "serif", bg = "white", fg = "white", mgp = c(2,0.2,0))
 plot(std.clim~sqrtArea, data = df, xlab = "", ylab = "", axes = F, xlim = c(0,sqrt(250000)), ylim = c(0,4))
 par(fg = "black")
 axis(1, lwd = LW, cex.axis = ALS, outer = T, hadj = 1,
@@ -41,10 +41,10 @@ axis(2, lwd = LW, cex.axis = ALS, las = 1)
 mtext("Climate Variability Index", side = 2, line = LL, cex = ALS, las = 0)
 points(df$sqrtArea, df$std.clim, pch = 16, cex = 0.9, col = "#000000")
 
-lines(sqrt(preds$area), preds$preds, col = "#67A9CF")
+lines(sqrt(preds$area), preds$preds, col = "black")
 upper = data.frame(se = preds$se_upper, area = sqrt(preds$area))
 lower = data.frame(se = preds$se_lower, area = sqrt(preds$area)) %>% arrange(desc(area))
 poly = bind_rows(upper,lower) %>% dplyr::select(area, se)
-polygon(poly, col = "#96969699", lty = 1, border = "#96969699")
+polygon(poly, col = "#96969699", lty = 1, border = NA)
 
 dev.off()

### 2017-05-30 Code_08 Supplementary_Plots.R
Added Forest Cover SI figures.

@@ -1,10 +1,57 @@
 #supplementary plots
-library(tidyverse);library(attenPlot)
+library(tidyverse);library(attenPlot);library(RColorBrewer);library(lme4)
+load("05_MonthDat_ClimForest.RData")
+load("05_AnnualDat_ClimForest.RData")
+
+#S1
+ggplot(Y.Data, aes(Year, Median.F, color = p5Harvest*100)) + geom_point() + 
+	facet_wrap(~Station.ID, scales = "free") + geom_smooth(method = "lm") +
+	scale_color_gradient2(name = expression(frac("%Harvest","5-Year")), midpoint=5, low="#67A9CF", mid="#F2F2F2",
+												high="#FF0000", space ="Lab") +
+	scale_x_continuous(name="Year", limits=c(1970, 2007), breaks = c(1970,1988,2007)) +
+	ylab(label = expression("Median Annual Flow (m"^3%.%"sec"^-1~")")) +
+	theme_classic(base_size = 9) +
+	theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"))
+folder = "~/sfuvault/Simon_Fraser_University/PhD_Research/Projects/River-Network-Flow-Trends/drafts/AGU_Journal/submission\ 2/"
+ggsave(paste(folder,"FigS1_SiteLogModels.pdf", sep = ""), width = 11, height = 8.5)
+
+#S2
+ggplot(M.Data, aes(p5Harvest*100, log(med.log.sd))) + geom_point(alpha = 0.5) + 
+	facet_wrap(~Month, scales = "free") + 
+	geom_smooth(aes(group = Station.ID, color = Station.ID), method ="lm", se = F) +
+	ylab(label = expression("log"["e"]~"Scaled Median Flow (m"^3%.%"sec"^-1*")"%.%"year"^-1)) +
+	xlab(expression("%Harvest"%.%"5-Year"^-1)) +
+	theme_classic(base_size = 9) +
+	theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"), legend.position = "none")
+ggsave(paste(folder,"FigS2_SiteMonthLogModels.pdf", sep = ""), width = 11, height = 8.5)
+
+#S3
+MmodGH = glmer(formula = med.log.sd~pHarvest+(1|Station.ID), family = Gamma(link = "log"), data = Y.Data)
[…]

### 2017-05-31 drafts/AGU_Journal/submission 2/ms_Geo_Let_SI.tex
Added timber harvest data reference.
Small manuscript updates to meet the requirments of the reviewer

@@
 These figures aim to demonstrate the complicated and unclear relationship between flow trends and forest cover change.
+Timber harvest data were obtained through British Columbia's Forest Practices Board, based on the Vegitation Resources Inventory (VRI) (\url{http:www.for.gov.bc.ca/hts/vri/}).
 Figures \ref{fig:S6} and \ref{fig:S7} are similar to figure 4 in the main text where monthly flow trends in the Fraser River basin are shown along with trend attenuation results.
@@
 Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed.
-(\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large), for four seasonally representative months.
+(\textbf{Right}) Trend estimates ($\hat{b}_{s}$) $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large), for four seasonally representative months.
 These reflect months in the prior columns and describe the variation in percent change per decade of minimum flow among sites.
@@
 Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed.
-(\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large),for four seasonally representative months.
+(\textbf{Right}) Trend estimates ($\hat{b}_{s}$) $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large), for four seasonally representative months.
 These reflect months in the prior columns and describe the variation in percent change per decade of median flow among sites.

### 2017-05-31 drafts/AGU_Journal/submission 2/Chezik_GeoLet_310517.tex
Small edits per reviewers comments

--- 2017-05-30 drafts/AGU_Journal/submission 2/ms_Geo_Let.tex
@@
 \begin{abstract}
-	River networks may dampen local hydrologic climate change through the aggregation of upstream climate portfolio assets.
+	River networks may dampen local hydrologic signals of climate change through the aggregation of upstream climate portfolio assets.
 Here we examine this hypothesis using flow and climate trend estimates (1970--2007) at 55 hydrometric gauge stations and across their contributing watersheds' within the Fraser River basin in British Columbia, Canada.
@@
 As billions are spent adapting to climate change by such means as building dikes or improving stormwater drainage \citep{Narain:2011}, certain habitats and processes provide a natural defense system that mitigates climate change impacts \citep{Jones:2012}.
-For example, coastal habitats such as oyster reefs, mangrove forests, and eelgrass beds shield 67\% of the United States coastline and 1.4 million people from sea level rise and storms \citep{Arkema:2013}.
+For example, coastal habitats such as oyster reefs, mangrove forests, and eelgrass beds shield 67\% of the United States coastline and 1.4 million people from sea level rise and storm impacts \citep{Arkema:2013}.
 Protection and enhancement of these natural systems has become a key component of many climate mitigation strategies \citep{Guerry:2015}, as they return multiple cost-effective services in contrast to engineered alternatives \citep{Jones:2012}.
@@
 
-Long-term changes in hydrology were less variable in larger catchments, as predicted, demonstrating between 92 and 96\% dampening of flow trends moving from headwater catchments to the network's confluence.
+Long-term changes in hydrology were less variable in larger catchments, as predicted, demonstrating between 92 and 96\% dampening of flow trends moving from headwater catchments to the network's outlet.
 In other words, the largest catchments were approximately 10 times more stable in their climate response than the smallest catchments.
@@
 	\caption{Annual flow trend attenuation within the Fraser River basin.
-(\textbf{Left}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large).
+(\textbf{Left}) Trend estimates ($\hat{b}_{s}$) $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large).
 Blue lines represent observed attenuation; orange and red represent simulated attenuation that is weaker and stronger than observed.
@@
 Decimal values represent the percent of simulated data exhibiting weaker attenuation (yellow) than observed.
-(\textbf{Right}) Trend estimates $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large), for four seasonally representative months.
+(\textbf{Right}) Trend estimates ($\hat{b}_{s}$) $\pm$ one standard error (SE, grey) plotted against watershed area (km\textsuperscript{2}), colored by climate portfolio strength (green = small, blue = large), for four seasonally representative months.
 These reflect months in the prior columns and describe the variation in percent change per decade of maximum flow among sites.
@@
 In the absence of climate dampening, the impacts of these climate extremes could have been considerably worse.
-In contrast to previously studied types of climate mitigation where habitats modify climate drivers (e.g., carbon storage) or physically absorb climate change impacts (e.g., mangroves), river network climate dampening smooths out extreme climatic trends thereby reducing the impact of local extremes.
+In contrast to previously studied types of climate mitigation where habitats modify climate drivers (e.g., carbon storage, \citep{Jones:2012}) or physically absorb climate change impacts (e.g., mangroves, \citep{Arkema:2013}), river network climate dampening smooths out extreme climatic trends thereby reducing the impact of local extremes.
 But just as the destruction of coastal habitats degrades their natural capacity to mitigate sea-level rise or storm events \citep{Arkema:2013}, the management of river basins likely impacts the ability for river networks to attenuate climate trends.
@@
 \bibitem[{\textit{Filmon}(2003)}]{Filmon:2003}
-Filmon, G. (0), Firestorm 2003.
+Filmon, G. (2003), Firestorm 2003.
 
@@
 \bibitem[{\textit{Tange}(2011)}]{Tange2011a}
-Tange, O. (2011), Gnu parallel - the command-line power tool, \textit{;login:
+Tange, O. (2011), {GNU} parallel - the command-line power tool, \textit{;login:
   The USENIX Magazine}, \textit{36}(1), 42--47,

### 2017-06-02 drafts/AGU_Journal/submission 2/ms_Geo_Let.tex
Removed 75+ words from MS to meet GPRL word limit and added a highlighted version of the MS to the .gitignore file.

@@
 \usepackage{gensymb}
+\usepackage{color,soul}
 
@@
 \begin{abstract}
-	River networks may dampen local hydrologic climate change through the aggregation of upstream climate portfolio assets.
+	River networks may dampen local hydrologic signals of climate change through the aggregation of upstream climate portfolio assets.
 Here we examine this hypothesis using flow and climate trend estimates (1970--2007) at 55 hydrometric gauge stations and across their contributing watersheds' within the Fraser River basin in British Columbia, Canada.
-Using a null hypothesis framework we compared our observed attenuation of river flow trends as a function of increasing area and climate trend diversity, with null-simulated estimates to gauge the likelihood and strength of our observations.
+Using a null hypothesis framework, we compared our observed attenuation of river flow trends as a function of increasing area and climate trend diversity, with null-simulated estimates to gauge the likelihood and strength of our observations.
 We found the Fraser River reduced variability in downstream long-term discharge by $>$91\%, with $>$3.1 times the attenuation than would be expected under null-simulation.
@@
 \includegraphics[width=35pc]{Fig1_Map.pdf}
-	\caption{Climate trends (1970 - 2007) in mean annual temperature (MAT) and mean annual precipitation (MAP) within the Fraser River basin overlayed on a digital elevation model of British Columbia, Canada.
+	\caption{Climate trends (1970 - 2007) in mean annual temperature (MAT) and mean annual precipitation (MAP) within the Fraser River basin overlaid on a digital elevation model of British Columbia, Canada.
 Flow gauge sites (dots) are scaled by the size of the contributing area.
@@
 As with other mid-latitude rivers \citep{Bindoff:2013}, the Fraser River's discharge volume and variability have increased over the last several decades \citep{Dery:2012,Morrison:2002}.
-Following projected pattens of climate change, the Fraser River basin is increasingly shifting from a snow dominated to rain-snow hybrid system \citep{Kang:2016}.
+Following projected patterns of climate change, the Fraser River basin is increasingly shifting from a snow dominated to rain-snow hybrid system \citep{Kang:2016}.
 Using 38 years of flow data collected concurrently at 55 hydrometric gauge stations throughout the watershed, we use a novel analytical approach to examine whether this large river exhibits signals of long-term climate dampening.
@@
 Because dams control flow and stabilize flow trends, we removed sites where daily flow exhibited unusually high autocorrelation and low variability relative to other sites of similar size, and sites that were observed downstream of, and atypically close to, a dam (\textit{n}=3).
-The downstream distance of gauge stations from dams was determined using GIS spatial layers available in the \textit{BC Data Catalog}, maintained by the provincial government of British Columbia.
+The downstream distance of gauge stations from dams was determined using GIS spatial layers in the \textit{BC Data Catalog}, maintained by the BC provincial goverment.
 
@@
 We used maximum- and minimum-flow estimates to capture trends in extreme discharge and median-flow estimates to capture general annual and seasonal trends.
-Because the timing of flow is particularly important to the phenology of many organisms, we also calculated the day-of-year in which half the annual flow was reached.
+Because timing of flow is particularly important to the phenology of many organisms, we also calculated the day-of-year in which half the annual flow was reached.
 
@@
 Digital elevation models were provided by the provincial government of British Columbia, Canada and state government of Washington, USA and were pre-processed with a breaching algorithm and stream burned to facilitate proper flow path and accumulation models \citep{Woodrow:2016}.
-We summarized the climate variability integrated at each flow gauge by calculating the standard deviation in climate trends within their catchments.
+We summarized the climate variability integrated at each flow gauge by calculating the standard deviation in climate trends within their catchments using the \textit{rasterstats} module (\url{https://github.com/perrygeo/python-rasterstats.git}) in Python v2.7.
 The standard deviations of these climate variables were scaled between 0 and 1 and summed as a general climate index for each catchment.
@@
 Captured in our climate portfolio proxy variable ($\mathrm{A}$), is the interaction between climate and landuse.
-Undoubtedly, timber harvest, the dominant source of land cover change in the Fraser River basin, scales with sub-basin area and contributes to changes in flow.
+Timber harvest, the dominant source of land cover change in the Fraser River basin, scales with sub-basin area and contributes to changes in flow.
 Although exploratory linear models at the site level do suggest an effect of timber harvest on long-term flow trends for some gauge stations (Figure~S1, S2), when information is shared across sites in a global model, climate effects dominate (Figure~S3).
@@
 
-Long-term changes in hydrology were less variable in larger catchments, as predicted, demonstrating between 92 and 96\% dampening of flow trends moving from headwater catchments to the network's confluence.
-In other words, the largest catchments were approximately 10 times more stable in their climate response than the smallest catchments.
+Long-term changes in hydrology were less variable in larger catchments, as predicted, demonstrating between 92 and 96\% dampening of flow trends moving from headwater catchments to the network's outlet.
+The largest catchments were approximately 10 times more stable in their climate response than the smallest catchments.
 For instance, trends ranged between a 6\% reduction and 19\% increase in median flow per decade among small watersheds, while large watersheds ranged between 0.8 and 1.6\% change per decade.
@@
 Our null model approach illustrates that support for a river network portfolio effect is greater than 90\% for all four annual flow trend metrics and as high as 98\% for median-annual-flow trends (Figure~\ref{fig:3}).
-Trend attenuation was; 3.1 (0.6--9.8 this and hereafter are 90\% intervals), 4.7 (1.1--14.7), 3.2 (0.9--9.0) and 5.2 (1.4--14.4) times greater than the null model for annual-flow timing, and minimum, maximum and median annual-flow, respectively.
+Trend attenuation was 3.1 (0.6--9.8 this and hereafter are 90\% intervals), 4.7 (1.1--14.7), 3.2 (0.9--9.0) and 5.2 (1.4--14.4) times greater than the null model for annual-flow timing, and minimum, maximum and median annual-flow, respectively.
 These statistics provide strong support for the hypothesis that river network portfolio effects contribute to climate dampening in large rivers.
@@
 Overall, climate portfolios tended to have greater attenuation during stable periods of the year and decreased during seasonal transitions.
-In the spring, the likelihood of network-driven attenuation of maximum flow trends was as low as 24\% (Figure~\ref{fig:4}).
-The deterioration of the network's attenuation strength was coincident with the spring freshet, when snowmelt drives high flows across the basin.
+In spring, the likelihood of network-driven attenuation of maximum flow trends was as low as 24\% (Figure~\ref{fig:4}).
+Deterioration of the network's attenuation strength was coincident with the spring freshet, when snowmelt drives high flows across the basin.
 Synchronization events such as the freshet may subvert river network dampening mechanisms by homogenizing the region's response to seasonally driven climate shifts.
@@
 The Fraser River basin has seen changes in landcover between 1970 and 2007, which have impacted river discharge \citep[e.g.,][]{Zhang:2014}.
-Over this period, timber harvest has been the dominant contributor to changes in forest cover but increasingly natural events are having a greater impact (Figure~S8).
+Over this period, timber harvest has been the dominant contributor to changes in forest cover (Figure~S8) but increasingly natural events are having a greater impact.
 For instance, abnormally hot, dry weather in 2003 led to fires that destroyed an unprecedented 2,600 km\textsuperscript{2} of forest largely within the Fraser River basin \citep{Filmon:2003}.
@@
 In the absence of climate dampening, the impacts of these climate extremes could have been considerably worse.
-In contrast to previously studied types of climate mitigation where habitats modify climate drivers (e.g., carbon storage) or physically absorb climate change impacts (e.g., mangroves), river network climate dampening smooths out extreme climatic trends thereby reducing the impact of local extremes.
+In contrast to previously studied types of climate mitigation where habitats modify climate drivers (e.g., carbon storage, \citep{Jones:2012}) or physically absorb climate change impacts (e.g., mangroves, \citep{Arkema:2013}), river network climate dampening smooths out extreme climatic trends thereby reducing the impact of local extremes.
 But just as the destruction of coastal habitats degrades their natural capacity to mitigate sea-level rise or storm events \citep{Arkema:2013}, the management of river basins likely impacts the ability for river networks to attenuate climate trends.
-For example, dams synchronize the flow regimes of rivers \citep{Poff:2007} and logging increases the magnitude and frequency of extreme events such as high and low flows through increased runoff rates \citep{Zhang:2014}.
+For example, dams synchronize the flow regimes of rivers \citep{Poff:2007}.
 These anthropogenic activities may magnify the impacts of climatic shifts such as the transition from snow- to rain-dominated precipitation in some basins.
@@
 \acknowledgments
-This paper was possible only through the extensive use of many open source, freely available tools.
-All flow data analysis was done using R v3.3.2 (\url{www.r-project.org/}) and all R code can be found at \url{http://github.com/kchezik/River-Network-Flow-Trends.git}.
-Spatial analysis of watersheds was done using GDAL/OGR v1.11.5, and Whitebox GAT v3.4 \citep{Lindsay:2016} with visualization help from QGIS v2.18.2 (\url{www.qgis.org/}).
-Climate data were gathered using ClimateWNA v5.4 \citep{Wang:2016} and summary statitics were calculated using R and the \textit{rasterstats} module (\url{https://github.com/perrygeo/python-rasterstats.git}) in Python v2.7.
+This paper relied on many open source tools including R v3.3.2 (\url{www.r-project.org/}), Python v2.7, QGIS v2.18 and WhiteboxGAT \citep{Lindsay:2016}.
+All R code can be found at \url{http://github.com/kchezik/River-Network-Flow-Trends.git}.
 Data processing was facilitated greatly by GNU Parallel \citep{Tange2011a}.
-This study could not have been done without the decades of flow data collected by those at Environment Canada.
+This study relied on decades of flow data collected by those at Environment Canada.
 Raw HYDAT data can be obtained via Environment Canada or a simplified version can be found on GitHub.

### 2017-06-29 Code_05 Monthly_Plot.R
Small aesthetic changes to plot for final submission to GeoPhysResLet

@@ -2,148 +2,149 @@
 ################## Plot Monthly Flow Trends Against Area and Display Climate Index Data #################
 #########################################################################################################
 
-library(attenPlot)#Funnel and Density plot function library.
+library(attenPlot)#funnel and density plot function library.
 
-#Density Plot Wrapper
+#density plot wrapper
 density_plot = function(data, response, type, xlimit){
 	
-	############################################ Prepare Data ###############################################
-	#Libraries to load.
+	############################################ prepare data ###############################################
+	#libraries to load.
 	library(tidyverse);library(wesanderson);library(RColorBrewer);library(lubridate)
-	#Control Panel
-	L.Width = 0.5; Axis.Lab.Size = 0.7; Lab.Loc = 2.1; vLine = 2
-	#Colour Control Panel
-	col1 = if_else(type == "int","#000000","#F2AD00"); col2= if_else(type=="int","#525252","#67A9CF")
-	col3="#F98400"; col4="#FF0000"; lcol="#969696"; shade = "99"
-	#Determine if we are plotting intercept (int) or exponent (exp) data and set boundaries accordingly.
+	#control panel
+	l.width = 0.5; axis.lab.size = 0.7; lab.loc = 2.1; vline = 2
+	#colour control panel
+	col1 = if_else(type == "int","#000000","#f2ad00"); col2= if_else(type=="int","#525252","#67a9cf")
+	col3="#f98400"; col4="#ff0000"; lcol="#969696"; shade = "99"
+	#determine if we are plotting intercept (int) or exponent (exp) data and set boundaries accordingly.
 	if(type == "int") par(oma = c(3,1.5,0.2,0.2), mar = c(0,0,0.1,0.3), mgp = c(3,0.5,0), family = "serif")
 	else par(mar = c(0,0.3,0.1,0), family = "serif")
-	#Create list of month labels to iterate through.
-	Months = as.character(lubridate::month(as.numeric(names(data)), label = T)); count = 1
+	#create list of month labels to iterate through.
+	months = as.character(lubridate::month(as.numeric(names(data)), label = T)); count = 1
 	
 	plyr::l_ply(data,function(x){
[…]

### 2017-06-29 Code_07 Area_ClimatePortfolioPlot.R

@@ -41,10 +41,10 @@ axis(2, lwd = LW, cex.axis = ALS, las = 1)
 mtext("Climate Variability Index", side = 2, line = LL, cex = ALS, las = 0)
 points(df$sqrtArea, df$std.clim, pch = 16, cex = 0.9, col = "#000000")
 
-lines(sqrt(preds$area), preds$preds, col = "black")
-upper = data.frame(se = preds$se_upper, area = sqrt(preds$area))
-lower = data.frame(se = preds$se_lower, area = sqrt(preds$area)) %>% arrange(desc(area))
-poly = bind_rows(upper,lower) %>% dplyr::select(area, se)
-polygon(poly, col = "#96969699", lty = 1, border = NA)
+#lines(sqrt(preds$area), preds$preds, col = "black")
+#upper = data.frame(se = preds$se_upper, area = sqrt(preds$area))
+#lower = data.frame(se = preds$se_lower, area = sqrt(preds$area)) %>% arrange(desc(area))
+#poly = bind_rows(upper,lower) %>% dplyr::select(area, se)
+#polygon(poly, col = "#96969699", lty = 1, border = NA)
 
 dev.off()

### 2017-06-29 Code_08 Supplementary_Plots.R

@@ -1,5 +1,5 @@
 #supplementary plots
-library(tidyverse);library(attenPlot);library(RColorBrewer);library(lme4)
+library(tidyverse);library(attenPlot);library(RColorBrewer);library(lme4);library(zoo);library(lubridate)
 load("05_MonthDat_ClimForest.RData")
 load("05_AnnualDat_ClimForest.RData")
 
@@ -13,7 +13,7 @@ ggplot(Y.Data, aes(Year, Median.F, color = p5Harvest*100)) + geom_point() +
 	theme_classic(base_size = 9) +
 	theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"))
 folder = "~/sfuvault/Simon_Fraser_University/PhD_Research/Projects/River-Network-Flow-Trends/drafts/AGU_Journal/submission\ 2/"
-ggsave(paste(folder,"FigS1_SiteLogModels.pdf", sep = ""), width = 11, height = 8.5)
+ggsave(paste(folder,"FigS1_SiteLogModels.pdf", sep = ""), width = 11, height = 7.5)
 
 #S2
 ggplot(M.Data, aes(p5Harvest*100, log(med.log.sd))) + geom_point(alpha = 0.5) + 
@@ -23,12 +23,12 @@ ggplot(M.Data, aes(p5Harvest*100, log(med.log.sd))) + geom_point(alpha = 0.5) +
 	xlab(expression("%Harvest"%.%"5-Year"^-1)) +
 	theme_classic(base_size = 9) +
 	theme(plot.margin = unit(c(0.1,0.1,0.1,0.1), "cm"), legend.position = "none")
-ggsave(paste(folder,"FigS2_SiteMonthLogModels.pdf", sep = ""), width = 11, height = 8.5)
+ggsave(paste(folder,"FigS2_SiteMonthLogModels.pdf", sep = ""), width = 11, height = 7.5)
 
 #S3
 MmodGH = glmer(formula = med.log.sd~pHarvest+(1|Station.ID), family = Gamma(link = "log"), data = Y.Data)
 temp = bind_cols(Y.Data, data.frame(fixed = predict(MmodGH, Y.Data)))
-ggplot(temp, aes(pHarvest*100, log(med.log.sd), color = Station.ID)) + geom_point(alpha = 0.5) + 
[…]

### 2017-06-29 drafts/AGU_Journal/submission 2/response.md
Final draft of GeoPhysResLet submission. Added reviewers comments, updated plots and added a reference.

# River networks dampen long-term hydrological signals of climate change
## K. A. Chezik, S. C. Anderson, J. W. Moore
### 26-Jun-2017

### Response to Reviewer 1

1. **Comment:** (Fig. 1 caption): "within the Fraser River basin overlaid on a digital elevation model of British Columbia" The DEM is not of British Columbia but only of part of it. I would instead suggest either, "within the Fraser River basin in British Columbia overlaid on a digital elevation "model" or "within the Fraser River basin overlaid on a digital elevation model within British Columbia"

**Response:** The reviewer's second suggestion for clarity was applied to the manuscript. We now make clear the DEM is *within* BC, Canada.

2. **Comment:** Figure 1: I still think it would be nice to show the location of dams on this figure, but maybe too distracting for the key message of the paper.

**Response:** We understand the reviewer's concern here but we do feel adding dams to the map would distract and even obscure the information pertinent to understanding the studies findings. Largely, plotting the location of dams adds little information beyond assuring the reader visually that dams of significant size are not in proximity to our sites, but in many instances, without being able to zoom to specific sites and streams, the map would be unclear if a site and dam are on the same flow path. We do feel we have provided well commented code on GitHub that will assure the curious/skeptical reader that we have done our due diligence on this topic. Furthermore, the location of our sites and the spatial layers are freely available and can be accessed for analysis.

3. **Comment:** L 145: "climate_s,t" I think should be "C_s,t" to parallel the description of equation 1.

**Response:** We thank the reviewer for catching this mistake. We have corrected the equation per the reviewer's suggestion.

4. **Comment:** L 164: "can have" would be more accurate, because one could think of a large catchment in a different landscape not having a diverse climate portfolio (agricultural Midwest). Or "here have" to be specific that you are talking about the Fraser Basin.

**Response:** We have made it clear that we are speaking about the Fraser River basin by using "*here have*" per the reviewers suggestion.

5. **Comment:** L167: say "GLS model" as you defined GLS earlier.

**Response:** Replaced generalized least squares with "*GLS*" per the reviewers suggestion.

6. **Comment:** L 172: Why 5700 km^2^? I assume this is the mean watershed area of the sites you are considering but the text does not make this clear.

**Response:** Yes the reviewer is correct, the y-intercept is at the mean watershed sqrt-area. We adjusted the language to make clear that the centered response variable corresponds to the y-intercept at ~5700 km^2^.

### Response to Reviewer 2

**General Comments:** The presented analysis represents a robust synthesis of substantial amounts of hydrological data. What is not always clear within the analysis though are the physical justifications for the chosen metrics, and instead the analysis seems to be almost entirely driven by regression of descriptive metrics that are not necessarily physically justified. As an example, if the yearly probability density function for the rivers is well approximated by a log-normal or similar functional form then the median is a physically justified value, however if the underlying probability density function is better represented by an anomalous distribution then the median does not really capture the physics of the process. This limits the transferability of the model to other catchments and the extent to which we may treat the analysis as predictive rather then descriptive. Another example of this is the use of monthly statistics which while informative and interesting to look at don't represent time periods relevant for river or climatic processes. In this case, figure 4 would be more informative if binned by season, rather then month (from the monthly data there appear to be two groups Oct-April and June-Sept., with May being a bit of an outlier). As a summary, the statistical regression models are only useful if the descriptive statistics they are based on capture the physics of the system in question. The chosen parameters may very well be the best parameters for this analysis, however it was not clear to me when reading the manuscript if they were indeed the correct parameters to describe the data. It would be beneficial if the authors could clarify their reasoning on this, either in the main text or supplementary material.

**Response:** Our choice of response metrics was largely driven by what we have observed in the literature and what metrics would be of interest to a wider audience. As described on lines 102 to 111, extremes in flow and the timing of flow are important metrics to humans and species. Because the Fraser is still a snow driven system, the shape of the annual discharge curve is largely driven by the spring freshet and dry summers, with cross site variation occurring largely in the late fall and winter months depending on coastal or interior location. Therefore, the reason we took the median flow value was that it was more robust to inter-annual fluctuations in precipitation than the mean across sites. We include a new plot in the supplementary information showing the distribution of flow for each year, by site (Figure S9). This should make clear that the distribution of the flow data underlying our summary metrics is largely normal or captured by a log-normal distribution because each year has a single distinct peak in flow.

1. **Comment:** Key point 2: How large is a 'large sub-catchment'? It might just be better to say 'larger' if the effect does indeed just generally scale with sub-catchment size.

**Response:** We made the change per the reviewers suggestion. '*Large*' has become '*Larger*' to demonstrate a relative sense of the word.

2. **Comment:** Ln. 89: Why a five day rolling window? Could you provide the reasoning behind this choice.

**Response:** The five day rolling average was used simply because it is a period of timing commonly applied as a smoothing function to reduce the impacts of extreme and unusual or erroneous events. An example reference by Dery et al. (2009) has been added.

3. **Comment:** Ln. 96: Why monthly response variables? If your time series has no structure on a monthly then this seems arbitrary.

**Response:** We did not bin by season because we feel the seasonal effect is more intuitive in a monthly format. In figure 4, we are able to more simply see the continuity of seasonal shifts if we have a more fine scale time variable than if we simply have two bins. Arguably, this format also allows for more informative cross site comparisons if another system were to be studied or if this system were to be studied at different periods, allowing for the observation of timing shifts in the seasonal waxing and waning of dampening.

4. **Comment:** Ln. 168. could you explain the reasoning behind taking the square root of catchment area? It is not clear why this was done. Is this supposed to represent a length scale for each site assuming an idealized geometry?

**Response:** We used the square root to spread out the catchments that are relatively small and reduce the leverage of the very large catchments. Also, given that area is in squared units, it makes sense to square root this response variable rather than log-transform or use some other transformation.

5. **Comment:** Equation 4: Please define delta.
6. **Comment:** Ln. 208: Please define delta^hat.

**Response:** Added a clarifying statement describing delta which should also make clear the meaning of delta_hat.

7. **Comment:** Ln. 234: It isn't clear what is meant by global climate here, the change for these sites is still regional (the data still represent a single river basin that is not particularly large by river standards).

**Response:** The reference to "*global climate*" is saying that while climate is variable from place to place, general shifts in global (i.e., general or mean temperature) climate (Holocene to Anthropocene), result in a variety of local climate responses. See list of changes to see how we have clarified this sentence.

8. **Comment:** Figure 2. What is the point of the fitted trend line? Is it just to guide the eye? The fitted trend line (if we are to take it at face value) suggests that the asymptotic behavior occurs for catchments substantially larger 40,0000 km^2^, closer to 250,000 km^2^. The data are more in line with a value of 40,000 km^2^, but then why include the fitted line?

**Response:** The fitted line is simply to guide the eye and provide for some sense of uncertainty in the observed relationship. This is not necessary and is relatively obvious. We have removed the fitted line.

### 2017-06-29 drafts/AGU_Journal/submission 2/ms_Geo_Let.tex
Final draft of GeoPhysResLet submission. Added reviewers comments, updated plots and added a reference.

@@
 %% To submit your paper:
-\documentclass[draft,linenumbers]{AGUJournal}
-\drafttrue
+%\documentclass[draft,linenumbers]{AGUJournal}
+%\drafttrue
+
+%% For final version.
+\documentclass{AGUJournal}
 
@@
 \usepackage{color,soul}
-
-%% For final version.
-% \documentclass{AGUJournal}
 
@@
 \item River networks dampen long-term flow trends by integrating varied climate responses across complex landscapes.
-\item Large sub-catchments integrate greater climate trend diversity, dampening the local extremes of climate change.
+\item \replaced{Large}{Larger} sub-catchments integrate greater climate trend diversity, dampening the local extremes of climate change.
 \item Rivers may possess an under-appreciated capacity, and provide a tool, for mitigating the impacts of climate change.
@@
 \includegraphics[width=35pc]{Fig1_Map.pdf}
-	\caption{Climate trends (1970 - 2007) in mean annual temperature (MAT) and mean annual precipitation (MAP) within the Fraser River basin overlaid on a digital elevation model of British Columbia, Canada.
+	\caption{Climate trends (1970 - 2007) in mean annual temperature (MAT) and mean annual precipitation (MAP) within the Fraser River basin overlaid on a digital elevation model \replaced{of}{within} British Columbia, Canada.
 Flow gauge sites (dots) are scaled by the size of the contributing area.
@@
 These daily data were smoothed using a five-day rolling average before summarizing into annual and monthly response variables.
-This smoothing technique was used to reduce the influence of erroneous and unusually extreme values.
+This smoothing technique was used to reduce the influence of erroneous and unusually extreme values \citep [e.g.,][]{Dery:2009}.
 Because dams control flow and stabilize flow trends, we removed sites where daily flow exhibited unusually high autocorrelation and low variability relative to other sites of similar size, and sites that were observed downstream of, and atypically close to, a dam (\textit{n}=3).
@@
 where $\mathrm{C}_{s,t}$ represents a climate variable and $\mathrm{Y}_{s,t}$ represents the time in years (e.g., 1970, 1971, \ldots, 2007).
-Parameters $\beta_{s}$ (slope) and $\alpha_{s}$ (intercept) represent the estimated mean effect of time on $\mathrm{C}_{s,t}$ and the estimate of $\mathrm{climate}_{s,t}$ at time zero, respectively.
+Parameters $\beta_{s}$ (slope) and $\alpha_{s}$ (intercept) represent the estimated mean effect of time on $\mathrm{C}_{s,t}$ and the estimate of $\mathrm{C}_{s,t}$ at time zero, respectively.
 Autocorrelated errors were incorporated as in eq.~\ref{eq1}.
@@
 We hypothesize that flow gauge sites with larger climate portfolios will exhibit greater long-term flow trend stability than sites with smaller climate portfolios.
-We used catchment area as a proxy for the climate portfolio diversity of a given site, given that larger catchments have more variable climate portfolios (Figure~\ref{fig:2}) as well as aggregate more water and carry more weight on downstream dampening.
-Therefore, to test whether networks dampen climate signals, we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a generalized least squares model:
+We used catchment area as a proxy for the climate portfolio diversity of a given site, given that larger \added{Fraser River sub-}catchments have more variable climate portfolios (Figure~\ref{fig:2}) as well as aggregate more water and carry more weight on downstream dampening.
+Therefore, to test whether networks dampen climate signals, we regressed site-specific trend estimates (e.g., \% change$\cdot$decade\textsuperscript{-1}) onto each site's catchment area using a \replaced{generalized least squares}{GLS} model:
 \begin{linenomath*}
@@
 We adjusted $\sqrt{\mathrm{A}_{s}}$ by subtracting the mean $\sqrt{\mathrm{A}}$, thereby centering our predictor.
-Fitted $d$ and $c$ parameters represent the mean effect of watershed area ($\sqrt{\mathrm{A}}$), on flow trends and the mean flow trend for a $\sim$5,700 km\textsuperscript{2} watershed, respectively.
+Fitted $d$ and $c$ parameters represent the mean effect of watershed area ($\sqrt{\mathrm{A}}$), on flow trends and the mean flow trend for \replaced{a}{an average sized watershed (i.e.,} $\sim$5,700 km\textsuperscript{2}\added{)} \deleted{watershed}, respectively.
 Captured in our climate portfolio proxy variable ($\mathrm{A}$), is the interaction between climate and landuse.
@@
 \end{linenomath*}
-where the variance of the estimated error ($\eta_{s}$) changes exponentially with increasing $\sqrt{\mathrm{A}}$.
-This variance function allowed us to predict the range of flow trend values that would be expected as watershed area increased.
+where the variance of the estimated error ($\eta_{s}$) \deleted{changes} exponentially \replaced{increasing}{increases} with $\sqrt{\mathrm{A}}$ \added{as defined by the estimated $\updelta$ parameter.} This variance function allowed us to predict the range of flow trend values that would be expected as watershed area increased.
 A decreasing range of flow trends with increasing watershed area indicates a dampening network effect.
@@
 Thus, rivers that drain larger catchments integrate a greater diversity of shifting climates.
-These findings suggest that larger areas contain greater landscape complexity resulting in a greater variety of climatic expressions that respond to changes in global climate differently.
+These findings suggest that larger areas contain greater landscape complexity, \replaced{resulting in a greater variety of climatic expressions that respond to changes in global climate differently.}{which produce a variety of climates locally and respond to global climate shifts differently.}
 
@@
 	\caption{Climate variability index as a function of catchment area.
-The mean trend line (black) was estimated using robust linear regression and standard error estimates (SE) (2$\cdot$SE = grey) were calculated using the weighted standard deviation estimate.}
+\replaced{The mean trend line (black) was estimated using robust linear regression and standard error estimates (SE)(2$\cdot$SE=grey) were calculated using the weighted standarad deviation estimate.}{As watershed area increases the variety of climatic trends within the watershed increases at a diminishing rate.}}
 \label{fig:2}
@@
 
+\bibitem[{\textit{D{\'e}ry et~al.}(2009)\textit{D{\'e}ry, Stahl, Moore,
+  Whitfield, Menounos, and Burford}}]{Dery:2009}
+D{\'e}ry, S.~J., K.~Stahl, R.~Moore, Whitfield, B.~Menounos, and J.~E. Burford
+  (2009), Detection of runoff timing changes in pluvial, nival, and glacial
+		rivers of western canada, \textit{Water Resour Res}, \textit{45}(4),
+  \doi{10.1029/2008WR006975}.
+
 \bibitem[{\textit{D{\'e}ry et~al.}(2012)\textit{D{\'e}ry,
@@
   \doi{10.1088/1748-9326/7/2/024019}.
-
-\bibitem[{\textit{Do~Hyuk~Kang et~al.}(2016)\textit{Do~Hyuk~Kang, Shi,
-  ul~Islam, and D{\'e}ry}}]{Kang:2016}
-Do~Hyuk~Kang, H.~G., X.~Shi, S.~ul~Islam, and S.~J. D{\'e}ry (2016), Impacts of
-		a rapidly declining mountain snowpack on streamflow timing in {C}anada{'}s
-  {Fraser River} basin, \textit{Scientific reports}, \textit{6}.
 
@@
   of snow to the hydrology of the {F}raser {R}iver basin, {C}anada.
+
+\bibitem[{\textit{Kang et~al.}(2016)\textit{Kang, Shi,
+  Islam, and D{\'e}ry}}]{Kang:2016}
+Kang, D., H.~Gao, X.~Shi, S.~ul~Islam, and S.~J. D{\'e}ry (2016), Impacts of
+		a rapidly declining mountain snowpack on streamflow timing in {C}anada{'}s
+  {Fraser River} basin, \textit{Scientific {R}eports}, \textit{6}.
 

### 2017-06-29 drafts/AGU_Journal/submission 2/ms_Geo_Let_SI.tex
Final draft of GeoPhysResLet submission. Added reviewers comments, updated plots and added a reference.

@@
-\documentclass[draft,grl]{agutexSI}
+%\documentclass[draft,grl]{agutexSI}
+\documentclass[grl]{agutexSI}
 
@@
 
-	The following supplementary material includes 8 additional figures.
+	The following supplementary material includes 9 additional figures.
 Figures~\ref{fig:S1}, \ref{fig:S2}, \ref{fig:S3} and \ref{fig:S8} explore the impact of timber harvest on flow in the Fraser River basin in British Columbia, CA.
@@
 Figures \ref{fig:S4} and \ref{fig:S5} are example plots demonstrating a single null-model iteration for the median flow metric.
+Figure \ref{fig:S9} shows annual discharge curves at each flow gauge station in the Fraser River basin.
+This figure demonstrates that our flow data are well approximated by normal and lognormal distributions which suggests median, maximum, minimum flow metrics and our flow timing variable (i.e., day-of-year to half annual flow), are descriptive of the system and unlikely to be biased.
 
@@
 	\setfigurenum{S1}
-	\noindent\includegraphics[width=40pc]{FigS1_SiteLogModels.pdf}
+	\noindent\includegraphics[width=55pc]{FigS1_SiteLogModels.pdf}
 	\caption{Median annual flow plotted at an annual time step between 1970 and 2007 for 55 flow gauge stations within the Fraser River basin.
@@
 	\setfigurenum{S2}
-	\noindent\includegraphics[width=40pc]{FigS2_SiteMonthLogModels.pdf}
+	\noindent\includegraphics[width=55pc]{FigS2_SiteMonthLogModels.pdf}
 	\caption{The response of monthly median flow trends (1970-2007) to logging in 55 sub-basins (colors) of the Fraser River watershed.
@@
 	\setfigurenum{S3}
-	\noindent\includegraphics[width=35pc]{FigS3_GlobalLogModel.pdf}
+	\noindent\includegraphics[width=45pc]{FigS3_GlobalLogModel.pdf}
 	\caption{The global response of median annual flow trends in the Fraser River basin to timber harvest between 1970 and 2007.
@@
 	\setfigurenum{S4}
-	\noindent\includegraphics[width=40pc]{FigS4_SiteSim.pdf}
+	\noindent\includegraphics[width=52pc]{FigS4_SiteSim.pdf}
 	\caption{Example site specific flow simulations.
@@
 	\setfigurenum{S5}
-	\noindent\includegraphics[width=35pc]{FigS5_BasinSim.pdf}
+	\noindent\includegraphics[width=45pc]{FigS5_BasinSim.pdf}
 	\caption{Example flow simulation results.
@@
 	\setfigurenum{S6}
-	\noindent\includegraphics[width=35pc]{FigS6_MinMonth.pdf}
+	\noindent\includegraphics[width=40pc]{FigS6_MinMonth.pdf}
 	\caption{Monthly minimum flow trend attenuation within the Fraser River basin.
@@
 	\setfigurenum{S7}
-	\noindent\includegraphics[width=35pc]{FigS7_MedMonth.pdf}
+	\noindent\includegraphics[width=40pc]{FigS7_MedMonth.pdf}
 	\caption{Monthly median flow trend attenuation within the Fraser River basin.
@@
 	\setfigurenum{S8}
-	\noindent\includegraphics[width=40pc]{FigS8_HarvestYear.pdf}
+	\noindent\includegraphics[width=55pc]{FigS8_HarvestYear.pdf}
 	\caption{Timber harvest between 1970 and 2007 within 55 sub-basins of the Fraser River watershed in British Columbia, CA.
@@
 
+\begin{sidewaysfigure}[h]
+	\centering
+	\setfigurenum{S9}
+	\noindent\includegraphics[width=55pc]{FigS9_FlowCurvesRaw.pdf}
+	\caption{Smoothed, five-day rolling average flow rates for each day in each year between 1970 and 2007 at 55 flow gauge stations in the Fraser River watershed.
+Years are colored oldest (red) to most recent (blue).
+The distributions demonstrate that flow in the Fraser River basin is defined by a spring melt and typically low summer flows.
+Flow timing, duration and magnitude vary between sites but are largely grouped by coastal and interior locations.
+For instance, sites 08MH006, 08MH029, 08MH076 and 08MH090 experience an earlier freshet and higher fall and winter flows due to their coastal proximity.
+}
+	\label{fig:S9}
+\end{sidewaysfigure}
+
 \end{document}
